### 2014-10-23 intro.tex
edits

\section{Introduction}

Many real-world web services --- such as Amazon, Facebook, Google, Twitter ---
replicate application state and logic across multiple \emph{replicas} within
and across data centers. Replication is intended not only to improve
application throughput and reduce user perceived latency, but also to tolerate
partial failures without compromising the service availability. Traditionally
programmers have relied on \emph{strong consistency} guarantees such as
Linerarizability~\cite{Herlihy1990} or serializability~\cite{Serializability}
in order to build correct applications. While strong consistency is simple to
reason about, the CAP theorem~\cite{Brewer2000,Gilbert} makes it impossible for
a system to achieve strong consistency, tolerate network partions and remain
highly available. Thus, modern web services, which aim to provide an "always
on" experience, overwhelmingly embrace availability and partition tolerance
over strong consistency. Several \emph{weak consistency} models such as
eventual consistency, causal consistency, session guarantees, and timeline
consistency have been proposed.

Under weak consistency, the programmer needs to be aware of concurrent
conflicting updates, and requires careful attention to avoid inconsistencies
such as negative balances in a bank account, and an item appearing in the
shopping cart after it has been removed~\cite{Dynamo}. Often times, the
inconsistency leaks from the application and is witnessed by the user. To
address this issue, many commercial~\cite{} and research~\cite{} systems
propose guarantees that are weaker than strong consistency, and hence need only
to contact a subset of replicas for discharging the operations. Indeed,
traditional database systems do provide a variety of transaction isolation
levels~\cite{}, each with a particular runtime overhead. With these
alternatives, it is upto the programmer to decide whether a weaker consistency
level is sufficient for a preserving the application-level consistency
properties. Understandably, this is an error prone process as the user not only
has to understand the intricate behavior of the application, but also the
semantics of the underlying store consistency levels, which are often presented
informally~\cite{}.

Moreover, weak consistency hinders compositional reasoning about programs.
While the application might be naturally expressed as maps, trees, counters,
queues, or graphs, geo-distributed stores are typically only provide a minimal
set of data types with in-built conflict resolution strategy such as
last-writer-wins (LWW) register, counters and sets~\cite{Cassandra,DynamoDB}.
Unlike traditional database systems, the geo-distributed stores typically lack
unrestricted transactional access to the data. Thus, the application developer
often has to coerce the application state to the data store capabilities.

To address these issues, we introduce Quelea, a declarative programming model
for eventually consistent stores. The key novelty of Quelea is a \emph{contract
language} to express the fine-grained application-level consistency properties.
The programmer uses the contract language to describe the set of legal
executions allowed over the replicated data types. The contracts are
constructed using primitive relations such as \emph{visibility} and
\emph{session order} and a carefully chosen set of logical operators. The
executions descriptions are similar to the high-level axiomatic decription of
relaxed memory models~\cite{Demange2013}. A contract enforcement system
logically analyses the contracts using an SMT solver~\cite{} and automatically
maps each operation to one of the store consistency levels. In addition, the
system statically reports an error if the given application-level consistency
property cannot be satisfied by the store. The paper makes the following
contributions:

\begin{itemize}

\item We introduce a data type definition language that allows the programmer
to describe their own convergent replicated data types. The data type
operations are defined as reductions over the set of all updates to a
particular object. New updates are simply appended to the existing set. This
expressive model lets the programmer implement their own convergence policy.

\item We present a contract language for describing fine-grained
application-level consistency properties, which is statically analyzed by a
contract enforcement system, and automatically assigns the operations to the
most efficient store consistency level, under which the application-level
contracts are guaranteed to hold.

\item We support coordination-free transactions over arbitrarty datatypes. We
extend our contract language to express fine-grained transaction isolation
guarantees, and utilize the contract enforcement system to automatically assign
the correct and most efficient isolation level for a transaction.

\item We present meta-theory that certifies the soundness of our contract
enforcement system, and ensures that an operation is only executed if the
required conditions on consistency are met.

\item We present extensive experimental evalation over a collection of
real-world applications, including a Twitter-like microblogging site and an
ebay-like auction site.

\end{itemize}

The rest of the paper is organized as follows. Section~\ref{sec:motivation}
motivates Quelea and introduces the system through a series of examples.
Section~\ref{sec:lang} presents the contract language, and
section~\ref{sec:opsem} presents the operational semantics of the Quelea
runtime system. Section~\ref{sec:txns} introduces transaction contracts and
classification. Section~\ref{sec:impl} describes the implementation and goes
into detail about the optimizations needed to make the system practical.
Section~\ref{sec:results} present the experimental evaluation.
Section~\ref{sec:related} and~\ref{sec:concl} discuss the related work and
conlusions.

### 2014-10-23 intro.tex
Major commit. First draft of intro and motivation.

@@
 reason about, the CAP theorem~\cite{Brewer2000,Gilbert} makes it impossible for
-a system to achieve strong consistency, tolerate network partions and remain
+a system to achieve strong consistency, tolerate network partitions and remain
 highly available. Thus, modern web services, which aim to provide an "always
@@
 over strong consistency. Several \emph{weak consistency} models such as
-eventual consistency, causal consistency, session guarantees, and timeline
+eventual consistency, causal consistency, session guarantees, and time line
 consistency have been proposed.
@@
 Under weak consistency, the programmer needs to be aware of concurrent
-conflicting updates, and requires careful attention to avoid inconsistencies
+conflicting updates, and has to pay careful attention to avoid inconsistencies
 such as negative balances in a bank account, and an item appearing in the
@@
 propose guarantees that are weaker than strong consistency, and hence need only
-to contact a subset of replicas for discharging the operations. Indeed,
-traditional database systems do provide a variety of transaction isolation
-levels~\cite{}, each with a particular runtime overhead. With these
-alternatives, it is upto the programmer to decide whether a weaker consistency
+to contact a subset of replicas for discharging the operations. With these
+alternatives, it is up to the programmer to decide whether a weaker consistency
 level is sufficient for a preserving the application-level consistency
@@
 
+%--KC--
+% The choice of correct consistency level is critical to the correctness and
+% scalability of the application. Choosing a weaker consistency level than
+% required introduces anomalies, while choosing a stronger than necessary level
+% affects program scalability.
+\KC{Do we include the LWW-Register number here to add emphasis to the fact that
+choosing the correct consistency level is essential for program scalability}.
+
 Moreover, weak consistency hinders compositional reasoning about programs.
@@
 last-writer-wins (LWW) register, counters and sets~\cite{Cassandra,DynamoDB}.
-Unlike traditional database systems, the geo-distributed stores typically lack
-unrestricted transactional access to the data. Thus, the application developer
-often has to coerce the application state to the data store capabilities.
+While traditional database systems enable composability through transactions,
+the state-of-the-art geo-distributed stores lack unrestricted transactional
+access to the data. Indeed, traditional database systems do offer a variety of
+transaction isolation levels~\cite{}, each with a particular runtime overhead,
+leaving the choice of correct consistency level to the programmer. Thus, the
+programmer working with geo-distributed stores has but to coerce the
+application state to the data store capabilities.
 
@@
 \emph{session order} and a carefully chosen set of logical operators. The
-executions descriptions are similar to the high-level axiomatic decription of
+executions descriptions are similar to the high-level axiomatic description of
 relaxed memory models~\cite{Demange2013}. A contract enforcement system
@@
 
-\item We introduce a data type definition language that allows the programmer
-to describe their own convergent replicated data types. The data type
-operations are defined as reductions over the set of all updates to a
-particular object. New updates are simply appended to the existing set. This
-expressive model lets the programmer implement their own convergence policy.
-
-\item We present a contract language for describing fine-grained
-application-level consistency properties, which is statically analyzed by a
-contract enforcement system, and automatically assigns the operations to the
-most efficient store consistency level, under which the application-level
+\item We introduce Quelea, a language for declaratively describing eventually
+consistent programs. Replicated data types in Quelea are defined as reductions
+over the set of previously seen updates on an object; new updates are simply
+appended to this set. Quelea is equipped with a contract language to describe
+fine-grained application-level consistency properties. The contracts are
+logically analyzed, and the corresponding operations are automatically assigned
+to the most efficient store consistency level under which the application-level
 contracts are guaranteed to hold.
 
-\item We support coordination-free transactions over arbitrarty datatypes. We
-extend our contract language to express fine-grained transaction isolation
+\item Quelea supports coordination-free transactions over arbitrary datatypes.
+We extend our contract language to express fine-grained transaction isolation
 guarantees, and utilize the contract enforcement system to automatically assign
@@
 
-\item We present extensive experimental evalation over a collection of
-real-world applications, including a Twitter-like microblogging site and an
-ebay-like auction site.
+\item Quelea is implemented as a shim layer over (and importantly without
+modifying) Cassandra, a general-purpose data store, illustrating the
+portability of our approach. An extensive experimental evaluation over a
+collection of real-world applications, including a Twitter-like micro-blogging
+site and an ebay-like auction site illustrates the practicality of our
+approach.
 
@@
 The rest of the paper is organized as follows. Section~\ref{sec:motivation}
-motivates Quelea and introduces the system through a series of examples.
+motivates Quelea and introduces the language through a series of examples.
 Section~\ref{sec:lang} presents the contract language, and
@@
 Section~\ref{sec:related} and~\ref{sec:concl} discuss the related work and
-conlusions.
+conclusions.

### 2014-10-23 motivation.tex
Major commit. First draft of intro and motivation.

\section{Motivation}

In this section, we motivate the need for declarative reasoning for eventually
consistent data store.

\subsection{System Model}

We assume a system model where the eventually consistent database is organized
as a collection of key-indexed, object-valued tables. The objects stored in the
tables are instances of the same replicated data type. The database is composed
of a number of \emph{replicas}, and each object is fully replicated across all
the replicas. Under this model, a client request for an operation with no
application-level consistency requirement can be serviced if at least one of
the replicas is reachable. Thus, with the addition of new replicas, and placing
them close to the clients, leads to improved throughput and reduced latencies.

\subsection{Bank account database}

Suppose our goal is to implement a bank account service on top an eventually
consistent data store, with the following schema,

\begin{codesql}
Create Table BankAccount (
  userID int PRIMARY KEY,
  userName string UNIQUE,
  balance float CHECK (balance >= 0))
\end{codesql}

\noindent with the following \emph{operations}.

\begin{codehaskell}
type UserName = String
-- returns false if the user name is already taken
addAccount     :: UserName -> Bool
getBalanceName :: UserName -> Float
depositName    :: UserName -> Float -> ()
-- returns false if the account has insufficient balance
withdrawName   :: UserName -> Float -> Bool
\end{codehaskell}

Notice that the operations manipulating the table take user name as input.
Since eventually consistent databases are non-relational~\cite{}, we need to
maintain a \emph{secondary index} to look up records by user name.

\begin{codesql}
Create Table UserIndex (
  userName string PRIMARY KEY,
  userID int
  FOREIGN KEY (userID)
    REFERENCES BankAccount (userID) )
\end{codesql}

\begin{figure}[t]
\centering
\subfigure[Negative Balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
\hfill
\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
\hfill
\subfigure[Monotonicity violation]{\label{fig:monotonicityAnomaly}\includegraphics[width=0.36\columnwidth]{Figures/Motivation3}}
\caption{Anomalies possible under eventual consistency for the get balance operation.}
\label{fig:cleanliness_examples}
\end{figure}

\subsection{Anomalies}

In a traditional database system, the programmer need only to define the
application-level integrity constraints along with the schema, and the database
management system automatically enforces the necessary coordination. However,
under eventual consistency, the onus is on the programmer to ensure not only
that the integrity constraints are preserved, but also to prevent any
inconsistencies being exposed to the user. For example, it is easy to see that
the withdraw operations must be strongly consistent operations to preserve the
integrity constraint. However, it is possible that the get balance operation
still returns a negative balance.

Figure~\ref{fig:negativeBalanceAnomaly} illustrates such a potential execution.
Assume that all operations are on the same object, and the balance on the
account was initially 0. Here, \cf{vis} edges capture the visibility
relation between the operations. In this case, the withdraw sees the deposit,
but the get balance operation only witnesses the withdraw. Such an execution is
possible under eventual consistency, where the \emph{effect} (update)
corresponding to the withdraw reaches the replica to which the get balance is
applied to before the causally preceding deposit effect. Alternatively, it is
possible that a get balance operation does not see the effects from its own
session (Figure~\ref{fig:missingUpdateAnomaly}), leading the user to
incorrectly conclude that the previous update failed to go through. Here, the
\cf{so} edge represents the session order that exists between the actions
on the same thread.

The anomalies discussed so far arise from weakly consistent behaviors on the
same object. With operations on multiple objects on distinct tables, the
programmer on eventually consistent databases has to take care of preserving
integrity constraints for foreign keys, materialized views and secondary
indexes. For example, Figure~\ref{fig:monotonicityAnomaly} illustrates the
anomaly that arises from the lack of monotonic update visibility. Adding a new
account involves a conditional insertion into the index table, followed by an
insertion into the bank account table. Many eventually consistent data stores
propose write-only transactions~\cite{}, and let us assume that add user
operation is marked as a write-only transaction. Despite this, under eventual
consistency, it is possible for a \cf{getBalanceName} operation to witness the
first insertion but not the second, exposing an inconsistent state. The two
reads may be served by different replicas, where the only the former has
witnessed the effects of add user transaction. Assigning the correct store
consistency level for the operations in this example such that the
application-level constraints are preserved is a non-trivial task.

\subsection{Implementation in Quelea}

Quelea automates this error-prone task; the programmer need only to express the
application-level consistency constraints in the contract language. Replicated
data types in Quelea are defined as reductions over the set of effects similar
to operation-based CRDTs~\cite{}. The following data type definitions capture
the operations and effects on the bank account and user index objects.

\begin{codehaskell}
data BankAcc = AddUser UserName | Deposit Float
             | Withdraw Float | GetBalance
data UserIdx = AddName UserID | GetName
\end{codehaskell}

Recall that the key of the bank account table is user id, which maps to a
particular bank account object. In Quelea, an object state is nothing but the
set of effects (called the \emph{context}) on this object. As such,
\cf{Deposit} and \cf{Withdraw} effects simply include the amount deposited or
withdrawn, and \cf{AddUser} effect includes the user name. Similarly, on the
user index table, where the key is user name, the value is simply an effect
(\cf{AddName}) binding some user id at this key. \cf{GetBalance} and
\cf{GetName} correspond to the corresponding read-only operations. The complete
definition of the operations on bank account and user index objects is given
below.

\begin{codehaskell}
-- Get balance takes no arguments and read-only
getBalance :: [BankAcc] {- context -} -> () {- args -}
  -> (() {- ret val -}, Maybe BankAcc {- effect -})
getBalance ctxt _ = (sum [v | Deposit v (- ctxt]
        - sum [v | Withdraw v (- ctxt], Nothing)

-- withdraw returns True on success
withdraw :: [BankAcc] -> Float -> (Bool, Maybe BankAcc)
withdraw c v = if sel1 $ getBalance c () >= v
               then (True, Just $ Withdraw v)
               else (False, Nothing)

deposit :: [BankAcc] -> v -> ((), Maybe BankAcc)
deposit _ v = ((), Just $ Deposit v)

addUser :: [BankAcc] -> UserName -> ((), Maybe BankAcc)
addUser _ n = ((), Just $ AddUser n)

addName :: [UserIdx] -> UserID -> (Bool, Maybe UserIdx)
addName [] uid = (True, Just $ AddName uid)
addName x:_ _ = (False, Nothing)

getName :: [UserIdx] -> ()
        -> (Maybe UserName, Maybe UserIdx)
getName [] _ = (Nothing, Nothing)
getName (AddUser uid:_) _ = (Just uid, Nothing)
\end{codehaskell}

The definitions are a straight forward encoding of the expected behavior.
The operation definitions presented here only capture the convergence
properties of the replicated data type, and are not concerned with the
consistency properties, which is expressed through the contract language.

\subsection{Contracts}

The contract for the strongly consistent \cf{addName} operation is given below:
\begin{smathpar}
\begin{array}{l}
\rsf{addNameCtrt} = \forall (a:\rsf{AddName}). \\
\qquad \sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} \vee \vis{\cureff}{a} \vee a = \cureff
\end{array}
\end{smathpar}

In the above definition, $\cureff$ represents the current effect --- the effect
emitted by the \cf{addName} operation. The contract simply states our
high-level observation; for any effect $a$ which is also an \cf{AddName} effect
on the same object, it must be the case that $a$ is visible to $\cureff$ or
vice verse, or $a$ is $\cureff$. Here, $\sameobjZ$ and $\visZ$ are primitive
relations. This contract ensures that any two \cf{addName} operations will be
seen by each other. The contract for withdraw is similar to the
\cf{addNameCtrt}.

Since the deposit operation does not have any restrictions, its contract is
simply $\true$. Same is the case for \cf{addUser} operation. The contract for
get balance operation is:
\begin{smathpar}
\begin{array}{l}
\rsf{getBalCtrt} = \forall (a:\rsf{Deposit}), (b:\rsf{Withdraw}). \\
\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\qquad \vee (\soZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff} \\
\qquad \vee (\soZ \cap \sameobjZ) (b,\cureff) \Rightarrow \vis{b}{\cureff}
\end{array}
\end{smathpar}

If a withdraw $b$ is visible to the get balance operation, then all deposit
operations $a$ visible to withdraw should be visible to the get balance
operation. This prevents negative balance anomalies. The rest of the contract
says that a get balance operation must witness previous deposit and withdraw
operations on the same object in the same session. This prevents missed update
anomalies.

Similar to contracts on operations, Quelea supports contracts on transactions.
The contract on the \cf{getBalanceName} transaction is given below:
\begin{smathpar}
\begin{array}{l}
\rsf{getBalanceNameCtrt} = \forall (a:\rsf{GetName}), (b:\rsf{GetBalance}), \\
\quad (c:\rsf{AddName}), (d:\rsf{AddUser}). ~\trans{a}{b}{c}{d} \wedge \so{a}{b} \\
\qquad \wedge ~\vis{c}{a} \wedge \sameobj{d}{b} \Rightarrow \vis{d}{b}
\end{array}
\end{smathpar}

$\trans{a}{b}{c}{d}$ says that the action pairs $a,b$ and $c,d$ are in the same
transaction, and the two transactions are distinct. The contract says
specifically forbids the anomaly presented in
Figure~\ref{fig:monotonicityAnomaly}. This is the desired semantics for the
\cf{getUserName} transaction.

The programmer simply defines such contracts on the operations and the
transactions. Quelea logically analyzes the contracts, and maps the
corresponding operation to the precise store consistency and isolation level.
Thus, Quelea equips the programmer with a declarative model for reasoning and
expressing eventually consistent programs.

### 2014-10-24 intro.tex
Intro modified

@@
 
-Many real-world web services --- such as Amazon, Facebook, Google, Twitter ---
-replicate application state and logic across multiple \emph{replicas} within
-and across data centers. Replication is intended not only to improve
-application throughput and reduce user perceived latency, but also to tolerate
-partial failures without compromising the service availability. Traditionally
-programmers have relied on \emph{strong consistency} guarantees such as
-Linerarizability~\cite{Herlihy1990} or serializability~\cite{Serializability}
-in order to build correct applications. While strong consistency is simple to
-reason about, the CAP theorem~\cite{Brewer2000,Gilbert} makes it impossible for
-a system to achieve strong consistency, tolerate network partitions and remain
-highly available. Thus, modern web services, which aim to provide an "always
-on" experience, overwhelmingly embrace availability and partition tolerance
-over strong consistency. Several \emph{weak consistency} models such as
-eventual consistency, causal consistency, session guarantees, and time line
-consistency have been proposed.
+Many real-world web applications --- such as Amazon, Facebook, Google, Twitter
+--- replicate application state and logic across multiple \emph{replicas} within
+and across data centers. Replication is intended not only to improve application
+throughput and reduce user perceived latency, but also to tolerate partial
+failures without compromising the service availability. Traditionally
+programmers have relied on \emph{strong consistency} guarantees, such as
+Linerizability~\cite{Herlihy1990} or serializability~\cite{Serializability},
+provided by the underlying data store in order to build correct applications.
+While strong consistency is simple to reason about, the CAP
+theorem~\cite{Brewer2000,Gilbert} makes it impossible for a replicated data
+store to achieve strong consistency, while tolerating network partitions and
+remaining highly available. Thus, modern web applications, which aim to provide
+an "always on" experience, overwhelmingly embrace availability and partition
+tolerance at the expense of strong consistency.
 
-Under weak consistency, the programmer needs to be aware of concurrent
-conflicting updates, and has to pay careful attention to avoid inconsistencies
-such as negative balances in a bank account, and an item appearing in the
-shopping cart after it has been removed~\cite{Dynamo}. Often times, the
-inconsistency leaks from the application and is witnessed by the user. To
-address this issue, many commercial~\cite{} and research~\cite{} systems
-propose guarantees that are weaker than strong consistency, and hence need only
-to contact a subset of replicas for discharging the operations. With these
-alternatives, it is up to the programmer to decide whether a weaker consistency
-level is sufficient for a preserving the application-level consistency
-properties. Understandably, this is an error prone process as the user not only
-has to understand the intricate behavior of the application, but also the
-semantics of the underlying store consistency levels, which are often presented
-informally~\cite{}.
+Nevetheless, to avoid exposing end user to absurdities such as negative balances
+in a bank account, or an item appearing in the shopping cart after it had been
+removed~\cite{Dynamo}, web applications \emph{do} require some level of
+consistency guarantees from the underlying data store. Accordingly, many
+commercial~\cite{} and research~\cite{} systems provide guarantees that are
+weaker than strong consistency, such as eventual consistency, causal
+consistency, session guarantees, and time line consistency. Each such \emph{weak
+consistency} guarantee imposes certain amount of penalty on the availability
+(hence, latency and throughput) of the system depending on the extent of
+inter-replica coordination required to ensure the guaranteed level of
+consistency. It is therefore incumbent on the application programmer to
+carefully reason about concurrent conflicting updates, and select the
+appropriate store-level weak consistency guarantee sufficient for preserving
+application-level integrity properties, while minimizing the adverse impact on
+the availability of the application. Understandably, this is an error prone
+process as the user not only has to understand the intricate behavior of the
+application, but also the semantics of the underlying store consistency levels,
+which are often presented informally~\cite{}.
 
@@
 Moreover, weak consistency hinders compositional reasoning about programs.
-While the application might be naturally expressed as maps, trees, counters,
-queues, or graphs, geo-distributed stores are typically only provide a minimal
-set of data types with in-built conflict resolution strategy such as
-last-writer-wins (LWW) register, counters and sets~\cite{Cassandra,DynamoDB}.
-While traditional database systems enable composability through transactions,
-the state-of-the-art geo-distributed stores lack unrestricted transactional
-access to the data. Indeed, traditional database systems do offer a variety of
-transaction isolation levels~\cite{}, each with a particular runtime overhead,
-leaving the choice of correct consistency level to the programmer. Thus, the
-programmer working with geo-distributed stores has but to coerce the
-application state to the data store capabilities.
+While applications are usually built around rich abstract data types, such as
+dictionaries, counters, queues, and graphs, geo-distributed stores typically
+only provide a minimal set of data types with in-built conflict resolution
+strategy such as last-writer-wins (LWW) register, counters and
+sets~\cite{Cassandra,DynamoDB}.  While traditional database systems enable
+composability through transactions, the state-of-the-art geo-distributed stores
+lack unrestricted transactional access to the data. Indeed, traditional database
+systems do offer a variety of transaction isolation levels~\cite{}, each with a
+particular runtime overhead, leaving the choice of correct consistency level to
+the programmer. Thus, the programmer working with geo-distributed stores has but
+to coerce the application state to the data store capabilities.
 
-To address these issues, we introduce Quelea, a declarative programming model
-for eventually consistent stores. The key novelty of Quelea is a \emph{contract
-language} to express the fine-grained application-level consistency properties.
-The programmer uses the contract language to describe the set of legal
-executions allowed over the replicated data types. The contracts are
-constructed using primitive relations such as \emph{visibility} and
-\emph{session order} and a carefully chosen set of logical operators. The
-executions descriptions are similar to the high-level axiomatic description of
-relaxed memory models~\cite{Demange2013}. A contract enforcement system
-logically analyses the contracts using an SMT solver~\cite{} and automatically
-maps each operation to one of the store consistency levels. In addition, the
-system statically reports an error if the given application-level consistency
-property cannot be satisfied by the store. The paper makes the following
-contributions:
+To address these issues, we introduce Quelea, a declarative programming system
+for eventually consistent data stores. Quelea's programming model is built
+around operation-based Commutative Replicated Data Types (CRDTs)
+~\cite{shapiroCRDT}. The state of an abstract data type in Quelea is modeled as
+a replicated \emph{append-only} log storing the history of all available
+\emph{update} operations\footnote{Available at each replica. Different replicas
+may have different logs as not all updates would have propagated to every
+replica.} performed on the data type . For instance, the state of a replicated
+counter ADT is stored on all replicas as a log of \emph{increment} and
+\emph{decrement} operations performed on the counter. An operation that indends
+to read the current value the counter at a replica only has to \emph{fold} over
+the history log at that replica. Our key observation is that under this model,
+consistency requirements of a replicated data type operation (eg:
+\emph{decrement}) can be expressed effectively in terms of assertions on
+\emph{membership} property of logs at various replicas. Quelea makes use of this
+observation to introduce a simple \emph{contract language} that is nevertheless
+capable of expressing fine-grained application-level consistency properties.
+The contract language is made up of a couple of primitive relations, namely
+\emph{visibility} and \emph{session order}, and a carefully choosen set of
+logical operators. Quelea expects application programmer to use the contract
+language to describe the set of legal executions allowed over the replicated
+data types.  
+% The executions descriptions are similar to the high-level axiomatic
+% description of relaxed memory models~\cite{Demange2013}.
+Subsequently, a contract enforcement system logically analyses contracts using
+an SMT solver~\cite{z3} and automatically tags each operation with the weakest
+store-level consistency guarantee that is required to satisfy application-level
+consistency requirements. In addition, the system statically reports an error if
+the given application-level consistency property cannot be satisfied by the
+store.
+
+The paper makes the following contributions:
 

### 2014-10-24 motivation.tex

@@
 The definitions are a straight forward encoding of the expected behavior.
-The operation definitions presented here only capture the convergence
-properties of the replicated data type, and are not concerned with the
-consistency properties, which is expressed through the contract language.
+Quelea programming model provides the programmer complete freedom regarding the
+semantics and desired convergence property of the replicated data type, and
+indeed, we can encode the well-known CRDTs~\cite{SSS} in a declarative fashion.
+Observe that the operation definitions presented here only capture the
+convergence properties of the replicated data type, and are not concerned with
+the consistency properties, which is expressed through the contract language.
 

### 2014-10-24 intro.tex
Merged changes in intro

@@
 provided by the underlying data store in order to build correct applications.
-While strong consistency is simple to reason about, the CAP
-theorem~\cite{Brewer2000,Gilbert} makes it impossible for a replicated data
-store to achieve strong consistency, while tolerating network partitions and
-remaining highly available. Thus, modern web applications, which aim to provide
+While strong consistency is an easily stated property, it
+masks the reality underlying large-scale distributed systems with respect to
+non-uniform latency (i.e., availability) and network
+partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web applications, which aim to provide
 an "always on" experience, overwhelmingly embrace availability and partition
@@
 Moreover, weak consistency hinders compositional reasoning about programs.
-While applications are usually built around rich abstract data types, such as
-dictionaries, counters, queues, and graphs, geo-distributed stores typically
-only provide a minimal set of data types with in-built conflict resolution
-strategy such as last-writer-wins (LWW) register, counters and
-sets~\cite{Cassandra,DynamoDB}.  While traditional database systems enable
-composability through transactions, the state-of-the-art geo-distributed stores
-lack unrestricted transactional access to the data. Indeed, traditional database
-systems do offer a variety of transaction isolation levels~\cite{}, each with a
-particular runtime overhead, leaving the choice of correct consistency level to
-the programmer. Thus, the programmer working with geo-distributed stores has but
-to coerce the application state to the data store capabilities.
+While an application might be naturally expressed in terms of well-understood
+abstract data types, such as dictionaries, counters, queues, and graphs,
+geo-distributed stores typically only provide a minimal set of data types with
+in-built conflict resolution strategy such as last-writer-wins (LWW) registers,
+counters and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while traditional
+database systems enable composability through transactions, existing
+geo-distributed stores lack unrestricted transactional access to the data.
+Working with geo-distributed stores thus requires application state to be
+suitably coerced to work within the capabilities of the store.
 
@@
 
-\item We introduce Quelea, a language for declaratively describing eventually
-consistent programs. Replicated data types in Quelea are defined as reductions
-over the set of previously seen updates on an object; new updates are simply
-appended to this set. Quelea is equipped with a contract language to describe
-fine-grained application-level consistency properties. The contracts are
-logically analyzed, and the corresponding operations are automatically assigned
-to the most efficient store consistency level under which the application-level
-contracts are guaranteed to hold.
+\item We introduce \name, a language for declaratively describing eventually
+  consistent programs that manipulate replicated data types. Contracts are
+  used to specify fine-grained application-level consistency properties, and
+  are analyzed to assign the most efficient and sound store consistency
+  level to the corresponding operation.
 
-\item Quelea supports coordination-free transactions over arbitrary datatypes.
-We extend our contract language to express fine-grained transaction isolation
-guarantees, and utilize the contract enforcement system to automatically assign
-the correct and most efficient isolation level for a transaction.
+\item \name supports coordination-free transactions over arbitrary
+  datatypes.  We extend our contract language to express fine-grained
+  transaction isolation guarantees, and utilize the contract enforcement
+  system to automatically assign the correct and most efficient isolation
+  level for a transaction.
 
-\item We present meta-theory that certifies the soundness of our contract
-enforcement system, and ensures that an operation is only executed if the
-required conditions on consistency are met.
+\item Metatheory that certifies the soundness of our contract enforcement
+  system, and ensures that an operation is only executed if the required
+  conditions on consistency are met is also given.
 
-\item Quelea is implemented as a shim layer over (and importantly without
-modifying) Cassandra, a general-purpose data store, illustrating the
-portability of our approach. An extensive experimental evaluation over a
-collection of real-world applications, including a Twitter-like micro-blogging
-site and an ebay-like auction site illustrates the practicality of our
-approach.
+\item \name is implemented as a shim layer over (and importantly without
+  modifying) Cassandra, a well-known general-purpose data store.  An
+  extensive experimental evaluation over a collection of real-world
+  applications, including a Twitter-like micro-blogging site and an
+  eBay-like auction site illustrates the practicality of our approach.
 
@@
 The rest of the paper is organized as follows. Section~\ref{sec:motivation}
-motivates Quelea and introduces the language through a series of examples.
-Section~\ref{sec:lang} presents the contract language, and
-section~\ref{sec:opsem} presents the operational semantics of the Quelea
-runtime system. Section~\ref{sec:txns} introduces transaction contracts and
+motivates \name and introduces the language through a series of examples.
+Section~\ref{sec:lang} describes the contract language, and
+Section~\ref{sec:opsem} presents an operational semantics.
+Section~\ref{sec:txns} introduces transaction contracts and
 classification. Section~\ref{sec:impl} describes the implementation and goes
 into detail about the optimizations needed to make the system practical.
-Section~\ref{sec:results} present the experimental evaluation.
-Section~\ref{sec:related} and~\ref{sec:concl} discuss the related work and
+Section~\ref{sec:results} discusses experimental evaluation.
+Section~\ref{sec:related} and~\ref{sec:concl} present related work and
 conclusions.

### 2014-10-24 intro.tex
Minor edits

@@
 
-To address these issues, we introduce Quelea, a declarative programming system
-for eventually consistent data stores. Quelea's programming model is built
-around operation-based Commutative Replicated Data Types (CRDTs)
-~\cite{shapiroCRDT}. The state of an abstract data type in Quelea is modeled as
-a replicated \emph{append-only} log storing the history of all available
-\emph{update} operations\footnote{Available at each replica. Different replicas
-may have different logs as not all updates would have propagated to every
-replica.} performed on the data type . For instance, the state of a replicated
-counter ADT is stored on all replicas as a log of \emph{increment} and
-\emph{decrement} operations performed on the counter. An operation that indends
-to read the current value the counter at a replica only has to \emph{fold} over
-the history log at that replica. Our key observation is that under this model,
-consistency requirements of a replicated data type operation (eg:
-\emph{decrement}) can be expressed effectively in terms of assertions on
-\emph{membership} property of logs at various replicas. Quelea makes use of this
-observation to introduce a simple \emph{contract language} that is nevertheless
-capable of expressing fine-grained application-level consistency properties.
-The contract language is made up of a couple of primitive relations, namely
-\emph{visibility} and \emph{session order}, and a carefully choosen set of
-logical operators. Quelea expects application programmer to use the contract
-language to describe the set of legal executions allowed over the replicated
-data types.  
+To address these issues, we introduce \name, a declarative programming system
+for eventually consistent data stores. \name's programming model is built around
+operation-based Commutative Replicated Data Types (CRDTs) ~\cite{shapiroCRDT}.
+The state of an abstract data type in \name is modeled as a replicated
+\emph{append-only} log storing the history of \emph{update} operations performed
+on the data type . For instance, the state of a replicated counter ADT is stored
+on all replicas as a log of \emph{increment} and \emph{decrement} operations
+performed on the counter. An operation that indends to read the current value
+the counter at a replica only has to \emph{fold} over the history log at that
+replica. Our key observation is that under this model, consistency requirements
+of a replicated data type operation (eg: \emph{decrement}) can be expressed
+effectively in terms of assertions on \emph{membership} property of logs at
+various replicas. \name makes use of this observation to introduce a simple
+\emph{contract language} that is nevertheless capable of expressing fine-grained
+application-level consistency properties.  The contract language is made up of a
+couple of primitive relations, namely \emph{visibility} and \emph{session
+order}, and a carefully choosen set of logical operators. \name expects
+application programmer to use the contract language to describe the set of legal
+executions allowed over the replicated data types.  
 % The executions descriptions are similar to the high-level axiomatic

### 2014-10-25 motivation.tex
M

@@
 \section{Motivation}
+\label{sec:motivation}
 
-In this section, we motivate the need for declarative reasoning for eventually
-consistent data store.
+In this section, with help of a bank account example, we motivate the
+need for a framework to express and enforce fine-grained
+application-level consistency requirements under weak consistency
+models. 
+% \GK{The following text needs to be polished}
 
-\subsection{System Model}
-
-We assume a system model where the eventually consistent database is organized
-as a collection of key-indexed, object-valued tables. The objects stored in the
-tables are instances of the same replicated data type. The database is composed
-of a number of \emph{replicas}, and each object is fully replicated across all
-the replicas. Under this model, a client request for an operation with no
-application-level consistency requirement can be serviced if at least one of
-the replicas is reachable. Thus, with the addition of new replicas, and placing
-them close to the clients, leads to improved throughput and reduced latencies.
+% We assume a system model where the eventually consistent database is organized
+% as a collection of key-indexed, object-valued tables. The objects stored in the
+% tables are instances of the same replicated data type. The database is composed
+% of a number of \emph{replicas}, and each object is fully replicated across all
+% the replicas. Under this model, a client request for an operation with no
+% application-level consistency requirement can be serviced if at least one of
+% the replicas is reachable. Thus, with the addition of new replicas, and placing
+% them close to the clients, leads to improved throughput and reduced latencies.
 
@@
 
-Suppose our goal is to implement a bank account service on top an eventually
-consistent data store, with the following schema,
+Suppose our goal is to implement a highly available bank account
+service on top of an eventually consistent data store. For the sake of
+presentation, we restrict ourselves to bank account objects (records)
+with following simple schema:
 
@@
 Create Table BankAccount (
-  userID int PRIMARY KEY,
-  userName string UNIQUE,
+  accountNo int PRIMARY KEY,
   balance float CHECK (balance >= 0))
@@
 
-\noindent with the following \emph{operations}.
+\noindent and with support for the following \emph{operations}:
 
 \begin{codehaskell}
-type UserName = String
--- returns false if the user name is already taken
-addAccount     :: UserName -> Bool
-getBalanceName :: UserName -> Float
-depositName    :: UserName -> Float -> ()
+type AccountNo = Int
+type Amount = Float
+getBalance :: AccountNo -> Amount
+deposit    :: AccountNo -> Amount -> ()
 -- returns false if the account has insufficient balance
-withdrawName   :: UserName -> Float -> Bool
+withdraw   :: AccountNo -> Amount -> Bool
 \end{codehaskell}
-
-Notice that the operations manipulating the table take user name as input.
-Since eventually consistent databases are non-relational~\cite{}, we need to
-maintain a \emph{secondary index} to look up records by user name.
-
-\begin{codesql}
-Create Table UserIndex (
-  userName string PRIMARY KEY,
-  userID int
-  FOREIGN KEY (userID)
-    REFERENCES BankAccount (userID) )
-\end{codesql}
 
@@
 \centering
+\subfigure[Unsafe Withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation4}}
+\hfill
 \subfigure[Negative Balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
@@
 \subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
-\hfill
-\subfigure[Monotonicity violation]{\label{fig:monotonicityAnomaly}\includegraphics[width=0.36\columnwidth]{Figures/Motivation3}}
+%\hfill
+%\subfigure[Monotonicity violation]{\label{fig:monotonicityAnomaly}\includegraphics[width=0.36\columnwidth]{Figures/Motivation3}}
 \caption{Anomalies possible under eventual consistency for the get balance operation.}
@@
 
+The sole attribute of a \cf{BankAccount} object, identified by its unique
+\cf{accountNo}, is its current \cf{balance}. A \cf{deposit} operation
+on a \cf{BankAccount} adds some \cf{Amount} to its \cf{balance},
+whereas a \cf{withdraw} operation subtracts. Operation \cf{getBalance}
+returns the current \cf{balance} in the account. A \cf{withdraw}
+operation succeeds and returns \cf{True} only if there is sufficient
+balance in the account. Otherwise, it fails and returns \cf{False}.
+
+If we were to use a traditional database system to store
+\cf{BankAccount} objects, all that needs to be done is to define the
+\cf{BankAccount} schema, as shown above. The database management
+system effectively sequentializes all accesses to the database, and
+automatically enforces the required integerity constraints on the
+data, so that it is never possible for a \cf{getBalance} operation to
+return a negative balance, or for a \cf{withdraw} to succeed without
+having sufficient balance in the account. However, under eventual
+consistency, such anomalies are bound to occur as demonstrated in the
+following subsection.
+
 \subsection{Anomalies}
 
-In a traditional database system, the programmer need only to define the
-application-level integrity constraints along with the schema, and the database
-management system automatically enforces the necessary coordination. However,
-under eventual consistency, the onus is on the programmer to ensure not only
-that the integrity constraints are preserved, but also to prevent any
-inconsistencies being exposed to the user. For example, it is easy to see that
-the withdraw operations must be strongly consistent operations to preserve the
-integrity constraint. However, it is possible that the get balance operation
-still returns a negative balance.
+Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}.
+Assume that all operations in the figure are on the same
+\cf{BankAccount} object. Execution contains two concurrent client
+sessions, where one of the sessions executes a \cf{withdraw} followed
+by a \cf{getBalance} operation. We say that \c{withdraw} precedes
+\cf{getBalance} in \emph{session order}, and visualize it as a
+directed \cf{so} edge from the former to the later. Edges labelled
+\cf{vis} in the figure capture the \emph{visibility} relation between
+operations; an operation is said to be \emph{visible} to other
+operation if the later can witness the effects of the former. Two
+operations that are not related either via \cf{vis} or \cf{so} are
+said to be concurrent. The figure shows an execution where concurrent
+\cf{withdraw}s in different sessions succeed independently, but a
+subsequent \cf{getBalance} operation that witnesses both the
+\cf{withdraw}s reports a negative balance.
 
-Figure~\ref{fig:negativeBalanceAnomaly} illustrates such a potential execution.
-Assume that all operations are on the same object, and the balance on the
-account was initially 0. Here, \cf{vis} edges capture the visibility
-relation between the operations. In this case, the withdraw sees the deposit,
-but the get balance operation only witnesses the withdraw. Such an execution is
-possible under eventual consistency, where the \emph{effect} (update)
-corresponding to the withdraw reaches the replica to which the get balance is
-applied to before the causally preceding deposit effect. Alternatively, it is
-possible that a get balance operation does not see the effects from its own
-session (Figure~\ref{fig:missingUpdateAnomaly}), leading the user to
-incorrectly conclude that the previous update failed to go through. Here, the
-\cf{so} edge represents the session order that exists between the actions
-on the same thread.
+It is easy to see that the only way to ensure that integrity
+constraint on \cf{balance} is to prevent concurrent withdrawls on the
+\cf{BankAccount}. This can be done by insisting that \cf{withdraw} be
+executed with strong consistency. Although this prevents unsafe
+withdrawls from going through, a strongly consistent \cf{withdraw}
+itself is not sufficient to prevent \cf{getBalance} from ever
+reporting a negative balance to the end user.
 
-The anomalies discussed so far arise from weakly consistent behaviors on the
-same object. With operations on multiple objects on distinct tables, the
-programmer on eventually consistent databases has to take care of preserving
-integrity constraints for foreign keys, materialized views and secondary
-indexes. For example, Figure~\ref{fig:monotonicityAnomaly} illustrates the
-anomaly that arises from the lack of monotonic update visibility. Adding a new
-account involves a conditional insertion into the index table, followed by an
-insertion into the bank account table. Many eventually consistent data stores
-propose write-only transactions~\cite{}, and let us assume that add user
-operation is marked as a write-only transaction. Despite this, under eventual
-consistency, it is possible for a \cf{getBalanceName} operation to witness the
-first insertion but not the second, exposing an inconsistent state. The two
-reads may be served by different replicas, where the only the former has
-witnessed the effects of add user transaction. Assigning the correct store
-consistency level for the operations in this example such that the
-application-level constraints are preserved is a non-trivial task.
+Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
+which consists of three concurrent sessions performing a \cf{deposit},
+a \cf{withdraw}, and a \cf{getBalance}, respectively, on the same
+\cf{BankAccount}. Assume that initial balance in the account is zero.
+As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in session
+2, can witnesses the effects of \cf{deposit(100)} from session 1,
+leading it to rightfully conclude that the current balance is 100.
+Consequently, \cf{withdraw(50)} is performed safely. The effect of
+this \cf{withdraw} operation subsequently becomes visible to the
+\cf{getBalance} in Session 3. However, \cf{getBalance}, which can only
+see the effect of \cf{withdraw} from session 2, but not \cf{deposit}
+from session 1, reports the balance of negative 50 to the user.
+
+Negative balance is not the only inconsistency that the user is
+exposed to under an eventually consistent bank account. Figure
+~\ref{fig:missingUpdateAnomaly} shows an execution where
+\cf{getBalance} operation in a session does not witness the effects of
+\cf{withdraw} operation performed in the same session, possibly
+because \cf{getBalance} was served by a replica that hasn't yet merged
+the \cf{withdraw} effect, which could have been served by a different
+replica. This anomaly leads the user to incorrectly conclude that the
+\cf{withdraw} operation failed to go through.
+
+Although it is easy to understand the reasons behind the occurance of
+the aforementioned negative balance and missing update anomalies, the
+fix is not quite apparent. Making \cf{getBalance} a strongly
+consistent operation is definitely sufficient to avert anomalies, but
+is it really necessary? Given that a strongly consistent operation is
+significantly expensive, both in terms of latency and monetary value, we ideally
+would not want to request strong consistency, unless there are
+no alternatives available. Unfortunately, this requires us to
+understand subtle differences in semantics of various kinds of weak
+consistency alternatives, such as timeline consistency and session
+guarantees, offered by off-the-shelf eventually consistent data
+stores. This process is error-prone, and becomes further complicated
+in presence of transactions, as demonstrated in Sec.
+~\ref{sec:transactions}.
 
@@
 
-Quelea automates this error-prone task; the programmer need only to express the
-application-level consistency constraints in the contract language. Replicated
-data types in Quelea are defined as reductions over the set of effects similar
-to operation-based CRDTs~\cite{}. The following data type definitions capture
-the operations and effects on the bank account and user index objects.
-
+\name automates this error-prone task by letting the programmer
+declare application-level consistency constraints as \emph{contracts},
+obviating the need to understand store-level consistency guarantees,
+and relieving programmer from having to match application
+requirements with appropriate store-level alternatives. The
+cornerstone of declarative reasoning in \name is the programming model
+built around operation-based CRDTs~\cite{shapiroCRDT}. A replicated
+data type in \name is defined as a tagged union type, where each tag
+is the name of a unique operation defined on the data type. 
+%types in Quelea are defined as reductions over the set of effects.
+%The following data type definitions capture the operations and effects
+%on the bank account and user index objects.
+For example, following is the definition of the \cf{BankAccount}
+data type in \name:
 \begin{codehaskell}
-data BankAcc = AddUser UserName | Deposit Float
-             | Withdraw Float | GetBalance
-data UserIdx = AddName UserID | GetName
+data BankAccount =  Deposit Float | Withdraw Float 
+                 | GetBalance
 \end{codehaskell}
 
-Recall that the key of the bank account table is user id, which maps to a
-particular bank account object. In Quelea, an object state is nothing but the
-set of effects (called the \emph{context}) on this object. As such,
-\cf{Deposit} and \cf{Withdraw} effects simply include the amount deposited or
-withdrawn, and \cf{AddUser} effect includes the user name. Similarly, on the
-user index table, where the key is user name, the value is simply an effect
-(\cf{AddName}) binding some user id at this key. \cf{GetBalance} and
-\cf{GetName} correspond to the corresponding read-only operations. The complete
-definition of the operations on bank account and user index objects is given
-below.
+In Quelea, an object state is nothing but the set of effects (called
+the \emph{context}) on this object. As such, \cf{Deposit} and
+\cf{Withdraw} effects simply include the amount deposited or
+withdrawn, and \cf{AddUser} effect includes the user name. Similarly,
+on the user index table, where the key is user name, the value is
+simply an effect (\cf{AddName}) binding some user id at this key.
+\cf{GetBalance} and \cf{GetName} correspond to the corresponding
+read-only operations. The complete definition of the operations on
+bank account and user index objects is given below.
 

### 2014-10-25 motivation.tex
Most changes to motivation done

@@
 
-Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}.
+Consider the execution shown in fig.~\ref{fig:unsafeWithdrawAnomaly}.
 Assume that all operations in the figure are on the same
-\cf{BankAccount} object. Execution contains two concurrent client
-sessions, where one of the sessions executes a \cf{withdraw} followed
-by a \cf{getBalance} operation. We say that \c{withdraw} precedes
-\cf{getBalance} in \emph{session order}, and visualize it as a
-directed \cf{so} edge from the former to the later. Edges labelled
-\cf{vis} in the figure capture the \emph{visibility} relation between
-operations; an operation is said to be \emph{visible} to other
-operation if the later can witness the effects of the former. Two
-operations that are not related either via \cf{vis} or \cf{so} are
-said to be concurrent. The figure shows an execution where concurrent
-\cf{withdraw}s in different sessions succeed independently, but a
-subsequent \cf{getBalance} operation that witnesses both the
-\cf{withdraw}s reports a negative balance.
+\cf{BankAccount} object with initial balance as 100. Execution
+contains two concurrent client sessions, where one of the sessions
+executes a \cf{withdraw} followed by a \cf{getBalance} operation. We
+say that \c{withdraw} precedes \cf{getBalance} in \emph{session
+order}, and visualize it as a directed \cf{so} edge from the former to
+the later. Edges labelled \cf{vis} in the figure capture the
+\emph{visibility} relation between operations; an operation is said to
+be \emph{visible} to other operation if the later can witness the
+effects of the former. Two operations that are not related either via
+\cf{vis} or \cf{so} are said to be concurrent. The figure shows an
+execution where concurrent \cf{withdraw}s in different sessions
+succeed independently, but a subsequent \cf{getBalance} operation that
+witnesses both the \cf{withdraw}s reports a negative balance.
 
@@
 \name automates this error-prone task by letting the programmer
-declare application-level consistency constraints as \emph{contracts},
-obviating the need to understand store-level consistency guarantees,
-and relieving programmer from having to match application
+declare application-level consistency constraints as \emph{contracts}.
+This obviates the need to understand store-level consistency guarantees,
+and relieves the programmer from having to match application
 requirements with appropriate store-level alternatives. The
 cornerstone of declarative reasoning in \name is the programming model
-built around operation-based CRDTs~\cite{shapiroCRDT}. A replicated
-data type in \name is defined as a tagged union type, where each tag
-is the name of a unique operation defined on the data type. 
+that draws largely on the concept of operation-based
+CRDTs~\cite{shapiroCRDT}. The state of a replicated data object in
+\name is nothing but the set of effects applied on the object. For
+example, to represent a \cf{BankAccount} object in \name, we first
+define an sum type of all possible effects on a \cf{BankAccount}
+object:
 %types in Quelea are defined as reductions over the set of effects.
@@
 %on the bank account and user index objects.
-For example, following is the definition of the \cf{BankAccount}
-data type in \name:
 \begin{codehaskell}
-data BankAccount =  Deposit Float | Withdraw Float 
+type Amount = Float
+data BankAccountEff = Deposit Amount
+                    | Withdraw Amount 
                     | GetBalance
 \end{codehaskell}
-
-In Quelea, an object state is nothing but the set of effects (called
-the \emph{context}) on this object. As such, \cf{Deposit} and
-\cf{Withdraw} effects simply include the amount deposited or
-withdrawn, and \cf{AddUser} effect includes the user name. Similarly,
-on the user index table, where the key is user name, the value is
-simply an effect (\cf{AddName}) binding some user id at this key.
-\cf{GetBalance} and \cf{GetName} correspond to the corresponding
-read-only operations. The complete definition of the operations on
-bank account and user index objects is given below.
-
+Subsequently, any \cf{BankAccount} object can be simply represented as a set of
+effects of type \cf{BankAccountEff} that took effect on the object.
+For example, a \cf{BankAccount} object obtained by performing a
+deposit of 100 and a withdrawl of 50 on a fresh bank account is
+represented as following:
+\begin{center}
+  \{\cf{Deposit(100), Withdraw(50)}\}
+\end{center}
+An operation on a replicated data object in \name is simply defined as
+a reduction over the set of effects representing the object, and is
+expected to produce a new effect along with a return value. For
+instance, following are the definitions of operations on
+\cf{BankAccount} type in \name (we represent sets as lists for
+clarity):
+\GK{Better define getBalance in terms of fold. It goes well with our
+narrative.}
 \begin{codehaskell}
--- Get balance takes no arguments and read-only
-getBalance :: [BankAcc] {- context -} -> () {- args -}
-  -> (() {- ret val -}, Maybe BankAcc {- effect -})
+type BankAccount = [BankAccountEff]
+getBalance ::  BankAccount {- object -}
+  -> () {- args -}
+  -> (Amount {- ret val -}, 
+      Maybe BankAccountEff {- effect -})
 getBalance ctxt _ = (sum [v | Deposit v (- ctxt]
@@
 
--- withdraw returns True on success
-withdraw :: [BankAcc] -> Float -> (Bool, Maybe BankAcc)
-withdraw c v = if sel1 $ getBalance c () >= v
-               then (True, Just $ Withdraw v)
+withdraw :: BankAccount -> Amount 
+  -> (Bool, Maybe BankAccountEff)
+withdraw c v = if fst (getBalance c ()) >= v
+               then (True, Just (Withdraw v))
                else (False, Nothing)
 
-deposit :: [BankAcc] -> v -> ((), Maybe BankAcc)
-deposit _ v = ((), Just $ Deposit v)
-
-addUser :: [BankAcc] -> UserName -> ((), Maybe BankAcc)
-addUser _ n = ((), Just $ AddUser n)
-
-addName :: [UserIdx] -> UserID -> (Bool, Maybe UserIdx)
-addName [] uid = (True, Just $ AddName uid)
-addName x:_ _ = (False, Nothing)
-
-getName :: [UserIdx] -> ()
-        -> (Maybe UserName, Maybe UserIdx)
-getName [] _ = (Nothing, Nothing)
-getName (AddUser uid:_) _ = (Just uid, Nothing)
+deposit :: BankAccount -> Amount
+  -> ((), Maybe BankAccountEff)
+deposit _ v = ((), Just (Deposit v))
 \end{codehaskell}
 
-The definitions are a straight forward encoding of the expected behavior.
-Quelea programming model provides the programmer complete freedom regarding the
-semantics and desired convergence property of the replicated data type, and
-indeed, we can encode the well-known CRDTs~\cite{SSS} in a declarative fashion.
-Observe that the operation definitions presented here only capture the
-convergence properties of the replicated data type, and are not concerned with
-the consistency properties, which is expressed through the contract language.
+A \cf{getBalance} operation accepts a \cf{BankAccount} object
+represented as a set of \cf{BankAccountEff}s, folds over the set
+adding deposits and subtracting withdraws, and returns the resultant
+amount, along with a new \cf{GetBalance} effect\footnote{wrapped
+inside a \cf{Maybe} type for technical reasons}. A \cf{withdraw}
+operation first checks if there is sufficient balance in the account.
+If the check passes, it returns \cf{True} along with a new
+\cf{Withdraw} effect. Otherwise, it returns \cf{False} and produces no
+effect. The definition of \cf{deposit} can be similarly understood.
+
+When a bank account operation is executed at one of the replicas in
+response to a client request, \cf{BankAccount} object is constructed
+using the effects available on the replica, and the definition of the
+current operation is applied on the object. Note the set of effects
+constituting the input \cf{BankAccount} object are the only effects
+that are \emph{visible} to the current operation.  Formally, we say
+that a \emph{visibility} relation exists between the input set of
+effects and the newly engendered effect. In contrast, a concurrent
+operation executing on a different replica can neither witness the
+effect of the current operation, not can its effect be visible to the
+current operation.
+
+It is worthwhile to note that the above definitions of
+\cf{BankAccount} operations are a straightforward encoding of the
+expected behavior.  \name programming model provides the programmer
+complete freedom regarding the semantics and desired convergence
+property of the replicated data type, and indeed, we can encode the
+well-known CRDTs~\cite{SSS} in a declarative fashion.  Observe that
+the operation definitions presented here only capture the convergence
+properties of the replicated data type, and are not concerned with the
+consistency properties, which is expressed through the contract
+language.
 
@@
 
-The contract for the strongly consistent \cf{addName} operation is given below:
+We have previously observed that two \cf{withdraw} operations cannot
+be executed concurrently for the integrity constraint on the account
+balance to hold. Informally, this means that for any two distinct
+\cf{withdraw} operations $a$ and $b$ on a \cf{BankAccount} object,
+either the effect of $a$ must be visible to $b$, or the effect of $b$
+must be visible to $a$. \name lets us declare this consistency
+requirement as a contract over \cf{withdraw}:
 \begin{smathpar}
 \begin{array}{l}
-\rsf{addNameCtrt} = \forall (a:\rsf{AddName}). \\
-\qquad \sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} \vee \vis{\cureff}{a} \vee a = \cureff
+\rsf{withdrawCtrt}(b) = \forall a. ~a \neq b \conj \rsf{isWithdraw}(a) \conj \\
+\qquad \sameobj{a}{b} \Rightarrow \vis{a}{b} \vee \vis{b}{a} 
+\end{array}
+\end{smathpar}
+The above definition of $\rsf{withdrawCtrt}$ is parameterized over the
+effect ($b:\cf{BankAccountEff}$) generated by the \cf{withdraw}
+operation. The predicate $\rsf{isWithdraw}$ is true for a
+\cf{Withdraw} effect, and false for any other effect. $\sameobj{a}{b}$
+is true if and only if $a$ and $b$ are effects over the same
+\cf{BankAccount} object; we only require \cf{withdraw} operations on
+same bank account to be totally ordered. Operations on other bank
+accounts can be executed concurrently. Rest of the contract is a
+straightforward translation of its informal presentation discussed
+earlier.
+
+For the \cf{getBalance}, we construct a contract  precisely to prevent
+the anomalies described previously. The contract is shown below:
+\begin{smathpar}
+\begin{array}{l}
+\rsf{getBalCtrt}(c) = \forall (a,b). ~ \rsf{isDeposit}(a) \conj
+  \rsf{isWithdraw}(b) \conj \\
+\qquad \vis{a}{b} \wedge \vis{b}{c} \Rightarrow \vis{a}{c} \\
+\qquad \vee (\soZ \cap \sameobjZ) (a,c) \Rightarrow \vis{a}{c} \\
+\qquad \vee (\soZ \cap \sameobjZ) (b,c) \Rightarrow \vis{b}{c}
 \end{array}
@@
 
-In the above definition, $\cureff$ represents the current effect --- the effect
-emitted by the \cf{addName} operation. The contract simply states our
-high-level observation; for any effect $a$ which is also an \cf{AddName} effect
-on the same object, it must be the case that $a$ is visible to $\cureff$ or
-vice verse, or $a$ is $\cureff$. Here, $\sameobjZ$ and $\visZ$ are primitive
-relations. This contract ensures that any two \cf{addName} operations will be
-seen by each other. The contract for withdraw is similar to the
-\cf{addNameCtrt}.
+\GK{It is better to split this contract into two parts.}
+If a withdraw $b$ is visible to the get balance operation, then all
+deposit operations $a$ visible to withdraw should be visible to the
+get balance operation. This prevents negative balance anomalies. The
+rest of the contract says that a get balance operation must witness
+previous deposit and withdraw operations on the same object in the
+same session. This prevents missed update anomalies.
 
-Since the deposit operation does not have any restrictions, its contract is
-simply $\true$. Same is the case for \cf{addUser} operation. The contract for
-get balance operation is:
-\begin{smathpar}
-\begin{array}{l}
-\rsf{getBalCtrt} = \forall (a:\rsf{Deposit}), (b:\rsf{Withdraw}). \\
-\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\qquad \vee (\soZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff} \\
-\qquad \vee (\soZ \cap \sameobjZ) (b,\cureff) \Rightarrow \vis{b}{\cureff}
-\end{array}
-\end{smathpar}
+Since the deposit operation does not have any restrictions, its
+contract is simply $\true$.
 
-If a withdraw $b$ is visible to the get balance operation, then all deposit
-operations $a$ visible to withdraw should be visible to the get balance
-operation. This prevents negative balance anomalies. The rest of the contract
-says that a get balance operation must witness previous deposit and withdraw
-operations on the same object in the same session. This prevents missed update
-anomalies.
+Note that the above contracts for \cf{withdraw} and \cf{getBalance}
+express their consistency requirements \emph{independent} of the
+semantics of the underlying store. To write such contracts, a
+programmer only needs to reason about the semantics of the application
+under the replication model professed by \name. The mapping of
+application-level consistency requirements to appropriate store-level
+guarantees is done automatically behind the screen. This process is
+described in the following two sections.
+% Similar to contracts on operations, Quelea supports contracts on transactions.
+% The contract on the \cf{getBalanceName} transaction is given below:
+% \begin{smathpar}
+% \begin{array}{l}
+% \rsf{getBalanceNameCtrt} = \forall (a:\rsf{GetName}), (b:\rsf{GetBalance}), \\
+% \quad (c:\rsf{AddName}), (d:\rsf{AddUser}). ~\trans{a}{b}{c}{d} \wedge \so{a}{b} \\
+% \qquad \wedge ~\vis{c}{a} \wedge \sameobj{d}{b} \Rightarrow \vis{d}{b}
+% \end{array}
+% \end{smathpar}
 
-Similar to contracts on operations, Quelea supports contracts on transactions.
-The contract on the \cf{getBalanceName} transaction is given below:
-\begin{smathpar}
-\begin{array}{l}
-\rsf{getBalanceNameCtrt} = \forall (a:\rsf{GetName}), (b:\rsf{GetBalance}), \\
-\quad (c:\rsf{AddName}), (d:\rsf{AddUser}). ~\trans{a}{b}{c}{d} \wedge \so{a}{b} \\
-\qquad \wedge ~\vis{c}{a} \wedge \sameobj{d}{b} \Rightarrow \vis{d}{b}
-\end{array}
-\end{smathpar}
+% $\trans{a}{b}{c}{d}$ says that the action pairs $a,b$ and $c,d$ are in the same
+% transaction, and the two transactions are distinct. The contract says
+% specifically forbids the anomaly presented in
+% Figure~\ref{fig:monotonicityAnomaly}. This is the desired semantics for the
+% \cf{getUserName} transaction.
 
-$\trans{a}{b}{c}{d}$ says that the action pairs $a,b$ and $c,d$ are in the same
-transaction, and the two transactions are distinct. The contract says
-specifically forbids the anomaly presented in
-Figure~\ref{fig:monotonicityAnomaly}. This is the desired semantics for the
-\cf{getUserName} transaction.
-
-The programmer simply defines such contracts on the operations and the
-transactions. Quelea logically analyzes the contracts, and maps the
-corresponding operation to the precise store consistency and isolation level.
-Thus, Quelea equips the programmer with a declarative model for reasoning and
-expressing eventually consistent programs.
+% The programmer simply defines such contracts on the operations and the
+% transactions. Quelea logically analyzes the contracts, and maps the
+% corresponding operation to the precise store consistency and isolation level.
+% Thus, Quelea equips the programmer with a declarative model for reasoning and
+% expressing eventually consistent programs.

### 2014-10-25 motivation.tex
Edits to motivation

@@
 declare application-level consistency constraints as \emph{contracts}.
-This obviates the need to understand store-level consistency guarantees,
-and relieves the programmer from having to match application
-requirements with appropriate store-level alternatives. The
-cornerstone of declarative reasoning in \name is the programming model
-that draws largely on the concept of operation-based
+% This obviates the need to understand store-level consistency guarantees,
+% and relieves the programmer from having to match application
+% requirements with appropriate store-level alternatives. 
+The cornerstone of declarative reasoning in \name is the programming
+model that draws largely on the concept of operation-based
 CRDTs~\cite{shapiroCRDT}. The state of a replicated data object in
@@
 clarity):
-\GK{Better define getBalance in terms of fold. It goes well with our
-narrative.}
+\GK{I think it is better define getBalance in terms of fold. It goes
+well with our narrative.}
 \begin{codehaskell}
@@
 
-\GK{It is better to split this contract into two parts.}
+\GK{It may be better to split this contract into two parts.}
 If a withdraw $b$ is visible to the get balance operation, then all

### 2014-10-27 intro.tex
Intro and Motivation
Motivation version 2

@@
 
-Many real-world web applications --- such as Amazon, Facebook, Google, Twitter
---- replicate application state and logic across multiple \emph{replicas} within
-and across data centers. Replication is intended not only to improve application
-throughput and reduce user perceived latency, but also to tolerate partial
-failures without compromising the service availability. Traditionally
-programmers have relied on \emph{strong consistency} guarantees, such as
-Linerizability~\cite{Herlihy1990} or serializability~\cite{Serializability},
-provided by the underlying data store in order to build correct applications.
-While strong consistency is an easily stated property, it
-masks the reality underlying large-scale distributed systems with respect to
-non-uniform latency (i.e., availability) and network
-partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web applications, which aim to provide
-an "always on" experience, overwhelmingly embrace availability and partition
-tolerance at the expense of strong consistency.
+Many real-world web services --- such as those built and maintained by Amazon,
+Facebook, Google, Twitter, etc. --- replicate application state and logic
+across multiple \emph{replicas} within and across data centers. Replication is
+intended not only to improve application throughput and reduce user-perceived
+latency, but also to tolerate partial failures without compromising overall
+service availability. Traditionally programmers have relied on \emph{strong
+consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
+serializability~\cite{Serializability} in order to build correct applications.
+While strong consistency is an easily stated property, it masks the reality
+underlying large-scale distributed systems with respect to non-uniform latency
+(i.e., availability) and network partitions~\cite{Brewer2000,Gilbert}. Indeed,
+modern web services, which aim to provide an "always on" experience,
+overwhelmingly favour availability and partition tolerance over strong
+consistency. To this end, several \emph{weak consistency} models such as
+eventual consistency, causal consistency, session guarantees, and timeline
+consistency have consequently been proposed.
 
-Nevetheless, to avoid exposing end user to absurdities such as negative balances
-in a bank account, or an item appearing in the shopping cart after it had been
-removed~\cite{Dynamo}, web applications \emph{do} require some level of
-consistency guarantees from the underlying data store. Accordingly, many
-commercial~\cite{} and research~\cite{} systems provide guarantees that are
-weaker than strong consistency, such as eventual consistency, causal
-consistency, session guarantees, and time line consistency. Each such \emph{weak
-consistency} guarantee imposes certain amount of penalty on the availability
-(hence, latency and throughput) of the system depending on the extent of
-inter-replica coordination required to ensure the guaranteed level of
-consistency. It is therefore incumbent on the application programmer to
-carefully reason about concurrent conflicting updates, and select the
-appropriate store-level weak consistency guarantee sufficient for preserving
-application-level integrity properties, while minimizing the adverse impact on
-the availability of the application. Understandably, this is an error prone
-process as the user not only has to understand the intricate behavior of the
-application, but also the semantics of the underlying store consistency levels,
-which are often presented informally~\cite{}.
+Under weak consistency, the developer needs to be aware of concurrent
+conflicting updates, and has to pay careful attention to avoid unwanted
+inconsistencies (e.g., negative balances in a bank account, or having an item
+appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
+the inconsistency leaks from the application and is witnessed by the user.
+Ultimately, the developer must decide the consistency level appropriate for a
+particular operation; this is understandably an error-prone process requiring
+intricate knowledge of both the application as well as the semantics and
+implementation of the underlying data store.  More often than not, only
+informal descriptions are available~\cite{}.  Picking the correct consistency
+level is critical not only for correctness but also for scalability of the
+application. While choosing a weaker consistency level than required may
+introduce program errors and anomalies, choosing a stronger one than necessary
+level can negatively impact program scalability.
 
@@
 % affects program scalability.
-\KC{Do we include the LWW-Register number here to add emphasis to the fact that
-choosing the correct consistency level is essential for program scalability}.
+%\KC{Do we include the LWW-Register number here to add emphasis to the fact that
+%choosing the correct consistency level is essential for program scalability}.
 
@@
 While an application might be naturally expressed in terms of well-understood
-abstract data types, such as dictionaries, counters, queues, and graphs,
+and expressive data types such as maps, trees, queues, or graphs,
 geo-distributed stores typically only provide a minimal set of data types with
-in-built conflict resolution strategy such as last-writer-wins (LWW) registers,
-counters and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while traditional
-database systems enable composability through transactions, existing
-geo-distributed stores lack unrestricted transactional access to the data.
-Working with geo-distributed stores thus requires application state to be
-suitably coerced to work within the capabilities of the store.
+in-built conflict resolution strategies such as last-writer-wins (LWW)
+registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor, while
+traditional database systems enable composability through transactions,
+existing geo-distributed stores lack unrestricted transactional access to the
+data.  Working with geo-distributed stores thus requires application state to
+be suitably coerced to work within the capabilities of the store.
 
-To address these issues, we introduce \name, a declarative programming system
-for eventually consistent data stores. \name's programming model is built around
-operation-based Commutative Replicated Data Types (CRDTs) ~\cite{shapiroCRDT}.
-The state of an abstract data type in \name is modeled as a replicated
-\emph{append-only} log storing the history of \emph{update} operations performed
-on the data type . For instance, the state of a replicated counter ADT is stored
-on all replicas as a log of \emph{increment} and \emph{decrement} operations
-performed on the counter. An operation that indends to read the current value
-the counter at a replica only has to \emph{fold} over the history log at that
-replica. Our key observation is that under this model, consistency requirements
-of a replicated data type operation (eg: \emph{decrement}) can be expressed
-effectively in terms of assertions on \emph{membership} property of logs at
-various replicas. \name makes use of this observation to introduce a simple
-\emph{contract language} that is nevertheless capable of expressing fine-grained
-application-level consistency properties.  The contract language is made up of a
-couple of primitive relations, namely \emph{visibility} and \emph{session
-order}, and a carefully choosen set of logical operators. \name expects
-application programmer to use the contract language to describe the set of legal
-executions allowed over the replicated data types.  
-% The executions descriptions are similar to the high-level axiomatic
-% description of relaxed memory models~\cite{Demange2013}.
-Subsequently, a contract enforcement system logically analyses contracts using
-an SMT solver~\cite{z3} and automatically tags each operation with the weakest
-store-level consistency guarantee that is required to satisfy application-level
-consistency requirements. In addition, the system statically reports an error if
-the given application-level consistency property cannot be satisfied by the
-store.
+To address these issues, we describe \name, a declarative programming model and
+implementation for eventually consistent stores. The key novelty of \name is an
+expressive \emph{contract language} to declare and verify fine-grained
+application-level consistency properties. The programmer uses the contract
+language to axiomatically describe the set of legal executions allowed over the
+replicated data type. Contracts are constructed using primitive consistency
+relations such as \emph{visibility} and \emph{session order} along with
+standard logical and relational operators. A \emph{contract enforcement system}
+automatically maps each operation to a particular consistency level available
+on the store. In addition, the system statically reports an error if the given
+application-level consistency property cannot be satisfied by the store.
 
@@
 \item We introduce \name, a language for declaratively describing eventually
-  consistent programs that manipulate replicated data types. Contracts are
-  used to specify fine-grained application-level consistency properties, and
-  are analyzed to assign the most efficient and sound store consistency
-  level to the corresponding operation.
+	consistent programs that manipulate replicated data types. Contracts are used
+	to specify fine-grained application-level consistency properties, and are
+	analyzed to assign the most efficient and sound store consistency level to
+	the corresponding operation.
 
-\item \name supports coordination-free transactions over arbitrary
-  datatypes.  We extend our contract language to express fine-grained
-  transaction isolation guarantees, and utilize the contract enforcement
-  system to automatically assign the correct and most efficient isolation
-  level for a transaction.
+\item \name supports coordination-free transactions over arbitrary datatypes.
+	We extend our contract language to express fine-grained transaction isolation
+	guarantees, and utilize the contract enforcement system to automatically
+	assign the correct and most efficient isolation level for a transaction.
 
@@
 
-\item \name is implemented as a shim layer over (and importantly without
-  modifying) Cassandra, a well-known general-purpose data store.  An
-  extensive experimental evaluation over a collection of real-world
-  applications, including a Twitter-like micro-blogging site and an
-  eBay-like auction site illustrates the practicality of our approach.
+\item \name is implemented as a shim layer over (and importantly, without
+	modifying) Cassandra~\cite{Cassandra}, a well-known general-purpose data
+	store.  An extensive experimental evaluation over a collection of real-world
+	applications, including a Twitter-like micro-blogging site and an eBay-like
+	auction site illustrates the practicality of our approach.
 
@@
 Section~\ref{sec:opsem} presents an operational semantics.
-Section~\ref{sec:txns} introduces transaction contracts and
-classification. Section~\ref{sec:impl} describes the implementation and goes
-into detail about the optimizations needed to make the system practical.
+Section~\ref{sec:txns} introduces transaction contracts and classification.
+Section~\ref{sec:impl} describes the implementation and goes into detail about
+the optimizations needed to make the system practical.
 Section~\ref{sec:results} discusses experimental evaluation.

### 2014-10-27 motivation.tex
Motivation version 2

--- 2014-10-25 motivation.tex
@@
 
-In this section, with help of a bank account example, we motivate the
-need for a framework to express and enforce fine-grained
-application-level consistency requirements under weak consistency
-models. 
-% \GK{The following text needs to be polished}
+In this section, we motivate the need and benefit of declarative reasoning under
+eventual consistency with the help of an example. We begin with a description of
+our system and programming model, and introduce the primitive relations that
+constitute an essential part of our contract language.
 
-% We assume a system model where the eventually consistent database is organized
-% as a collection of key-indexed, object-valued tables. The objects stored in the
-% tables are instances of the same replicated data type. The database is composed
-% of a number of \emph{replicas}, and each object is fully replicated across all
-% the replicas. Under this model, a client request for an operation with no
-% application-level consistency requirement can be serviced if at least one of
-% the replicas is reachable. Thus, with the addition of new replicas, and placing
-% them close to the clients, leads to improved throughput and reduced latencies.
+\subsection{System Model}
+\label{sec:sysmod}
 
-\subsection{Bank account database}
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{Figures/SystemModel}
+\caption{Quelea system model.}
+\label{fig:sysmod}
+\end{figure}
 
-Suppose our goal is to implement a highly available bank account
-service on top of an eventually consistent data store. For the sake of
-presentation, we restrict ourselves to bank account objects (records)
-with following simple schema:
+Figure~\ref{fig:sysmod} provides a schematic diagram of our system model. The
+distributed store is composed of a collection of \emph{replicas}, each of which
+stores a set of \emph{objects} ($x,y,\ldots$). We assume that every object is
+replicated at every replica in the store. The state of an object at any replica
+is the set of all updates (\emph{effects}) performed on the object. For
+example, the state of $x$ at replica 1 is the set composed of effects $w^x_1$
+and $w^x_2$.
 
-\begin{codesql}
-Create Table BankAccount (
-  accountNo int PRIMARY KEY,
-  balance float CHECK (balance >= 0))
-\end{codesql}
+Each object is associated with a set of \emph{operations}. The clients interact
+with the store by invoking the operations on objects. The sequence of operations
+invoked by a particular client on the store is called a \emph{session}. The
+data store is typically accessed by a large number of clients (and hence
+sessions) concurrently. Importantly, the clients are oblivious to which replica
+an operation is applied to; the data store may choose to route the operation to
+the closest replica for minimizing latency, or route it to a different replica
+in order to balance the load in the system. For example, the operations $foo$
+and $bar$ invoked by the same session on the same object, might end up being
+applied to different replicas because replica 1 to which $foo$ was applied to
+might be unreachable when invoking $bar$.
 
-\noindent and with support for the following \emph{operations}:
+Suppose an operation $foo$ is invoked on a object $x$ with arguments $arg_1$ at
+replica 1. The operation $foo$ simply \emph{reduces} over the current set of
+effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
+$v1$ that is sent back to the client, and emits a \emph{single} new effect
+$w^x_4$, which is appended to the state $x$ at replica 1. Thus, every operation
+is evaluated over a \emph{snapshot} of the state of the object on which it is
+invoked. In this case, we say that the effect $w^x_4$ \emph{witnessed} $w^x_1$
+and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
+written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
+$\visZ$ is the visibility relation between the effects. Visibility is an
+irreflexive (an operation cannot see its own effect) and asymmetric relation,
+and only relates effects produced by operations on the same object.
 
-\begin{codehaskell}
-type AccountNo = Int
-type Amount = Float
-getBalance :: AccountNo -> Amount
-deposit    :: AccountNo -> Amount -> ()
--- returns false if the account has insufficient balance
-withdraw   :: AccountNo -> Amount -> Bool
-\end{codehaskell}
+Read-only operation execution is similar except that they do not produce a new
+effect. The effect added to a particular replica is asynchronously sent to
+other replicas. We assume that \emph{eventually}, an effect applied at one
+replica will be merged into all other replicas. The two effects $w^x_4$ and
+$w^x_5$ that arise from the same session are said to be in \emph{session order}
+(written logically as $\so{w^x_4}{w^x_5}$). Session order is a irreflexive,
+transitive relation. The effects $w^x_4$ and $w^x_5$ arising from operations
+applied to the same object $x$ are said to be under the \emph{same object}
+relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every
+effect with the operation that generated the effect with the help of a relation
+$\operZ$. In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$
+hold. For simplicity, we assume that all the operation names across all object
+types are distinct.
 
-\begin{figure}[t]
+The system model described allows all the inconsistencies associated with
+eventual consistency. The goal of our work is to identify the precise
+consistency level for each operation such that the application-level
+constraints are not violated. In the next section, we will concretely describe
+the challenges associated with constructing a consistent bank account on top of
+eventually consistent data stores. Subsequently, we will illustrate how our
+contract language, armed with the primitive relations $\visZ$, $\soZ$,
+$\sameobjZ$ and $\operZ$, mitigates these challenges.
+
+\subsection{Eventually Consistent Bank Account}
+
+Suppose our goal is to implement a highly available bank account on top of an
+eventually consistent data store, with the \emph{integrity} constraint that the
+balance should never drop below 0. Let us assume that the bank account object
+provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
+with the assumption that the withdraw fails if the account has insufficient
+balance to complete the withdraw operation. Our goal is to choose the correct
+consistency level for each operation such that (1) the balance never drops
+below zero and (2) the \cf{getBalance} operation never incorrectly show a
+negative balance. To be able to do this, we need to consider the anomalies that
+could arise under eventual consistency.
+
+\begin{figure}
 \centering
-\subfigure[Unsafe Withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation4}}
+\subfigure[Unsafe Withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
 \hfill
@@
 \subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
-%\hfill
-%\subfigure[Monotonicity violation]{\label{fig:monotonicityAnomaly}\includegraphics[width=0.36\columnwidth]{Figures/Motivation3}}
 \caption{Anomalies possible under eventual consistency for the get balance operation.}
@@
 
-The sole attribute of a \cf{BankAccount} object, identified by its unique
-\cf{accountNo}, is its current \cf{balance}. A \cf{deposit} operation
-on a \cf{BankAccount} adds some \cf{Amount} to its \cf{balance},
-whereas a \cf{withdraw} operation subtracts. Operation \cf{getBalance}
-returns the current \cf{balance} in the account. A \cf{withdraw}
-operation succeeds and returns \cf{True} only if there is sufficient
-balance in the account. Otherwise, it fails and returns \cf{False}.
+Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
+that all operations in the figure are on the same bank account object with
+initial balance as 0. Session 1 performs a \cf{deposit} of 100, followed by a
+\cf{withdraw} of 80 in the same session. The \cf{withdraw} operation witnesses
+the deposit and succeeds\footnote{Although visibility and session order
+relations relate effects, we have abused the notation in these examples to
+relate operations, with the idea that the relations relate the effect emitted
+by those operations}. Subsequently, session 2 perform a \cf{withdraw}
+operation, but importantly, due to eventual consistency, only witnesses the
+\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
+\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
+constraint. A subsequent \cf{getBalance} operation, that witnesses all the
+previous operations and reports a negative balance.
 
-If we were to use a traditional database system to store
-\cf{BankAccount} objects, all that needs to be done is to define the
-\cf{BankAccount} schema, as shown above. The database management
-system effectively sequentializes all accesses to the database, and
-automatically enforces the required integerity constraints on the
-data, so that it is never possible for a \cf{getBalance} operation to
-return a negative balance, or for a \cf{withdraw} to succeed without
-having sufficient balance in the account. However, under eventual
-consistency, such anomalies are bound to occur as demonstrated in the
-following subsection.
+It is easy to see that the only way to ensure that integrity constraint on
+\cf{balance} is to prevent concurrent withdrawals on the same bank account
+object. This can be done by insisting that \cf{withdraw} be executed with
+strong consistency. Although this prevents the integrity constraint from being
+violated, a strongly consistent \cf{withdraw} itself is not sufficient to
+prevent \cf{getBalance} from ever reporting a negative balance to the end user.
 
-\subsection{Anomalies}
+Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}, which
+consists of three concurrent sessions performing a \cf{deposit}, a
+\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
+account object. Assume that initial balance in the account is zero. As the
+\cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2, can
+witnesses the effects of \cf{deposit(100)} from session 1, leading it to
+rightfully conclude that the current balance is 100. Consequently,
+\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
+operation subsequently becomes visible to the \cf{getBalance} in Session 3.
+However, \cf{getBalance}, which only witnesses the effect of \cf{withdraw} from
+session 2, but not the \cf{deposit} from session 1, reports the balance of
+negative 50 to the user.
 
-Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}.
-Assume that all operations in the figure are on the same
-\cf{BankAccount} object. Execution contains two concurrent client
-sessions, where one of the sessions executes a \cf{withdraw} followed
-by a \cf{getBalance} operation. We say that \c{withdraw} precedes
-\cf{getBalance} in \emph{session order}, and visualize it as a
-directed \cf{so} edge from the former to the later. Edges labelled
-\cf{vis} in the figure capture the \emph{visibility} relation between
-operations; an operation is said to be \emph{visible} to other
-operation if the later can witness the effects of the former. Two
-operations that are not related either via \cf{vis} or \cf{so} are
-said to be concurrent. The figure shows an execution where concurrent
-\cf{withdraw}s in different sessions succeed independently, but a
-subsequent \cf{getBalance} operation that witnesses both the
-\cf{withdraw}s reports a negative balance.
-
-It is easy to see that the only way to ensure that integrity
-constraint on \cf{balance} is to prevent concurrent withdrawls on the
-\cf{BankAccount}. This can be done by insisting that \cf{withdraw} be
-executed with strong consistency. Although this prevents unsafe
-withdrawls from going through, a strongly consistent \cf{withdraw}
-itself is not sufficient to prevent \cf{getBalance} from ever
-reporting a negative balance to the end user.
-
-Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
-which consists of three concurrent sessions performing a \cf{deposit},
-a \cf{withdraw}, and a \cf{getBalance}, respectively, on the same
-\cf{BankAccount}. Assume that initial balance in the account is zero.
-As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in session
-2, can witnesses the effects of \cf{deposit(100)} from session 1,
-leading it to rightfully conclude that the current balance is 100.
-Consequently, \cf{withdraw(50)} is performed safely. The effect of
-this \cf{withdraw} operation subsequently becomes visible to the
-\cf{getBalance} in Session 3. However, \cf{getBalance}, which can only
-see the effect of \cf{withdraw} from session 2, but not \cf{deposit}
-from session 1, reports the balance of negative 50 to the user.
-
-Negative balance is not the only inconsistency that the user is
-exposed to under an eventually consistent bank account. Figure
-~\ref{fig:missingUpdateAnomaly} shows an execution where
-\cf{getBalance} operation in a session does not witness the effects of
-\cf{withdraw} operation performed in the same session, possibly
-because \cf{getBalance} was served by a replica that hasn't yet merged
-the \cf{withdraw} effect, which could have been served by a different
+Negative balance is not the only inconsistency that the user is exposed to
+under an eventually consistent bank account. Figure
+~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
+operation in a session does not witness the effects of \cf{withdraw} operation
+performed in the same session, possibly because \cf{getBalance} was served by a
+replica that has not yet merged the \cf{withdraw} effect, served by a different
 replica. This anomaly leads the user to incorrectly conclude that the
@@
 
-Although it is easy to understand the reasons behind the occurance of
-the aforementioned negative balance and missing update anomalies, the
-fix is not quite apparent. Making \cf{getBalance} a strongly
-consistent operation is definitely sufficient to avert anomalies, but
-is it really necessary? Given that a strongly consistent operation is
-significantly expensive, both in terms of latency and monetary value, we ideally
-would not want to request strong consistency, unless there are
-no alternatives available. Unfortunately, this requires us to
-understand subtle differences in semantics of various kinds of weak
-consistency alternatives, such as timeline consistency and session
-guarantees, offered by off-the-shelf eventually consistent data
-stores. This process is error-prone, and becomes further complicated
-in presence of transactions, as demonstrated in Sec.
-~\ref{sec:transactions}.
-
-\subsection{Implementation in Quelea}
-
-\name automates this error-prone task by letting the programmer
-declare application-level consistency constraints as \emph{contracts},
-obviating the need to understand store-level consistency guarantees,
-and relieving programmer from having to match application
-requirements with appropriate store-level alternatives. The
-cornerstone of declarative reasoning in \name is the programming model
-built around operation-based CRDTs~\cite{shapiroCRDT}. A replicated
-data type in \name is defined as a tagged union type, where each tag
-is the name of a unique operation defined on the data type. 
-%types in Quelea are defined as reductions over the set of effects.
-%The following data type definitions capture the operations and effects
-%on the bank account and user index objects.
-For example, following is the definition of the \cf{BankAccount}
-data type in \name:
-\begin{codehaskell}
-data BankAccount =  Deposit Float | Withdraw Float 
-                 | GetBalance
-\end{codehaskell}
-
-In Quelea, an object state is nothing but the set of effects (called
-the \emph{context}) on this object. As such, \cf{Deposit} and
-\cf{Withdraw} effects simply include the amount deposited or
-withdrawn, and \cf{AddUser} effect includes the user name. Similarly,
-on the user index table, where the key is user name, the value is
-simply an effect (\cf{AddName}) binding some user id at this key.
-\cf{GetBalance} and \cf{GetName} correspond to the corresponding
-read-only operations. The complete definition of the operations on
-bank account and user index objects is given below.
-
-\begin{codehaskell}
--- Get balance takes no arguments and read-only
-getBalance :: [BankAcc] {- context -} -> () {- args -}
-  -> (() {- ret val -}, Maybe BankAcc {- effect -})
-getBalance ctxt _ = (sum [v | Deposit v (- ctxt]
-        - sum [v | Withdraw v (- ctxt], Nothing)
-
--- withdraw returns True on success
-withdraw :: [BankAcc] -> Float -> (Bool, Maybe BankAcc)
-withdraw c v = if sel1 $ getBalance c () >= v
-               then (True, Just $ Withdraw v)
-               else (False, Nothing)
-
-deposit :: [BankAcc] -> v -> ((), Maybe BankAcc)
-deposit _ v = ((), Just $ Deposit v)
-
-addUser :: [BankAcc] -> UserName -> ((), Maybe BankAcc)
-addUser _ n = ((), Just $ AddUser n)
-
-addName :: [UserIdx] -> UserID -> (Bool, Maybe UserIdx)
-addName [] uid = (True, Just $ AddName uid)
-addName x:_ _ = (False, Nothing)
-
-getName :: [UserIdx] -> ()
-        -> (Maybe UserName, Maybe UserIdx)
-getName [] _ = (Nothing, Nothing)
-getName (AddUser uid:_) _ = (Just uid, Nothing)
-\end{codehaskell}
-
-The definitions are a straight forward encoding of the expected behavior.
-Quelea programming model provides the programmer complete freedom regarding the
-semantics and desired convergence property of the replicated data type, and
-indeed, we can encode the well-known CRDTs~\cite{SSS} in a declarative fashion.
-Observe that the operation definitions presented here only capture the
-convergence properties of the replicated data type, and are not concerned with
-the consistency properties, which is expressed through the contract language.
+Although it is easy to understand the reasons behind the occurrence of the
+aforementioned negative balance and missing update anomalies, the fix is not
+quite apparent. Making \cf{getBalance} a strongly consistent operation is
+definitely sufficient to avert anomalies, but is it really necessary? Given
+that a strongly consistent operation is significantly expensive, both in terms
+of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally would
+not want to request strong consistency, unless there are no alternatives
+available. Unfortunately, this requires us to understand subtle differences in
+semantics of various kinds of weak consistency alternatives, offered by the
+data stores.
 
@@
 
-The contract for the strongly consistent \cf{addName} operation is given below:
+\name automates this error-prone task by letting the programmer declare
+application-level consistency constraints as \emph{contracts}. The key
+intuition is that, in the case of the bank account example, any execution that
+does not exhibit the anomalies described in the previous section is a
+\emph{well-formed} execution on the bank account object. In the same vein as
+axiomatic memory models for shared memory multicore
+processors~\cite{Demange2013,Sarkar2011}, the programmer axiomatically
+describes the set of legal executions on top of each replicated date type over
+a trace of operation invocations on that type. Quelea ensures that any
+execution on top of such datatypes always belongs to the set of well-formed
+executions on that type.
+
+In the bank account example, we observed that, in order to preserve the
+integrity constraint, the \cf{withdraw} operation must be strongly consistent.
+That is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible
+to $b$ or vice versa. Utilizing the primitive relations described in
+Section~\ref{sec:sysmod}, we express this application-level consistency
+requirement as a contract over \cf{withdraw}:
 \begin{smathpar}
 \begin{array}{l}
-\rsf{addNameCtrt} = \forall (a:\rsf{AddName}). \\
-\qquad \sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} \vee \vis{\cureff}{a} \vee a = \cureff
+\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
+\end{array}
+\end{smathpar}
+In the above contract, $\cureff$ stands for the effect emitted by the
+\cf{withdraw} operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an
+effect which was emitted by a \cf{withdraw} operation i.e)
+$\oper{a}{\rcf{withdraw}}$ holds. The contract says that, if the current
+operation emits an effect $\cureff$, then for any operation $a$ which was
+emitted by a \cf{withdraw} operation, it is the case that $a = \cureff$ or $a$
+is visible to $\cureff$ or vice versa. Any execution on a bank account object
+that preserves the above contract for a \cf{withdraw} operation is said to have
+a correct implementation of \cf{withdraw}.
+
+For the \cf{getBalance}, we construct a contract  precisely to prevent the
+anomalies described previously. The contract is shown below:
+\begin{smathpar}
+\begin{array}{l}
+\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
+\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
+\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
 \end{array}
@@
 
-In the above definition, $\cureff$ represents the current effect --- the effect
-emitted by the \cf{addName} operation. The contract simply states our
-high-level observation; for any effect $a$ which is also an \cf{AddName} effect
-on the same object, it must be the case that $a$ is visible to $\cureff$ or
-vice verse, or $a$ is $\cureff$. Here, $\sameobjZ$ and $\visZ$ are primitive
-relations. This contract ensures that any two \cf{addName} operations will be
-seen by each other. The contract for withdraw is similar to the
-\cf{addNameCtrt}.
+If a \cf{withdraw} $b$ is visible to the get balance operation, then all
+\cf{deposit} operations $a$ visible to the \cf{withdraw} should be visible to
+the get balance operation. This prevents negative balance anomalies. The syntax
+$c : \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
+emitted either by a \cf{deposit} operation or by a \cf{withdraw} operation. The
+contract language provides relational operators to compose relations. The
+syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
+last line of the above contract says that if a \cf{deposit} or a \cf{withdraw}
+precedes a \cf{getBalance} operation in session order, and is applied on the
+same object as the \cf{getBalance} operation, then it must be the case that the
+\cf{getBalance} operation witnesses the effects of the preceding operations.
+Since the deposit operation does not have any restrictions, its contract is
+simply $\true$.
 
-Since the deposit operation does not have any restrictions, its contract is
-simply $\true$. Same is the case for \cf{addUser} operation. The contract for
-get balance operation is:
-\begin{smathpar}
-\begin{array}{l}
-\rsf{getBalCtrt} = \forall (a:\rsf{Deposit}), (b:\rsf{Withdraw}). \\
-\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\qquad \vee (\soZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff} \\
-\qquad \vee (\soZ \cap \sameobjZ) (b,\cureff) \Rightarrow \vis{b}{\cureff}
-\end{array}
-\end{smathpar}
+\subsection{From contracts to implementation}
 
-If a withdraw $b$ is visible to the get balance operation, then all deposit
-operations $a$ visible to withdraw should be visible to the get balance
-operation. This prevents negative balance anomalies. The rest of the contract
-says that a get balance operation must witness previous deposit and withdraw
-operations on the same object in the same session. This prevents missed update
-anomalies.
+Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
+the application-level consistency requirements, and make no reference to the
+semantics of the underlying store. To write such contracts, a programmer only
+needs to reason about the semantics of the application under the \name system
+model. The mapping of application-level consistency requirements to appropriate
+store-level guarantees is done automatically behind the screen.
 
-Similar to contracts on operations, Quelea supports contracts on transactions.
-The contract on the \cf{getBalanceName} transaction is given below:
-\begin{smathpar}
-\begin{array}{l}
-\rsf{getBalanceNameCtrt} = \forall (a:\rsf{GetName}), (b:\rsf{GetBalance}), \\
-\quad (c:\rsf{AddName}), (d:\rsf{AddUser}). ~\trans{a}{b}{c}{d} \wedge \so{a}{b} \\
-\qquad \wedge ~\vis{c}{a} \wedge \sameobj{d}{b} \Rightarrow \vis{d}{b}
-\end{array}
-\end{smathpar}
+How might one go about ensuring that any execution adheres to the contract? The
+challenge is that contract is a declarative characterization of correctness of
+execution, while what we need is an operational procedure for \emph{enforcing}
+the contracts. One strategy is to adopt speculative execution as espoused by
+other eventually consistent systems, where operations are tentatively applied
+as they are received from the client or other replicas. We can maintain a
+runtime manifestation of the axiomatic characterization and check the
+well-formedness at runtime, and rollback the execution if
+ill-formed~\cite{rxcml}. However, the overhead of state maintenance and the
+complexity of user-defined contracts (unlike simple tests such as cyclic
+dependence) make this technique unviable.
 
-$\trans{a}{b}{c}{d}$ says that the action pairs $a,b$ and $c,d$ are in the same
-transaction, and the two transactions are distinct. The contract says
-specifically forbids the anomaly presented in
-Figure~\ref{fig:monotonicityAnomaly}. This is the desired semantics for the
-\cf{getUserName} transaction.
-
-The programmer simply defines such contracts on the operations and the
-transactions. Quelea logically analyzes the contracts, and maps the
-corresponding operation to the precise store consistency and isolation level.
-Thus, Quelea equips the programmer with a declarative model for reasoning and
-expressing eventually consistent programs.
+Instead, Quelea statically analyzes the contracts using an SMT solver, and maps
+each operation to a particular store-level consistency under which the contract
+would be preserved. In particular, the key idea is that the system implementor
+axiomatically describes the store semantics in the same contract language as
+the one used to express the contract on the operation. We translate the
+contracts down to decidable subset of first-order logic~\cite{}, and discharge
+proof obligation of checking whether a particular store consistency level is
+sufficient for the given contract with the help of an SMT solver. The next
+section describes the contract language and classification.

### 2014-10-27 motivation.tex

@@
 \centering
-\subfigure[Unsafe Withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
+\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
 \hfill
-\subfigure[Negative Balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
+\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
 \hfill

### 2014-10-27 contract-language.tex
Sections 3 and 4.

\section{Contract Language}
\label{sec:contract-lang}

% Contract Language Syntax
% ------------------------
\input{contract-syntax.tex}

In this section, we formalize the contract language of \name, and
describe our contract classification scheme, which analyzes a contract
and maps it to the weakest store consistency level sufficient to
satisfy its consistency requirements.

% classifies contracts on the basis of the weakest
% store-level consistency guarantee

\subsection{Syntax}

The syntax of our core contract language is shown in Fig.
~\ref{fig:contract-lang}. The language is based on first-order logic
(FOL), and admits prenex quantification over typed effect variables.
We use a special effect variable ($\cureff$) to denote the effect of
\emph{current operation} - the operation for which a contract is being
written. The type of an effect is simply the label of the operation
(eg: \cf{withdraw}) that engendered the effect. We admit disjuntion in
types to let an effect variable range over effects of multiple
operations. Since each replicated data type comes with a unique set of
operations (therefore, unique set of types for effects), our contract
language is implictly parameterized over a replicated data type
specification.
\GK{I think sec 2. is the right place to introduce the idea of rdt
spec via the bank account example.}

Quantifier-free propositions in our contract language are
conjunctions, disjunctions and implications of predicates expressing
relations between pairs of effect variables. Syntactic class of
relations is seeded with primitive $\visZ$, $\soZ$, and $\sameobjZ$
relations, and also admits derived relations that are expressible as
union, intersection, or transitive closure\footnote{Strictly speaking,
$R^{+}$ is not the transitive closure of $R$, as transitive closure is
not expressible in FOL. Instead, $R^{+}$ in our language denotes
\emph{a} superset of transitive closure of $R$. Formally, $R^{+}$ is
any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow
R'(x,z)$} of primitive relations. Commonly used derived relations are
the \emph{same object session order} ($\sooZ$), and the the
\emph{happens before order} ($\hboZ$) defined as following:
\begin{smathpar}
\begin{array}{lcl}
{\sf soo} & = &  {\sf so} \cap {\sf sameobj}\\
{\sf hbo} & = & ({\sf soo} \cup {\sf vis})^{+}\\
\end{array}
\end{smathpar}

Note that our contract language (stripped of its syntactic sugar) is a
carefully chosen subset of first-order logic that is known to be
decidable~\cite{epr}. In particular, for any two contracts ($\cv_1$
and $\cv_2$) in our language, the implication check $\cv_1 \Rightarrow
\cv_2$ is decidable, allowing us to automatically deduce that $\cv_2$
is weaker than $\cv_1$. We formally define the \emph{weaker than}
relation among contracts as following: 
\begin{definition}
A contract $\cv_2$ is said to be weaker than $\cv_1$ (denoted $\cv_2
\le \cv_1$ ) if and only if:
\begin{center}
$\Delta \vdash \cv_1 \Rightarrow \cv_2$
\end{center}
\end{definition}
\noindent The context $\Delta$ in the above definition captures assumptions
about the nature of primitive relations, such as the assumption that
$\Rvis$ and $\Rso$ are irreflexive.

This ability to statically compare contracts is key to contract
classification and subsequent enforcement. This is described in the
following subsection.

\subsection{Contract Classification}

Application-level consistency requirements expressed by \name
contracts can have varying impact on the available of the data store
trying to meet the requirements. We identify three major classes of
contracts that differ on this parameter: 
% For instance, total-order requirement
% expressed by the \cf{withdraw} contract (Sec. ~\ref{sec:motivation})
% imposes global coordination among all replicas, and consequently makes
% the store unavailable for any concurrent session. In contrast, the
% trivial (\cf{true}) contract of the \cf{deposit} operation expresses
% no special consistency requirements; any replica can serve a deposit
% request immediately, and the store remains available to concurrent
% sessions. Based on how a contract affects the availability of the data
% store, we classify it into one of the following three categories:

\begin{itemize}
\item \textbf{Unavailable}: These are the contracts that are
satisfiable only under strong consistency. The need for global
coordination to satisfy such contracts makes the store unavailable for
concurrent sessions. An example is the total-order contract on
\cf{withdraw} (Sec. ~\ref{sec:motivation}).
\item \textbf{Sticky-available}: These are the contracts that require
see a \emph{causally consistent} snapshot of the store that
\emph{must} include some or all of the operations from the same
session. By applying all operations of a session to the same logical
replica (colloquially, \emph{sticking} a session to a replica), it is
possible for a store to satisfy such contracts, while remaining highly
available. The \cf{getBalance} contract that is sufficient to prevent
the missing update anomaly is a sticky available contract.
\item \textbf{Highly-available}: Contracts that only require to see
\emph{some} causally consistent snapshop of the store are said to
require \emph{causal visiblity} guarantee. This guarantee can be
achieved by forcing every replica to reveal only those effects whose
preceding effects are already included in the replica. Consequently,
such contracts can always be satisfied by the store, letting it remain
highly available for other requests.
\end{itemize}

It is important to note that, although finer distinctions can be made
between availability implications of contracts, the above
classification scheme is the one that is known to have practical
significance\cite{bailisHAT}. More importantly, the consistency levels
described in the classification scheme, namely strong consistency
({\sf sc}), causal consistency ({\sc cc}), and causal visibility ({\sf
cv}), can be implemented efficiently on most off-the-shelf eventually
consistent data stores\cite{...}. By determining which of the
aformentioned classes a contract falls into, the corresponding
implementation of the consistency level on the data store can be used
to satisfy the contract efficiently. Nonetheless, our classification
scheme is extendible, meaning that a new availability class can be
carved out of existing ones with the advent of new data stores
supporting efficient enforcement of finer consistency levels.

To classify contracts into one of the aforementioned availability
classes, we need to determine if the corresponding consistency
guarantee is sufficient to satisfy the contract. Towards this end, we
encode the {\sf sc}, {\sf cc}, and {\sf cv} consistency guarantees as
formulas in our contract language ($\hboZ$, the happens-before order,
was defined previously in the same section):
\begin{smathpar}
\stretcharraybig
\begin{array}{rcl}
{\sf sc} & = 		 & \forall a. (\hbo{a}{\cureff} \vee \hbo{\cureff}{a} \vee a = \cureff) \\
         & & \conj (\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}) \\
				 & & \conj (\hbo{\cureff}{a} \Rightarrow \vis{\cureff}{a}) \\
{\sf cc} & = & \forall a. \hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff} \\
{\sf cv} & = & \forall a,b. \hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
\end{array}
\end{smathpar}
Comparing ($\le$) a contract formula against the above formulas of
{\sf sc}, {\sf cc}, and {\sf cv} will let us determine if the contract
can be satisfied under strong consistency, causal consistency, and
causal visiblity consistency guarantees, respectively. Note that {\sf
sc}, {\sf cc} and {\sf cv} are themselves totally ordered with respect
to the $\le$ relation:
\begin{center}
$\cvc \le \ccc \le \scc$
\end{center}
This concurs with the intuition that any contract satisfiable under
$\ccc$ or $\cvc$ is satisfiable under $\scc$ , and any contract that
is satisfiable under $\cvc$ is satisfiable under $\ccc$. Nonetheless,
we determine the availability class of a contract based on the
\emph{weakest} guarantee (among $\cvc$, $\ccc$, and $\scc$) required
to satisfy the contract. The classification scheme is presented
formally as rules in Figure ~\ref{sem:classfiy}.
% Contract classification rules
% ------------------------------
\input{contract-classification}
The rules to classify contracts into availability classes are
straightforward. The remaining rule judges a contract as {\sf
WellFormed} if and only if it can be satisfied under strong
consistency. We only consider well-formed contracts for classification
and subsequent enforcement at run-time. The rest are reported as
errors.

As a result of our automatic classification scheme, a store that can
enforce $\scc$, $\ccc$, and $\cvc$ contracts can also enforce any
well-formed \name contract by determining its availability class.  The
following section formally describes operational semantics of one such
store, along with a meta-theoretic result that certifies the soundness
of our classification-based contract enforcement strategy.

### 2014-10-27 motivation.tex

@@
 
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{Graphs/BankAccount_Th_vs_Lat}
+\caption{Bank Account: Throughput vs Latency. Quelea is our implementation.
+Strong and Quelea are correct. High. Avail. and causal are incorrect (withdraw
+not strong). 20\% withdraw, 20\% deposit, 60\% getBalance.}
+\label{grf:BankAcc_Th_vs_Lat}
+\end{figure}
+
 Although it is easy to understand the reasons behind the occurrence of the
@@
 definitely sufficient to avert anomalies, but is it really necessary? Given
-that a strongly consistent operation is significantly expensive, both in terms
-of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally would
-not want to request strong consistency, unless there are no alternatives
-available. Unfortunately, this requires us to understand subtle differences in
-semantics of various kinds of weak consistency alternatives, offered by the
-data stores.
+that a strongly consistent operation is significantly expensive
+(Figure~\ref{grf:BankAcc_Th_vs_Lat} \KC{Should I include the numbers here? If
+yes, I'll expand the text.}), both in terms of availability and monetary
+overheads~\cite{DynamoDB,Pileus}, we ideally would not want to request strong
+consistency, unless there are no alternatives available. Unfortunately, this
+requires us to understand subtle differences in semantics of various kinds of
+weak consistency alternatives, offered by the data stores.
 
@@
 
-\subsection{From contracts to implementation}
+\subsection{From Contracts to Implementation}
 

### 2014-10-27 core-lang.tex

\section{Operational Semantics}
\label{sec:core-opsem}

We demonstrate our ideas using a replicated data store formalized as a
triple $\E;\Theta;\Sigma$, where $\E$ denotes the execution state,
$\Theta$ maps replicas to their local state, and $\Sigma$, the session
soup, is the set of client sessions interacting with the store.
Operational semantics given in Figure~\ref{sem:oper} succinctly
capture the behaviour of the store that is relevant to demonstrate the
operational significance of contract classification. For the sake of
clarity, we only consider a single replicated object of well-defined
type (for eg: a replicated object of type {\sf bank\_account}) in our
formalization.  Our semantics are parametric over the specification of
this replicated data type.  Figure~\ref{sem:oper} formalizes
replicated data type (RDT) specification as tuple
$(\tau,\Ops,\Ctrts)$, where $\tau$ is the data type, $\Ops$ maps
labels ($op$) of operations on $\tau$ to their definitions, while
$\Ctrts$ maps them to their consistency contracts ($\cv$). The language of
contracts is described in Fig.~\ref{lang:contract}. The definition of
an operation is expected to be a lambda expression in the server-side
language\footnote{Since a replicated object is maintained by a
server, we call the language in which replicated data type operations
are defined as \emph{server-side language}.}, although we do not
enforce this in our formalization. We intentionally leave the
server-side language underspecified to demonstrate that our ideas are
independent of the choice of the programming language used to
implement the replicated data type.

% Core Operational Semantics
% --------------------------
\input{core-opsem.tex}

Clients interact with the store via multiple concurrent sessions. For
technical reasons, we tag every session in the session soup ($\Sigma$)
with a unique {\sf SessID} and the sequence number of next operation
in that session. As shown in Fig.~\ref{sem:oper}, we model a session
simply as a sequence of actions on the replicated object maintained by
the store, where an action is defined as an application of replicated
data type operation to a value. As with the server-side language, our
development is also orthogonal to the choice of the client-side
language, but we require both the languages to agree on the syntactic
class of values that replicated data type operations ($op$) produce
and consume.

An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
of \emph{effect}s , and $\visZ,\soZ \subseteq \EffSoup \times
\EffSoup$ are visibility and session order relations over effects,
respectively. An effect ($\eff$), modeled as a tuple $(s,i,op,v)$,
records the fact that $i^{th}$ action in session with {\sf SessID}
$s$, which is an operation $op$ on the replicated object, has been
successfully executed on some replica yielding a return value $v$.
Note that the combination of $s$ and $i$ uniquely identifies the
effect. Session order relation ($\soZ$) relates effects generated by
the same session. An effect $\eff = (s,i,op,v)$ is said to precede
another effect $\eff' = (s',i',op',v')$ in session order if and
only if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
\emph{available} at rest of the replicas eventually. The local state
of a replica $r$ (i.e., $\Theta(r)$) is a set of effects that are
currently available at $r$. An operation $op$ performed at replica $r$
can only witness the set of effects ($\Theta(r)$) visible at $r$. If we denote the
effect generated by the operation $op$ as $\eff_{op}$, then $\Theta(r)
\times \{\eff_{op}\} ~\subseteq~ \visZ$. Note that in presence of
multiple replicated objects, $\visZ$ only relates effects of
operations performed on a single object. For instance, a {\sf
withdraw} operation can witness (via $\visZ$) a {\sf deposit}
operation on the same bank account, but not a {\sf deposit} on a
different account. Often, in our formalization, we use $\visZ$ and
$\soZ$ binary relations to obtain a set of effects visible to a given
effect $\eff$, or set of effects that precede a given effect $\eff$ in
the session order. As a syntactic convenience, whenever $R$ is a
binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$
to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.

Rule $\rulelabel{Comm}$ captures inter-replica communication. The
basic guarantee provided by the store is causal consistency. That is,
an effect ($\eff$) will become \emph{available} at a replica $r$ 
only after all the effects that causally precede $\eff$ become
available at $r$. Observe that this condition is applicable even for
the effects that originate at replica $r$. It is helpful to think of
such effects being stored in a \emph{write buffer} until causal
consistency condition is met, at which point they can be merged with
the local state ($\Theta(r)$). Nevertheless, it is important to note
that that enforcing causal consistency does not require inter-replica
coordination.

Rule $\rulelabel{Oper}$ is an auxiliary reduction of the form

\begin{smathpar}
\auxred{\Theta} {(\E,\langle s,i,op(v) \rangle)} {r} {(\E',\eff)}
\end{smathpar}

\noindent Under the store configuration $\Theta$, the rule captures
the progress in execution (from $\E$ to $\E'$) due to the application
of operation $op$ to replica $r$ corresponding to a new effect $\eff$.
The rule first constructs a \emph{context} for the application from
the local state ($\Theta(r)$) of the replica, by
projecting\footnote{{\textsf{ctxt*}} is auxiliary function
\textsf{ctxt} extended straightforwardly to set of effects} relevant information from
effects in $\Theta(r)$. It then substitutes the definition
($\Ops(op)$) of the operation for its label ($op$), and relies on the
reduction relation ($\rdtredsto$) of the server-side language to
reduce the application $\Ops(op)(v,~ctxt)$ to a value $v'$.
Subsequently, the attributes of the new effect $\eff$ in the execution
state. In particular, we record that the effects in $\Theta(r)$ are
visible to $\eff$, and $\eff$ follows in session order ($\soZ$) every
effect in the same session $s$ with a smaller sequence number.

If the operation $op$ is highly-available (rule $\rulelabel{HA}$), we
simply apply the operation to any replica $r$. Since the store
provides causal visibility, highly available operations are
satisfiable under any replica. If the operation is sticky-available
(rule $\rulelabel{SA}$), the operation can be applied to a replica $r$
only if it already contains the effects of all the previous performed
in the same session. This guarantee can be satisfied by applying all
operations from the same session to the same logical copy of the
database. If such a logical copy is untenable, then the operation
might block. Since the store is assumed to converge eventually, the
blocked sticky-available operation guaranteed to unblock eventually.

An unavailable operation expects strong (sequential) consistency.
That is, universe of all effects ($\EffSoup$) in an execution ($\E$)
must be partitionable into a set of effects that \emph{happened
before} $\eff$ and another set that \emph{happened after} $\eff$,
where $\eff$ is the effect generated by an unavailable operation. The
rule $\rulelabel{UA}$ enforces this sequencing in two steps; firstly,
it insists that the the unavailable operation ($op$) witness effects
of all operations executed so far by requiring the global set of
effects $\EffSoup$ to be a subset of local state ($\Theta(r)$) of the
replica ($r$) executing $op$. Secondly, the rule requires the effect
($\eff$) generated by $op$ to be added to the local state of every
other replica in the store, so that further operations on these
replicas can witness the effect of $op$. Since both these steps
require global coordination among replicas, the store is
\emph{unavailable} during the time it is executing $op$. However, it
must be noted that executing unavailable operation does not entail the
convergence of local states of all replicas; it simply means that
there is one replica ($r$) whose local state is the upper bound of
local states of all replicas, and there exists a set containing
atleast one effect ($\eff$), which is their lower bound.

\subsection{Metatheory}

To help us state our main meta-theoretic result, we first define a
well-formedness condition over execution states:

\begin{definition}[Well-formed execution]
An execution state $\E = \Exec$ corresponding to an execution of
operations over a replicated datatype with specification $\rdtspec =
(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if 
$\forall \eff \in \EffSoup. ~\E \models \De{\cv[\eff/\cureff]}$, where $\cv =
\Ctrts(\operZ(\eff))$.
\end{definition}

Symbol $\models$ denotes the standard \emph{models} relation of
first-order logic.  A valid model for a first-order formula defines
the universe of discourse, and provides interpretations for free
symbols in the formula such that the formula is true under given
interpretations.  An execution state ($\E$) provides interpretations
for $\Rvis$ and $\Rso$ relations, while fixing the universe of
discourse to be the set of all generated effects ($\EffSoup$).
Intuitively, the well-formedness definition captures the idea that for
every generated effect, the contract on the corresponding operation is
satisfied under the given execution state.

Our main meta-theoretic result guarantees the soundness of our
operational semantics in enforcing all contractual obligations by
ensuring that every reduction step preserves the well-formedness of
execution state.

\begin{theorem}[Well-formedness preservation]
\label{lem:core-preservation}
If $\E$ is well-formed and $\E; \Theta; \Sigma \goesto \E'; \Theta';
\Sigma'$, then $\E'$ is well-formed.
\end{theorem}
% Lemmas
%\input{core-lemmas.tex}

The proof of the theorem, which we could not include here owing to
space constraints, can be found in our tech report~\cite{techrep}.

% It should be noted that Theorem ~\ref{lem:core-preservation} only
% proves the safety of our system. Proving its liveness requires us to
% axiomatize eventual consistency.

### 2014-10-27 intro.tex
intro and motive

@@
 
-Many real-world web services --- such as those built and maintained by Amazon,
-Facebook, Google, Twitter, etc. --- replicate application state and logic
-across multiple \emph{replicas} within and across data centers. Replication is
-intended not only to improve application throughput and reduce user-perceived
-latency, but also to tolerate partial failures without compromising overall
-service availability. Traditionally programmers have relied on \emph{strong
-consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
-serializability~\cite{Serializability} in order to build correct applications.
-While strong consistency is an easily stated property, it masks the reality
-underlying large-scale distributed systems with respect to non-uniform latency
-(i.e., availability) and network partitions~\cite{Brewer2000,Gilbert}. Indeed,
-modern web services, which aim to provide an "always on" experience,
-overwhelmingly favour availability and partition tolerance over strong
-consistency. To this end, several \emph{weak consistency} models such as
-eventual consistency, causal consistency, session guarantees, and timeline
-consistency have consequently been proposed.
+Many real-world web services --- such as those built and maintained by
+Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
+logic across multiple \emph{replicas} within and across data
+centers. Replication is intended not only to improve application throughput
+and reduce user-perceived latency, but also to tolerate partial failures
+without compromising overall service availability. Traditionally programmers
+have relied on \emph{strong consistency} guarantees such as
+linerarizability~\cite{Herlihy1990} or
+serializability~\cite{Serializability} in order to build correct
+applications.  While strong consistency is an easily stated property, it
+masks the reality underlying large-scale distributed systems with respect to
+non-uniform latency (i.e., availability) and network
+partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
+to provide an "always on" experience, overwhelmingly favour availability and
+partition tolerance over strong consistency. To this end, several \emph{weak
+  consistency} models such as eventual consistency, causal consistency,
+session guarantees, and timeline consistency have been proposed.
 
@@
 conflicting updates, and has to pay careful attention to avoid unwanted
-inconsistencies (e.g., negative balances in a bank account, or having an item
-appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
-the inconsistency leaks from the application and is witnessed by the user.
-Ultimately, the developer must decide the consistency level appropriate for a
-particular operation; this is understandably an error-prone process requiring
-intricate knowledge of both the application as well as the semantics and
-implementation of the underlying data store.  More often than not, only
-informal descriptions are available~\cite{}.  Picking the correct consistency
-level is critical not only for correctness but also for scalability of the
+inconsistencies (e.g., negative balances in a bank account, or having an
+item appear in a shopping cart after it has been
+removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
+application and is witnessed by the user.  Ultimately, the developer must
+decide the consistency level appropriate for a particular operation; this is
+understandably an error-prone process requiring intricate knowledge of both
+the application as well as the semantics and implementation of the
+underlying data store, which typically have only informal
+descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
+critical not only for correctness but also for scalability of the
 application. While choosing a weaker consistency level than required may
-introduce program errors and anomalies, choosing a stronger one than necessary
-level can negatively impact program scalability.
+introduce program errors and anomalies, choosing a stronger one than
+necessary can negatively impact program scalability.
 
@@
 
-Moreover, weak consistency hinders compositional reasoning about programs.
-While an application might be naturally expressed in terms of well-understood
-and expressive data types such as maps, trees, queues, or graphs,
-geo-distributed stores typically only provide a minimal set of data types with
-in-built conflict resolution strategies such as last-writer-wins (LWW)
-registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor, while
+Weak consistency also hinders compositional reasoning about programs.  While
+an application might be naturally expressed in terms of well-understood and
+expressive data types such as maps, trees, queues, or graphs,
+geo-distributed stores typically only provide a minimal set of data types
+with in-built conflict resolution strategies such as last-writer-wins (LWW)
+registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while
 traditional database systems enable composability through transactions,
-existing geo-distributed stores lack unrestricted transactional access to the
-data.  Working with geo-distributed stores thus requires application state to
-be suitably coerced to work within the capabilities of the store.
+geo-distributed stores typically lack unrestricted transactional access to
+the data.  Working in this environment thus requires application state to be
+suitably coerced to function using only the capabilities of the store.
 
-To address these issues, we describe \name, a declarative programming model and
-implementation for eventually consistent stores. The key novelty of \name is an
-expressive \emph{contract language} to declare and verify fine-grained
-application-level consistency properties. The programmer uses the contract
-language to axiomatically describe the set of legal executions allowed over the
-replicated data type. Contracts are constructed using primitive consistency
-relations such as \emph{visibility} and \emph{session order} along with
-standard logical and relational operators. A \emph{contract enforcement system}
-automatically maps each operation to a particular consistency level available
-on the store. In addition, the system statically reports an error if the given
-application-level consistency property cannot be satisfied by the store.
-
-The paper makes the following contributions:
+To address these issues, we describe \name, a declarative programming model
+and implementation for eventually consistent geo-distributed data
+stores. The key novelty of \name is an expressive \emph{contract language}
+to declare and verify fine-grained application-level consistency
+properties. The programmer uses the contract language to axiomatically
+specify the set of legal executions allowed over the replicated data
+type. Contracts are constructed using primitive consistency relations such
+as \emph{visibility} and \emph{session order} along with standard logical
+and relational operators. A \emph{contract enforcement system} automatically
+maps operations over the datatype to a particular consistency level
+available on the store, and provably validates the correctness of the
+mapping.  The paper makes the following contributions:
 
 \begin{itemize}
-
+\setlength{\itemsep}{2pt}
 \item We introduce \name, a language for declaratively describing eventually
@@
 	the corresponding operation.
-
 \item \name supports coordination-free transactions over arbitrary datatypes.
@@
 	assign the correct and most efficient isolation level for a transaction.
-
-\item Metatheory that certifies the soundness of our contract enforcement
+\item We provide metatheory that certifies the soundness of our contract enforcement
 	system, and ensures that an operation is only executed if the required
-	conditions on consistency are met is also given.
-
-\item \name is implemented as a shim layer over (and importantly, without
-	modifying) Cassandra~\cite{Cassandra}, a well-known general-purpose data
-	store.  An extensive experimental evaluation over a collection of real-world
-	applications, including a Twitter-like micro-blogging site and an eBay-like
-	auction site illustrates the practicality of our approach.
-
+	conditions on consistency are met.
+\item \name is implemented as a transparent shim layer over
+  Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
+  extensive experimental evaluation over a collection of real-world
+  applications, including a Twitter-like micro-blogging site and an
+  eBay-like auction site illustrates the practicality of our approach.
 \end{itemize}

### 2014-10-27 motivation.tex
intro and motive

--- 2014-10-27 motivation.tex
@@
 
-In this section, we motivate the need and benefit of declarative reasoning under
-eventual consistency with the help of an example. We begin with a description of
-our system and programming model, and introduce the primitive relations that
-constitute an essential part of our contract language.
+In this section, we motivate the need and benefit of declarative reasoning
+for eventually consistent systems.  We begin with a description of our
+system and programming model, and introduce the primitive relations that
+constitute the essential part of our contract language.
 
@@
 
-Each object is associated with a set of \emph{operations}. The clients interact
-with the store by invoking the operations on objects. The sequence of operations
-invoked by a particular client on the store is called a \emph{session}. The
-data store is typically accessed by a large number of clients (and hence
-sessions) concurrently. Importantly, the clients are oblivious to which replica
-an operation is applied to; the data store may choose to route the operation to
-the closest replica for minimizing latency, or route it to a different replica
-in order to balance the load in the system. For example, the operations $foo$
-and $bar$ invoked by the same session on the same object, might end up being
-applied to different replicas because replica 1 to which $foo$ was applied to
-might be unreachable when invoking $bar$.
+Each object is associated with a set of \emph{operations}. The clients
+interact with the store by invoking operations on objects. The sequence of
+operations invoked by a particular client on the store is called a
+\emph{session}. The data store is typically accessed by a large number of
+clients (and hence sessions) concurrently. Importantly, the clients are
+oblivious to which replica an operation is applied to; the data store may
+choose to route the operation to the closest replica for minimizing latency,
+or route it to a different replica in order to balance the load in the
+system. For example, the operations \emph{foo} and \emph{bar} invoked by the same
+session on the same object, might end up being applied to different replicas
+because replica 1 (to which \emph{foo} was applied) might be unreachable when
+the client invokes \emph{bar}.
 
-Suppose an operation $foo$ is invoked on a object $x$ with arguments $arg_1$ at
-replica 1. The operation $foo$ simply \emph{reduces} over the current set of
+When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
+replica 1, it simply \emph{reduces} over the current set of
 effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
 $v1$ that is sent back to the client, and emits a \emph{single} new effect
-$w^x_4$, which is appended to the state $x$ at replica 1. Thus, every operation
+$w^x_4$ that is appended to the state $x$ at replica 1. Thus, every operation
 is evaluated over a \emph{snapshot} of the state of the object on which it is
-invoked. In this case, we say that the effect $w^x_4$ \emph{witnessed} $w^x_1$
+invoked. In this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$
 and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
 written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
-$\visZ$ is the visibility relation between the effects. Visibility is an
+$\visZ$ is the visibility relation between effects. Visibility is an
 irreflexive (an operation cannot see its own effect) and asymmetric relation,
 and only relates effects produced by operations on the same object.
+Executing a read-only operation is similar except
+that no new effects are produced.
 
-Read-only operation execution is similar except that they do not produce a new
-effect. The effect added to a particular replica is asynchronously sent to
-other replicas. We assume that \emph{eventually}, an effect applied at one
-replica will be merged into all other replicas. The two effects $w^x_4$ and
-$w^x_5$ that arise from the same session are said to be in \emph{session order}
-(written logically as $\so{w^x_4}{w^x_5}$). Session order is a irreflexive,
+The effect added to a particular replica is asynchronously sent to other
+replicas. We assume that \emph{eventually} an effect applied at one replica
+will be merged into all other replicas. Two effects $w^x_4$ and $w^x_5$ that
+arise from the same session are said to be in \emph{session order} (written
+logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive,
 transitive relation. The effects $w^x_4$ and $w^x_5$ arising from operations
@@
 relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every
-effect with the operation that generated the effect with the help of a relation
-$\operZ$. In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$
-hold. For simplicity, we assume that all the operation names across all object
-types are distinct.
+effect with the operation that generated the effect with the help of a
+relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
+$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names
+across all object types are distinct.
 
-The system model described allows all the inconsistencies associated with
-eventual consistency. The goal of our work is to identify the precise
-consistency level for each operation such that the application-level
-constraints are not violated. In the next section, we will concretely describe
-the challenges associated with constructing a consistent bank account on top of
+This model admits all the inconsistencies associated with eventual
+consistency. The goal of this work is to identify the precise consistency
+level for each operation such that application-level constraints are not
+violated. In the next section, we will concretely describe the challenges
+associated with constructing a consistent bank account on top of an
 eventually consistent data stores. Subsequently, we will illustrate how our
-contract language, armed with the primitive relations $\visZ$, $\soZ$,
-$\sameobjZ$ and $\operZ$, mitigates these challenges.
+contract and specification language, armed with the primitive relations
+$\visZ$, $\soZ$, $\sameobjZ$ and $\operZ$, mitigates these challenges.
 
@@
 
-Suppose our goal is to implement a highly available bank account on top of an
-eventually consistent data store, with the \emph{integrity} constraint that the
-balance should never drop below 0. Let us assume that the bank account object
-provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
-with the assumption that the withdraw fails if the account has insufficient
-balance to complete the withdraw operation. Our goal is to choose the correct
-consistency level for each operation such that (1) the balance never drops
+Consider how we might implement a highly available bank account on top of an
+eventually consistent data store, with the \emph{integrity} constraint that
+the balance should never drop below 0. Let us assume that the bank account
+object provides three operations: \cf{deposit}, \cf{withdraw} and
+\cf{getBalance}, with the assumption that the withdraw fails if the account
+has insufficient balance. Our goal is to choose the correct consistency
+level for each of these operations such that (1) the balance never drops
 below zero and (2) the \cf{getBalance} operation never incorrectly show a
-negative balance. To be able to do this, we need to consider the anomalies that
-could arise under eventual consistency.
+negative balance. To be able to do this, we need to consider the anomalies
+that could arise under eventual consistency.
 
@@
 that all operations in the figure are on the same bank account object with
-initial balance as 0. Session 1 performs a \cf{deposit} of 100, followed by a
+the initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed by a
 \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation witnesses
@@
 \cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
-constraint. A subsequent \cf{getBalance} operation, that witnesses all the
-previous operations and reports a negative balance.
+constraint. A subsequent \cf{getBalance} operation, that happens to witness all the
+previous operations,  would report a negative balance.
 
-It is easy to see that the only way to ensure that integrity constraint on
+It is easy to see that the only way to ensure the integrity constraint on
 \cf{balance} is to prevent concurrent withdrawals on the same bank account
 object. This can be done by insisting that \cf{withdraw} be executed with
-strong consistency. Although this prevents the integrity constraint from being
-violated, a strongly consistent \cf{withdraw} itself is not sufficient to
-prevent \cf{getBalance} from ever reporting a negative balance to the end user.
+strong consistency properties. Although this prevents the integrity
+constraint from being violated, a strongly consistent \cf{withdraw} itself
+is not sufficient to prevent \cf{getBalance} from ever reporting a negative
+balance to the end user.
 
@@
 consists of three concurrent sessions performing a \cf{deposit}, a
-\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
-account object. Assume that initial balance in the account is zero. As the
-\cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2, can
-witnesses the effects of \cf{deposit(100)} from session 1, leading it to
-rightfully conclude that the current balance is 100. Consequently,
-\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
-operation subsequently becomes visible to the \cf{getBalance} in Session 3.
-However, \cf{getBalance}, which only witnesses the effect of \cf{withdraw} from
-session 2, but not the \cf{deposit} from session 1, reports the balance of
-negative 50 to the user.
+\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same
+bank account object. As before, assume that the initial balance in the
+account is zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)}
+in session 2, can witnesses the effects of \cf{deposit(100)} from session 1,
+leading it to rightfully conclude that the current balance is
+50. Consequently, \cf{withdraw(50)} is performed safely. The effect of this
+\cf{withdraw} operation subsequently becomes visible to the \cf{getBalance}
+in Session 3.  However, because \cf{getBalance} only witnesses the effect of
+the \cf{withdraw} operation from session 2, but not the \cf{deposit} from
+session 1, reports the balance of negative 50 to the user.
 
-Negative balance is not the only inconsistency that the user is exposed to
-under an eventually consistent bank account. Figure
+Producing a negative balance is not the only inconsistency that the user is
+exposed to under an eventually consistent bank account. Figure
 ~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
-operation in a session does not witness the effects of \cf{withdraw} operation
-performed in the same session, possibly because \cf{getBalance} was served by a
-replica that has not yet merged the \cf{withdraw} effect, served by a different
-replica. This anomaly leads the user to incorrectly conclude that the
-\cf{withdraw} operation failed to go through.
+operation in a session does not witness the effects of an earlier
+\cf{withdraw} operation performed in the same session, possibly because it
+was served by a replica that has not yet merged the \cf{withdraw} effect,
+served by a different replica. This anomaly leads the user to incorrectly
+conclude that the \cf{withdraw} operation failed to go through.
+
+%% \begin{figure}
+%% \centering
+%% \includegraphics[width=\columnwidth]{Graphs/BankAccount_Th_vs_Lat}
+%% \caption{Bank Account: Throughput vs Latency. Quelea is our implementation.
+%% Strong and Quelea are correct. High. Avail. and causal are incorrect (withdraw
+%% not strong). 20\% withdraw, 20\% deposit, 60\% getBalance.}
+%% \label{grf:BankAcc_Th_vs_Lat}
+%% \end{figure}
 
 Although it is easy to understand the reasons behind the occurrence of the
-aforementioned negative balance and missing update anomalies, the fix is not
-quite apparent. Making \cf{getBalance} a strongly consistent operation is
-definitely sufficient to avert anomalies, but is it really necessary? Given
-that a strongly consistent operation is significantly expensive, both in terms
-of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally would
-not want to request strong consistency, unless there are no alternatives
-available. Unfortunately, this requires us to understand subtle differences in
-semantics of various kinds of weak consistency alternatives, offered by the
-data stores.
+aforementioned negative balance and missing update anomalies, finding the
+appropriate fixes is not readily apparent. Making \cf{getBalance} a strongly
+consistent operation is definitely sufficient to avert anomalies, but is it
+really necessary? Given the cost of enforcing strong
+consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid unless there
+are no viable alternatives.  Exploring the space of these alternatives
+requires understanding the subtle differences in semantics of various kinds
+of weak consistency alternatives.
+
+%%  a strongly consistent operation is
+%% significantly expensive (Figure~\ref{grf:BankAcc_Th_vs_Lat} \KC{Should I
+%%   include the numbers here? If yes, I'll expand the text.}), both in terms
+%% of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally
+%% would not want to request strong consistency, unless there are no
+%% alternatives available.
 
@@
 
-\name automates this error-prone task by letting the programmer declare
-application-level consistency constraints as \emph{contracts}. The key
-intuition is that, in the case of the bank account example, any execution that
-does not exhibit the anomalies described in the previous section is a
-\emph{well-formed} execution on the bank account object. In the same vein as
-axiomatic memory models for shared memory multicore
-processors~\cite{Demange2013,Sarkar2011}, the programmer axiomatically
-describes the set of legal executions on top of each replicated date type over
-a trace of operation invocations on that type. Quelea ensures that any
-execution on top of such datatypes always belongs to the set of well-formed
-executions on that type.
+\name helps facilitate the mapping of operations to appropriate consistency
+levels by letting the programmer declare application-level consistency
+constraints as \emph{contracts} that axiomatically specify the set of
+allowed executions involving this operation.  In the case of the bank
+account example, any execution that does not exhibit the anomalies described
+in the previous section is a \emph{well-formed} execution on the bank
+account object.  By specifying the set of legal executions for each data
+type in terms of a trace of operation invocations on that type,
+\name\ ensures that all executions over that type are well-formed.
 
-In the bank account example, we observed that, in order to preserve the
-integrity constraint, the \cf{withdraw} operation must be strongly consistent.
-That is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible
-to $b$ or vice versa. Utilizing the primitive relations described in
+In our running example, it is clear that in order to preserve the integrity
+constraint, the \cf{withdraw} operation must be strongly consistent.  That
+is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to
+$b$ or vice versa. Utilizing the primitive relations described in
 Section~\ref{sec:sysmod}, we express this application-level consistency
-requirement as a contract over \cf{withdraw}:
+requirement as a contract or specification over \cf{withdraw}:
 \begin{smathpar}
@@
 \end{smathpar}
-In the above contract, $\cureff$ stands for the effect emitted by the
-\cf{withdraw} operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an
-effect which was emitted by a \cf{withdraw} operation i.e)
-$\oper{a}{\rcf{withdraw}}$ holds. The contract says that, if the current
-operation emits an effect $\cureff$, then for any operation $a$ which was
-emitted by a \cf{withdraw} operation, it is the case that $a = \cureff$ or $a$
-is visible to $\cureff$ or vice versa. Any execution on a bank account object
-that preserves the above contract for a \cf{withdraw} operation is said to have
-a correct implementation of \cf{withdraw}.
+Here, $\cureff$ stands for the effect emitted by the \cf{withdraw}
+operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an effect
+which was emitted by a \cf{withdraw} operation i.e.,
+$\oper{a}{\rcf{withdraw}}$ holds.  The contract specifies that if the
+current operation emits an effect $\cureff$, then for any operation $a$
+which was emitted by a \cf{withdraw} operation, it is the case that $a =
+\cureff$ or $a$ is visible to $\cureff$, or vice versa. Any execution on a
+bank account object that preserves the above contract for a \cf{withdraw}
+operation is said to be derived from a correct implementation of
+\cf{withdraw}.
 
-For the \cf{getBalance}, we construct a contract  precisely to prevent the
-anomalies described previously. The contract is shown below:
+For \cf{getBalance}, we construct the following contract:
 \begin{smathpar}
@@
 
-If a \cf{withdraw} $b$ is visible to the get balance operation, then all
-\cf{deposit} operations $a$ visible to the \cf{withdraw} should be visible to
-the get balance operation. This prevents negative balance anomalies. The syntax
-$c : \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
-emitted either by a \cf{deposit} operation or by a \cf{withdraw} operation. The
-contract language provides relational operators to compose relations. The
-syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
-last line of the above contract says that if a \cf{deposit} or a \cf{withdraw}
-precedes a \cf{getBalance} operation in session order, and is applied on the
-same object as the \cf{getBalance} operation, then it must be the case that the
-\cf{getBalance} operation witnesses the effects of the preceding operations.
-Since the deposit operation does not have any restrictions, its contract is
-simply $\true$.
+If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
+operations $a$ visible to the \cf{withdraw} should also be visible to
+\cf{getBalance}. This prevents negative balance anomalies. The expression $c
+: \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
+emitted either by a \cf{deposit} or a \cf{withdraw} operation.  Additonally,
+our specification language provides relational operators to compose
+relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
+\wedge R_2(a,b)$. The last line of the above contract says that if a
+\cf{deposit} or a \cf{withdraw} precedes a \cf{getBalance} operation in
+session order, and is applied on the same object as the \cf{getBalance}
+operation, then it must be the case that the \cf{getBalance} operation
+witnesses the effects of the preceding operations.  Since there are no
+restrictions on when and how a \cf{deposit} operation can execute, its
+contract is simply $\true$.
 
-\subsection{From contracts to implementation}
+\subsection{From Contracts to Implementation}
 
 Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
-the application-level consistency requirements, and make no reference to the
-semantics of the underlying store. To write such contracts, a programmer only
-needs to reason about the semantics of the application under the \name system
-model. The mapping of application-level consistency requirements to appropriate
-store-level guarantees is done automatically behind the screen.
+application-level consistency requirements, and make no reference to the
+semantics of the underlying store. To write contracts, a programmer only
+needs to reason about the semantics of the application under the \name
+system model. The mapping of application-level consistency requirements to
+appropriate store-level guarantees is done automatically behind-the-scene.
 
-How might one go about ensuring that any execution adheres to the contract? The
-challenge is that contract is a declarative characterization of correctness of
-execution, while what we need is an operational procedure for \emph{enforcing}
-the contracts. One strategy is to adopt speculative execution as espoused by
-other eventually consistent systems, where operations are tentatively applied
-as they are received from the client or other replicas. We can maintain a
-runtime manifestation of the axiomatic characterization and check the
-well-formedness at runtime, and rollback the execution if
-ill-formed~\cite{rxcml}. However, the overhead of state maintenance and the
-complexity of user-defined contracts (unlike simple tests such as cyclic
-dependence) make this technique unviable.
+How might one go about ensuring that any execution adheres to the contract?
+The challenge is that a contract provides a declarative (axiomatic)
+specification of an execution, while what is required is is an operational
+procedure for \emph{enforcing} its implicit constraints. One strategy would
+be to execute operations speculatively.  Here, operations are tentatively
+applied as they are received from the client or other replicas. We can
+maintain a runtime manifestation of executions, and check well-formedness
+conditions at runtime, rolling back executions if they are
+ill-formed. However, the overhead of state maintenance and the complexity of
+user-defined contracts make this technique infeasible in practice.
 
-Instead, Quelea statically analyzes the contracts using an SMT solver, and maps
-each operation to a particular store-level consistency under which the contract
-would be preserved. In particular, the key idea is that the system implementor
-axiomatically describes the store semantics in the same contract language as
-the one used to express the contract on the operation. We translate the
-contracts down to decidable subset of first-order logic~\cite{}, and discharge
-proof obligation of checking whether a particular store consistency level is
-sufficient for the given contract with the help of an SMT solver. The next
-section describes the contract language and classification.
+We devise a static approach instead.  Contracts are analyzed with the help
+of a theorem prover, and statically mapped to a particular store-level
+consistency property that the prover guarantees preserves contract
+semantics. The key idea is that the system implementor axiomatically
+describes the store semantics in the \emph{same} contract language as the
+one used to express the contract on the operation.

### 2014-10-27 motivation.tex
motivation
Edits

@@
 
-In this section, we motivate the need and benefit of declarative reasoning
-for eventually consistent systems.  We begin with a description of our
-system and programming model, and introduce the primitive relations that
-constitute the essential part of our contract language.
+In this section, we motivate the need and benefit of declarative reasoning for
+eventually consistent systems.  We begin with a description of our system and
+programming model, and introduce the primitive relations that constitute the
+essential part of our contract language.
 
@@
 \centering
-\includegraphics[width=\columnwidth]{Figures/SystemModel}
+\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
 \caption{Quelea system model.}
@@
 
-Each object is associated with a set of \emph{operations}. The clients
-interact with the store by invoking operations on objects. The sequence of
-operations invoked by a particular client on the store is called a
-\emph{session}. The data store is typically accessed by a large number of
-clients (and hence sessions) concurrently. Importantly, the clients are
-oblivious to which replica an operation is applied to; the data store may
-choose to route the operation to the closest replica for minimizing latency,
-or route it to a different replica in order to balance the load in the
-system. For example, the operations \emph{foo} and \emph{bar} invoked by the same
-session on the same object, might end up being applied to different replicas
-because replica 1 (to which \emph{foo} was applied) might be unreachable when
-the client invokes \emph{bar}.
+Each object is associated with a set of \emph{operations}. The clients interact
+with the store by invoking operations on objects. The sequence of operations
+invoked by a particular client on the store is called a \emph{session}. The
+data store is typically accessed by a large number of clients (and hence
+sessions) concurrently. Importantly, the clients are oblivious to which replica
+an operation is applied to; the data store may choose to route the operation to
+the closest replica for minimizing latency, or route it to a different replica
+in order to balance the load in the system. For example, the operations
+\emph{foo} and \emph{bar} invoked by the same session on the same object, might
+end up being applied to different replicas because replica 1 (to which
+\emph{foo} was applied) might be unreachable when the client invokes
+\emph{bar}.
 
 When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
-replica 1, it simply \emph{reduces} over the current set of
-effects at that replica on that object ($w^x_1$ and $w^x_2$), produces a result
-$v1$ that is sent back to the client, and emits a \emph{single} new effect
-$w^x_4$ that is appended to the state $x$ at replica 1. Thus, every operation
-is evaluated over a \emph{snapshot} of the state of the object on which it is
-invoked. In this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$
-and $w^x_2$. Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
-written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
-$\visZ$ is the visibility relation between effects. Visibility is an
-irreflexive (an operation cannot see its own effect) and asymmetric relation,
-and only relates effects produced by operations on the same object.
-Executing a read-only operation is similar except
-that no new effects are produced.
+replica 1, it simply \emph{reduces} over the current set of effects at that
+replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
+sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
+appended to the state of $x$ at replica 1. Thus, every operation is evaluated
+over a \emph{snapshot} of the state of the object on which it is invoked. In
+this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$ and $w^x_2$.
+Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$, written
+logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where $\visZ$ is
+the visibility relation between effects. Visibility is an irreflexive (an
+operation cannot see its own effect) and asymmetric relation, and only relates
+effects produced by operations on the same object. Executing a read-only
+operation is similar except that no new effects are produced.
 
@@
 arise from the same session are said to be in \emph{session order} (written
-logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive,
-transitive relation. The effects $w^x_4$ and $w^x_5$ arising from operations
-applied to the same object $x$ are said to be under the \emph{same object}
-relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every
-effect with the operation that generated the effect with the help of a
-relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
-$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names
-across all object types are distinct.
+logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive, transitive
+relation. The effects $w^x_4$ and $w^x_5$ arising from operations applied to
+the same object $x$ are said to be under the \emph{same object} relation,
+written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every effect with
+the operation that generated the effect with the help of a relation $\operZ$.
+In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$ hold. For
+simplicity, we assume all operation names across all object types are distinct.
 
-This model admits all the inconsistencies associated with eventual
-consistency. The goal of this work is to identify the precise consistency
-level for each operation such that application-level constraints are not
-violated. In the next section, we will concretely describe the challenges
-associated with constructing a consistent bank account on top of an
-eventually consistent data stores. Subsequently, we will illustrate how our
-contract and specification language, armed with the primitive relations
-$\visZ$, $\soZ$, $\sameobjZ$ and $\operZ$, mitigates these challenges.
+This model admits all the inconsistencies associated with eventual consistency.
+The goal of this work is to identify the precise consistency level for each
+operation such that application-level constraints are not violated. In the next
+section, we will concretely describe the challenges associated with
+constructing a consistent bank account on top of an eventually consistent data
+store. Subsequently, we will illustrate how our contract and specification
+language, armed with the primitive relations $\visZ$, $\soZ$, $\sameobjZ$ and
+$\operZ$, mitigates these challenges.
 
-\subsection{Eventually Consistent Bank Account}
+\subsection{Example}
 
 Consider how we might implement a highly available bank account on top of an
-eventually consistent data store, with the \emph{integrity} constraint that
-the balance should never drop below 0. Let us assume that the bank account
-object provides three operations: \cf{deposit}, \cf{withdraw} and
-\cf{getBalance}, with the assumption that the withdraw fails if the account
-has insufficient balance. Our goal is to choose the correct consistency
-level for each of these operations such that (1) the balance never drops
-below zero and (2) the \cf{getBalance} operation never incorrectly show a
-negative balance. To be able to do this, we need to consider the anomalies
-that could arise under eventual consistency.
+eventually consistent data store, with the \emph{integrity} constraint that the
+balance should never drop below 0. Let us assume that the bank account object
+provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
+with the assumption that the withdraw fails if the account has insufficient
+balance. Our goal is to choose the correct consistency level for each of these
+operations such that (1) the balance never drops below zero and (2) the
+\cf{getBalance} operation never incorrectly returns a negative balance. To be
+able to do this, we need to consider the anomalies that could arise under
+eventual consistency.
 
@@
 Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
-that all operations in the figure are on the same bank account object with
-the initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed by a
-\cf{withdraw} of 80 in the same session. The \cf{withdraw} operation witnesses
-the deposit and succeeds\footnote{Although visibility and session order
-relations relate effects, we have abused the notation in these examples to
-relate operations, with the idea that the relations relate the effect emitted
-by those operations}. Subsequently, session 2 perform a \cf{withdraw}
+that all operations in the figure are on the same bank account object with the
+initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
+by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
+witnesses the deposit and succeeds\footnote{Although visibility and session
+order relations relate effects, we have abused the notation in these examples
+to relate operations, with the idea that the relations relate the effect
+emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
 operation, but importantly, due to eventual consistency, only witnesses the
@@
 \cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
-constraint. A subsequent \cf{getBalance} operation, that happens to witness all the
-previous operations,  would report a negative balance.
+constraint. A subsequent \cf{getBalance} operation, that happens to witness all
+the previous operations,  would report a negative balance.
 
@@
 object. This can be done by insisting that \cf{withdraw} be executed with
-strong consistency properties. Although this prevents the integrity
-constraint from being violated, a strongly consistent \cf{withdraw} itself
-is not sufficient to prevent \cf{getBalance} from ever reporting a negative
-balance to the end user.
+strong consistency properties. Although this prevents the integrity constraint
+from being violated, a strongly consistent \cf{withdraw} itself is not
+sufficient to prevent \cf{getBalance} from ever reporting a negative balance to
+the end user.
 
@@
 consists of three concurrent sessions performing a \cf{deposit}, a
-\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same
-bank account object. As before, assume that the initial balance in the
-account is zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)}
-in session 2, can witnesses the effects of \cf{deposit(100)} from session 1,
-leading it to rightfully conclude that the current balance is
-50. Consequently, \cf{withdraw(50)} is performed safely. The effect of this
-\cf{withdraw} operation subsequently becomes visible to the \cf{getBalance}
-in Session 3.  However, because \cf{getBalance} only witnesses the effect of
-the \cf{withdraw} operation from session 2, but not the \cf{deposit} from
-session 1, reports the balance of negative 50 to the user.
+\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
+account object. As before, assume that the initial balance in the account is
+zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2,
+can witnesses the effects of \cf{deposit(100)} from session 1, leading it to
+rightfully conclude that the current balance is 50. Consequently,
+\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
+operation subsequently becomes visible to the \cf{getBalance} in Session 3.
+However, because \cf{getBalance} only witnesses the effect of the \cf{withdraw}
+operation from session 2, but not the \cf{deposit} from session 1, reports the
+balance of negative 50 to the user.
 
@@
 ~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
-operation in a session does not witness the effects of an earlier
-\cf{withdraw} operation performed in the same session, possibly because it
-was served by a replica that has not yet merged the \cf{withdraw} effect,
-served by a different replica. This anomaly leads the user to incorrectly
-conclude that the \cf{withdraw} operation failed to go through.
-
-%% \begin{figure}
-%% \centering
-%% \includegraphics[width=\columnwidth]{Graphs/BankAccount_Th_vs_Lat}
-%% \caption{Bank Account: Throughput vs Latency. Quelea is our implementation.
-%% Strong and Quelea are correct. High. Avail. and causal are incorrect (withdraw
-%% not strong). 20\% withdraw, 20\% deposit, 60\% getBalance.}
-%% \label{grf:BankAcc_Th_vs_Lat}
-%% \end{figure}
+operation in a session does not witness the effects of an earlier \cf{withdraw}
+operation performed in the same session, possibly because it was served by a
+replica that has not yet merged the \cf{withdraw} effect, served by a different
+replica. This anomaly leads the user to incorrectly conclude that the
+\cf{withdraw} operation failed to go through.
 
@@
 really necessary? Given the cost of enforcing strong
-consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid unless there
-are no viable alternatives.  Exploring the space of these alternatives
-requires understanding the subtle differences in semantics of various kinds
-of weak consistency alternatives.
+consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid unless there are
+no viable alternatives.  Exploring the space of these alternatives requires
+understanding the subtle differences in semantics of various kinds of weak
+consistency alternatives.
 
-%%  a strongly consistent operation is
-%% significantly expensive (Figure~\ref{grf:BankAcc_Th_vs_Lat} \KC{Should I
-%%   include the numbers here? If yes, I'll expand the text.}), both in terms
-%% of availability and monetary overheads~\cite{DynamoDB,Pileus}, we ideally
-%% would not want to request strong consistency, unless there are no
-%% alternatives available.
-
-\subsection{Contracts}
+\subsection{Specifications and Contracts}
 
@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} that axiomatically specify the set of
-allowed executions involving this operation.  In the case of the bank
-account example, any execution that does not exhibit the anomalies described
-in the previous section is a \emph{well-formed} execution on the bank
-account object.  By specifying the set of legal executions for each data
-type in terms of a trace of operation invocations on that type,
-\name\ ensures that all executions over that type are well-formed.
+constraints as \emph{contracts} that axiomatically specify the set of allowed
+executions involving this operation.  In the case of the bank account example,
+any execution that does not exhibit the anomalies described in the previous
+section is a \emph{well-formed} execution on the bank account object.  By
+specifying the set of legal executions for each data type in terms of a trace
+of operation invocations on that type, \name\ ensures that all executions over
+that type are well-formed.
 
 In our running example, it is clear that in order to preserve the integrity
-constraint, the \cf{withdraw} operation must be strongly consistent.  That
-is, given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to
-$b$ or vice versa. Utilizing the primitive relations described in
+constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
+given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
+vice versa. Utilizing the primitive relations described in
 Section~\ref{sec:sysmod}, we express this application-level consistency
 requirement as a contract or specification over \cf{withdraw}:
+
 \begin{smathpar}
@@
 \end{smathpar}
-Here, $\cureff$ stands for the effect emitted by the \cf{withdraw}
-operation. The syntax $a : \rcf{withdraw}$ states that $a$ is an effect
-which was emitted by a \cf{withdraw} operation i.e.,
-$\oper{a}{\rcf{withdraw}}$ holds.  The contract specifies that if the
-current operation emits an effect $\cureff$, then for any operation $a$
-which was emitted by a \cf{withdraw} operation, it is the case that $a =
-\cureff$ or $a$ is visible to $\cureff$, or vice versa. Any execution on a
-bank account object that preserves the above contract for a \cf{withdraw}
-operation is said to be derived from a correct implementation of
+
+Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
+The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
+by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
+contract specifies that if the current operation emits an effect $\cureff$,
+then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
+is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
+Any execution on a bank account object that preserves the above contract for a
+\cf{withdraw} operation is said to be derived from a correct implementation of
 \cf{withdraw}.
@@
 For \cf{getBalance}, we construct the following contract:
+
 \begin{smathpar}
@@
 
+The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
+effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
 If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
 operations $a$ visible to the \cf{withdraw} should also be visible to
-\cf{getBalance}. This prevents negative balance anomalies. The expression $c
-: \rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an effect that was
-emitted either by a \cf{deposit} or a \cf{withdraw} operation.  Additonally,
-our specification language provides relational operators to compose
-relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
-\wedge R_2(a,b)$. The last line of the above contract says that if a
-\cf{deposit} or a \cf{withdraw} precedes a \cf{getBalance} operation in
-session order, and is applied on the same object as the \cf{getBalance}
-operation, then it must be the case that the \cf{getBalance} operation
-witnesses the effects of the preceding operations.  Since there are no
-restrictions on when and how a \cf{deposit} operation can execute, its
-contract is simply $\true$.
+\cf{getBalance}. This prevents negative balance anomalies.
+
+Let us consider the last line in the above contract. Our specification language
+provides operators to for composing relations. Thus, the syntax $(R_1 \cap
+R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
+above contract says that if a \cf{deposit} or a \cf{withdraw} precedes a
+\cf{getBalance} operation in session order, and is applied on the same object
+as the \cf{getBalance} operation, then it must be the case that the
+\cf{getBalance} operation witnesses the effects of the preceding operations.
+
+Since there are no restrictions on when and how a \cf{deposit} operation can
+execute, its contract is simply $\true$.
 
@@
 application-level consistency requirements, and make no reference to the
-semantics of the underlying store. To write contracts, a programmer only
-needs to reason about the semantics of the application under the \name
-system model. The mapping of application-level consistency requirements to
-appropriate store-level guarantees is done automatically behind-the-scene.
+semantics of the underlying store. To write contracts, a programmer only needs
+to reason about the semantics of the application under the \name system model.
+The mapping of application-level consistency requirements to appropriate
+store-level guarantees is done automatically behind-the-scene.
 
-How might one go about ensuring that any execution adheres to the contract?
-The challenge is that a contract provides a declarative (axiomatic)
-specification of an execution, while what is required is is an operational
-procedure for \emph{enforcing} its implicit constraints. One strategy would
-be to execute operations speculatively.  Here, operations are tentatively
-applied as they are received from the client or other replicas. We can
-maintain a runtime manifestation of executions, and check well-formedness
-conditions at runtime, rolling back executions if they are
-ill-formed. However, the overhead of state maintenance and the complexity of
-user-defined contracts make this technique infeasible in practice.
+How might one go about ensuring that any execution adheres to the contract? The
+challenge is that a contract provides a declarative (axiomatic) specification
+of an execution, while what is required is is an operational procedure for
+\emph{enforcing} its implicit constraints. One strategy would be to execute
+operations speculatively.  Here, operations are tentatively applied as they are
+received from the client or other replicas. We can maintain a runtime
+manifestation of executions, and check well-formedness conditions at runtime,
+rolling back executions if they are ill-formed. However, the overhead of state
+maintenance and the complexity of user-defined contracts make this technique
+infeasible in practice.
 
-We devise a static approach instead.  Contracts are analyzed with the help
-of a theorem prover, and statically mapped to a particular store-level
-consistency property that the prover guarantees preserves contract
-semantics. The key idea is that the system implementor axiomatically
-describes the store semantics in the \emph{same} contract language as the
-one used to express the contract on the operation.
+We devise a static approach instead. Contracts are analyzed with the help of a
+theorem prover, and statically mapped to a particular store-level consistency
+property that the prover guarantees preserves contract semantics. We call this
+procedure \emph{contract classification}. Given the variety and complexity of
+store level consistency properties, the idea is that the system implementor
+parameterizes the classification procedure by describing the store semantics in
+the \emph{same} contract language as the one used to express the contract on
+the operations. In the next section, we describe the contract language in
+detail and describe the classification procedure for a particular store
+semantics.

### 2014-10-27 core-lang.tex
Minor edits

@@
 clarity, we only consider a single replicated object of well-defined
-type (for eg: a replicated object of type {\sf bank\_account}) in our
+type (for eg: a replicated object of type \cf{BankAccount}) in our
 formalization.  Our semantics are parametric over the specification of
@@
 labels ($op$) of operations on $\tau$ to their definitions, while
-$\Ctrts$ maps them to their consistency contracts ($\cv$). The language of
-contracts is described in Fig.~\ref{lang:contract}. The definition of
-an operation is expected to be a lambda expression in the server-side
-language\footnote{Since a replicated object is maintained by a
-server, we call the language in which replicated data type operations
-are defined as \emph{server-side language}.}, although we do not
-enforce this in our formalization. We intentionally leave the
-server-side language underspecified to demonstrate that our ideas are
-independent of the choice of the programming language used to
+$\Ctrts$ maps them to their consistency contracts ($\cv$). The
+definition of an operation is expected to be a lambda expression in
+the server-side language\footnote{Since a replicated object is
+maintained by a server, we call the language in which replicated data
+type operations are defined as \emph{server-side language}.}, although
+we do not enforce this in our formalization. We intentionally leave
+the server-side language underspecified to demonstrate that our ideas
+are independent of the choice of the programming language used to
 implement the replicated data type.
@@
 An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
-of \emph{effect}s , and $\visZ,\soZ \subseteq \EffSoup \times
+of effects, and $\visZ,\soZ \subseteq \EffSoup \times
 \EffSoup$ are visibility and session order relations over effects,
@@
 the same session. An effect $\eff = (s,i,op,v)$ is said to precede
-another effect $\eff' = (s',i',op',v')$ in session order if and
-only if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
-\emph{available} at rest of the replicas eventually. The local state
-of a replica $r$ (i.e., $\Theta(r)$) is a set of effects that are
-currently available at $r$. An operation $op$ performed at replica $r$
-can only witness the set of effects ($\Theta(r)$) visible at $r$. If we denote the
-effect generated by the operation $op$ as $\eff_{op}$, then $\Theta(r)
-\times \{\eff_{op}\} ~\subseteq~ \visZ$. Note that in presence of
-multiple replicated objects, $\visZ$ only relates effects of
-operations performed on a single object. For instance, a {\sf
-withdraw} operation can witness (via $\visZ$) a {\sf deposit}
+another effect $\eff' = (s',i',op',v')$ in session order if and only
+if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
+visible at rest of the replicas eventually. The local state of a
+replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
+visible at $r$. An operation $op$ performed at replica $r$ can only
+witness the set of effects ($\Theta(r)$) visible at $r$.  Note that
+in presence of multiple replicated objects, $\visZ$ only relates
+effects of operations performed on a single object. For instance, a
+{\sf withdraw} operation can witness (via $\visZ$) a {\sf deposit}
 operation on the same bank account, but not a {\sf deposit} on a
@@
 Rule $\rulelabel{Comm}$ captures inter-replica communication. The
-basic guarantee provided by the store is causal consistency. That is,
-an effect ($\eff$) will become \emph{available} at a replica $r$ 
-only after all the effects that causally precede $\eff$ become
-available at $r$. Observe that this condition is applicable even for
-the effects that originate at replica $r$. It is helpful to think of
-such effects being stored in a \emph{write buffer} until causal
-consistency condition is met, at which point they can be merged with
-the local state ($\Theta(r)$). Nevertheless, it is important to note
-that that enforcing causal consistency does not require inter-replica
-coordination.
+basic guarantee provided by the store is causal visibility. That is,
+an effect ($\eff$) is made visible at a replica $r$ only after all
+the effects that causally precede $\eff$ are made visible at $r$.
+Observe that this condition is applicable even for the effects that
+originate at replica $r$. 
+% It is helpful to think of effects being
+% stored in a \emph{write buffer} until causal consistency condition is
+% met, at which point they can be merged with the local state
+% ($\Theta(r)$). 
+It is important to note that that enforcing causal visibility does not
+require any inter-replica coordination.
 
@@
 the progress in execution (from $\E$ to $\E'$) due to the application
-of operation $op$ to replica $r$ corresponding to a new effect $\eff$.
+of operation $op$ to replica $r$ resulting in a new effect $\eff$.
 The rule first constructs a \emph{context} for the application from
@@
 projecting\footnote{{\textsf{ctxt*}} is auxiliary function
-\textsf{ctxt} extended straightforwardly to set of effects} relevant information from
-effects in $\Theta(r)$. It then substitutes the definition
-($\Ops(op)$) of the operation for its label ($op$), and relies on the
-reduction relation ($\rdtredsto$) of the server-side language to
-reduce the application $\Ops(op)(v,~ctxt)$ to a value $v'$.
-Subsequently, the attributes of the new effect $\eff$ in the execution
-state. In particular, we record that the effects in $\Theta(r)$ are
-visible to $\eff$, and $\eff$ follows in session order ($\soZ$) every
-effect in the same session $s$ with a smaller sequence number.
+\textsf{ctxt} extended straightforwardly to set of effects} relevant
+information from effects in $\Theta(r)$. It then substitutes the
+definition ($\Ops(op)$) of the operation for its label ($op$), and
+relies on the reduction relation ($\rdtredsto$) of the server-side
+language to reduce the application $\Ops(op)(v,~ctxt)$ to a value
+$v'$.  Subsequently, the the attributes of execution state, namely
+$\EffSoup$, $\visZ$ and $\soZ$, are extended to incorporate
+the new effect ($\eff$).
 
@@
 (rule $\rulelabel{SA}$), the operation can be applied to a replica $r$
-only if it already contains the effects of all the previous performed
-in the same session. This guarantee can be satisfied by applying all
+only if it already contains the effects of all the previous operations
+from the same session. This guarantee can be satisfied by applying all
 operations from the same session to the same logical copy of the

### 2014-10-27 core-lang.tex
Fixed typos

@@
 
-We demonstrate our ideas using a replicated data store formalized as a
-triple $\E;\Theta;\Sigma$, where $\E$ denotes the execution state,
-$\Theta$ maps replicas to their local state, and $\Sigma$, the session
-soup, is the set of client sessions interacting with the store.
-Operational semantics given in Figure~\ref{sem:oper} succinctly
-capture the behaviour of the store that is relevant to demonstrate the
-operational significance of contract classification. For the sake of
-clarity, we only consider a single replicated object of well-defined
-type (for eg: a replicated object of type \cf{BankAccount}) in our
-formalization.  Our semantics are parametric over the specification of
-this replicated data type.  Figure~\ref{sem:oper} formalizes
-replicated data type (RDT) specification as tuple
-$(\tau,\Ops,\Ctrts)$, where $\tau$ is the data type, $\Ops$ maps
-labels ($op$) of operations on $\tau$ to their definitions, while
-$\Ctrts$ maps them to their consistency contracts ($\cv$). The
-definition of an operation is expected to be a lambda expression in
-the server-side language\footnote{Since a replicated object is
-maintained by a server, we call the language in which replicated data
-type operations are defined as \emph{server-side language}.}, although
-we do not enforce this in our formalization. We intentionally leave
-the server-side language underspecified to demonstrate that our ideas
-are independent of the choice of the programming language used to
-implement the replicated data type.
+We formalize a replicated data store as a triple $\E;\Theta;\Sigma$,
+where $\E$ denotes the execution state, $\Theta$ maps replicas to
+their local state, and $\Sigma$, the session soup, is the set of
+concurrent client sessions interacting with the store.  Operational
+semantics given in Figure~\ref{sem:oper} succinctly capture the
+behaviour of the store that is relevant to demonstrate the operational
+significance of contract classification. For the sake of clarity, we
+only consider a single replicated object of well-defined type (for eg:
+a replicated object of type \cf{BankAccount}) in our formalization.
+Our semantics are parametric over the specification of this replicated
+data type.  Figure~\ref{sem:oper} formalizes replicated data type
+(RDT) specification as tuple $(\tau,\Ops,\Ctrts)$, where $\tau$ is the
+data type, $\Ops$ maps labels ($op$) of operations on $\tau$ to their
+definitions, while $\Ctrts$ maps them to their consistency contracts
+($\cv$). The definition of an operation is expected to be a lambda
+expression in the server-side language\footnote{Since a replicated
+object is maintained by a server, we call the language in which
+replicated data type operations are defined as \emph{server-side
+language}.}, although we do not enforce this in our formalization. We
+intentionally leave the server-side language underspecified to
+demonstrate that our ideas are independent of the choice of the
+programming language used to implement the replicated data type.
 
@@
 
-Clients interact with the store via multiple concurrent sessions. For
+A session is modeled simply as a sequence of actions on the replicated
+object maintained by the store, where an action is defined as an
+application of replicated data type operation to a value. For
 technical reasons, we tag every session in the session soup ($\Sigma$)
 with a unique {\sf SessID} and the sequence number of next operation
-in that session. As shown in Fig.~\ref{sem:oper}, we model a session
-simply as a sequence of actions on the replicated object maintained by
-the store, where an action is defined as an application of replicated
-data type operation to a value. As with the server-side language, our
-development is also orthogonal to the choice of the client-side
-language, but we require both the languages to agree on the syntactic
-class of values that replicated data type operations ($op$) produce
-and consume.
+in that session. As with the server-side language, our development is
+also orthogonal to the choice of the client-side language, but we
+require both the languages to agree on the syntactic class of values
+that replicated data type operations ($op$) produce and consume.
 
@@
 visible at $r$. An operation $op$ performed at replica $r$ can only
-witness the set of effects ($\Theta(r)$) visible at $r$.  Note that
-in presence of multiple replicated objects, $\visZ$ only relates
-effects of operations performed on a single object. For instance, a
-{\sf withdraw} operation can witness (via $\visZ$) a {\sf deposit}
-operation on the same bank account, but not a {\sf deposit} on a
-different account. Often, in our formalization, we use $\visZ$ and
+witness the set of effects ($\Theta(r)$) visible at $r$.  
+% Note that in presence of multiple replicated objects, $\visZ$ only
+% relates effects of operations performed on a single object. For
+% instance, a {\sf withdraw} operation can witness (via $\visZ$) a {\sf
+% deposit} operation on the same bank account, but not a {\sf deposit}
+% on a different account. 
+Often, in our formalization, we use $\visZ$ and
 $\soZ$ binary relations to obtain a set of effects visible to a given
@@
 
-Rule $\rulelabel{Oper}$ is an auxiliary reduction of the form
-
+Rule $\rulelabel{Oper}$ is an auxiliary reduction of the
+form:\vspace{-1.7mm}
 \begin{smathpar}
 \auxred{\Theta} {(\E,\langle s,i,op(v) \rangle)} {r} {(\E',\eff)}
+\vspace{-1.7mm}
 \end{smathpar}
-
 \noindent Under the store configuration $\Theta$, the rule captures
@@
 operations over a replicated datatype with specification $\rdtspec =
-(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if 
-$\forall \eff \in \EffSoup. ~\E \models \De{\cv[\eff/\cureff]}$, where $\cv =
+(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if $\forall
+\eff \in \EffSoup. ~\E \models \DDe{[\eff/\cureff]\cv}$, where $\cv =
 \Ctrts(\operZ(\eff))$.
@@
 
-Symbol $\models$ denotes the standard \emph{models} relation of
+$\models$ denotes the standard \emph{models} relation of
 first-order logic.  A valid model for a first-order formula defines

### 2014-10-27 core-lang.tex
Renamed Comm to EffVis

@@
 
-Rule $\rulelabel{Comm}$ captures inter-replica communication. The
-basic guarantee provided by the store is causal visibility. That is,
-an effect ($\eff$) is made visible at a replica $r$ only after all
-the effects that causally precede $\eff$ are made visible at $r$.
-Observe that this condition is applicable even for the effects that
-originate at replica $r$. 
+Basic gurantee provided by the store is causal visibility, which is
+captured by the rule $\rulelabel{EffVis}$ as a condition for an effect
+to be visible at a replica. The rule makes an effect ($\eff$) visible
+at a replica $r$ only after all the effects that causally precede
+$\eff$ are made visible at $r$. 
+% Observe that this condition is
+% applicable even for the effects that originate at replica $r$. 
 % It is helpful to think of effects being
@@
 It is important to note that that enforcing causal visibility does not
-require any inter-replica coordination.
+require any inter-replica coordination. Any eventually consistent
+store can provide causal visibility while being highly available.
+Therefore, we do not lose any generality by assuming that the store
+provides causal visiblity.
 

### 2014-10-28 contract-language.tex

@@
 
-In this section, we formalize the contract language of \name, and
-describe our contract classification scheme, which analyzes a contract
-and maps it to the weakest store consistency level sufficient to
-satisfy its consistency requirements.
+In this section, we formalize the contract language of \name, and describe our
+contract classification scheme, which analyzes a contract and maps it to the
+weakest store consistency level sufficient to satisfy its consistency
+requirements.
 
@@
 The syntax of our core contract language is shown in Fig.
-~\ref{fig:contract-lang}. The language is based on first-order logic
-(FOL), and admits prenex quantification over typed effect variables.
-We use a special effect variable ($\cureff$) to denote the effect of
-\emph{current operation} - the operation for which a contract is being
-written. The type of an effect is simply the label of the operation
-(eg: \cf{withdraw}) that engendered the effect. We admit disjuntion in
-types to let an effect variable range over effects of multiple
-operations. Since each replicated data type comes with a unique set of
-operations (therefore, unique set of types for effects), our contract
-language is implictly parameterized over a replicated data type
-specification.
-\GK{I think sec 2. is the right place to introduce the idea of rdt
-spec via the bank account example.}
+~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
+admits prenex quantification over typed effect variables. We use a special
+effect variable ($\cureff$) to denote the effect of \emph{current operation} -
+the operation for which a contract is being written. The type of an effect is
+simply the name of the operation (eg: \cf{withdraw}) that engendered the
+effect. We admit disjuntion in types to let an effect variable range over
+multiple operation names.
 
-Quantifier-free propositions in our contract language are
-conjunctions, disjunctions and implications of predicates expressing
-relations between pairs of effect variables. Syntactic class of
-relations is seeded with primitive $\visZ$, $\soZ$, and $\sameobjZ$
-relations, and also admits derived relations that are expressible as
-union, intersection, or transitive closure\footnote{Strictly speaking,
-$R^{+}$ is not the transitive closure of $R$, as transitive closure is
-not expressible in FOL. Instead, $R^{+}$ in our language denotes
-\emph{a} superset of transitive closure of $R$. Formally, $R^{+}$ is
-any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
-\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow
-R'(x,z)$} of primitive relations. Commonly used derived relations are
-the \emph{same object session order} ($\sooZ$), and the the
-\emph{happens before order} ($\hboZ$) defined as following:
+Quantifier-free propositions in our contract language are conjunctions,
+disjunctions and implications of predicates expressing relations between pairs
+of effect variables. Syntactic class of relations is seeded with primitive
+$\visZ$, $\soZ$, and $\sameobjZ$ relations, and also admits derived relations
+that are expressible as union, intersection, or transitive
+closure\footnote{Strictly speaking, $R^{+}$ is not the transitive closure of
+$R$, as transitive closure is not expressible in FOL. Instead, $R^{+}$ in our
+language denotes \emph{a} superset of transitive closure of $R$. Formally,
+$R^{+}$ is any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
+\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$} of
+primitive relations. Commonly used derived relations are the \emph{same object
+session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), the \emph{happens
+before order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$), and the per-object
+happens-before order ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
+
+\subsection{Capturing Store Semantics}
+\label{sec:store_sem}
+
+An important aspect of our contract classification system is that the store
+semantics is also captured using the same contract language used to describe
+application-level consistency. In this regard, similar
+to~\cite{Burckhardt2014}, we can rigorously define a wide variety of store
+semantics including those that combine any subset of session and causality
+guarantees, and multiple consistency levels. For example, a store that offers
+strong consistency is captured by the contract:
+
+\vspace{-1em}
 \begin{smathpar}
-\begin{array}{lcl}
-{\sf soo} & = &  {\sf so} \cap {\sf sameobj}\\
-{\sf hbo} & = & ({\sf soo} \cup {\sf vis})^{+}\\
-\end{array}
+\scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
 \end{smathpar}
 
-Note that our contract language (stripped of its syntactic sugar) is a
+Similarly, a store that offers per-object causal consistency is captured by the
+contract:
+
+\vspace{-1em}
+\begin{smathpar}
+\ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
+\end{smathpar}
+
+This ability to represent store semantics and application-level consistency in
+the same language is vital to contract classification. Observe that out
+contract language does not incorporate real (i.e., wall-clock) time. Hence, the
+contract language cannot describe store semantics based on real time such as
+recency or bounded-staleness guarantees offered by certain
+stores~\cite{Pileus}.
+
+\subsection{Contract Comparison}
+
+Our goal is to classify the contracts, and map the corresponding operation to
+the \emph{weakest} store level consistency. To this end, we need to be able to
+compare the strengths of contracts. Let $\cv_{op}$ be a contract for a
+particular operation $op$, and $\cv_{st}$ capture a particular store
+consistency level. We would like to determine whether $op$ can be \emph{safely
+discharged} at the store consistency level $\cv_{st}$ such that the resulting
+execution does not violate $\cv_{op}$.
+
+Since our contracts represent axiomatic definition of program executions, let
+$\Mod{\cv}$ be the set of all executions under which $\cv$ is satisfied. If
+every execution $\E \in \Mod{\cv_{st}}$ is also a member of $\Mod{\cv_{op}}$,
+then $op$ can be safely discharged under the store consistency level
+$\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq \Mod{\cv_{op}}$. This is the
+model-theoretic consequence relation, written as $\cv_{st} \models_m \cv_{op}$.
+
+Observe that our contract language (stripped of its syntactic sugar) is a
 carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}. In particular, for any two contracts ($\cv_1$
-and $\cv_2$) in our language, the implication check $\cv_1 \Rightarrow
-\cv_2$ is decidable, allowing us to automatically deduce that $\cv_2$
-is weaker than $\cv_1$. We formally define the \emph{weaker than}
-relation among contracts as following: 
+decidable~\cite{epr}. Due to the completeness theorem~\cite{completeness},
+$\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st} \Rightarrow \cv_{op}$.
+Due to the decidability of our contract language this implication check is
+automatically discharged with the help of a theorem prover. In this case, we
+say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$ (written $\cv_{op} \le
+\cv_{st}$). Formally,
+
 \begin{definition}
-A contract $\cv_2$ is said to be weaker than $\cv_1$ (denoted $\cv_2
-\le \cv_1$ ) if and only if:
+An contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
+\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
 \begin{center}
-$\Delta \vdash \cv_1 \Rightarrow \cv_2$
 \end{center}
 \end{definition}
-\noindent The context $\Delta$ in the above definition captures assumptions
-about the nature of primitive relations, such as the assumption that
-$\Rvis$ and $\Rso$ are irreflexive.
+\vspace{-1em}
 
-This ability to statically compare contracts is key to contract
-classification and subsequent enforcement. This is described in the
-following subsection.
+\noindent where $\Delta$ captures assumptions about the nature of primitive
+relations, such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
+relation $\hbZ$ is acyclic, preventing thin-air reads. \KC{Ideally, we need to
+list all the axioms somewhere.}
 
@@
 
-Application-level consistency requirements expressed by \name
-contracts can have varying impact on the available of the data store
-trying to meet the requirements. We identify three major classes of
-contracts that differ on this parameter: 
-% For instance, total-order requirement
-% expressed by the \cf{withdraw} contract (Sec. ~\ref{sec:motivation})
-% imposes global coordination among all replicas, and consequently makes
-% the store unavailable for any concurrent session. In contrast, the
-% trivial (\cf{true}) contract of the \cf{deposit} operation expresses
-% no special consistency requirements; any replica can serve a deposit
-% request immediately, and the store remains available to concurrent
-% sessions. Based on how a contract affects the availability of the data
-% store, we classify it into one of the following three categories:
+While the store semantics offered by commercial and research data stores vary
+widely, for our exposition, we identify three particular consistency levels --
+eventual, causal and strong, commonly offered by many distributed store with
+tunable consistency, with increasing overhead in terms of latency and
+availability. Indeed, the techniques presented here can be extended to other
+consistency stratifications in a straight forward manner. We assign each
+application-level contract into one of these following classes:
 
 \begin{itemize}
-\item \textbf{Unavailable}: These are the contracts that are
-satisfiable only under strong consistency. The need for global
-coordination to satisfy such contracts makes the store unavailable for
-concurrent sessions. An example is the total-order contract on
-\cf{withdraw} (Sec. ~\ref{sec:motivation}).
-\item \textbf{Sticky-available}: These are the contracts that require
-see a \emph{causally consistent} snapshot of the store that
-\emph{must} include some or all of the operations from the same
-session. By applying all operations of a session to the same logical
-replica (colloquially, \emph{sticking} a session to a replica), it is
-possible for a store to satisfy such contracts, while remaining highly
-available. The \cf{getBalance} contract that is sufficient to prevent
-the missing update anomaly is a sticky available contract.
-\item \textbf{Highly-available}: Contracts that only require to see
-\emph{some} causally consistent snapshop of the store are said to
-require \emph{causal visiblity} guarantee. This guarantee can be
-achieved by forcing every replica to reveal only those effects whose
-preceding effects are already included in the replica. Consequently,
-such contracts can always be satisfied by the store, letting it remain
-highly available for other requests.
+\setlength{\itemsep}{2pt}
+
+\item \textbf{Eventually consistency}: Eventually consistent operations can be
+satisfied as long as the client can reach at least one replica. For example,
+\cf{deposit} is a highly-available operation. While eventually consistent data
+store typically offer \emph{basic} eventual consistency with all possible
+anomalies, we assume that our store provides stronger semantics that remains
+highly-available~\cite{BailisHAT,COPS}; the store always exposes causal cut of
+the updates. This semantics can be formally captured as:
+
+\vspace{-1em}
+\begin{smathpar}
+\ecc = \forall a,b. \hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
+\end{smathpar}
+
+\noindent where $\small \hboZ = ((\soZ \cap \sameobjZ) \cup \visZ)^+$.
+
+\item \textbf{Causal consistency}: Operations with causally consistent
+contracts are required to see a causally consistent snapshot of the object
+state, including the actions performed on the same session.  The latter
+requirement entails that if two operations $o_1$ and $o_2$ from the same
+session are applied to two different replicas $r_1$ and $r_2$, the second
+operation cannot be discharged until the effect of $o_1$ is merged with $r_2$.
+We assume that causality is only tracked through operations on the same object;
+two operations in the same session but on different objects are considered
+causally unrelated under this definition. Stores typically avoid tracking
+causaility across objects to mitigate the overhead due to unintended causality.
+The corresponding store semantics is captured by the contract $\ccc$ presented
+in Section~\ref{sec:store_sem}.
+
+\item \textbf{Strong Consistency}: Strongly consistent operations may be block
+indefinitely under network partitions. An example is the total-order contract
+on \cf{withdraw} operation. The corresponding store semantics is captured by
+the contract $\scc$ presented in Section~\ref{sec:store_sem}.
+
 \end{itemize}
 
-It is important to note that, although finer distinctions can be made
-between availability implications of contracts, the above
-classification scheme is the one that is known to have practical
-significance\cite{bailisHAT}. More importantly, the consistency levels
-described in the classification scheme, namely strong consistency
-({\sf sc}), causal consistency ({\sc cc}), and causal visibility ({\sf
-cv}), can be implemented efficiently on most off-the-shelf eventually
-consistent data stores\cite{...}. By determining which of the
-aformentioned classes a contract falls into, the corresponding
-implementation of the consistency level on the data store can be used
-to satisfy the contract efficiently. Nonetheless, our classification
-scheme is extendible, meaning that a new availability class can be
-carved out of existing ones with the advent of new data stores
-supporting efficient enforcement of finer consistency levels.
+Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves totally
+ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$. This
+concurs with the intuition that any contract satisfiable under $\ecc$ or $\ccc$
+is satisfiable under $\scc$ , and any contract that is satisfiable under $\ecc$
+is satisfiable under $\ccc$. Nonetheless, we determine the class of a contract
+based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$)
+required to satisfy the contract. The classification scheme is presented
+formally as rules in Figure~\ref{sem:classify}.
 
-To classify contracts into one of the aforementioned availability
-classes, we need to determine if the corresponding consistency
-guarantee is sufficient to satisfy the contract. Towards this end, we
-encode the {\sf sc}, {\sf cc}, and {\sf cv} consistency guarantees as
-formulas in our contract language ($\hboZ$, the happens-before order,
-was defined previously in the same section):
-\begin{smathpar}
-\stretcharraybig
-\begin{array}{rcl}
-{\sf sc} & = 		 & \forall a. (\hbo{a}{\cureff} \vee \hbo{\cureff}{a} \vee a = \cureff) \\
-         & & \conj (\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}) \\
-				 & & \conj (\hbo{\cureff}{a} \Rightarrow \vis{\cureff}{a}) \\
-{\sf cc} & = & \forall a. \hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-{\sf cv} & = & \forall a,b. \hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\end{array}
-\end{smathpar}
-Comparing ($\le$) a contract formula against the above formulas of
-{\sf sc}, {\sf cc}, and {\sf cv} will let us determine if the contract
-can be satisfied under strong consistency, causal consistency, and
-causal visiblity consistency guarantees, respectively. Note that {\sf
-sc}, {\sf cc} and {\sf cv} are themselves totally ordered with respect
-to the $\le$ relation:
-\begin{center}
-$\cvc \le \ccc \le \scc$
-\end{center}
-This concurs with the intuition that any contract satisfiable under
-$\ccc$ or $\cvc$ is satisfiable under $\scc$ , and any contract that
-is satisfiable under $\cvc$ is satisfiable under $\ccc$. Nonetheless,
-we determine the availability class of a contract based on the
-\emph{weakest} guarantee (among $\cvc$, $\ccc$, and $\scc$) required
-to satisfy the contract. The classification scheme is presented
-formally as rules in Figure ~\ref{sem:classfiy}.
 % Contract classification rules
@@
 \input{contract-classification}
-The rules to classify contracts into availability classes are
-straightforward. The remaining rule judges a contract as {\sf
-WellFormed} if and only if it can be satisfied under strong
-consistency. We only consider well-formed contracts for classification
-and subsequent enforcement at run-time. The rest are reported as
-errors.
 
-As a result of our automatic classification scheme, a store that can
-enforce $\scc$, $\ccc$, and $\cvc$ contracts can also enforce any
-well-formed \name contract by determining its availability class.  The
-following section formally describes operational semantics of one such
-store, along with a meta-theoretic result that certifies the soundness
-of our classification-based contract enforcement strategy.
+The rules to classify contracts into availability classes are straightforward.
+The remaining rule judges a contract as {\sf WellFormed} if and only if it can
+be satisfied under strong consistency. The rest are reported as errors.

### 2014-10-28 contract-language.tex
contract language

@@
 ~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
-admits prenex quantification over typed effect variables. We use a special
+admits prenex universal quantification over typed effect variables. We use a special
 effect variable ($\cureff$) to denote the effect of \emph{current operation} -
 the operation for which a contract is being written. The type of an effect is
-simply the name of the operation (eg: \cf{withdraw}) that engendered the
+simply the name of the operation (eg: \cf{withdraw}) that induced the
 effect. We admit disjuntion in types to let an effect variable range over
@@
 disjunctions and implications of predicates expressing relations between pairs
-of effect variables. Syntactic class of relations is seeded with primitive
+of effect variables. The syntactic class of relations is seeded with primitive
 $\visZ$, $\soZ$, and $\sameobjZ$ relations, and also admits derived relations
@@
 
-Similarly, a store that offers per-object causal consistency is captured by the
+\noindent Similarly, a store that offers per-object causal consistency is captured by the
 contract:
@@
 
-This ability to represent store semantics and application-level consistency in
+\noindent This ability to represent store semantics and application-level consistency in
 the same language is vital to contract classification. Observe that out
@@
 
-Our goal is to classify the contracts, and map the corresponding operation to
-the \emph{weakest} store level consistency. To this end, we need to be able to
-compare the strengths of contracts. Let $\cv_{op}$ be a contract for a
+Our goal is to classify contracts, and map the operation they describe to
+the \emph{weakest} store-level consistency level that  nonetheless
+satisfies the contract's constraints. To this end, we need a mechanism to 
+compare the ``strength'' of a contract. Let $\cv_{op}$ be a contract for a
 particular operation $op$, and $\cv_{st}$ capture a particular store
-consistency level. We would like to determine whether $op$ can be \emph{safely
-discharged} at the store consistency level $\cv_{st}$ such that the resulting
-execution does not violate $\cv_{op}$.
+consistency level. We would like to determine whether $op$ can be
+\emph{safely discharged} at the store consistency level $\cv_{st}$ such that
+the resulting execution does not violate $\cv_{op}$.
 
 Since our contracts represent axiomatic definition of program executions, let
-$\Mod{\cv}$ be the set of all executions under which $\cv$ is satisfied. If
+$\Mod{\cv}$ be the set of all executions under which contract $\cv$ is satisfied. If
 every execution $\E \in \Mod{\cv_{st}}$ is also a member of $\Mod{\cv_{op}}$,
@@
 carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}. Due to the completeness theorem~\cite{completeness},
-$\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st} \Rightarrow \cv_{op}$.
-Due to the decidability of our contract language this implication check is
+decidable~\cite{epr}.  Completeness of the logic~\cite{completeness} allows
+us to claim that $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st} \Rightarrow \cv_{op}$.
+Due to the decidability of our contract language, this implication check is
 automatically discharged with the help of a theorem prover. In this case, we
 say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$ (written $\cv_{op} \le
-\cv_{st}$). Formally,
-
-\begin{definition}
-An contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
-\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
-\begin{center}
-\end{center}
-\end{definition}
-\vspace{-1em}
-
-\noindent where $\Delta$ captures assumptions about the nature of primitive
-relations, such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
-relation $\hbZ$ is acyclic, preventing thin-air reads. \KC{Ideally, we need to
-list all the axioms somewhere.}
+\cv_{st}$).
 
@@
 
-While the store semantics offered by commercial and research data stores vary
-widely, for our exposition, we identify three particular consistency levels --
-eventual, causal and strong, commonly offered by many distributed store with
-tunable consistency, with increasing overhead in terms of latency and
-availability. Indeed, the techniques presented here can be extended to other
-consistency stratifications in a straight forward manner. We assign each
+While the store semantics offered by commercial and research data stores
+vary widely, for our purposes, we identify three particular consistency
+levels -- eventual, causal, and strong, commonly offered by many distributed
+stores with tunable consistency, with increasing overhead in terms of
+latency and availability. Indeed, the techniques presented here can be
+extended to other consistency stratifications. We assign each
 application-level contract into one of these following classes:
@@
 satisfied as long as the client can reach at least one replica. For example,
-\cf{deposit} is a highly-available operation. While eventually consistent data
+\cf{deposit} is an eventually consistent operation. While eventually consistent data
 store typically offer \emph{basic} eventual consistency with all possible
-anomalies, we assume that our store provides stronger semantics that remains
-highly-available~\cite{BailisHAT,COPS}; the store always exposes causal cut of
-the updates. This semantics can be formally captured as:
-
+anomalies, we assume that our store provides stronger semantics that remain
+highly-available~\cite{BailisHAT,COPS}; the store always exposes a causal cut of
+the updates. This semantics can be formally captured in terms of the
+following contract definition:
 \vspace{-1em}
@@
 \end{smathpar}
-
 \noindent where $\small \hboZ = ((\soZ \cap \sameobjZ) \cup \visZ)^+$.
@@
   session are applied to two different replicas $r_1$ and $r_2$, the second
-operation cannot be discharged until the effect of $o_1$ is merged with $r_2$.
-We assume that causality is only tracked through operations on the same object;
-two operations in the same session but on different objects are considered
-causally unrelated under this definition. Stores typically avoid tracking
-causaility across objects to mitigate the overhead due to unintended causality.
-The corresponding store semantics is captured by the contract $\ccc$ presented
-in Section~\ref{sec:store_sem}.
+  operation cannot be discharged until the effect of $o_1$ is merged with
+  $o_2$ in both $r_1$ and $r_2$.  We assume that causality is only tracked
+  through operations on the same object; two operations in the same session
+  but on different objects are considered causally unrelated under this
+  definition. Stores typically avoid tracking causality across objects to
+  mitigate overheads when causality tracking is unnecessary.  The corresponding
+  store semantics is captured by the contract $\ccc$ presented in
+  Section~\ref{sec:store_sem}.
 
-\item \textbf{Strong Consistency}: Strongly consistent operations may be block
-indefinitely under network partitions. An example is the total-order contract
-on \cf{withdraw} operation. The corresponding store semantics is captured by
-the contract $\scc$ presented in Section~\ref{sec:store_sem}.
+\item \textbf{Strong Consistency}: Strongly consistent operations may block
+  indefinitely under network partitions. An example is the total-order
+  contract on \cf{withdraw} operation. The corresponding store semantics is
+  captured by the contract $\scc$ presented in Section~\ref{sec:store_sem}.
 
@@
 
-Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves totally
-ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$. This
-concurs with the intuition that any contract satisfiable under $\ecc$ or $\ccc$
-is satisfiable under $\scc$ , and any contract that is satisfiable under $\ecc$
-is satisfiable under $\ccc$. Nonetheless, we determine the class of a contract
-based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$)
-required to satisfy the contract. The classification scheme is presented
-formally as rules in Figure~\ref{sem:classify}.
+\noindent Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are
+themselves totally ordered with respect to the $\le$ relation: $\ecc \le
+\ccc \le \scc$. This concurs with the intuition that any contract
+satisfiable under $\ecc$ or $\ccc$ is satisfiable under $\scc$ , and any
+contract that is satisfiable under $\ecc$ is satisfiable under
+$\ccc$. Nonetheless, we determine the class of a contract based on the
+\emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$) required to
+satisfy the contract. The classification scheme is presented formally 
+in Figure~\ref{sem:classify}; here, $\Delta$ captures assumptions about the nature of primitive
+relations, such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
+relation $\hbZ$ is acyclic, preventing thin-air reads. \KC{Ideally, we need to
+list all the axioms somewhere.}  The following definition thus precisely
+captures our intuition about contract strength in terms of these rules:
+
+\begin{definition}
+A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
+\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
+\begin{center}
+\end{center}
+\end{definition}
+\vspace{-1em}
 
@@
 
-The rules to classify contracts into availability classes are straightforward.
-The remaining rule judges a contract as {\sf WellFormed} if and only if it can
-be satisfied under strong consistency. The rest are reported as errors.
+%% The rules to classify contracts into availability classes are straightforward.
+%% The remaining rule judges a contract as {\sf WellFormed} if and only if it can
+%% be satisfied under strong consistency. The rest are reported as errors.

### 2014-10-28 intro.tex
contract language

@@
 
-The rest of the paper is organized as follows. Section~\ref{sec:motivation}
-motivates \name and introduces the language through a series of examples.
-Section~\ref{sec:lang} describes the contract language, and
+The rest of the paper is organized as follows. The next section describes
+the system model. Section~\ref{sec:motivation} describes the challenges in
+programming with eventually consistent data stores, and introduces
+\name\ contracts as a proposed solution to overcome these issues.
+Section~\ref{sec:lang} provides more details the contract language.
 Section~\ref{sec:opsem} presents an operational semantics.
 Section~\ref{sec:txns} introduces transaction contracts and classification.
-Section~\ref{sec:impl} describes the implementation and goes into detail about
-the optimizations needed to make the system practical.
+Section~\ref{sec:impl} describes the implementation and provides details 
+about the optimizations needed to make the system practical.
 Section~\ref{sec:results} discusses experimental evaluation.

### 2014-10-28 motivation.tex
edits
motivation.tex
contract language

@@
-\section{Motivation}
-\label{sec:motivation}
-
-In this section, we motivate the need and benefit of declarative reasoning for
-eventually consistent systems.  We begin with a description of our system and
-programming model, and introduce the primitive relations that constitute the
-essential part of our contract language.
-
-\subsection{System Model}
+\section{System Model}
 \label{sec:sysmod}
@@
 
-\subsection{Example}
+\section{Motivation}
+\label{sec:motivation}
 
@@
 really necessary? Given the cost of enforcing strong
-consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid unless there are
-no viable alternatives.  Exploring the space of these alternatives requires
-understanding the subtle differences in semantics of various kinds of weak
-consistency alternatives.
+consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid it unless
+there are no viable alternatives.  Exploring the space of these alternatives
+requires understanding the subtle differences in semantics of various kinds
+of weak consistency alternatives.
 
@@
 requirement as a contract or specification over \cf{withdraw}:
-
+\vspace{-.5em}
 \begin{smathpar}
@@
 \end{smathpar}
-
-Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
+\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
 The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
@@
 
-For \cf{getBalance}, we construct the following contract:
-
+\noindent For \cf{getBalance}, we construct the following contract:
 \begin{smathpar}
@@
 \end{smathpar}
-
-The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
+\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
 effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
@@
 
-Let us consider the last line in the above contract. Our specification language
-provides operators to for composing relations. Thus, the syntax $(R_1 \cap
-R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
-above contract says that if a \cf{deposit} or a \cf{withdraw} precedes a
-\cf{getBalance} operation in session order, and is applied on the same object
-as the \cf{getBalance} operation, then it must be the case that the
-\cf{getBalance} operation witnesses the effects of the preceding operations.
+Our specification language provides operators to compose
+relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
+\wedge R_2(a,b)$. The last line of the above contract says that if a
+\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation in
+session order, and is applied on the same object as the \cf{getBalance}
+operation, then it must be the case that the \cf{getBalance} operation
+witnesses the effects of the preceding operations.
 

### 2014-10-28 core-lang.tex
edits
Formatting operational semantics

@@
 
-We formalize a replicated data store as a triple $\E;\Theta;\Sigma$,
-where $\E$ denotes the execution state, $\Theta$ maps replicas to
-their local state, and $\Sigma$, the session soup, is the set of
-concurrent client sessions interacting with the store.  Operational
-semantics given in Figure~\ref{sem:oper} succinctly capture the
-behaviour of the store that is relevant to demonstrate the operational
-significance of contract classification. For the sake of clarity, we
-only consider a single replicated object of well-defined type (for eg:
-a replicated object of type \cf{BankAccount}) in our formalization.
-Our semantics are parametric over the specification of this replicated
-data type.  Figure~\ref{sem:oper} formalizes replicated data type
-(RDT) specification as tuple $(\tau,\Ops,\Ctrts)$, where $\tau$ is the
-data type, $\Ops$ maps labels ($op$) of operations on $\tau$ to their
-definitions, while $\Ctrts$ maps them to their consistency contracts
-($\cv$). The definition of an operation is expected to be a lambda
-expression in the server-side language\footnote{Since a replicated
-object is maintained by a server, we call the language in which
-replicated data type operations are defined as \emph{server-side
-language}.}, although we do not enforce this in our formalization. We
-intentionally leave the server-side language underspecified to
-demonstrate that our ideas are independent of the choice of the
-programming language used to implement the replicated data type.
+In this section, we describe the operational semantics of a store that safely
+discharges operations based on the classification given in the previous
+section. We also present the meta-theoretic result that certifies the soundness
+of our classification-based contract enforcement strategy.
+
+We formalize a replicated data store as a triple $\E;\Theta;\Sigma$, where $\E$
+denotes the execution state, $\Theta$ maps replicas to their local state, and
+$\Sigma$, the session soup, is the set of concurrent client sessions
+interacting with the store.  Operational semantics given in
+Figure~\ref{sem:oper} succinctly capture the behaviour of the store that is
+relevant to demonstrate the operational significance of contract
+classification. For the sake of clarity, we only consider a single replicated
+object of well-defined type (for eg: a replicated object of type
+\cf{BankAccount}) in our formalization. Our semantics are parametric over the
+specification of this replicated data type.  Figure~\ref{sem:oper} formalizes
+replicated data type (RDT) specification as tuple $(\tau,\Ops,\Ctrts)$, where
+$\tau$ is the data type, $\Ops$ maps labels ($op$) of operations on $\tau$ to
+their definitions, while $\Ctrts$ maps them to their consistency contracts
+($\cv$). The definition of an operation is expected to be a lambda expression
+in the server-side language\footnote{Since a replicated object is maintained by
+a server, we call the language in which replicated data type operations are
+defined as \emph{server-side language}.}, although we do not enforce this in
+our formalization. We intentionally leave the server-side language
+underspecified to demonstrate that our ideas are independent of the choice of
+the programming language used to implement the replicated data type.
 
@@
 
-A session is modeled simply as a sequence of actions on the replicated
-object maintained by the store, where an action is defined as an
-application of replicated data type operation to a value. For
-technical reasons, we tag every session in the session soup ($\Sigma$)
-with a unique {\sf SessID} and the sequence number of next operation
-in that session. As with the server-side language, our development is
-also orthogonal to the choice of the client-side language, but we
-require both the languages to agree on the syntactic class of values
-that replicated data type operations ($op$) produce and consume.
+A session is modeled simply as a sequence of actions on the replicated object
+maintained by the store, where an action is defined as an application of
+replicated data type operation to a value. For technical reasons, we tag every
+session in the session soup ($\Sigma$) with a unique {\sf SessID} and the
+sequence number of next operation in that session. As with the server-side
+language, our development is also orthogonal to the choice of the client-side
+language, but we require both the languages to agree on the syntactic class of
+values that replicated data type operations ($op$) produce and consume.
 
-An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
-of effects, and $\visZ,\soZ \subseteq \EffSoup \times
-\EffSoup$ are visibility and session order relations over effects,
-respectively. An effect ($\eff$), modeled as a tuple $(s,i,op,v)$,
-records the fact that $i^{th}$ action in session with {\sf SessID}
-$s$, which is an operation $op$ on the replicated object, has been
-successfully executed on some replica yielding a return value $v$.
-Note that the combination of $s$ and $i$ uniquely identifies the
-effect. Session order relation ($\soZ$) relates effects generated by
-the same session. An effect $\eff = (s,i,op,v)$ is said to precede
-another effect $\eff' = (s',i',op',v')$ in session order if and only
-if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
-visible at rest of the replicas eventually. The local state of a
-replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
-visible at $r$. An operation $op$ performed at replica $r$ can only
-witness the set of effects ($\Theta(r)$) visible at $r$.  
+An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set of
+effects, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are visibility and
+session order relations over effects, respectively. An effect ($\eff$), modeled
+as a tuple $(s,i,op,v)$, records the fact that $i^{th}$ action in session with
+{\sf SessID} $s$, which is an operation $op$ on the replicated object, has been
+successfully executed on some replica yielding a return value $v$. Note that
+the combination of $s$ and $i$ uniquely identifies the effect. Session order
+relation ($\soZ$) relates effects generated by the same session. An effect
+$\eff = (s,i,op,v)$ is said to precede another effect $\eff' = (s',i',op',v')$
+in session order if and only if $s'=s$ and $i'\ge i$. An effect generated at a
+replica becomes visible at rest of the replicas eventually. The local state of
+a replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
+visible at $r$. An operation $op$ performed at replica $r$ can only witness the
+set of effects ($\Theta(r)$) visible at $r$.
 % Note that in presence of multiple replicated objects, $\visZ$ only
@@
 % on a different account.
-Often, in our formalization, we use $\visZ$ and
-$\soZ$ binary relations to obtain a set of effects visible to a given
-effect $\eff$, or set of effects that precede a given effect $\eff$ in
-the session order. As a syntactic convenience, whenever $R$ is a
-binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
-such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$
-to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.
+Often, in our formalization, we use $\visZ$ and $\soZ$ binary relations to
+obtain a set of effects visible to a given effect $\eff$, or set of effects
+that precede a given effect $\eff$ in the session order. As a syntactic
+convenience, whenever $R$ is a binary relation, we write $R(\eff)$ to denote
+the set of all $\eff'$ such that $(\eff,\eff') \in R$.  Conversely, we write
+$R^{-1}(\eff)$ to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.
 
-Basic gurantee provided by the store is causal visibility, which is
-captured by the rule $\rulelabel{EffVis}$ as a condition for an effect
-to be visible at a replica. The rule makes an effect ($\eff$) visible
-at a replica $r$ only after all the effects that causally precede
-$\eff$ are made visible at $r$. 
+Basic gurantee provided by the store is causal visibility, which is captured by
+the rule $\rulelabel{EffVis}$ as a condition for an effect to be visible at a
+replica. The rule makes an effect ($\eff$) visible at a replica $r$ only after
+all the effects that causally precede $\eff$ are made visible at $r$.
 % Observe that this condition is
@@
 % ($\Theta(r)$).
-It is important to note that that enforcing causal visibility does not
-require any inter-replica coordination. Any eventually consistent
-store can provide causal visibility while being highly available.
-Therefore, we do not lose any generality by assuming that the store
-provides causal visiblity.
+It is important to note that that enforcing causal visibility does not require
+any inter-replica coordination. Any eventually consistent store can provide
+causal visibility while being highly available. Therefore, we do not lose any
+generality by assuming that the store provides causal visiblity.
 
@@
 \end{smathpar}
-\noindent Under the store configuration $\Theta$, the rule captures
-the progress in execution (from $\E$ to $\E'$) due to the application
-of operation $op$ to replica $r$ resulting in a new effect $\eff$.
-The rule first constructs a \emph{context} for the application from
-the local state ($\Theta(r)$) of the replica, by
-projecting\footnote{{\textsf{ctxt*}} is auxiliary function
+\noindent Under the store configuration $\Theta$, the rule captures the
+progress in execution (from $\E$ to $\E'$) due to the application of operation
+$op$ to replica $r$ resulting in a new effect $\eff$. The rule first constructs
+a \emph{context} for the application from the local state ($\Theta(r)$) of the
+replica, by projecting\footnote{{\textsf{ctxt*}} is auxiliary function
 \textsf{ctxt} extended straightforwardly to set of effects} relevant
-information from effects in $\Theta(r)$. It then substitutes the
-definition ($\Ops(op)$) of the operation for its label ($op$), and
-relies on the reduction relation ($\rdtredsto$) of the server-side
-language to reduce the application $\Ops(op)(v,~ctxt)$ to a value
-$v'$.  Subsequently, the the attributes of execution state, namely
-$\EffSoup$, $\visZ$ and $\soZ$, are extended to incorporate
-the new effect ($\eff$).
+information from effects in $\Theta(r)$. It then substitutes the definition
+($\Ops(op)$) of the operation for its label ($op$), and relies on the reduction
+relation ($\rdtredsto$) of the server-side language to reduce the application
+$\Ops(op)(v,~ctxt)$ to a value $v'$.  Subsequently, the the attributes of
+execution state, namely $\EffSoup$, $\visZ$ and $\soZ$, are extended to
+incorporate the new effect ($\eff$).
 
-If the operation $op$ is highly-available (rule $\rulelabel{HA}$), we
-simply apply the operation to any replica $r$. Since the store
-provides causal visibility, highly available operations are
-satisfiable under any replica. If the operation is sticky-available
-(rule $\rulelabel{SA}$), the operation can be applied to a replica $r$
-only if it already contains the effects of all the previous operations
-from the same session. This guarantee can be satisfied by applying all
-operations from the same session to the same logical copy of the
-database. If such a logical copy is untenable, then the operation
-might block. Since the store is assumed to converge eventually, the
-blocked sticky-available operation guaranteed to unblock eventually.
+If the operation $op$ is highly-available (rule $\rulelabel{HA}$), we simply
+apply the operation to any replica $r$. Since the store provides causal
+visibility, highly available operations are satisfiable under any replica. If
+the operation is sticky-available (rule $\rulelabel{SA}$), the operation can be
+applied to a replica $r$ only if it already contains the effects of all the
+previous operations from the same session. This guarantee can be satisfied by
+applying all operations from the same session to the same logical copy of the
+database. If such a logical copy is untenable, then the operation might block.
+Since the store is assumed to converge eventually, the blocked sticky-available
+operation guaranteed to unblock eventually.
 
-An unavailable operation expects strong (sequential) consistency.
-That is, universe of all effects ($\EffSoup$) in an execution ($\E$)
-must be partitionable into a set of effects that \emph{happened
-before} $\eff$ and another set that \emph{happened after} $\eff$,
-where $\eff$ is the effect generated by an unavailable operation. The
-rule $\rulelabel{UA}$ enforces this sequencing in two steps; firstly,
-it insists that the the unavailable operation ($op$) witness effects
-of all operations executed so far by requiring the global set of
-effects $\EffSoup$ to be a subset of local state ($\Theta(r)$) of the
-replica ($r$) executing $op$. Secondly, the rule requires the effect
-($\eff$) generated by $op$ to be added to the local state of every
-other replica in the store, so that further operations on these
-replicas can witness the effect of $op$. Since both these steps
-require global coordination among replicas, the store is
-\emph{unavailable} during the time it is executing $op$. However, it
-must be noted that executing unavailable operation does not entail the
-convergence of local states of all replicas; it simply means that
-there is one replica ($r$) whose local state is the upper bound of
-local states of all replicas, and there exists a set containing
-atleast one effect ($\eff$), which is their lower bound.
+An unavailable operation expects strong (sequential) consistency. That is,
+universe of all effects ($\EffSoup$) in an execution ($\E$) must be
+partitionable into a set of effects that \emph{happened before} $\eff$ and
+another set that \emph{happened after} $\eff$, where $\eff$ is the effect
+generated by an unavailable operation. The rule $\rulelabel{UA}$ enforces this
+sequencing in two steps; firstly, it insists that the the unavailable operation
+($op$) witness effects of all operations executed so far by requiring the
+global set of effects $\EffSoup$ to be a subset of local state ($\Theta(r)$) of
+the replica ($r$) executing $op$. Secondly, the rule requires the effect
+($\eff$) generated by $op$ to be added to the local state of every other
+replica in the store, so that further operations on these replicas can witness
+the effect of $op$. Since both these steps require global coordination among
+replicas, the store is \emph{unavailable} during the time it is executing $op$.
+However, it must be noted that executing unavailable operation does not entail
+the convergence of local states of all replicas; it simply means that there is
+one replica ($r$) whose local state is the upper bound of local states of all
+replicas, and there exists a set containing atleast one effect ($\eff$), which
+is their lower bound.
 
@@
 
-$\models$ denotes the standard \emph{models} relation of
-first-order logic.  A valid model for a first-order formula defines
-the universe of discourse, and provides interpretations for free
-symbols in the formula such that the formula is true under given
-interpretations.  An execution state ($\E$) provides interpretations
-for $\Rvis$ and $\Rso$ relations, while fixing the universe of
-discourse to be the set of all generated effects ($\EffSoup$).
-Intuitively, the well-formedness definition captures the idea that for
-every generated effect, the contract on the corresponding operation is
-satisfied under the given execution state.
+$\models$ denotes the standard \emph{models} relation of first-order logic.  A
+valid model for a first-order formula defines the universe of discourse, and
+provides interpretations for free symbols in the formula such that the formula
+is true under given interpretations.  An execution state ($\E$) provides
+interpretations for $\Rvis$ and $\Rso$ relations, while fixing the universe of
+discourse to be the set of all generated effects ($\EffSoup$). Intuitively, the
+well-formedness definition captures the idea that for every generated effect,
+the contract on the corresponding operation is satisfied under the given
+execution state.
 
-Our main meta-theoretic result guarantees the soundness of our
-operational semantics in enforcing all contractual obligations by
-ensuring that every reduction step preserves the well-formedness of
-execution state.
+Our main meta-theoretic result guarantees the soundness of our operational
+semantics in enforcing all contractual obligations by ensuring that every
+reduction step preserves the well-formedness of execution state.
 
@@
 
-The proof of the theorem, which we could not include here owing to
-space constraints, can be found in our tech report~\cite{techrep}.
+The proof of the theorem, which we could not include here owing to space
+constraints, can be found in our tech report~\cite{techrep}.
 

### 2014-10-28 contract-language.tex
contract language
Formatting operational semantics

@@
 carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}.  Completeness of the logic~\cite{completeness} allows
-us to claim that $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st} \Rightarrow \cv_{op}$.
-Due to the decidability of our contract language, this implication check is
-automatically discharged with the help of a theorem prover. In this case, we
-say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$ (written $\cv_{op} \le
-\cv_{st}$).
+decidable~\cite{epr}.  Completeness of the logic~\cite{completeness} allows us
+to claim that $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
+\Rightarrow \cv_{op}$. Due to the decidability of our contract language, this
+implication check is automatically discharged with the help of a theorem
+prover. In this case, we say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$
+(written $\cv_{op} \le \cv_{st}$).
 
@@
 
-\item \textbf{Eventually consistency}: Eventually consistent operations can be
-satisfied as long as the client can reach at least one replica. For example,
-\cf{deposit} is an eventually consistent operation. While eventually consistent data
-store typically offer \emph{basic} eventual consistency with all possible
-anomalies, we assume that our store provides stronger semantics that remain
-highly-available~\cite{BailisHAT,COPS}; the store always exposes a causal cut of
-the updates. This semantics can be formally captured in terms of the
-following contract definition:
+\item \textbf{Eventually consistency}: Eventually consistent operations can
+  be satisfied as long as the client can reach at least one replica. For
+  example, \cf{deposit} is an eventually consistent operation; its semantics
+  does not require its action to manifest on all replicas before other
+  operations in its session are allowed to proceed. While eventually
+  consistent data store typically offer \emph{basic} eventual consistency
+  with all possible anomalies, we assume that our store provides stronger
+  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
+  always exposes a causal cut of the updates. This semantics can be formally
+  captured in terms of the following contract definition:
 \vspace{-1em}
 \begin{smathpar}
-\ecc = \forall a,b. \hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
+\ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
 \end{smathpar}

### 2014-10-28 intro1.tex

--- 2014-10-27 intro.tex
@@
 
-Many real-world web services --- such as those built and maintained by Amazon,
-Facebook, Google, Twitter, etc. --- replicate application state and logic
-across multiple \emph{replicas} within and across data centers. Replication is
-intended not only to improve application throughput and reduce user-perceived
-latency, but also to tolerate partial failures without compromising overall
-service availability. Traditionally programmers have relied on \emph{strong
-consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
-serializability~\cite{Serializability} in order to build correct applications.
-While strong consistency is an easily stated property, it masks the reality
-underlying large-scale distributed systems with respect to non-uniform latency
-(i.e., availability) and network partitions~\cite{Brewer2000,Gilbert}. Indeed,
-modern web services, which aim to provide an "always on" experience,
-overwhelmingly favour availability and partition tolerance over strong
-consistency. To this end, several \emph{weak consistency} models such as
-eventual consistency, causal consistency, session guarantees, and timeline
-consistency have consequently been proposed.
+Many real-world web services --- such as those built and maintained by
+Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
+logic across multiple \emph{replicas} within and across data
+centers. Replication is intended not only to improve application throughput
+and reduce user-perceived latency, but also to tolerate partial failures
+without compromising overall service availability. Traditionally programmers
+have relied on \emph{strong consistency} guarantees such as
+linerarizability~\cite{Herlihy1990} or
+serializability~\cite{Serializability} in order to build correct
+applications. While strong consistency is an easily stated property, it
+masks the reality underlying large-scale distributed systems with respect to
+non-uniform latency (i.e., availability) and network
+partitions~\cite{Brewer2000,Gilbert}.  Indeed, modern web services, which
+aim to provide an "always on" experience, overwhelmingly favour availability
+and partition tolerance over strong consistency. Several \emph{weak
+  consistency} models such as eventual consistency, causal consistency,
+session guarantees, and timeline consistency have consequently been
+proposed.
 
@@
 conflicting updates, and has to pay careful attention to avoid unwanted
-inconsistencies (e.g., negative balances in a bank account, or having an item
-appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
-the inconsistency leaks from the application and is witnessed by the user.
-Ultimately, the developer must decide the consistency level appropriate for a
-particular operation; this is understandably an error-prone process requiring
-intricate knowledge of both the application as well as the semantics and
-implementation of the underlying data store.  More often than not, only
-informal descriptions are available~\cite{}.  Picking the correct consistency
-level is critical not only for correctness but also for scalability of the
-application. While choosing a weaker consistency level than required may
-introduce program errors and anomalies, choosing a stronger one than necessary
-level can negatively impact program scalability.
+inconsistencies (e.g., negative balances in a bank account, or having an
+item appear in a shopping cart after it has been
+removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
+application and is witnessed by the user.  Ultimately, the developer must
+decide the consistency level appropriate for a particular operation; this is
+understandably an error-prone process requiring intricate knowledge of both
+the application as well as the semantics and implementation of the
+underlying data store.  More often than not, only informal descriptions are
+available~\cite{}.  Picking the correct consistency level is critical not
+only for correctness but also for scalability of the application. While
+choosing a weaker consistency level than required may introduce program
+errors and anomalies, choosing a stronger one than necessary level can
+negatively impact program scalability.
 
@@
 Moreover, weak consistency hinders compositional reasoning about programs.
-While an application might be naturally expressed in terms of well-understood
-and expressive data types such as maps, trees, queues, or graphs,
-geo-distributed stores typically only provide a minimal set of data types with
-in-built conflict resolution strategies such as last-writer-wins (LWW)
-registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor, while
-traditional database systems enable composability through transactions,
-existing geo-distributed stores lack unrestricted transactional access to the
-data.  Working with geo-distributed stores thus requires application state to
-be suitably coerced to work within the capabilities of the store.
+While an application might be naturally expressed in terms of
+well-understood and expressive data types such as maps, trees, queues, or
+graphs, geo-distributed stores typically only provide a minimal set of data
+types with in-built conflict resolution strategies such as last-writer-wins
+(LWW) registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermor,
+while traditional database systems enable composability through
+transactions, existing geo-distributed stores lack unrestricted
+transactional access to the data.  Working with geo-distributed stores thus
+requires application state to be suitably coerced to work within the
+capabilities of the store.
 
-To address these issues, we describe \name, a declarative programming model and
-implementation for eventually consistent stores. The key novelty of \name is an
-expressive \emph{contract language} to declare and verify fine-grained
-application-level consistency properties. The programmer uses the contract
-language to axiomatically describe the set of legal executions allowed over the
-replicated data type. Contracts are constructed using primitive consistency
-relations such as \emph{visibility} and \emph{session order} along with
-standard logical and relational operators. A \emph{contract enforcement system}
-automatically maps each operation to a particular consistency level available
-on the store. In addition, the system statically reports an error if the given
-application-level consistency property cannot be satisfied by the store.
+To address these issues, we describe \name, a declarative programming model
+and implementation for eventually consistent stores. The key novelty of
+\name\ is an expressive \emph{contract language} to declare and verify
+fine-grained application-level consistency properties.  The programmer uses
+the contract language to axiomatically describe the set of legal executions
+allowed over the replicated data type.  Contracts are constructed using
+primitive consistency relations such as \emph{visibility} and \emph{session
+  order} along with standard logical and relational operators.  A
+\emph{contract enforcement system} automatically maps each operation to a
+particular consistency level available on the store.  In addition, the
+system statically reports an error if the given application-level
+consistency property cannot be satisfied by the store.
 
@@
 \item We introduce \name, a language for declaratively describing eventually
-	consistent programs that manipulate replicated data types. Contracts are used
-	to specify fine-grained application-level consistency properties, and are
-	analyzed to assign the most efficient and sound store consistency level to
-	the corresponding operation.
+  consistent programs that manipulate replicated data types. Contracts are
+  used to specify fine-grained application-level consistency properties, and
+  are analyzed to assign the most efficient and sound store consistency
+  level to the corresponding operation.
 
-\item \name supports coordination-free transactions over arbitrary datatypes.
-	We extend our contract language to express fine-grained transaction isolation
-	guarantees, and utilize the contract enforcement system to automatically
-	assign the correct and most efficient isolation level for a transaction.
+\item \name supports coordination-free transactions over arbitrary
+  datatypes.  We extend our contract language to express fine-grained
+  transaction isolation guarantees, and utilize the contract enforcement
+  system to automatically assign the correct and most efficient isolation
+  level for a transaction.
 
@@
 
-\item \name is implemented as a shim layer over (and importantly, without
-	modifying) Cassandra~\cite{Cassandra}, a well-known general-purpose data
-	store.  An extensive experimental evaluation over a collection of real-world
-	applications, including a Twitter-like micro-blogging site and an eBay-like
-	auction site illustrates the practicality of our approach.
+\item \name is implemented as a shim layer over (and importantly without
+  modifying) Cassandra, a well-known general-purpose data store.  An
+  extensive experimental evaluation over a collection of real-world
+  applications, including a Twitter-like micro-blogging site and an
+  eBay-like auction site illustrates the practicality of our approach.
 
@@
 Section~\ref{sec:opsem} presents an operational semantics.
-Section~\ref{sec:txns} introduces transaction contracts and classification.
-Section~\ref{sec:impl} describes the implementation and goes into detail about
-the optimizations needed to make the system practical.
+Section~\ref{sec:txns} introduces transaction contracts and
+classification. Section~\ref{sec:impl} describes the implementation and goes
+into detail about the optimizations needed to make the system practical.
 Section~\ref{sec:results} discusses experimental evaluation.

### 2014-10-28 transactions.tex
Adding text in transaction section

\section{Transaction Contracts}

While contracts on individual operations offer the programmer object-level
declarative reasoning, real-world scenarios often involve operations that span
multiple objects. In order to address this problem, several recent
systems~\cite{COPS,BurckhardtESOP,BailisHAT} have proposed a variety of
transactions, with varying semantics, in order to compose operations on
multiple objects. However, given that classical transaction models such as
serializability~\cite{} and snapshot isolation~\cite{} require inter-replica
coordination, these systems espouse \emph{coordination-free transactions} that
remain available under network partitions, but only provide weaker isolation
guarantees.

Coordination-free transactions have intricate consistency semantics and widely
varying runtime overheads. As with operation-level consistency, the onus is on
the programmer to pick the correct transaction kind, the choice of which is
complicated by the consistency requirement of the operations it composes. \name
automates the choice of assigning the correct and most efficient transaction
isolation level.

### 2014-10-28 contract-language.tex
Minor edit
Adding text in transaction section
Edits to Motiv. Removed opsem. Added metatheory.

@@
 The syntax of our core contract language is shown in Fig.
-~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
-admits prenex universal quantification over typed effect variables. We use a special
-effect variable ($\cureff$) to denote the effect of \emph{current operation} -
-the operation for which a contract is being written. The type of an effect is
-simply the name of the operation (eg: \cf{withdraw}) that induced the
-effect. We admit disjuntion in types to let an effect variable range over
-multiple operation names.
+~\ref{fig:contract-lang}. The language is based on first-order logic
+(FOL), and admits prenex universal quantification over typed effect
+variables. We use a special effect variable ($\cureff$) to denote the
+effect of \emph{current operation} - the operation for which a
+contract is being written. The type of an effect is simply the name of
+the operation (eg: \cf{withdraw}) that induced the effect. We admit
+disjuntion in types to let an effect variable range over multiple
+operation names.
 
-Quantifier-free propositions in our contract language are conjunctions,
-disjunctions and implications of predicates expressing relations between pairs
-of effect variables. The syntactic class of relations is seeded with primitive
-$\visZ$, $\soZ$, and $\sameobjZ$ relations, and also admits derived relations
-that are expressible as union, intersection, or transitive
-closure\footnote{Strictly speaking, $R^{+}$ is not the transitive closure of
-$R$, as transitive closure is not expressible in FOL. Instead, $R^{+}$ in our
-language denotes \emph{a} superset of transitive closure of $R$. Formally,
-$R^{+}$ is any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
-\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$} of
-primitive relations. Commonly used derived relations are the \emph{same object
-session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), the \emph{happens
-before order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$), and the per-object
-happens-before order ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
+Quantifier-free propositions in our contract language are
+conjunctions, disjunctions and implications of predicates expressing
+relations between pairs of effect variables. The syntactic class of
+relations is seeded with primitive $\visZ$, $\soZ$, and $\sameobjZ$
+relations, and also admits derived relations that are expressible as
+union, intersection, or transitive closure\footnote{Strictly speaking,
+$R^{+}$ is not the transitive closure of $R$, as transitive closure is
+not expressible in FOL.  Instead, $R^{+}$ in our language denotes
+\emph{a} superset of transitive closure of $R$. Formally, $R^{+}$ is
+any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
+\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow
+R'(x,z)$} of primitive relations.  Commonly used derived relations are
+the \emph{same object session order} ($\small \sooZ = \soZ ~\cap~
+\sameobjZ$), 
+%the \emph{happens before order} ($\small \hbZ = (\soZ
+%~\cup~ \visZ)^+$), 
+and the \emph{same object happens-before order} ($\small \hboZ = (\sooZ
+~\cup~ \visZ)^+$).
 
@@
 
-Since our contracts represent axiomatic definition of program executions, let
-$\Mod{\cv}$ be the set of all executions under which contract $\cv$ is satisfied. If
-every execution $\E \in \Mod{\cv_{st}}$ is also a member of $\Mod{\cv_{op}}$,
-then $op$ can be safely discharged under the store consistency level
-$\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq \Mod{\cv_{op}}$. This is the
-model-theoretic consequence relation, written as $\cv_{st} \models_m \cv_{op}$.
+Since our contracts represent axiomatic definition of program
+executions, let $\Mod{\cv}$ be the set of all executions under which
+contract $\cv$ is satisfied. If every execution $\E \in
+\Mod{\cv_{st}}$ is also a member of $\Mod{\cv_{op}}$, then $op$ can be
+safely discharged under the store consistency level $\cv_{st}$.
+Formally, $\Mod{\cv_{st}} \subseteq \Mod{\cv_{op}}$. This is the
+model-theoretic consequence relation, written as $\cv_{st} \models_m
+\cv_{op}$. In this case, we say that $\cv_{op}$ is \emph{weaker than}
+$\cv_{st}$ (written $\cv_{op} \le \cv_{st}$).
 
@@
 carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}.  Completeness of the logic~\cite{completeness} allows us
-to claim that $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
-\Rightarrow \cv_{op}$. Due to the decidability of our contract language, this
-implication check is automatically discharged with the help of a theorem
-prover. In this case, we say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$
-(written $\cv_{op} \le \cv_{st}$).
+decidable~\cite{epr}.  Since first-order logic is sound and complete
+~\cite{completeness}, $\cv_{st} \models_m \cv_{op}$ if and only if
+$\cv_{st} \Rightarrow \cv_{op}$.  Due to the decidability of our
+contract language, this implication check is automatically discharged
+with the help of a theorem prover.
 
@@
   captured in terms of the following contract definition:
-\vspace{-1em}
+\vspace{-0.6em}
 \begin{smathpar}
-\ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
+\ecc = \forall a,b. \hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
 \end{smathpar}
@@
   session are applied to two different replicas $r_1$ and $r_2$, the second
-  operation cannot be discharged until the effect of $o_1$ is merged with
-  $o_2$ in both $r_1$ and $r_2$.  We assume that causality is only tracked
-  through operations on the same object; two operations in the same session
-  but on different objects are considered causally unrelated under this
-  definition. Stores typically avoid tracking causality across objects to
-  mitigate overheads when causality tracking is unnecessary.  The corresponding
-  store semantics is captured by the contract $\ccc$ presented in
-  Section~\ref{sec:store_sem}.
+  operation cannot be discharged until the effect of $o_1$ is merged
+  with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance} operation
+  requires causal consistency, as it requires the operations from the
+  same session to be visible, which cannot be guaranteed under
+  eventual consistency. We assume that causality is only tracked
+  through operations on the same object; two operations in the same
+  session but on different objects are considered causally unrelated
+  under this definition. Stores typically avoid tracking causality
+  across objects to mitigate overheads when causality tracking is
+  unnecessary. The corresponding store semantics is captured by the
+  contract $\ccc$ presented in Section~\ref{sec:store_sem}.
 
@@
 \ccc \le \scc$. This concurs with the intuition that any contract
-satisfiable under $\ecc$ or $\ccc$ is satisfiable under $\scc$ , and any
+satisfiable under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any
 contract that is satisfiable under $\ecc$ is satisfiable under
@@
 satisfy the contract. The classification scheme is presented formally 
-in Figure~\ref{sem:classify}; here, $\Delta$ captures assumptions about the nature of primitive
-relations, such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
-relation $\hbZ$ is acyclic, preventing thin-air reads. \KC{Ideally, we need to
-list all the axioms somewhere.}  The following definition thus precisely
-captures our intuition about contract strength in terms of these rules:
+in Figure~\ref{sem:classify}. Along with three straightforward rules
+that classify contracts into consistency classes, the classification
+scheme also presents a rule that judge well-formedness of a contract.
+A contract is well-formed if and only if it is satisfiable under
+$\scc$ - the strongest possible consistency guarantee any store can
+provide. The $\le$ relation used by the rules is defined formally
+below:
 
@@
 \vspace{-1em}
+The $\Delta$ referred in the above defintion captures assumptions
+about the nature of primitive relations, such as $\Rvis$ and $\Rso$
+are irreflexive, and the happens-before relation $\hbZ$ is acyclic,
+preventing thin-air reads. \KC{Ideally, we need to list all the axioms
+somewhere.}
 
@@
 \input{contract-classification}
-
-%% The rules to classify contracts into availability classes are straightforward.
-%% The remaining rule judges a contract as {\sf WellFormed} if and only if it can
-%% be satisfied under strong consistency. The rest are reported as errors.

### 2014-10-28 contract-language.tex
More text in transaction section

--- 2014-10-28 contract-language.tex
@@
 ~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
-admits prenex universal quantification over typed effect variables. We use a special
-effect variable ($\cureff$) to denote the effect of \emph{current operation} -
-the operation for which a contract is being written. The type of an effect is
-simply the name of the operation (eg: \cf{withdraw}) that induced the
-effect. We admit disjuntion in types to let an effect variable range over
-multiple operation names.
+admits prenex universal quantification over typed effect variables. We use a
+special effect variable ($\cureff$) to denote the effect of \emph{current
+operation} - the operation for which a contract is being written. The type of
+an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
+the effect. We admit disjuntion in types to let an effect variable range over
+multiple operation names. The contract $\small \forall (a : \tau_1 \vee
+\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
+\vee \oper{a}{\tau_2}) \Rightarrow \psi$.
 
@@
 Since our contracts represent axiomatic definition of program executions, let
-$\Mod{\cv}$ be the set of all executions under which contract $\cv$ is satisfied. If
-every execution $\E \in \Mod{\cv_{st}}$ is also a member of $\Mod{\cv_{op}}$,
-then $op$ can be safely discharged under the store consistency level
-$\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq \Mod{\cv_{op}}$. This is the
-model-theoretic consequence relation, written as $\cv_{st} \models_m \cv_{op}$.
+$\Mod{\cv}$ be the set of all executions under which contract $\cv$ is
+satisfied. If every execution $\E \in \Mod{\cv_{st}}$ is also a member of
+$\Mod{\cv_{op}}$, then $op$ can be safely discharged under the store
+consistency level $\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq
+\Mod{\cv_{op}}$. This is the model-theoretic consequence relation, written as
+$\cv_{st} \models_m \cv_{op}$.
 
@@
 	does not require its action to manifest on all replicas before other
-  operations in its session are allowed to proceed. While eventually
-  consistent data store typically offer \emph{basic} eventual consistency
-  with all possible anomalies, we assume that our store provides stronger
-  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
-  always exposes a causal cut of the updates. This semantics can be formally
-  captured in terms of the following contract definition:
+	operations in its session are allowed to proceed. While eventually consistent
+	data store typically offer \emph{basic} eventual consistency with all
+	possible anomalies, we assume that our store provides stronger semantics that
+	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
+	causal cut of the updates. This semantics can be formally captured in terms
+	of the following contract definition:
+
 \vspace{-1em}
@@
 	session are applied to two different replicas $r_1$ and $r_2$, the second
-  operation cannot be discharged until the effect of $o_1$ is merged with
-  $o_2$ in both $r_1$ and $r_2$.  We assume that causality is only tracked
-  through operations on the same object; two operations in the same session
-  but on different objects are considered causally unrelated under this
-  definition. Stores typically avoid tracking causality across objects to
-  mitigate overheads when causality tracking is unnecessary.  The corresponding
-  store semantics is captured by the contract $\ccc$ presented in
+	operation cannot be discharged until the effect of $o_1$ is merged with $o_2$
+	in both $r_1$ and $r_2$.  \cf{getBalance} operation is classified as a
+	causally consistent operation. We assume that causality is only tracked
+	through operations on the same object; two operations in the same session but
+	on different objects are considered causally unrelated under this definition.
+	Stores typically avoid tracking causality across objects to mitigate
+	overheads when causality tracking is unnecessary.  The corresponding store
+	semantics is captured by the contract $\ccc$ presented in
 	Section~\ref{sec:store_sem}.
@@
 
-\noindent Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are
-themselves totally ordered with respect to the $\le$ relation: $\ecc \le
-\ccc \le \scc$. This concurs with the intuition that any contract
-satisfiable under $\ecc$ or $\ccc$ is satisfiable under $\scc$ , and any
-contract that is satisfiable under $\ecc$ is satisfiable under
-$\ccc$. Nonetheless, we determine the class of a contract based on the
-\emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$) required to
-satisfy the contract. The classification scheme is presented formally
-in Figure~\ref{sem:classify}; here, $\Delta$ captures assumptions about the nature of primitive
-relations, such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
-relation $\hbZ$ is acyclic, preventing thin-air reads. \KC{Ideally, we need to
-list all the axioms somewhere.}  The following definition thus precisely
-captures our intuition about contract strength in terms of these rules:
+\noindent Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
+totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$.
+This concurs with the intuition that any contract satisfiable under $\ecc$ or
+$\ccc$ is satisfiable under $\scc$ , and any contract that is satisfiable under
+$\ecc$ is satisfiable under $\ccc$. Nonetheless, we determine the class of a
+contract based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and
+$\scc$) required to satisfy the contract. The classification scheme is
+presented formally in Figure~\ref{sem:classify}; here, $\Delta$ captures
+assumptions about the nature of primitive relations, such as $\Rvis$ and $\Rso$
+are irreflexive, and the happens-before relation $\hbZ$ is acyclic, preventing
+thin-air reads. \KC{Ideally, we need to list all the axioms somewhere.}  The
+following definition thus precisely captures our intuition about contract
+strength in terms of these rules:
 

### 2014-10-28 transactions.tex
More text in transaction section

@@
 remain available under network partitions, but only provide weaker isolation
-guarantees.
+guarantees. Coordination-free transactions have intricate consistency semantics
+and widely varying runtime overheads. As with operation-level consistency, the
+onus is on the programmer to pick the correct transaction kind, the choice of
+which is complicated by the consistency requirement of the operations it
+composes.
 
-Coordination-free transactions have intricate consistency semantics and widely
-varying runtime overheads. As with operation-level consistency, the onus is on
-the programmer to pick the correct transaction kind, the choice of which is
-complicated by the consistency requirement of the operations it composes. \name
-automates the choice of assigning the correct and most efficient transaction
-isolation level.
+\subsection{Syntax Extension}
+
+\name automates the choice of assigning the correct and most efficient
+transaction isolation level. Similar to contracts on individual operations, the
+programmer associates contracts with transactions, declaratively expressing its
+consistency specification. We extend the contract language with a new term
+under quantifier-free propositions ${\small \txnZ}~S_1~S_2$, where $S_1$ and
+$S_2$ are sets of effects, and introduce a new primitive equivalence relation
+$\small \sametxnZ$ that holds for effects from the same transaction. $\small
+\txn{a,b}{c,d}$ is just syntactic sugar for $\small \sametxn{a}{b} ~\wedge~
+\sametxn{c}{d} ~\wedge~ \neg\sametxn{a}{c}$, where $a$ and $b$ considered to
+belong to the \emph{current} transaction. We assume that operations not part of
+any transaction to belong to their own unique transaction. While the
+transactions may have varying isolation guarantees, we make the standard
+assumption that all transactions provide atomicity. Hence, we include the
+following axiom in $\Delta$: $\small \forall a,b,c.~\txn{a}{b,c} ~\wedge~
+\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
+\vis{c}{a}$.
+
+\subsection{Transactional Bank Account}
+
+In order to illustrate the utility of declarative reasoning for transactions,
+let us extend our running bank account example with use two accounts (objects)
+-- current ($c$) and savings ($s$). Each account provides operations
+\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
+defined previously. We consider two transactions -- \cf{save(amt)}, which
+transfers \cf{amt} from current to savings, and \cf{totalBalance} returns
+the sum of the balances of individual accounts. The pseudo code for the
+transactions is given below:
+
+\noindent \begin{minipage}[t]{0.5\columnwidth}
+\begin{codepython}
+def save(amt):
+  atomically psi1:
+    b = c.withdraw(amt)
+    if (b): s.deposit(amt)
+\end{codepython}
+\end{minipage}
+\begin{minipage}[t]{0.5\columnwidth}
+\begin{codepython}
+def totalBalance():
+  atomically psi2:
+    b1 = c.getBalance()
+    b2 = s.getBalance()
+    return b1 + b2
+\end{codepython}
+\end{minipage}
+
+\noindent where $\cv_1$ and $\cv_2$ are the contracts on the corresponding
+transactions. Our goal is to ensure that \cf{totalBalance} returns the result
+obtained from a consistent snapshot of the object states.
+
+While making both transactions serializable would ensure correctness,
+distributed stores rarely offer serializable transactions. Moreover,
+serializability is unavailable and hinders scalability. As we will see, these
+transactions can be satisfied with much weaker isolation guarantees. Despite
+the atomicity offered by the transaction, anomalies are still possible.

### 2014-10-28 contract-language.tex

@@
 primitive relations.  Commonly used derived relations are the \emph{same object
-session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), the \emph{happens
-before order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$), and the per-object
-happens-before order ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
+session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), and the \emph{same
+object happens-before order} ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
 
@@
 carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}.  Completeness of the logic~\cite{completeness} allows us
-to claim that $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
+decidable~\cite{epr}.  Since first-order logic is sound and complete
+~\cite{completeness}, $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
 \Rightarrow \cv_{op}$.  Due to the decidability of our contract language, this
 implication check is automatically discharged with the help of a theorem
-prover. In this case, we say that $\cv_{op}$ is \emph{weaker than} $\cv_{st}$
-(written $\cv_{op} \le \cv_{st}$).
+prover.
 
@@
   does not require its action to manifest on all replicas before other
-	operations in its session are allowed to proceed. While eventually consistent
-	data store typically offer \emph{basic} eventual consistency with all
-	possible anomalies, we assume that our store provides stronger semantics that
-	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
-	causal cut of the updates. This semantics can be formally captured in terms
-	of the following contract definition:
+  operations in its session are allowed to proceed. While eventually
+  consistent data store typically offer \emph{basic} eventual consistency
+  with all possible anomalies, we assume that our store provides stronger
+  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
+  always exposes a causal cut of the updates. This semantics can be formally
+  captured in terms of the following contract definition:
 
-\vspace{-1em}
 \begin{smathpar}
@@
 	operation cannot be discharged until the effect of $o_1$ is merged with $o_2$
-	in both $r_1$ and $r_2$.  \cf{getBalance} operation is classified as a
-	causally consistent operation. We assume that causality is only tracked
-	through operations on the same object; two operations in the same session but
-	on different objects are considered causally unrelated under this definition.
-	Stores typically avoid tracking causality across objects to mitigate
-	overheads when causality tracking is unnecessary.  The corresponding store
-	semantics is captured by the contract $\ccc$ presented in
-	Section~\ref{sec:store_sem}.
+	in both $r_1$ and $r_2$. The \cf{getBalance} operation requires causal
+	consistency, as it requires the operations from the same session to be
+	visible, which cannot be guaranteed under eventual consistency. We assume
+	that causality is only tracked through operations on the same object; two
+	operations in the same session but on different objects are considered
+	causally unrelated under this definition. Stores typically avoid tracking
+	causality across objects to mitigate overheads when causality tracking is
+	unnecessary. The corresponding store semantics is captured by the contract
+	$\ccc$ presented in Section~\ref{sec:store_sem}.
 
@@
 This concurs with the intuition that any contract satisfiable under $\ecc$ or
-$\ccc$ is satisfiable under $\scc$ , and any contract that is satisfiable under
+$\ccc$ is satisfiable under $\scc$, and any contract that is satisfiable under
 $\ecc$ is satisfiable under $\ccc$. Nonetheless, we determine the class of a
 contract based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and
-$\scc$) required to satisfy the contract. The classification scheme is
-presented formally in Figure~\ref{sem:classify}; here, $\Delta$ captures
-assumptions about the nature of primitive relations, such as $\Rvis$ and $\Rso$
-are irreflexive, and the happens-before relation $\hbZ$ is acyclic, preventing
-thin-air reads. \KC{Ideally, we need to list all the axioms somewhere.}  The
-following definition thus precisely captures our intuition about contract
-strength in terms of these rules:
+$\scc$) required to satisfy the contract.
+
+% Contract classification rules
+% ------------------------------
+\input{contract-classification}
+
+The classification scheme is presented formally in Figure~\ref{sem:classify}.
+Along with three straightforward rules that classify contracts into consistency
+classes, the classification scheme also presents a rule that judge
+well-formedness of a contract. A contract is well-formed if and only if it is
+satisfiable under $\scc$ - the strongest possible consistency guarantee any
+store can provide. The $\le$ relation used by the rules is defined formally
+below:
 
@@
 
-% Contract classification rules
-% ------------------------------
-\input{contract-classification}
-
-%% The rules to classify contracts into availability classes are straightforward.
-%% The remaining rule judges a contract as {\sf WellFormed} if and only if it can
-%% be satisfied under strong consistency. The rest are reported as errors.
+The $\Delta$ referred in the above defintion captures assumptions about the
+nature of primitive relations, such as $\Rvis$ and $\Rso$ are irreflexive, and
+the happens-before relation $\hbZ$ is acyclic, preventing thin-air reads.
+\KC{Ideally, we need to list all the axioms somewhere.}

### 2014-10-28 core-lang.tex

--- 2014-10-27 core-lang.tex
@@
-\section{Operational Semantics}
+\section{Meta-theory}
 \label{sec:core-opsem}
 
-We formalize a replicated data store as a triple $\E;\Theta;\Sigma$,
-where $\E$ denotes the execution state, $\Theta$ maps replicas to
-their local state, and $\Sigma$, the session soup, is the set of
-concurrent client sessions interacting with the store.  Operational
-semantics given in Figure~\ref{sem:oper} succinctly capture the
-behaviour of the store that is relevant to demonstrate the operational
-significance of contract classification. For the sake of clarity, we
-only consider a single replicated object of well-defined type (for eg:
-a replicated object of type \cf{BankAccount}) in our formalization.
-Our semantics are parametric over the specification of this replicated
-data type.  Figure~\ref{sem:oper} formalizes replicated data type
-(RDT) specification as tuple $(\tau,\Ops,\Ctrts)$, where $\tau$ is the
-data type, $\Ops$ maps labels ($op$) of operations on $\tau$ to their
-definitions, while $\Ctrts$ maps them to their consistency contracts
-($\cv$). The definition of an operation is expected to be a lambda
-expression in the server-side language\footnote{Since a replicated
-object is maintained by a server, we call the language in which
-replicated data type operations are defined as \emph{server-side
-language}.}, although we do not enforce this in our formalization. We
-intentionally leave the server-side language underspecified to
-demonstrate that our ideas are independent of the choice of the
-programming language used to implement the replicated data type.
+In this section, we present the meta-theoretic result that certifies
+the soundness of our classification-based contract enforcement
+strategy. 
+% We show that any data store can correctly enforce $\scc$,
+% $\ccc$, and $\ecc$ guarantees can satisfy application-level
+% consistency constraints of all \name contracts.
+
+Fig.~\ref{sem:oper} summarizes the artifacts needed to state our
+meta-theoretic result. We formalize an abstract model of a replicated
+data store as a tuple $\E;\Sigma$, where $\E$ denotes the execution
+state, and $\Sigma$, the session soup, is the set of concurrent client
+sessions interacting with the store. The execution state ($\E$) of a
+data store is modeled as a tuple $\Exec$ where $\EffSoup$ is a set of
+effects during the execution, and $\visZ,\soZ \subseteq \EffSoup
+\times \EffSoup$ are visibility and session order relations,
+respectively, witnessed over generated effects. A session is modeled
+simply as a sequence of replicated data type operations ($op$), tagged
+with the consistency class ($\tau$) of their contracts (as determined
+by the contract classification scheme). We abbreviate contract classes
+($\tau$) as uppercase {\sf SC}, {\sf CC}, and {\sf EC} for brevity.
+Their representative contracts ($\cv(\tau)$) are, as usual, written
+$\scc$, $\ccc$, and $\ecc$, respectively.  Our contract language and
+classification scheme is independent of store semantics. Accordingly,
+we abstain from making any specific assumptions about semantics of the
+store. However, we do assume the existence of a reduction relation of
+form:
+\begin{smathpar} 
+  \auxred{} {(\E,\Sigma)} {\eff} {(\E',\Sigma')}
+\end{smathpar}
+\noindent The relation captures store's progress in execution (from
+$\E$ to $\E'$) due to the application of an operation from one of the
+sessions in $\Sigma$ to a replica, generating a new effect ($\eff$).
+
+Note that an execution state ($\E$) provides interpretations for
+primitive relations (eg: $\visZ$) that occur free in contract
+formulas, and also fixes the domain of quantification to set of all
+effects ($\EffSoup$) observed during the execution. Therefore, $\E$ is
+a potential model for any contract formula ($\cv$). If $\E$ is indeed
+a model of $\cv$ (i.e., $\E \models \cv$), we say that the store
+enforced the contract $\cv$ in execution $\E$.
 
@@
 
-A session is modeled simply as a sequence of actions on the replicated
-object maintained by the store, where an action is defined as an
-application of replicated data type operation to a value. For
-technical reasons, we tag every session in the session soup ($\Sigma$)
-with a unique {\sf SessID} and the sequence number of next operation
-in that session. As with the server-side language, our development is
-also orthogonal to the choice of the client-side language, but we
-require both the languages to agree on the syntactic class of values
-that replicated data type operations ($op$) produce and consume.
+We now state the soundness of contract enforcement as following:
 
-An execution state $\E$ is a tuple $\Exec$ where $\EffSoup$ is a set
-of effects, and $\visZ,\soZ \subseteq \EffSoup \times
-\EffSoup$ are visibility and session order relations over effects,
-respectively. An effect ($\eff$), modeled as a tuple $(s,i,op,v)$,
-records the fact that $i^{th}$ action in session with {\sf SessID}
-$s$, which is an operation $op$ on the replicated object, has been
-successfully executed on some replica yielding a return value $v$.
-Note that the combination of $s$ and $i$ uniquely identifies the
-effect. Session order relation ($\soZ$) relates effects generated by
-the same session. An effect $\eff = (s,i,op,v)$ is said to precede
-another effect $\eff' = (s',i',op',v')$ in session order if and only
-if $s'=s$ and $i'\ge i$. An effect generated at a replica becomes
-visible at rest of the replicas eventually. The local state of a
-replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
-visible at $r$. An operation $op$ performed at replica $r$ can only
-witness the set of effects ($\Theta(r)$) visible at $r$.  
-% Note that in presence of multiple replicated objects, $\visZ$ only
-% relates effects of operations performed on a single object. For
-% instance, a {\sf withdraw} operation can witness (via $\visZ$) a {\sf
-% deposit} operation on the same bank account, but not a {\sf deposit}
-% on a different account. 
-Often, in our formalization, we use $\visZ$ and
-$\soZ$ binary relations to obtain a set of effects visible to a given
-effect $\eff$, or set of effects that precede a given effect $\eff$ in
-the session order. As a syntactic convenience, whenever $R$ is a
-binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
-such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$
-to denote the set of all $\eff'$ such that $(\eff',\eff) \in R$.
+\begin{theorem}[Soundness of Contract Enforcement]
+\label{lem:core-preservation}
+Let $\cv$ be the contract of a replicated data type operation $op$,
+and let $\tau$ denote the consistency class of $\cv$ 
+as determined by the contract classification scheme. Forall execution
+states $\E$, $\E'$ such that 
+$\auxred{} {(\E,\langle (op,\tau);\,\sigma \rangle \pll \Sigma)} {\eff} 
+ {(\E',\langle \sigma \rangle \pll \Sigma)}$
+if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$ 
+\end{theorem}
 
-Rule $\rulelabel{Comm}$ captures inter-replica communication. The
-basic guarantee provided by the store is causal visibility. That is,
-an effect ($\eff$) is made visible at a replica $r$ only after all
-the effects that causally precede $\eff$ are made visible at $r$.
-Observe that this condition is applicable even for the effects that
-originate at replica $r$. 
-% It is helpful to think of effects being
-% stored in a \emph{write buffer} until causal consistency condition is
-% met, at which point they can be merged with the local state
-% ($\Theta(r)$). 
-It is important to note that that enforcing causal visibility does not
-require any inter-replica coordination.
-
-Rule $\rulelabel{Oper}$ is an auxiliary reduction of the
-form:\vspace{-1.7mm}
-\begin{smathpar}
-\auxred{\Theta} {(\E,\langle s,i,op(v) \rangle)} {r} {(\E',\eff)}
-\vspace{-1.7mm}
-\end{smathpar}
-\noindent Under the store configuration $\Theta$, the rule captures
-the progress in execution (from $\E$ to $\E'$) due to the application
-of operation $op$ to replica $r$ resulting in a new effect $\eff$.
-The rule first constructs a \emph{context} for the application from
-the local state ($\Theta(r)$) of the replica, by
-projecting\footnote{{\textsf{ctxt*}} is auxiliary function
-\textsf{ctxt} extended straightforwardly to set of effects} relevant
-information from effects in $\Theta(r)$. It then substitutes the
-definition ($\Ops(op)$) of the operation for its label ($op$), and
-relies on the reduction relation ($\rdtredsto$) of the server-side
-language to reduce the application $\Ops(op)(v,~ctxt)$ to a value
-$v'$.  Subsequently, the the attributes of execution state, namely
-$\EffSoup$, $\visZ$ and $\soZ$, are extended to incorporate
-the new effect ($\eff$).
-
-If the operation $op$ is highly-available (rule $\rulelabel{HA}$), we
-simply apply the operation to any replica $r$. Since the store
-provides causal visibility, highly available operations are
-satisfiable under any replica. If the operation is sticky-available
-(rule $\rulelabel{SA}$), the operation can be applied to a replica $r$
-only if it already contains the effects of all the previous operations
-from the same session. This guarantee can be satisfied by applying all
-operations from the same session to the same logical copy of the
-database. If such a logical copy is untenable, then the operation
-might block. Since the store is assumed to converge eventually, the
-blocked sticky-available operation guaranteed to unblock eventually.
-
-An unavailable operation expects strong (sequential) consistency.
-That is, universe of all effects ($\EffSoup$) in an execution ($\E$)
-must be partitionable into a set of effects that \emph{happened
-before} $\eff$ and another set that \emph{happened after} $\eff$,
-where $\eff$ is the effect generated by an unavailable operation. The
-rule $\rulelabel{UA}$ enforces this sequencing in two steps; firstly,
-it insists that the the unavailable operation ($op$) witness effects
-of all operations executed so far by requiring the global set of
-effects $\EffSoup$ to be a subset of local state ($\Theta(r)$) of the
-replica ($r$) executing $op$. Secondly, the rule requires the effect
-($\eff$) generated by $op$ to be added to the local state of every
-other replica in the store, so that further operations on these
-replicas can witness the effect of $op$. Since both these steps
-require global coordination among replicas, the store is
-\emph{unavailable} during the time it is executing $op$. However, it
-must be noted that executing unavailable operation does not entail the
-convergence of local states of all replicas; it simply means that
-there is one replica ($r$) whose local state is the upper bound of
-local states of all replicas, and there exists a set containing
-atleast one effect ($\eff$), which is their lower bound.
-
-\subsection{Metatheory}
-
-To help us state our main meta-theoretic result, we first define a
-well-formedness condition over execution states:
-
-\begin{definition}[Well-formed execution]
-An execution state $\E = \Exec$ corresponding to an execution of
-operations over a replicated datatype with specification $\rdtspec =
-(\tau,\Ops,\Ctrts)$ is said to be well-formed if and only if $\forall
-\eff \in \EffSoup. ~\E \models \DDe{[\eff/\cureff]\cv}$, where $\cv =
-\Ctrts(\operZ(\eff))$.
-\end{definition}
-
-$\models$ denotes the standard \emph{models} relation of
-first-order logic.  A valid model for a first-order formula defines
-the universe of discourse, and provides interpretations for free
-symbols in the formula such that the formula is true under given
-interpretations.  An execution state ($\E$) provides interpretations
-for $\Rvis$ and $\Rso$ relations, while fixing the universe of
-discourse to be the set of all generated effects ($\EffSoup$).
-Intuitively, the well-formedness definition captures the idea that for
-every generated effect, the contract on the corresponding operation is
-satisfied under the given execution state.
-
-Our main meta-theoretic result guarantees the soundness of our
-operational semantics in enforcing all contractual obligations by
-ensuring that every reduction step preserves the well-formedness of
-execution state.
-
-\begin{theorem}[Well-formedness preservation]
-\label{lem:core-preservation}
-If $\E$ is well-formed and $\E; \Theta; \Sigma \goesto \E'; \Theta';
-\Sigma'$, then $\E'$ is well-formed.
-\end{theorem}
-% Lemmas
-%\input{core-lemmas.tex}
-
-The proof of the theorem, which we could not include here owing to
-space constraints, can be found in our tech report~\cite{techrep}.
-
-% It should be noted that Theorem ~\ref{lem:core-preservation} only
-% proves the safety of our system. Proving its liveness requires us to
-% axiomatize eventual consistency.
+The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ guarantees in all executions, then contract
+classification scheme can be used to satisfy any \name contract on
+that data store. The proof of the theorem has been included in our
+tech report~\cite{techrep}. Tech report also contains operational
+semantics of our implementation of the data store, which concretizes
+the reduction relation ($\xhookrightarrow{}$), along with the proof
+that it correctly enforces all \name contracts.

### 2014-10-28 contract-language.tex
Edits

@@
 
+% Contract classification rules
+% ------------------------------
+\input{contract-classification}
+
 \noindent Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
@@
 contract based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and
-$\scc$) required to satisfy the contract.
-
-% Contract classification rules
-% ------------------------------
-\input{contract-classification}
-
-The classification scheme is presented formally in Figure~\ref{sem:classify}.
-Along with three straightforward rules that classify contracts into consistency
-classes, the classification scheme also presents a rule that judge
-well-formedness of a contract. A contract is well-formed if and only if it is
-satisfiable under $\scc$ - the strongest possible consistency guarantee any
-store can provide. The $\le$ relation used by the rules is defined formally
-below:
+$\scc$) required to satisfy the contract. The classification scheme is
+presented formally in Figure~\ref{sem:classify}. The $\le$ relation used by the
+rules is defined formally below:
 
@@
 the happens-before relation $\hbZ$ is acyclic, preventing thin-air reads.
-\KC{Ideally, we need to list all the axioms somewhere.}
+\KC{Ideally, we need to list all the axioms somewhere.} Along with three
+straightforward rules that classify contracts into consistency classes, the
+classification scheme also presents a rule that judge well-formedness of a
+contract. A contract is well-formed if and only if it is satisfiable under
+$\scc$ - the strongest possible consistency guarantee any store can provide.

### 2014-10-29 contract-language.tex
Edits to Secs 4 and 5.

--- 2014-10-28 contract-language.tex
@@
 
+% \subsection{Capturing Store Semantics}
+% \label{sec:store_sem}
+
+% An important aspect of our contract classification system is that the store
+% semantics is also captured using the same contract language used to describe
+% application-level consistency. In this regard, similar
+% to~\cite{Burckhardt2014}, we can rigorously define a wide variety of store
+% semantics including those that combine any subset of session and causality
+% guarantees, and multiple consistency levels. For example, a store that offers
+% strong consistency is captured by the contract:
+
+% \vspace{-1em}
+% \begin{smathpar}
+% \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
+% \end{smathpar}
+
+% \noindent Similarly, a store that offers per-object causal consistency is captured by the
+% contract:
+
+% \vspace{-1em}
+% \begin{smathpar}
+% \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
+% \end{smathpar}
+
+% \noindent This ability to represent store semantics and application-level consistency in
+% the same language is vital to contract classification.
+
+% \subsection{Contract Comparison}
+
+% Our goal is to classify contracts, and map the operation they describe to
+% the \emph{weakest} store-level consistency level that  nonetheless
+% satisfies the contract's constraints. To this end, we need a mechanism to
+% compare the ``strength'' of a contract. Let $\cv_{op}$ be a contract for a
+% particular operation $op$, and $\cv_{st}$ capture a particular store
+% consistency level. We would like to determine whether $op$ can be
+% \emph{safely discharged} at the store consistency level $\cv_{st}$ such that
+% the resulting execution does not violate $\cv_{op}$.
+
+% Since our contracts represent axiomatic definition of program executions, let
+% $\Mod{\cv}$ be the set of all executions under which contract $\cv$ is
+% satisfied. If every execution $\E \in \Mod{\cv_{st}}$ is also a member of
+% $\Mod{\cv_{op}}$, then $op$ can be safely discharged under the store
+% consistency level $\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq
+% \Mod{\cv_{op}}$. This is the model-theoretic consequence relation, written as
+% $\cv_{st} \models_m \cv_{op}$.
+
+% Observe that our contract language (stripped of its syntactic sugar) is a
+% carefully chosen subset of first-order logic that is known to be
+% decidable~\cite{epr}.  Since first-order logic is sound and complete
+% ~\cite{completeness}, $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
+% \Rightarrow \cv_{op}$.  Due to the decidability of our contract language, this
+% implication check is automatically discharged with the help of a theorem
+% prover.
+
 \subsection{Capturing Store Semantics}
-\label{sec:store_sem}
 
-An important aspect of our contract classification system is that the store
-semantics is also captured using the same contract language used to describe
-application-level consistency. In this regard, similar
-to~\cite{Burckhardt2014}, we can rigorously define a wide variety of store
-semantics including those that combine any subset of session and causality
-guarantees, and multiple consistency levels. For example, a store that offers
-strong consistency is captured by the contract:
+An important aspect of our contract language is its ability to capture
+store-level consistency guarantees, along with application-level
+consistency requirements. Similar to~\cite{Burckhardt2014}, we can
+rigorously define a wide variety of store semantics including those
+that combine any subset of session and causality guarantees, and
+multiple consistency levels.  However, for our purposes, we identify
+three particular consistency levels -- eventual, causal, and strong,
+commonly offered by many distributed stores with tunable consistency,
+with increasing overhead in terms of latency and availability. For
+each of the these three consistency levels, we capture the semantics
+as a formula in our contract language, and informally describe the kind
+of application-level consistency requirements that can met under the
+consistency level:
 
-\vspace{-1em}
-\begin{smathpar}
-\scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
-\end{smathpar}
-
-\noindent Similarly, a store that offers per-object causal consistency is captured by the
-contract:
-
-\vspace{-1em}
-\begin{smathpar}
-\ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
-\end{smathpar}
-
-\noindent This ability to represent store semantics and application-level consistency in
-the same language is vital to contract classification. Observe that out
-contract language does not incorporate real (i.e., wall-clock) time. Hence, the
-contract language cannot describe store semantics based on real time such as
-recency or bounded-staleness guarantees offered by certain
-stores~\cite{Pileus}.
-
-\subsection{Contract Comparison}
-
-Our goal is to classify contracts, and map the operation they describe to
-the \emph{weakest} store-level consistency level that  nonetheless
-satisfies the contract's constraints. To this end, we need a mechanism to
-compare the ``strength'' of a contract. Let $\cv_{op}$ be a contract for a
-particular operation $op$, and $\cv_{st}$ capture a particular store
-consistency level. We would like to determine whether $op$ can be
-\emph{safely discharged} at the store consistency level $\cv_{st}$ such that
-the resulting execution does not violate $\cv_{op}$.
-
-Since our contracts represent axiomatic definition of program executions, let
-$\Mod{\cv}$ be the set of all executions under which contract $\cv$ is
-satisfied. If every execution $\E \in \Mod{\cv_{st}}$ is also a member of
-$\Mod{\cv_{op}}$, then $op$ can be safely discharged under the store
-consistency level $\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq
-\Mod{\cv_{op}}$. This is the model-theoretic consequence relation, written as
-$\cv_{st} \models_m \cv_{op}$.
-
-Observe that our contract language (stripped of its syntactic sugar) is a
-carefully chosen subset of first-order logic that is known to be
-decidable~\cite{epr}.  Since first-order logic is sound and complete
-~\cite{completeness}, $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
-\Rightarrow \cv_{op}$.  Due to the decidability of our contract language, this
-implication check is automatically discharged with the help of a theorem
-prover.
-
-\subsection{Contract Classification}
-
-While the store semantics offered by commercial and research data stores
-vary widely, for our purposes, we identify three particular consistency
-levels -- eventual, causal, and strong, commonly offered by many distributed
-stores with tunable consistency, with increasing overhead in terms of
-latency and availability. Indeed, the techniques presented here can be
-extended to other consistency stratifications. We assign each
-application-level contract into one of these following classes:
+% Indeed, the techniques presented here can be
+% extended to other consistency stratifications. We assign each
+% application-level contract into one of these following classes:
 
@@
   captured in terms of the following contract definition:
-
+  \vspace{-0.8em}
   \begin{smathpar}
@@
 
-\item \textbf{Causal consistency}: Operations with causally consistent
-	contracts are required to see a causally consistent snapshot of the object
-	state, including the actions performed on the same session.  The latter
-	requirement entails that if two operations $o_1$ and $o_2$ from the same
-	session are applied to two different replicas $r_1$ and $r_2$, the second
-	operation cannot be discharged until the effect of $o_1$ is merged with $o_2$
-	in both $r_1$ and $r_2$. The \cf{getBalance} operation requires causal
-	consistency, as it requires the operations from the same session to be
-	visible, which cannot be guaranteed under eventual consistency. We assume
-	that causality is only tracked through operations on the same object; two
-	operations in the same session but on different objects are considered
-	causally unrelated under this definition. Stores typically avoid tracking
-	causality across objects to mitigate overheads when causality tracking is
-	unnecessary. The corresponding store semantics is captured by the contract
-	$\ccc$ presented in Section~\ref{sec:store_sem}.
+\item \textbf{Causal consistency}: Causally consistent operations
+  require to see a causally consistent snapshot of the object state,
+  including the actions performed on the same session.  The latter
+  requirement entails that if two operations $o_1$ and $o_2$ from the
+  same session are applied to two different replicas $r_1$ and $r_2$,
+  the second operation cannot be discharged until the effect of $o_1$ is
+  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
+  operation requires causal consistency, as it requires the operations
+  from the same session to be visible, which cannot be guaranteed under
+  eventual consistency. We assume that causality is only tracked through
+  operations on the same object; two operations in the same session but
+  on different objects are considered causally unrelated under this
+  definition. Stores typically avoid tracking causality across objects
+  to mitigate overheads when causality tracking is unnecessary. The
+  corresponding store semantics is captured by the contract $\ccc$
+  defined below:
+  \vspace{-0.8em}
+  \begin{smathpar}
+  \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
+  \end{smathpar}
 
@@
   contract on \cf{withdraw} operation. The corresponding store semantics is
-  captured by the contract $\scc$ presented in Section~\ref{sec:store_sem}.
+  captured by the $\scc$ contract definition:
+  \vspace{-0.8em}
+  \begin{smathpar}
+  \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
+  \end{smathpar}
 
@@
 
-\noindent Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
-totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$.
-This concurs with the intuition that any contract satisfiable under $\ecc$ or
-$\ccc$ is satisfiable under $\scc$, and any contract that is satisfiable under
-$\ecc$ is satisfiable under $\ccc$. Nonetheless, we determine the class of a
-contract based on the \emph{weakest} guarantee (among $\ecc$, $\ccc$, and
-$\scc$) required to satisfy the contract.
+Observe that out contract language does not incorporate real (i.e.,
+wall-clock) time. Hence, the contract language cannot describe store
+semantics based on real time such as recency or bounded-staleness
+guarantees offered by certain stores~\cite{Pileus}.
 
@@
 
-The classification scheme is presented formally in Figure~\ref{sem:classify}.
-Along with three straightforward rules that classify contracts into consistency
-classes, the classification scheme also presents a rule that judge
-well-formedness of a contract. A contract is well-formed if and only if it is
-satisfiable under $\scc$ - the strongest possible consistency guarantee any
-store can provide. The $\le$ relation used by the rules is defined formally
-below:
+\subsection{Contract Comparison and Classification}
 
+Our goal is to map application-level consistency constraints of
+operations to appropriate store-level consistency guarantees capable
+of satisfying constraints.  The ability to express both of them as
+contracts in our contract language lets us compare and determine if
+contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
+satisfied under a store consistency level identified by the contract
+$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
+relation for our contract language as following:
 \begin{definition}
@@
 \end{definition}
-\vspace{-1em}
 
-The $\Delta$ referred in the above defintion captures assumptions about the
-nature of primitive relations, such as $\Rvis$ and $\Rso$ are irreflexive, and
-the happens-before relation $\hbZ$ is acyclic, preventing thin-air reads.
-\KC{Ideally, we need to list all the axioms somewhere.}
+\vspace{-2em} \noindent The $\Delta$ referred in the above defintion captures
+assumptions about the nature of primitive relations, such as $\Rvis$
+and $\Rso$ are irreflexive, and the happens-before relation $\hbZ$ is
+acyclic, preventing thin-air reads.
+
+If the contract ($\cv_{op}$) of an operation ($op$) is \emph{weaker
+than} a store contract ($\cv_{st}$), then constraints expressed by the
+former are implied by guarantees provided by the later. Consequently,
+$\cv_{op}$ is satisfiable under $\cv_{st}$. Observe that the contracts
+$\scc$, $\ccc$ and $\ecc$ are themselves totally ordered with respect
+to the $\le$ relation: $\ecc \le \ccc \le \scc$.  This concurs with
+the intuition that any contract satisfiable under $\ecc$ or $\ccc$ is
+satisfiable under $\scc$, and any contract that is satisfiable under
+$\ecc$ is satisfiable under $\ccc$. However, we are interested in the
+\emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$) required
+to satisfy the contract. We define the corresponding consistency level
+as the \emph{consistency class of the contract}. The classification
+scheme is presented formally in Figure~\ref{sem:classify}.
+
+Along with three straightforward rules that classify contracts into
+consistency classes, the classification scheme also presents a rule
+that judges well-formedness of a contract. A contract is well-formed
+if and only if it is satisfiable under $\scc$ - the strongest possible
+consistency guarantee any store can provide. Otherwise, it is
+considered ill-formed, and rejected statically.

### 2014-10-29 core-lang.tex
Edits
Edits to Secs 4 and 5.

@@
 sessions interacting with the store. The execution state ($\E$) of a
-data store is modeled as a tuple $\Exec$ where $\EffSoup$ is a set of
-effects during the execution, and $\visZ,\soZ \subseteq \EffSoup
-\times \EffSoup$ are visibility and session order relations,
-respectively, witnessed over generated effects. A session is modeled
-simply as a sequence of replicated data type operations ($op$), tagged
-with the consistency class ($\tau$) of their contracts (as determined
-by the contract classification scheme). We abbreviate contract classes
-($\tau$) as uppercase {\sf SC}, {\sf CC}, and {\sf EC} for brevity.
-Their representative contracts ($\cv(\tau)$) are, as usual, written
-$\scc$, $\ccc$, and $\ecc$, respectively.  Our contract language and
-classification scheme is independent of store semantics. Accordingly,
-we abstain from making any specific assumptions about semantics of the
-store. However, we do assume the existence of a reduction relation of
-form:
+data store is modeled as a tuple $\Exec$ where $\EffSoup$, called the
+\emph{effect soup}, is a set of effects generated during the
+execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
+visibility and session order relations, respectively, witnessed over
+generated effects. A session is modeled simply as a sequence of
+replicated data type operations ($op$), tagged with the consistency
+class ($\tau$) of their contracts (as determined by the contract
+classification scheme). Our contract language and classification
+scheme is independent of store semantics. Accordingly, we abstain from
+making any specific assumptions about semantics of the store. However,
+we do assume the existence of a reduction relation of form:
+
 \begin{smathpar}
-  \auxred{} {(\E,\Sigma)} {\eff} {(\E',\Sigma')}
+  \auxred{} {\E;\langle (op,\tau);\sigma \rangle \pll \Sigma} {\eff} 
+    {\E';\langle \sigma \rangle \pll \Sigma'}
 \end{smathpar}
+
 \noindent The relation captures store's progress in execution (from
-$\E$ to $\E'$) due to the application of an operation from one of the
-sessions in $\Sigma$ to a replica, generating a new effect ($\eff$).
+$\E$ to $\E'$) due to the application of an operation ($op$) from one
+of the sessions in $\Sigma$ to a replica, generating a new effect
+($\eff$).
 
@@
 states $\E$, $\E'$ such that
-$\auxred{} {(\E,\langle (op,\tau);\,\sigma \rangle \pll \Sigma)} {\eff} 
- {(\E',\langle \sigma \rangle \pll \Sigma)}$
+$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
+ {\E';\langle \sigma \rangle \pll \Sigma}$
 if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
@@
 semantics of our implementation of the data store, which concretizes
-the reduction relation ($\xhookrightarrow{}$), along with the proof
+the reduction relation ($\xrightarrow{}$), along with the proof
 that it correctly enforces all \name contracts.

### 2014-10-29 contract-language.tex
Spacing edits to motivation, contract-language. More text in transactions.

@@
   captured in terms of the following contract definition:
-  \vspace{-0.8em}
+
+  \vspace{-1em}
   \begin{smathpar}
@@
   defined below:
-  \vspace{-0.8em}
+
+  \vspace{-1em}
   \begin{smathpar}
@@
   captured by the $\scc$ contract definition:
-  \vspace{-0.8em}
+
+  \vspace{-1em}
   \begin{smathpar}
@@
 relation for our contract language as following:
+
 \begin{definition}
@@
 
-\vspace{-2em} \noindent The $\Delta$ referred in the above defintion captures
+\vspace{-1em}
+\noindent The $\Delta$ referred in the above defintion captures
 assumptions about the nature of primitive relations, such as $\Rvis$

### 2014-10-29 transactions.tex
Spacing edits to motivation, contract-language. More text in transactions.

@@
 transactions can be satisfied with much weaker isolation guarantees. Despite
-the atomicity offered by the transaction, anomalies are still possible.
+the atomicity offered by the transaction, anomalies are still possible. For
+example, the two \cf{getBalance} operations in \cf{totalBalance} transactions
+might be served by different replicas with distinct set of committed \cf{save}
+transactions. If the first(second) \cf{getBalance} operation witness a
+\cf{save} transaction that is not witnessed by the second(first)
+\cf{getBalance} operation, then the balance returned will be less(greater) than
+the actual balance. It is not immediately apparent which weakest isolation
+guaratee will be sufficient to prevent the anomaly.
+
+Instead, \name requires the programmer to simply state the consistency
+requirement as a contract. Since we would like both the \cf{getBalance}
+operations to witness the same set of \cf{save} transactions, we define
+$\psi_2$ as:
+
+\begin{smathpar}
+\begin{array}{l}
+\cv_{2} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, (c:\rcf{withdraw} \vee \rcf{deposit}), d. \\
+\qquad \txn{a,b}{c,d} ~\wedge~ \vis{d}{b} ~\wedge~ \sameobj{a}{c} \Rightarrow \vis{c}{a}
+\end{array}
+\end{smathpar}

### 2014-10-29 transactions.tex
Added transaction section

@@
 
-\subsection{Syntax Extension}
+\subsection{Syntax and Semantics Extension}
+\label{sec:syn_sem_ext}
 
@@
 \sametxn{c}{d} ~\wedge~ \neg\sametxn{a}{c}$, where $a$ and $b$ considered to
-belong to the \emph{current} transaction. We assume that operations not part of
-any transaction to belong to their own unique transaction. While the
-transactions may have varying isolation guarantees, we make the standard
-assumption that all transactions provide atomicity. Hence, we include the
-following axiom in $\Delta$: $\small \forall a,b,c.~\txn{a}{b,c} ~\wedge~
-\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
-\vis{c}{a}$.
+belong to the \emph{current} transaction.
+
+We assume that operations not part of any transaction to belong to their own
+unique transaction. While the transactions may have varying isolation
+guarantees, we make the standard assumption that all transactions provide
+atomicity. Hence, we include the following axiom in $\Delta$: $\small \forall
+a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
+\vis{c}{a}$. The semantics of this contract is illustrated represented in
+Figure~\ref{fig:txn_atomicity}.
 
@@
 transactions is given below:
+\vspace{-1em}
 
@@
 the actual balance. It is not immediately apparent which weakest isolation
-guaratee will be sufficient to prevent the anomaly.
+guarantee will be sufficient to prevent the anomaly.
 
@@
 requirement as a contract. Since we would like both the \cf{getBalance}
-operations to witness the same set of \cf{save} transactions, we define
-$\psi_2$ as:
+operations to witness the same set of \cf{save} transactions, we define the
+constraint on \cf{totalBalance} transaction $\psi_2$ as:
 
+\vspace{-1em}
 \begin{smathpar}
 \begin{array}{l}
-\cv_{2} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, (c:\rcf{withdraw} \vee \rcf{deposit}), d. \\
-\qquad \txn{a,b}{c,d} ~\wedge~ \vis{d}{b} ~\wedge~ \sameobj{a}{c} \Rightarrow \vis{c}{a}
+\cv_{2} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, \\
+\qquad (c:\rcf{withdraw} \vee \rcf{deposit}), (c:\rcf{withdraw} \vee \rcf{deposit}). \\
+\qquad \txn{a,b}{c,d} ~\wedge~ \vis{c}{a} ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
 \end{array}
 \end{smathpar}
+
+\noindent The key idea in the above definition is that the $\txnZ$ primitive
+allows us to relate operations on different objects.
+
+The \cf{save} transaction only needs to ensure that the two writes it performs
+are made visible atomically. Since this is ensured by combining them in a
+transaction, \cf{save} does not require any additional constraints, and
+$\psi_1$ is simply $\small \true$.
+
+\subsection{Coordination-free Transactions}
+
+\begin{figure}
+\centering
+\subfigure[Atomicity]{\label{fig:txn_atomicity}\includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}}
+\hfill
+\subfigure[Monotonic Atomic View]{\label{fig:txn_mav}\includegraphics[width=0.25\columnwidth]{Figures/TxnMAV}}
+\hfill
+\subfigure[Repeatable Read]{\label{fig:txn_rr}\includegraphics[width=0.25\columnwidth]{Figures/TxnRR}}
+\caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
+The dotted line represents the visibility requested by the contracts.}
+\label{fig:cleanliness_examples}
+\end{figure}
+
+In order to illustrate the utility of transaction contract classification, we
+identify three well-understood coordination-free transaction semantics -- Read
+Committed (RC)~\cite{}, Monotonic Atomic View (MAV)~\cite{} and Repeatable Read
+(RR)~\cite{}, and illustrate the classification strategy. Our technique can
+indeed be applied to a different isolation level lattice.
+
+A transaction with ANSI RC semantics only witnesses committed operations. Let
+us assume that the store buffers updates until all the updates from the
+transaction are available at a replica. If the transaction commits, the
+buffered updates are made visible. Otherwise, the buffered updates are
+discarded. In particular, a store implementing RC does not require
+inter-replica coordination. RC does not entail any further isolation
+guarantees. We can express RC as follows:
+
+\vspace{-1em}
+\begin{smathpar}
+\rcc = \forall a,b,c.~\txn{a}{b,c} \wedge \sameobj{b}{c} \wedge \vis{b}{a} \Rightarrow \vis{c}{a}
+\end{smathpar}
+
+\noindent Notice that the above definition is the same as the atomicity
+guarantee of transaction described in Section~\ref{sec:syn_sem_ext}. The
+\cf{save} is an example for RC transaction.
+
+MAV semantics ensures that if some operation in a transaction $T_1$ witnesses
+the effects of another transaction $T_2$, then subsequent operations in $T_1$
+will also witness the effects of $T_2$. MAV semantics is useful for
+maintaining the integrity of foreign key constraints, materialized views and
+secondary updates. In order to implement MAV, a store only needs to keep track
+of the set of transactions $S_t$ witnessed by the running transaction, and
+before performing an operation at some replica, ensure that the replica
+includes all the transactions in $S_t$. Hence, MAV is coordination-free. MAV
+semantics is captured with the following contract:
+
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\mavc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \so{a}{b} ~\wedge~ \vis{c}{a} \\
+\qquad\qquad\qquad\qquad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
+\end{array}
+\end{smathpar}
+
+\noindent whose semantics is illustrated in the Figure~\ref{fig:txn_mav}.
+
+RR semantics requires that the transaction witness a snapshot of the data store
+state. Importantly, this snapshot can be obtained from any replica, and hence
+RR is coordination-free. An example for such a transaction is the
+\cf{totalBalance} transaction. The semantics of RR is captured by the following
+contract:
+
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\rrc = \forall a,b,c,d.~\txn{a,b}{c,d} ~\wedge~ \vis{c}{a} \\
+\qquad\qquad\qquad\quad ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
+\end{array}
+\end{smathpar}
+
+\noindent whose semantics is illustrated in the Figure~\ref{fig:txn_rr}.
+
+\subsection{Classification}
+
+Similar to operation-level contracts, with respect to $\le$ relation, the
+coordination-free transaction semantics described here form a total order:
+$\rcc \le \mavc \le \rrc$. The transaction classification is also similar to
+the operation-level contract classification presented in
+Figure~\ref{sem:classify}; given a contract $\psi$ on a transaction, we start
+from the weakest transaction contract $\rcc$, and progressively compare its
+strength to the known transaction contracts until we find a isolation level
+under which $\psi$ can be safely discharged. Otherwise, we report a type error.

### 2014-10-29 contract-language.tex
Minor edit/comment
edits

@@
 \item \textbf{Causal consistency}: Causally consistent operations
-  require to see a causally consistent snapshot of the object state,
+  are required to see a causally consistent snapshot of the object state,
   including the actions performed on the same session.  The latter
@@
 
-Observe that out contract language does not incorporate real (i.e.,
-wall-clock) time. Hence, the contract language cannot describe store
-semantics based on real time such as recency or bounded-staleness
-guarantees offered by certain stores~\cite{Pileus}.
+% SJ: not sure this is necessary, or appropriate here.
+%% \noindent Observe that out contract language does not incorporate real (i.e.,
+%% wall-clock) time. Hence, it cannot describe store
+%% semantics such as recency or bounded-staleness
+%% guarantees offered by certain stores~\cite{Pileus}.
 
@@
 
-Our goal is to map application-level consistency constraints of
+Our goal is to map application-level consistency constraints on
 operations to appropriate store-level consistency guarantees capable
-of satisfying constraints.  The ability to express both of them as
+of satisfying these constraints.  The ability to express both of them as
 contracts in our contract language lets us compare and determine if

### 2014-10-29 core-lang.tex
edits

@@
-\section{Meta-theory}
+\section{Soundness}
 \label{sec:core-opsem}
 
-In this section, we present the meta-theoretic result that certifies
-the soundness of our classification-based contract enforcement
+In this section, we present a soundness result that certifies
+the correctness of our classification-based contract enforcement
 strategy.
@@
 
-Fig.~\ref{sem:oper} summarizes the artifacts needed to state our
-meta-theoretic result. We formalize an abstract model of a replicated
-data store as a tuple $\E;\Sigma$, where $\E$ denotes the execution
-state, and $\Sigma$, the session soup, is the set of concurrent client
-sessions interacting with the store. The execution state ($\E$) of a
-data store is modeled as a tuple $\Exec$ where $\EffSoup$, called the
-\emph{effect soup}, is a set of effects generated during the
+Fig.~\ref{sem:oper} summarizes the artifacts needed. We formalize an
+abstract model of a replicated data store as a tuple $\E;\Sigma$, where $\E$
+denotes the execution state, and $\Sigma$, the session soup, is the set of
+concurrent client sessions interacting with the store. The execution state
+($\E$) of a data store is modeled as a tuple $\Exec$ where $\EffSoup$,
+called the \emph{effect soup}, is a set of effects generated during the
 execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
 visibility and session order relations, respectively, witnessed over
-generated effects. A session is modeled simply as a sequence of
-replicated data type operations ($op$), tagged with the consistency
-class ($\tau$) of their contracts (as determined by the contract
-classification scheme). Our contract language and classification
-scheme is independent of store semantics. Accordingly, we abstain from
-making any specific assumptions about semantics of the store. However,
-we do assume the existence of a reduction relation of form:
+generated effects. A session is modeled simply as a sequence of replicated
+data type operations ($op$), tagged with the consistency class ($\tau$) of
+their contracts (as determined by the contract classification scheme). Our
+contract language and classification scheme is independent of store
+semantics. Accordingly, we avoid making any specific assumptions about the
+semantics of the store. However, we do assume the existence of a reduction
+relation of the form:
 
@@
 
-\noindent The relation captures store's progress in execution (from
+\noindent The relation captures the store's execution progress (from
 $\E$ to $\E'$) due to the application of an operation ($op$) from one
@@
 
-We now state the soundness of contract enforcement as following:
-
 \begin{theorem}[Soundness of Contract Enforcement]
@@
 classification scheme can be used to satisfy any \name contract on
-that data store. The proof of the theorem has been included in our
-tech report~\cite{techrep}. Tech report also contains operational
-semantics of our implementation of the data store, which concretizes
+that data store. The proof of the theorem is available in the supplementary
+material, which also contains an operational
+semantics of our implementation of the data store that concretizes
 the reduction relation ($\xrightarrow{}$), along with the proof

### 2014-10-29 contract-language.tex
Integrated formalization with Sec 4.

--- 2014-10-29 contract-language.tex
@@
 
-% \subsection{Capturing Store Semantics}
-% \label{sec:store_sem}
+\subsection{Semantics}
 
-% An important aspect of our contract classification system is that the store
-% semantics is also captured using the same contract language used to describe
-% application-level consistency. In this regard, similar
-% to~\cite{Burckhardt2014}, we can rigorously define a wide variety of store
-% semantics including those that combine any subset of session and causality
-% guarantees, and multiple consistency levels. For example, a store that offers
-% strong consistency is captured by the contract:
+\begin{figure}
+\begin{smathpar}
+\stretcharraybig
+\begin{array}{lclcl}
+\multicolumn{5}{l}{
+  {\eff} \in \mathtt{Effect} \qquad
+  {\cv} \in \mathtt{Contract} \qquad
+  \set{\eff} \in \mathtt{Effect\; Set}
+}\\
+\EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
+\visZ, \soZ, &	\in & \mathtt{Relations} & \coloneqq & 
+  \set{\eff}\times\set{\eff} \\
+\sameobjZ		&     &  & \\
+{\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
+\end{array}
+\end{smathpar}
 
-% \vspace{-1em}
-% \begin{smathpar}
-% \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
-% \end{smathpar}
+\caption{An Axiomatic Execution}
+\label{sem:contracts}
+\end{figure}
 
-% \noindent Similarly, a store that offers per-object causal consistency is captured by the
-% contract:
+In this section, we formally define what it means for an execution to
+satisfy a contract.
 
-% \vspace{-1em}
-% \begin{smathpar}
-% \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
-% \end{smathpar}
+\name contracts are constraints over axiomatic definitions of program
+executions.Fig.~\ref{sem:contracts} summarizes artifacts relevant to
+define an axiomatic execution. We formalize an axiomatic execution as
+a tuple $\Exec$, where $\EffSoup$, called the \emph{effect soup}, is
+the set of all effects generated during the program execution, and
+$\visZ,\soZ, \sameobjZ \subseteq \EffSoup \times \EffSoup$ are
+\emph{visibility}, \emph{session order}, and \emph{same object}
+relations, respectively, witnessed over generated effects at run-time.
 
-% \noindent This ability to represent store semantics and application-level consistency in
-% the same language is vital to contract classification.
-
-% \subsection{Contract Comparison}
-
-% Our goal is to classify contracts, and map the operation they describe to
-% the \emph{weakest} store-level consistency level that  nonetheless
-% satisfies the contract's constraints. To this end, we need a mechanism to
-% compare the ``strength'' of a contract. Let $\cv_{op}$ be a contract for a
-% particular operation $op$, and $\cv_{st}$ capture a particular store
-% consistency level. We would like to determine whether $op$ can be
-% \emph{safely discharged} at the store consistency level $\cv_{st}$ such that
-% the resulting execution does not violate $\cv_{op}$.
-
-% Since our contracts represent axiomatic definition of program executions, let
-% $\Mod{\cv}$ be the set of all executions under which contract $\cv$ is
-% satisfied. If every execution $\E \in \Mod{\cv_{st}}$ is also a member of
-% $\Mod{\cv_{op}}$, then $op$ can be safely discharged under the store
-% consistency level $\cv_{st}$. Formally, $\Mod{\cv_{st}} \subseteq
-% \Mod{\cv_{op}}$. This is the model-theoretic consequence relation, written as
-% $\cv_{st} \models_m \cv_{op}$.
-
-% Observe that our contract language (stripped of its syntactic sugar) is a
-% carefully chosen subset of first-order logic that is known to be
-% decidable~\cite{epr}.  Since first-order logic is sound and complete
-% ~\cite{completeness}, $\cv_{st} \models_m \cv_{op}$ if and only if $\cv_{st}
-% \Rightarrow \cv_{op}$.  Due to the decidability of our contract language, this
-% implication check is automatically discharged with the help of a theorem
-% prover.
+Note that the axiomatic definition an execution ($\E$) provides
+interpretations for primitive relations (eg: $\visZ$) that occur free
+in contract formulas, and also fixes the domain of quantification to
+set of all effects ($\EffSoup$) observed during the program execution.
+As such, $\E$ is a potential model for any first-order formula ($\cv$)
+expressible in our contract language. If $\E$ is indeed a valid model
+for $\cv$ (expressed using model-theoretic consequence relation as $\E
+\models \cv$), we say that the execution $\E$ satisfied the contract
+$\cv$:
+\begin{definition}
+An axiomatic execution $\E$ is said to satisfy a contract $\cv$ if and
+only if $E \models \cv$.
+\end{definition}
 
@@
 \vspace{-1em}
-\noindent The $\Delta$ referred in the above defintion captures
-assumptions about the nature of primitive relations, such as $\Rvis$
-and $\Rso$ are irreflexive, and the happens-before relation $\hbZ$ is
-acyclic, preventing thin-air reads.
+\noindent The $\Delta$ referred in the above defintion is a
+conjunction of assumptions about the nature of primitive relations,
+such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
+relation $\hbZ$ is acyclic. A \emph{well-formed} axiomatic execution
+($\E$) is expected to satisfy these assumptions (i.e., $\E \models
+\Delta$).
 
@@
 than} a store contract ($\cv_{st}$), then constraints expressed by the
-former are implied by guarantees provided by the later. Consequently,
-$\cv_{op}$ is satisfiable under $\cv_{st}$. Observe that the contracts
-$\scc$, $\ccc$ and $\ecc$ are themselves totally ordered with respect
-to the $\le$ relation: $\ecc \le \ccc \le \scc$.  This concurs with
-the intuition that any contract satisfiable under $\ecc$ or $\ccc$ is
-satisfiable under $\scc$, and any contract that is satisfiable under
-$\ecc$ is satisfiable under $\ccc$. However, we are interested in the
-\emph{weakest} guarantee (among $\ecc$, $\ccc$, and $\scc$) required
-to satisfy the contract. We define the corresponding consistency level
-as the \emph{consistency class of the contract}. The classification
-scheme is presented formally in Figure~\ref{sem:classify}.
+former are implied by guarantees provided by the later. Due to the
+completeness of first-order logic\cite{completeness}, this means that
+any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
+$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
+\cv_{op}$). Consequently, it is safe to execute operation $op$ under
+store consistency level captured by $\cv_{st}$.
+
+Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
+totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le
+\scc$.  This concurs with the intuition that any contract satisfiable
+under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any contract
+that is satisfiable under $\ecc$ is satisfiable under $\ccc$. However,
+we are interested in the \emph{weakest} guarantee (among $\ecc$,
+$\ccc$, and $\scc$) required to satisfy the contract. We define the
+corresponding consistency level as the \emph{consistency class of the
+contract}. The classification scheme is presented formally in
+Figure~\ref{sem:classify}.
 
@@
 considered ill-formed, and rejected statically.
+
+\subsection{Soundness of Contract Classification}
+
+In this section, we present a meta-theoretic result that certifies the
+soundness of classification-based contract enforcement. To help us
+state the result, we present an axiomatic model of the our system
+described informally in Sec.~\ref{sec:sysmod}:
+
+\begin{smathpar}
+\begin{array}{lclcl}
+\multicolumn{5}{l}{
+  {op} \in \mathtt{Operation} \qquad
+  \tau \in \mathtt{Consistency\; Class}
+}\\
+\cv(\tau) & \in & \mathtt{Store\; Contract\; of \; \tau} & \coloneqq & \scc,
+  \ccc, \ecc\\
+{\sigma} & \in & \mathtt{Session} & \coloneqq & \cdot \ALT (op,\tau); \sigma \\
+\Sigma 	& \in & \mathtt{Session\;Soup}   & \coloneqq & 
+  \langle s,{\sigma} \rangle \pll \Sigma \ALT \emptyset \\
+				&			&	\mathtt{Store}		  & \coloneqq & \E;\Sigma \\
+\end{array}
+\end{smathpar}
+
+We model the system as a tuple $\E;\Sigma$, where the execution $\E$
+captures the data store's current state of the execution, and
+$\Sigma$, the session soup, is the set of concurrent client sessions
+interacting with the store. A session is modeled simply as a sequence
+of replicated data type operations ($op$), tagged with the consistency
+class ($\tau$) of their contracts (as determined by the contract
+classification scheme). We assume rewrite relation of form:
+
+\vspace{-1.7em}
+\begin{smathpar}
+  \auxred{} {\E;\langle (op,\tau);\sigma \rangle \pll \Sigma} {\eff} 
+    {\E';\langle \sigma \rangle \pll \Sigma'}
+\end{smathpar}
+
+\noindent on the system state. The relation captures store's progress
+in execution (from $\E$ to $\E'$) due to the application of an
+operation ($op$) from one of the sessions in $\Sigma$ to a replica,
+generating a new effect ($\eff$). If the resultant execution ($\E'$)
+satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
+\models \cv(\tau)$), the we say that the store has \emph{enforced} the
+contract $\cv(\tau)$ in the execution $\E'$.
+
+With help of the axiomatic model, we now state the soundness of
+contract enforcement as following:
+
+\begin{theorem}[Soundness of Contract Enforcement]
+\label{lem:core-preservation}
+Let $\cv$ be a well-formed contract of a replicated data type operation $op$,
+and let $\tau$ denote the consistency class of $\cv$ as determined by
+the contract classification scheme. Forall well-formed execution
+states $\E$, $\E'$ such that
+$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
+ {\E';\langle \sigma \rangle \pll \Sigma}$
+if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
+\end{theorem}
+
+The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the
+same store, extended with classification scheme, can enforce all
+well-formed \name contracts. The proof of the theorem has been
+relegated to our tech report~\cite{techrep} owing to space
+constraints \footnote{Tech report also contains operational semantics of our
+implementation of the data store, which concretizes the rewrite
+relation ($\xrightarrow{}$), along with the proof that it correctly
+enforces all \name contracts.}.

### 2014-10-29 implementation.tex
Added RDT example in motivation

\section{Implementation}

In this section, we describe the instantiation of \name on top of off-the-shelf
eventually consistent distributed store Cassandra. We begin with a description
of the programming interface.

The programmer simply pairs up the operation implementation with the
corresponding contract to complete the RDT implementation. The contracts are
statically checked, and mapped to the corresponding store consistency level.
The store is exposed as a collection key-indexed, object-valued tables, where
each table holds objects of the same type. The transaction language operates
over this view, and has primitives to invoke an operation on an object, and
invoke a transaction on a sequence of operation invocations. The transaction
contract is also statically checked and classified.

### 2014-10-29 model.tex
Added RDT example in motivation

--- 2014-10-28 motivation.tex
@@
 $\operZ$, mitigates these challenges.
-
-\section{Motivation}
-\label{sec:motivation}
-
-Consider how we might implement a highly available bank account on top of an
-eventually consistent data store, with the \emph{integrity} constraint that the
-balance should never drop below 0. Let us assume that the bank account object
-provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
-with the assumption that the withdraw fails if the account has insufficient
-balance. Our goal is to choose the correct consistency level for each of these
-operations such that (1) the balance never drops below zero and (2) the
-\cf{getBalance} operation never incorrectly returns a negative balance. To be
-able to do this, we need to consider the anomalies that could arise under
-eventual consistency.
-
-\begin{figure}
-\centering
-\subfigure[Unsafe withdraw]{\label{fig:unsafeWithdrawAnomaly}\includegraphics[width=0.34\columnwidth]{Figures/Motivation4}}
-\hfill
-\subfigure[Negative balance]{\label{fig:negativeBalanceAnomaly}\includegraphics[width=0.31\columnwidth]{Figures/Motivation2}}
-\hfill
-\subfigure[Missing update]{\label{fig:missingUpdateAnomaly}\includegraphics[width=0.26\columnwidth]{Figures/Motivation1}}
-\caption{Anomalies possible under eventual consistency for the get balance operation.}
-\label{fig:cleanliness_examples}
-\end{figure}
-
-Consider the execution shown in Figure~\ref{fig:unsafeWithdrawAnomaly}. Assume
-that all operations in the figure are on the same bank account object with the
-initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
-by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
-witnesses the deposit and succeeds\footnote{Although visibility and session
-order relations relate effects, we have abused the notation in these examples
-to relate operations, with the idea that the relations relate the effect
-emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
-operation, but importantly, due to eventual consistency, only witnesses the
-\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
-\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
-constraint. A subsequent \cf{getBalance} operation, that happens to witness all
-the previous operations,  would report a negative balance.
-
-It is easy to see that the only way to ensure the integrity constraint on
-\cf{balance} is to prevent concurrent withdrawals on the same bank account
-object. This can be done by insisting that \cf{withdraw} be executed with
-strong consistency properties. Although this prevents the integrity constraint
-from being violated, a strongly consistent \cf{withdraw} itself is not
-sufficient to prevent \cf{getBalance} from ever reporting a negative balance to
-the end user.
-
-Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}, which
-consists of three concurrent sessions performing a \cf{deposit}, a
-\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
-account object. As before, assume that the initial balance in the account is
-zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2,
-can witnesses the effects of \cf{deposit(100)} from session 1, leading it to
-rightfully conclude that the current balance is 50. Consequently,
-\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
-operation subsequently becomes visible to the \cf{getBalance} in Session 3.
-However, because \cf{getBalance} only witnesses the effect of the \cf{withdraw}
-operation from session 2, but not the \cf{deposit} from session 1, reports the
-balance of negative 50 to the user.
-
-Producing a negative balance is not the only inconsistency that the user is
-exposed to under an eventually consistent bank account. Figure
-~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
-operation in a session does not witness the effects of an earlier \cf{withdraw}
-operation performed in the same session, possibly because it was served by a
-replica that has not yet merged the \cf{withdraw} effect, served by a different
-replica. This anomaly leads the user to incorrectly conclude that the
-\cf{withdraw} operation failed to go through.
-
-Although it is easy to understand the reasons behind the occurrence of the
-aforementioned negative balance and missing update anomalies, finding the
-appropriate fixes is not readily apparent. Making \cf{getBalance} a strongly
-consistent operation is definitely sufficient to avert anomalies, but is it
-really necessary? Given the cost of enforcing strong
-consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid it unless
-there are no viable alternatives.  Exploring the space of these alternatives
-requires understanding the subtle differences in semantics of various kinds
-of weak consistency alternatives.
-
-\subsection{Specifications and Contracts}
-
-\name helps facilitate the mapping of operations to appropriate consistency
-levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} that axiomatically specify the set of allowed
-executions involving this operation.  In the case of the bank account example,
-any execution that does not exhibit the anomalies described in the previous
-section is a \emph{well-formed} execution on the bank account object.  By
-specifying the set of legal executions for each data type in terms of a trace
-of operation invocations on that type, \name\ ensures that all executions over
-that type are well-formed.
-
-In our running example, it is clear that in order to preserve the integrity
-constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
-given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. Utilizing the primitive relations described in
-Section~\ref{sec:sysmod}, we express this application-level consistency
-requirement as a contract or specification over \cf{withdraw}:
-\vspace{-.5em}
-\begin{smathpar}
-\begin{array}{l}
-\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
-\end{array}
-\end{smathpar}
-\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
-The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
-by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
-contract specifies that if the current operation emits an effect $\cureff$,
-then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
-is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
-Any execution on a bank account object that preserves the above contract for a
-\cf{withdraw} operation is said to be derived from a correct implementation of
-\cf{withdraw}.
-
-\noindent For \cf{getBalance}, we construct the following contract:
-\begin{smathpar}
-\begin{array}{l}
-\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
-\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
-\end{array}
-\end{smathpar}
-\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
-effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
-If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
-operations $a$ visible to the \cf{withdraw} should also be visible to
-\cf{getBalance}. This prevents negative balance anomalies.
-
-Our specification language provides operators to compose
-relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
-\wedge R_2(a,b)$. The last line of the above contract says that if a
-\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation in
-session order, and is applied on the same object as the \cf{getBalance}
-operation, then it must be the case that the \cf{getBalance} operation
-witnesses the effects of the preceding operations.
-
-Since there are no restrictions on when and how a \cf{deposit} operation can
-execute, its contract is simply $\true$.
-
-\subsection{From Contracts to Implementation}
-
-Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
-application-level consistency requirements, and make no reference to the
-semantics of the underlying store. To write contracts, a programmer only needs
-to reason about the semantics of the application under the \name system model.
-The mapping of application-level consistency requirements to appropriate
-store-level guarantees is done automatically behind-the-scene.
-
-How might one go about ensuring that any execution adheres to the contract? The
-challenge is that a contract provides a declarative (axiomatic) specification
-of an execution, while what is required is is an operational procedure for
-\emph{enforcing} its implicit constraints. One strategy would be to execute
-operations speculatively.  Here, operations are tentatively applied as they are
-received from the client or other replicas. We can maintain a runtime
-manifestation of executions, and check well-formedness conditions at runtime,
-rolling back executions if they are ill-formed. However, the overhead of state
-maintenance and the complexity of user-defined contracts make this technique
-infeasible in practice.
-
-We devise a static approach instead. Contracts are analyzed with the help of a
-theorem prover, and statically mapped to a particular store-level consistency
-property that the prover guarantees preserves contract semantics. We call this
-procedure \emph{contract classification}. Given the variety and complexity of
-store level consistency properties, the idea is that the system implementor
-parameterizes the classification procedure by describing the store semantics in
-the \emph{same} contract language as the one used to express the contract on
-the operations. In the next section, we describe the contract language in
-detail and describe the classification procedure for a particular store
-semantics.

### 2014-10-29 motivation.tex
Spacing edits to motivation, contract-language. More text in transactions.
Added transaction section
Added RDT example in motivation

@@
-\section{System Model}
-\label{sec:sysmod}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
-\caption{Quelea system model.}
-\label{fig:sysmod}
-\end{figure}
-
-Figure~\ref{fig:sysmod} provides a schematic diagram of our system model. The
-distributed store is composed of a collection of \emph{replicas}, each of which
-stores a set of \emph{objects} ($x,y,\ldots$). We assume that every object is
-replicated at every replica in the store. The state of an object at any replica
-is the set of all updates (\emph{effects}) performed on the object. For
-example, the state of $x$ at replica 1 is the set composed of effects $w^x_1$
-and $w^x_2$.
-
-Each object is associated with a set of \emph{operations}. The clients interact
-with the store by invoking operations on objects. The sequence of operations
-invoked by a particular client on the store is called a \emph{session}. The
-data store is typically accessed by a large number of clients (and hence
-sessions) concurrently. Importantly, the clients are oblivious to which replica
-an operation is applied to; the data store may choose to route the operation to
-the closest replica for minimizing latency, or route it to a different replica
-in order to balance the load in the system. For example, the operations
-\emph{foo} and \emph{bar} invoked by the same session on the same object, might
-end up being applied to different replicas because replica 1 (to which
-\emph{foo} was applied) might be unreachable when the client invokes
-\emph{bar}.
-
-When \emph{foo} is invoked on a object $x$ with arguments \emph{arg}$_1$ at
-replica 1, it simply \emph{reduces} over the current set of effects at that
-replica on that object ($w^x_1$ and $w^x_2$), produces a result $v1$ that is
-sent back to the client, and emits a \emph{single} new effect $w^x_4$ that is
-appended to the state of $x$ at replica 1. Thus, every operation is evaluated
-over a \emph{snapshot} of the state of the object on which it is invoked. In
-this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$ and $w^x_2$.
-Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$, written
-logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where $\visZ$ is
-the visibility relation between effects. Visibility is an irreflexive (an
-operation cannot see its own effect) and asymmetric relation, and only relates
-effects produced by operations on the same object. Executing a read-only
-operation is similar except that no new effects are produced.
-
-The effect added to a particular replica is asynchronously sent to other
-replicas. We assume that \emph{eventually} an effect applied at one replica
-will be merged into all other replicas. Two effects $w^x_4$ and $w^x_5$ that
-arise from the same session are said to be in \emph{session order} (written
-logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive, transitive
-relation. The effects $w^x_4$ and $w^x_5$ arising from operations applied to
-the same object $x$ are said to be under the \emph{same object} relation,
-written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every effect with
-the operation that generated the effect with the help of a relation $\operZ$.
-In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$ hold. For
-simplicity, we assume all operation names across all object types are distinct.
-
-This model admits all the inconsistencies associated with eventual consistency.
-The goal of this work is to identify the precise consistency level for each
-operation such that application-level constraints are not violated. In the next
-section, we will concretely describe the challenges associated with
-constructing a consistent bank account on top of an eventually consistent data
-store. Subsequently, we will illustrate how our contract and specification
-language, armed with the primitive relations $\visZ$, $\soZ$, $\sameobjZ$ and
-$\operZ$, mitigates these challenges.
-
 \section{Motivation}
@@
 eventually consistent data store, with the \emph{integrity} constraint that the
-balance should never drop below 0. Let us assume that the bank account object
-provides three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance},
-with the assumption that the withdraw fails if the account has insufficient
-balance. Our goal is to choose the correct consistency level for each of these
-operations such that (1) the balance never drops below zero and (2) the
+balance should never drop below 0. We will begin by implementing the bank
+account replicated data type (RDT) in \name, and then describe the mechanisms
+to obtain the desired correctness guarantees.
+
+\subsection{RDT specification}
+
+A key novelty	in \name is that it allows the addition of new RDTs to the store,
+immensely improving the programmability of data stores. The programmer no
+longer has to coerce the application logic to utilize the store provided data
+types. In addition, \name treats the convergence semantics of the data type
+separately from the consistency properties. The separation of concern allows
+the programmer to \emph{operationally} reason about conflict resolution, while
+permitting \emph{declarative} reasoning for consistency. As mentioned in the
+previous section, RDTs are described in terms of operations on a RDT object,
+where each operation is described as a reduction over the state of the object
+i.e., the list of effects. Let us assume that the bank account object provides
+three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the
+assumption that the withdraw fails if the account has insufficient balance.
+Every operation in \name is of the following type, written in Haskell syntax:
+
+\begin{codehaskell}
+type Operation = (Effect e, Arg a, Res r)
+               => [e] -> a -> (r, Maybe e)
+\end{codehaskell}
+
+\noindent It takes a list of effects (the \emph{context} for the operation), an
+argument and returns a result along with an optional effect. Read-only
+operations do not return any effect. If an effect is returned, it is added to
+the state of the object at the current replica, and asynchronously sent to
+other replicas. The implementation of the bank account operations is given
+below:
+
+\begin{codehaskell}
+data Acc = Deposit_ Int | Withdraw_ Int | GetBalance_
+
+getBalance :: [Acc] -> () -> (Int, Maybe Acc)
+getBalance ctxt _ =
+  let res = sum [x | Deposit_ x <- ctxt]
+						- sum [x | Withdraw_ x <- ctxt]
+	in (res, Nothing)
+
+deposit :: [Acc] -> Int -> ((), Maybe Acc)
+deposit _ amt = (amt, Just $ Deposit_ amt)
+
+withdraw :: [Acc] -> Int -> (Bool, Maybe Acc)
+withdraw ctxt v =
+	if sel1 $ getBalance ctxt () >= v
+  then (True, Just $ Withdraw_ v)
+	else (False, Nothing)
+\end{codehaskell}
+
+The datatype \cf{Acc} represents the effect type for the bank account. The
+context of the operations is a snapshot of the state of the object at some
+replica. In this sense, every operation on the RDT is atomic, and thus
+permitting sequential reasoning for implementing eventually consistent data
+types. We have implemented a large corpus of RDTs for realistic benchmarks
+including shopping carts, auction and micro-blogging sites in few tens of lines
+of code.
+
+\subsection{Anomalies under Eventual Consistency}
+
+Our goal is to choose the correct consistency level for each of the bank
+account operations such that (1) the balance never drops below zero and (2) the
 \cf{getBalance} operation never incorrectly returns a negative balance. To be
@@
 requirement as a contract or specification over \cf{withdraw}:
-\vspace{-.5em}
+
+\vspace{-1em}
 \begin{smathpar}
@@
 \end{smathpar}
+
 \noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
@@
 \noindent For \cf{getBalance}, we construct the following contract:
+
+\vspace{-1em}
 \begin{smathpar}
@@
 \end{smathpar}
+
 \noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
@@
 Since there are no restrictions on when and how a \cf{deposit} operation can
-execute, its contract is simply $\true$.
+execute, its contract is simply $\small \true$.
 

### 2014-10-29 contract-language.tex
edits

@@
 the set of all effects generated during the program execution, and
-$\visZ,\soZ, \sameobjZ \subseteq \EffSoup \times \EffSoup$ are
+$\visZ,\soZ,\sameobjZ \subseteq \EffSoup \times \EffSoup$ are
 \emph{visibility}, \emph{session order}, and \emph{same object}

### 2014-10-29 implementation.tex
edits

@@
 In this section, we describe the instantiation of \name on top of off-the-shelf
-eventually consistent distributed store Cassandra. We begin with a description
-of the programming interface.
-
-The programmer simply pairs up the operation implementation with the
-corresponding contract to complete the RDT implementation. The contracts are
-statically checked, and mapped to the corresponding store consistency level.
-The store is exposed as a collection key-indexed, object-valued tables, where
-each table holds objects of the same type. The transaction language operates
-over this view, and has primitives to invoke an operation on an object, and
-invoke a transaction on a sequence of operation invocations. The transaction
-contract is also statically checked and classified.
+eventually consistent distributed store Cassandra. In particular, we focus on
+the challenges in efficiently realizing \name programming model.

### 2014-10-29 transactions.tex
Added RDT example in motivation
Minor edit

@@
 def save(amt):
-  atomically psi1:
+  atomically psi1 :
     b = c.withdraw(amt)
@@
 def totalBalance():
-  atomically psi2:
+  atomically psi2 :
     b1 = c.getBalance()
@@
 The dotted line represents the visibility requested by the contracts.}
-\label{fig:cleanliness_examples}
+\label{fig:transaction}
 \end{figure}

### 2014-10-29 motivation.tex
Minor edit

@@
 \begin{codehaskell}
-data Acc = Deposit_ Int | Withdraw_ Int | GetBalance_
+data Acc = Deposit Int | Withdraw Int | GetBalance
 
@@
 getBalance ctxt _ =
-  let res = sum [x | Deposit_ x <- ctxt]
-						- sum [x | Withdraw_ x <- ctxt]
+  let res = sum [x | Deposit x <- ctxt]
+						- sum [x | Withdraw x <- ctxt]
 	in (res, Nothing)
@@
 deposit :: [Acc] -> Int -> ((), Maybe Acc)
-deposit _ amt = (amt, Just $ Deposit_ amt)
+deposit _ amt = (amt, Just $ Deposit amt)
 
@@
 	if sel1 $ getBalance ctxt () >= v
-  then (True, Just $ Withdraw_ v)
+  then (True, Just $ Withdraw v)
 	else (False, Nothing)

### 2014-10-29 motivation.tex
example

@@
 
-A key novelty	in \name is that it allows the addition of new RDTs to the store,
-immensely improving the programmability of data stores. The programmer no
-longer has to coerce the application logic to utilize the store provided data
+A key novelty of \name is that it allows the addition of new RDTs to the
+store, greatly improving programmability.  Specifically, programmers no
+longer have to coerce application logic to conform to to store-provided data
 types. In addition, \name treats the convergence semantics of the data type
-separately from the consistency properties. The separation of concern allows
-the programmer to \emph{operationally} reason about conflict resolution, while
-permitting \emph{declarative} reasoning for consistency. As mentioned in the
-previous section, RDTs are described in terms of operations on a RDT object,
-where each operation is described as a reduction over the state of the object
-i.e., the list of effects. Let us assume that the bank account object provides
-three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the
-assumption that the withdraw fails if the account has insufficient balance.
-Every operation in \name is of the following type, written in Haskell syntax:
+separately from consistency properties. This separation of concerns allows
+the programmer to \emph{operationally} reason about conflict resolution,
+while permitting \emph{declarative} reasoning for consistency. As mentioned
+in the previous section, RDTs are described in terms of operations on a RDT
+object, where each operation is described as a reduction over the state of
+the object i.e., the list of effects.
+
+Assume that the bank account object provides three operations: \cf{deposit},
+\cf{withdraw} and \cf{getBalance}, with the assumption that the \cf{withdraw}
+fails if the account has insufficient balance.  Every operation in \name is
+of the following type, written in Haskell syntax:
 
@@
 
-\noindent It takes a list of effects (the \emph{context} for the operation), an
+\noindent An operation takes a list of effects (the \emph{context} for the operation), an
 argument and returns a result along with an optional effect. Read-only

### 2014-10-29 motivation.tex
Minor edits. Before introducing Haskell contracts.
Motivation uses Haskell contracts

--- 2014-10-29 motivation.tex
@@
 the state of the object at the current replica, and asynchronously sent to
-other replicas. The implementation of the bank account operations is given
-below:
+other replicas. The implementation of the bank account operations in \name is
+given below:
 
 \begin{codehaskell}
-data Acc = Deposit_ Int | Withdraw_ Int | GetBalance_
+data Acc = Deposit_ Int | Withdraw_ Int
+				 | GetBalance_
 
@@
 	else (False, Nothing)
+
+mkOperName ''Acc
 \end{codehaskell}
@@
 of code.
+
+The last line in the above code snippet uses Template Haskell~\cite{} to
+generate the following datatype declaration:
+
+\begin{codehaskell}
+data OperName = Deposit | Withdraw | GetBalance
+\end{codehaskell}
+
+\noindent that declares a nullary constructor for each of the constructors in
+the given type, dropping the tailing underscore from the original constructor
+names. This declaration will come in handy for writing the contracts.
 
@@
 
+\begin{figure}
+\begin{codehaskell}
+data Effect, Rel, Prop, Fol
+type Contract = Effect -> Fol
+
+Vis,So,SameObj :: Rel
+(cup),(cap)      :: Rel -> Rel -> Rel
+transClosure :: Rel -> Rel
+((-)          :: (Effect, Effect) -> Rel -> Prop
+true         :: Prop
+(==),vis,so,sameObj :: Effect -> Effect -> Prop
+(wedge),(vee),(=>) :: Prop -> Prop -> Prop
+lift         :: Prop -> Fol
+(forall) :: [OperName] -> (Effect -> Fol) -> Fol
+\end{codehaskell}
+\caption{Quelea contract language.}
+\label{code:ctrt_lang}
+\end{figure}
+
+\name contract language is presented in Figure~\ref{code:ctrt_lang}. The
+language includes primitive relations \cf{Vis},\cf{So} and \cf{SameObj},
+operators to compose relations ($\cup,\cap$ and \cf{transClosure}), and check
+inclusion ($\in$). The language also includes standard propositional logic
+operators, and a prenex quantified $\forall$ operator. The functions \cf{vis},
+\cf{so} and \cf{sameObj} are derived entities. For example, \cf{vis a b = (a,b)
+$\in$ Vis}.
+
 In our running example, it is clear that in order to preserve the integrity
@@
 given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. Utilizing the primitive relations described in
-Section~\ref{sec:sysmod}, we express this application-level consistency
-requirement as a contract or specification over \cf{withdraw}:
+vice versa. We express this application-level consistency as the following
+contract:
 
-\vspace{-1em}
-\begin{smathpar}
-\begin{array}{l}
-\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
-\end{array}
-\end{smathpar}
+\begin{codehaskell}
+withdrawCtrt x = (forall) [Withdraw] $ \a -> lift $
+	sameObj a x => a == e vee vis a x vee vis x a
+\end{codehaskell}
 
-\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
-The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
-by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
-contract specifies that if the current operation emits an effect $\cureff$,
-then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
-is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
-Any execution on a bank account object that preserves the above contract for a
-\cf{withdraw} operation is said to be derived from a correct implementation of
-\cf{withdraw}.
+\noindent Here, \cf{x} stands for the effect emitted by the current
+(\cf{withdraw}) operation. The syntax \cf{\small $(\forall)
+$[Withdraw]($\backslash$a $\rightarrow \ldots$)} states that $a$ is an effect
+which was emitted by a \cf{withdraw} operation i.e., $\small
+\oper{a}{\rcf{withdraw}}$ holds. The contract specifies that if the withdraw
+operation emits an effect \cf{x}, then for any operation \cf{a} which was
+emitted by a \cf{withdraw} operation, it is the case that \cf{a} and \cf{x} are
+the same effect or \cf{a} is visible to \cf{x}, or vice versa. Any execution on
+a bank account object that preserves the above contract for a \cf{withdraw}
+operation is said to be derived from a correct implementation of \cf{withdraw}.
 
@@
 
-\vspace{-1em}
-\begin{smathpar}
-\begin{array}{l}
-\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
-\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
-\end{array}
-\end{smathpar}
+\begin{codehaskell}
+getBalanceCtrt x = (forall) [Deposit, Withdraw] $ \a ->
+	(forall) [Deposit] $ \b -> (forall) [Withdraw] $ \c ->
+		lift $ ((a,x) (- (So cap SameObj) => vis a x) wedge
+  				 (vis b c wedge vis c x => vis b x)
+\end{codehaskell}
 
-\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
-effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
-If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
-operations $a$ visible to the \cf{withdraw} should also be visible to
+\noindent Here, \cf{a} is an effect emitted by either a \cf{deposit} or a
+\cf{withdraw} operation. The fourth line in the above snippet says that if a
+\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation
+in session order, and is applied on the same object as the \cf{getBalance}
+operation, then it must be the case that the \cf{getBalance} operation
+witnesses the effects of that preceding operations. The last line says that if
+a \cf{withdraw} $c$ is visible to \cf{getBalance}, then all \cf{deposit}
+operations $b$ visible to the \cf{withdraw} should also be visible to
 \cf{getBalance}. This prevents negative balance anomalies.
 
-Our specification language provides operators to compose
-relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
-\wedge R_2(a,b)$. The last line of the above contract says that if a
-\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation in
-session order, and is applied on the same object as the \cf{getBalance}
-operation, then it must be the case that the \cf{getBalance} operation
-witnesses the effects of the preceding operations.
+There are no restrictions on when and how a \cf{deposit} operation can execute.
+Hence, the deposit contract imposes no consistency restrictions.
 
-Since there are no restrictions on when and how a \cf{deposit} operation can
-execute, its contract is simply $\small \true$.
+\begin{codehaskell}
+depositCtrt x = lift true
+\end{codehaskell}
 

### 2014-10-29 motivation.tex

@@
 getBalance ctxt _ =
-  let res = sum [x | Deposit_ x <- ctxt]
-						- sum [x | Withdraw_ x <- ctxt]
+  let res = sum [x | Deposit x <- ctxt]
+          - sum [x | Withdraw x <- ctxt]
 	in (res, Nothing)

### 2014-10-30 motivation.tex
edits
Edits

@@
 \begin{figure}
+%Careful while adjsting spaces in the following code; screws "::" alignment.
 \begin{codehaskell}
@@
 There are no restrictions on when and how a \cf{deposit} operation can execute.
-Hence, the deposit contract imposes no consistency restrictions.
+Hence, the deposit contract imposes no consistency restrictions. The deposit
+contract is:
 
@@
 depositCtrt x = lift true
+\end{codehaskell}
+
+\subsection{Putting it together}
+
+The combines the RDT implementation along with the contracts to instantiate a
+new RDT as follows:
+
+\begin{codehaskell}
+mkRDT [(Deposit, deposit, $(chk depositCtrt)),
+  (Withdraw, withdraw, $(chk withdrawCtrt)),
+	(GetBalance, getBalance, $(chk getBalanceCtrt))]
+\end{codehaskell}
+
+\noindent The function \cf{chk} analyzes the contract and maps it to one of the
+available store consistency levels. From the perspective of the client of \name
+store, the store is organized as a collection of key-indexed, object-valued
+tables, where each object in a table has the same type. Clients invoke
+operations using the following primitive:
+
+\begin{codehaskell}
+data Quelea a
+instance Monad Quelea
+invoke :: (Arg a, Res r)
+			 => Key -> OperName -> a -> Quelea r
 \end{codehaskell}

### 2014-10-30 transactions.tex
Edits throughout

@@
 defined previously. We consider two transactions -- \cf{save(amt)}, which
-transfers \cf{amt} from current to savings, and \cf{totalBalance} returns
-the sum of the balances of individual accounts. The pseudo code for the
-transactions is given below:
+transfers \cf{amt} from current to savings, and \cf{totalBalance} returns the
+sum of the balances of individual accounts. The \name code for the transactions
+is given below:
 \vspace{-1em}
@@
 \noindent \begin{minipage}[t]{0.5\columnwidth}
-\begin{codepython}
-def save(amt):
-  atomically psi1 :
-    b = c.withdraw(amt)
-    if (b): s.deposit(amt)
-\end{codepython}
+\begin{codehaskell}
+save :: Quelea ()
+save amt =
+  x <- $(chkTxn psi1)
+  atomically x $ do
+    b <- withdraw c amt
+    when b $ deposit s amt
+\end{codehaskell}
 \end{minipage}
 \begin{minipage}[t]{0.5\columnwidth}
-\begin{codepython}
-def totalBalance():
-  atomically psi2 :
-    b1 = c.getBalance()
-    b2 = s.getBalance()
+\begin{codehaskell}
+totalBalance :: Quelea Int
+totalBalance =
+  x <- $(chkTxn psi2)
+  atomically x $ do
+    b1 <- getBalance s
+    b2 <- getBalance s
     return b1 + b2
-\end{codepython}
+\end{codehaskell}
 \end{minipage}
 
-\noindent where $\cv_1$ and $\cv_2$ are the contracts on the corresponding
+\noindent In the above code example, the operations on bank account are simply
+wrappers around the store interface. For example,
+
+\begin{codehaskell}
+withdraw :: Key -> Int -> Quelea Bool
+withdraw k amt = invoke k Withdraw amt
+\end{codehaskell}
+
+\noindent $\cv_1$ and $\cv_2$ are the contracts on the corresponding
 transactions. Our goal is to ensure that \cf{totalBalance} returns the result

### 2014-10-30 intro.tex
Formatting operational semantics
edits

@@
 \setlength{\itemsep}{2pt}
-\item We introduce \name, a language for declaratively describing eventually
-	consistent programs that manipulate replicated data types. Contracts are used
-	to specify fine-grained application-level consistency properties, and are
-	analyzed to assign the most efficient and sound store consistency level to
-	the corresponding operation.
+\item We introduce \name, a Haskell library that supports the description
+  and validation of replicated data types found in programs intended to
+  execute on eventually consistent stores. Contracts are used to specify
+  fine-grained application-level consistency properties, and are analyzed to
+  assign the most efficient and sound store consistency level to the
+  corresponding operation.
 \item \name supports coordination-free transactions over arbitrary datatypes.
@@
 	assign the correct and most efficient isolation level for a transaction.
-\item We provide metatheory that certifies the soundness of our contract enforcement
-	system, and ensures that an operation is only executed if the required
-	conditions on consistency are met.
-\item \name is implemented as a transparent shim layer over
+\item We provide metatheory that certifies the soundness of our contract
+	enforcement system, and ensures that an operation is only executed if the
+	required conditions on consistency are met.
+\item An implementation of \name as a transparent shim layer over
 	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
 	extensive experimental evaluation over a collection of real-world
-  applications, including a Twitter-like micro-blogging site and an
-  eBay-like auction site illustrates the practicality of our approach.
+	applications, including a Twitter-like micro-blogging site and an eBay-like
+	auction site illustrates the practicality of our approach.
 \end{itemize}

### 2014-10-30 motivation.tex
Edits
edits

@@
 the state of the object at the current replica, and asynchronously sent to
-other replicas. The implementation of the bank account operations in \name is
-given below:
+other replicas. The implementation of these operations is
+given in Figure~\ref{fig:ex}.
 
+\begin{figure}[htb]
 \begin{codehaskell}
@@
 \end{codehaskell}
+\caption{Definition of a replicated datatype in Quelea written in Haskell syntax.}
+\label{fig:ex}
+\end{figure}
 
@@
 replica. In this sense, every operation on the RDT is atomic, and thus
-permitting sequential reasoning for implementing eventually consistent data
+permits sequential reasoning for implementing eventually consistent data
 types. We have implemented a large corpus of RDTs for realistic benchmarks
-including shopping carts, auction and micro-blogging sites in few tens of lines
-of code.
+including shopping carts, auction and micro-blogging sites, all in few tens
+of lines of code.
 
-The last line in the above code snippet uses Template Haskell~\cite{} to
+The last line in the code snippet uses Template Haskell~\cite{} to
 generate the following datatype declaration:
@@
 \name helps facilitate the mapping of operations to appropriate consistency
-levels by letting the programmer declare application-level consistency
+levels by letting the programmer additionally declare application-level consistency
 constraints as \emph{contracts} that axiomatically specify the set of allowed
@@
 
-\name contract language is presented in Figure~\ref{code:ctrt_lang}. The
+\name The contract language is presented in Figure~\ref{code:ctrt_lang}. The
 language includes primitive relations \cf{Vis},\cf{So} and \cf{SameObj},
@@
 
-In our running example, it is clear that in order to preserve the integrity
-constraint, the \cf{withdraw} operation must be strongly consistent. That is,
-given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. We express this application-level consistency as the following
-contract:
+In our running example, it is clear that in order to preserve the critical
+integrity constraint, the \cf{withdraw} operation must be strongly
+consistent. That is, given two \cf{withdraw} operations $a$ and $b$, either
+$a$ is visible to $b$ or vice versa. We express this application-level
+consistency as the following contract:
 
@@
 \noindent For \cf{getBalance}, we construct the following contract:
-
 \begin{codehaskell}
@@
 \end{codehaskell}
-
 \noindent Here, \cf{a} is an effect emitted by either a \cf{deposit} or a
@@
 instance Monad Quelea
-invoke :: (Arg a, Res r)
-			 => Key -> OperName -> a -> Quelea r
+invoke :: (Table t, Key k, Arg a, Res r)
+			 => t -> k -> OperName -> a -> Quelea r
 \end{codehaskell}

### 2014-10-30 intro.tex
intro

@@
 \setlength{\itemsep}{2pt}
-\item We introduce \name, a Haskell library that supports the description
+\item We introduce \name, a shallow extension of Haskell that supports the description
   and validation of replicated data types found in programs intended to

### 2014-10-30 transactions.tex
Edits
Transactions. Minor edit

@@
 \begin{codehaskell}
-save :: Quelea ()
 save amt =
-  x <- $(chkTxn psi1)
+  x <- $(classify psi1)
   atomically x $ do
@@
 \begin{codehaskell}
-totalBalance :: Quelea Int
 totalBalance =
-  x <- $(chkTxn psi2)
+  x <- $(classify psi2)
   atomically x $ do
@@
 \end{minipage}
-
-\noindent In the above code example, the operations on bank account are simply
-wrappers around the store interface. For example,
-
-\begin{codehaskell}
-withdraw :: Key -> Int -> Quelea Bool
-withdraw k amt = invoke k Withdraw amt
-\end{codehaskell}
 

### 2014-10-30 implementation.tex

@@
 In this section, we describe the instantiation of \name on top of off-the-shelf
-eventually consistent distributed store Cassandra. In particular, we focus on
-the challenges in efficiently realizing \name programming model.
+eventually consistent distributed store Cassandra. We particularly focus on the
+challenges in realizing an efficient implementation of our model.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\columnwidth]{Figures/ImplModel}
+\end{center}
+\caption{Implementation Model.}
+\label{fig:impl_mod}
+\end{figure}
+
+Figure~\ref{fig:impl_mod} describes the overall system architecture. The main
+idea in our implementation of \name is to offload the majority of concerns
+related to data management such as replication, fault tolerance, availability
+and convergence to Cassandra, an off-the-shelf eventually consistent store
+(called \emph{backing store} in the sequel). Replicated data type
+implementation and the stronger consistency semantics are enforced in the
+\emph{shim layer}. Our implementation supports eventual, causal and strong
+consistency for operations, and RC, MAV and RR transaction semantics. All of
+this functionality is implemented on top of the standard interface exposed by
+Cassandra, and illustrates the portability of \name programming model. Complex
+business logic is implemented with the help of transactions on the server side,
+with clients interacting with the web-service over REST interface.
+
+\subsection{Shim Layer}
+
+The shim layer maintains a causally consistent in-memory snapshot of a subset
+of objects in the backing store, by explicitly tracking dependencies introduced
+between the effects due to visibility, session and same transaction relations.
+The dependence tracking is similar to the techniques presented in~\cite{BoltOn}
+and~\cite{Eiger}, with the usual optimizations making use of transitivity
+properties for minimizing the number of dependencies. Shim layer performs the
+reductions associated with replicated datatype operations corresponding to
+client requests. As the backing store provides durability, convergence and
+fault tolerance, each shim layer node simply acts as a soft-state cache, and
+can safely be terminated at any instant. Similarly, more shim layer nodes can
+be spawned on demand.
+
+Every effect generated as a result of an effectful operation on an object
+inserts a new row $(o,e,vis,txn,val)$ into the backing store, where $o$ and $e$
+are object and (unique) effect identifiers, $vis$ is the set of identifiers of
+effects visible to this operation, $txn$ is an optional transaction identifier,
+and $val$ is the value associated with the effect. The shim layer periodically
+fetches updates from the backing store for those objects which were accessed in
+the previous time slice. Since causally consistent operations require an
+up-to-date view of the current session, the shim layer node synchronously
+fetches operations if the causally preceding operations in the current session
+are not available in the cache. Strongly consistent operations are performed
+after obtaining exclusive leases on objects. The lease mechanism is implemented
+with the help of Cassandra's support for conditional updates and expiring
+columns.

### 2014-10-30 motivation.tex
Added reference to Fig. 3 in Sec 3.

--- 2014-10-29 motivation.tex
@@
 
-A key novelty of \name is that it allows the addition of new RDTs to the
-store, greatly improving programmability.  Specifically, programmers no
-longer have to coerce application logic to conform to to store-provided data
+A key novelty	in \name is that it allows the addition of new RDTs to the store,
+immensely improving the programmability of data stores. The programmer no
+longer has to coerce the application logic to utilize the store provided data
 types. In addition, \name treats the convergence semantics of the data type
-separately from consistency properties. This separation of concerns allows
-the programmer to \emph{operationally} reason about conflict resolution,
-while permitting \emph{declarative} reasoning for consistency. As mentioned
-in the previous section, RDTs are described in terms of operations on a RDT
-object, where each operation is described as a reduction over the state of
-the object i.e., the list of effects.
-
-Assume that the bank account object provides three operations: \cf{deposit},
-\cf{withdraw} and \cf{getBalance}, with the assumption that the \cf{withdraw}
-fails if the account has insufficient balance.  Every operation in \name is
-of the following type, written in Haskell syntax:
+separately from the consistency properties. The separation of concern allows
+the programmer to \emph{operationally} reason about conflict resolution, while
+permitting \emph{declarative} reasoning for consistency. As mentioned in the
+previous section, RDTs are described in terms of operations on a RDT object,
+where each operation is described as a reduction over the state of the object
+i.e., the list of effects. Let us assume that the bank account object provides
+three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the
+assumption that the withdraw fails if the account has insufficient balance.
+Every operation in \name is of the following type, written in Haskell syntax:
 
@@
 
-\noindent An operation takes a list of effects (the \emph{context} for the operation), an
+\noindent It takes a list of effects (the \emph{context} for the operation), an
 argument and returns a result along with an optional effect. Read-only
@@
 \begin{codehaskell}
-data Acc = Deposit_ Int | Withdraw_ Int
-				 | GetBalance_
+data Acc = Deposit_ Int | Withdraw_ Int | GetBalance_
 
@@
 getBalance ctxt _ =
-  let res = sum [x | Deposit x <- ctxt]
-          - sum [x | Withdraw x <- ctxt]
+  let res = sum [x | Deposit_ x <- ctxt]
+						- sum [x | Withdraw_ x <- ctxt]
 	in (res, Nothing)
@@
 	else (False, Nothing)
-
-mkOperName ''Acc
 \end{codehaskell}
@@
 of code.
-
-The last line in the above code snippet uses Template Haskell~\cite{} to
-generate the following datatype declaration:
-
-\begin{codehaskell}
-data OperName = Deposit | Withdraw | GetBalance
-\end{codehaskell}
-
-\noindent that declares a nullary constructor for each of the constructors in
-the given type, dropping the tailing underscore from the original constructor
-names. This declaration will come in handy for writing the contracts.
 
@@
 
-\begin{figure}
-\begin{codehaskell}
-data Effect, Rel, Prop, Fol
-type Contract = Effect -> Fol
-
-Vis,So,SameObj :: Rel
-(cup),(cap)      :: Rel -> Rel -> Rel
-transClosure :: Rel -> Rel
-((-)          :: (Effect, Effect) -> Rel -> Prop
-true         :: Prop
-(==),vis,so,sameObj :: Effect -> Effect -> Prop
-(wedge),(vee),(=>) :: Prop -> Prop -> Prop
-lift         :: Prop -> Fol
-(forall) :: [OperName] -> (Effect -> Fol) -> Fol
-\end{codehaskell}
-\caption{Quelea contract language.}
-\label{code:ctrt_lang}
-\end{figure}
-
-\name contract language is presented in Figure~\ref{code:ctrt_lang}. The
-language includes primitive relations \cf{Vis},\cf{So} and \cf{SameObj},
-operators to compose relations ($\cup,\cap$ and \cf{transClosure}), and check
-inclusion ($\in$). The language also includes standard propositional logic
-operators, and a prenex quantified $\forall$ operator. The functions \cf{vis},
-\cf{so} and \cf{sameObj} are derived entities. For example, \cf{vis a b = (a,b)
-$\in$ Vis}.
-
 In our running example, it is clear that in order to preserve the integrity
@@
 given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. We express this application-level consistency as the following
-contract:
+vice versa. Utilizing the primitive relations described in
+Section~\ref{sec:sysmod}, we express this application-level consistency
+requirement as a contract or specification over \cf{withdraw}:
 
-\begin{codehaskell}
-withdrawCtrt x = (forall) [Withdraw] $ \a -> lift $
-	sameObj a x => a == e vee vis a x vee vis x a
-\end{codehaskell}
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
+\end{array}
+\end{smathpar}
 
-\noindent Here, \cf{x} stands for the effect emitted by the current
-(\cf{withdraw}) operation. The syntax \cf{\small $(\forall)
-$[Withdraw]($\backslash$a $\rightarrow \ldots$)} states that $a$ is an effect
-which was emitted by a \cf{withdraw} operation i.e., $\small
-\oper{a}{\rcf{withdraw}}$ holds. The contract specifies that if the withdraw
-operation emits an effect \cf{x}, then for any operation \cf{a} which was
-emitted by a \cf{withdraw} operation, it is the case that \cf{a} and \cf{x} are
-the same effect or \cf{a} is visible to \cf{x}, or vice versa. Any execution on
-a bank account object that preserves the above contract for a \cf{withdraw}
-operation is said to be derived from a correct implementation of \cf{withdraw}.
+\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
+The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
+by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
+contract specifies that if the current operation emits an effect $\cureff$,
+then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
+is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
+Any execution on a bank account object that preserves the above contract for a
+\cf{withdraw} operation is said to be derived from a correct implementation of
+\cf{withdraw}.
 
@@
 
-\begin{codehaskell}
-getBalanceCtrt x = (forall) [Deposit, Withdraw] $ \a ->
-	(forall) [Deposit] $ \b -> (forall) [Withdraw] $ \c ->
-		lift $ ((a,x) (- (So cap SameObj) => vis a x) wedge
-  				 (vis b c wedge vis c x => vis b x)
-\end{codehaskell}
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
+\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
+\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
+\end{array}
+\end{smathpar}
 
-\noindent Here, \cf{a} is an effect emitted by either a \cf{deposit} or a
-\cf{withdraw} operation. The fourth line in the above snippet says that if a
-\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance} operation
-in session order, and is applied on the same object as the \cf{getBalance}
-operation, then it must be the case that the \cf{getBalance} operation
-witnesses the effects of that preceding operations. The last line says that if
-a \cf{withdraw} $c$ is visible to \cf{getBalance}, then all \cf{deposit}
-operations $b$ visible to the \cf{withdraw} should also be visible to
+\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
+effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
+If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
+operations $a$ visible to the \cf{withdraw} should also be visible to
 \cf{getBalance}. This prevents negative balance anomalies.
 
-There are no restrictions on when and how a \cf{deposit} operation can execute.
-Hence, the deposit contract imposes no consistency restrictions.
+Our specification language (Fig~\ref{fig:contract-lang}) provides
+operators to compose relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$
+is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
+above contract says that if a \cf{deposit} or a \cf{withdraw}
+operation precedes a \cf{getBalance} operation in session order, and
+is applied on the same object as the \cf{getBalance} operation, then
+it must be the case that the \cf{getBalance} operation witnesses the
+effects of the preceding operations.
 
-\begin{codehaskell}
-depositCtrt x = lift true
-\end{codehaskell}
+Since there are no restrictions on when and how a \cf{deposit}
+operation can execute, its contract is simply $\small \true$.
 

### 2014-10-30 motivation.tex
The placement of ref to Fig. 3 changed

@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} that axiomatically specify the set of allowed
-executions involving this operation.  In the case of the bank account example,
-any execution that does not exhibit the anomalies described in the previous
-section is a \emph{well-formed} execution on the bank account object.  By
-specifying the set of legal executions for each data type in terms of a trace
-of operation invocations on that type, \name\ ensures that all executions over
-that type are well-formed.
+constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}) that axiomatically
+specify the set of allowed executions involving this operation. In the case of
+the bank account example, any execution that does not exhibit the anomalies
+described in the previous section is a \emph{well-formed} execution on the bank
+account object.  By specifying the set of legal executions for each data type in
+terms of a trace of operation invocations on that type, \name\ ensures that all
+executions over that type are well-formed.
 
@@
 
-Our specification language (Fig~\ref{fig:contract-lang}) provides
-operators to compose relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$
-is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
-above contract says that if a \cf{deposit} or a \cf{withdraw}
-operation precedes a \cf{getBalance} operation in session order, and
-is applied on the same object as the \cf{getBalance} operation, then
-it must be the case that the \cf{getBalance} operation witnesses the
-effects of the preceding operations.
+Our specification language  provides operators to compose relations.
+Thus, the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b)
+\wedge R_2(a,b)$. The last line of the above contract says that if a
+\cf{deposit} or a \cf{withdraw} operation precedes a \cf{getBalance}
+operation in session order, and is applied on the same object as the
+\cf{getBalance} operation, then it must be the case that the
+\cf{getBalance} operation witnesses the effects of the preceding
+operations.
 

### 2014-10-30 contract-language.tex
Edits
edits

@@
 
-In this section, we formalize the contract language of \name, and describe our
+In this section, we formalize \name's contract language, and describe our
 contract classification scheme, which analyzes a contract and maps it to the
@@
 \begin{smathpar}
-\stretcharraybig
 \begin{array}{lclcl}
@@
 
-In this section, we formally define what it means for an execution to
-satisfy a contract.
-
 \name contracts are constraints over axiomatic definitions of program
@@
 
-Note that the axiomatic definition an execution ($\E$) provides
+Note that the axiomatic definition of an execution ($\E$) provides
 interpretations for primitive relations (eg: $\visZ$) that occur free
@@
 expressible in our contract language. If $\E$ is indeed a valid model
-for $\cv$ (expressed using model-theoretic consequence relation as $\E
+for $\cv$ (expressed using the model-theoretic consequence relation as $\E
 \models \cv$), we say that the execution $\E$ satisfied the contract
@@
   operations in its session are allowed to proceed. While eventually
-  consistent data store typically offer \emph{basic} eventual consistency
+  consistent data stores typically offer \emph{basic} eventual consistency
   with all possible anomalies, we assume that our store provides stronger
@@
 
-\item \textbf{Causal consistency}: Causally consistent operations
-  are required to see a causally consistent snapshot of the object state,
+\item \textbf{Causal consistency}: Causally consistent operations are
+  required to see a causally consistent snapshot of the object state,
   including the actions performed on the same session.  The latter
-  requirement entails that if two operations $o_1$ and $o_2$ from the
-  same session are applied to two different replicas $r_1$ and $r_2$,
-  the second operation cannot be discharged until the effect of $o_1$ is
-  merged with $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance}
-  operation requires causal consistency, as it requires the operations
-  from the same session to be visible, which cannot be guaranteed under
-  eventual consistency. We assume that causality is only tracked through
-  operations on the same object; two operations in the same session but
-  on different objects are considered causally unrelated under this
-  definition. Stores typically avoid tracking causality across objects
-  to mitigate overheads when causality tracking is unnecessary. The
-  corresponding store semantics is captured by the contract $\ccc$
-  defined below:
+  requirement entails that if two operations $o_1$ and $o_2$ from the same
+  session are applied to two different replicas $r_1$ and $r_2$, the second
+  operation cannot be discharged until the effect of $o_1$ is merged with
+  $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance} operation requires
+  causal consistency, as it requires the operations from the same session to
+  be visible, which cannot be guaranteed under just eventual consistency. We
+  assume that causality is only tracked through operations on the same
+  object; two operations in the same session but on different objects are
+  considered causally unrelated under this definition. Stores typically
+  avoid tracking causality across objects to mitigate overheads when
+  causality tracking is unnecessary. The corresponding store semantics is
+  captured by the contract $\ccc$ defined below:
 
@@
 class ($\tau$) of their contracts (as determined by the contract
-classification scheme). We assume rewrite relation of form:
+classification scheme). We assume a rewrite relation of form:
 
@@
 
-\noindent on the system state. The relation captures store's progress
-in execution (from $\E$ to $\E'$) due to the application of an
+\noindent on the system state. The relation captures the progress of
+the execution (from $\E$ to $\E'$)  due to the application of an
 operation ($op$) from one of the sessions in $\Sigma$ to a replica,
@@
 satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
-\models \cv(\tau)$), the we say that the store has \emph{enforced} the
+\models \cv(\tau)$), then we say that the store has \emph{enforced} the
 contract $\cv(\tau)$ in the execution $\E'$.
@@
 With help of the axiomatic model, we now state the soundness of
-contract enforcement as following:
+contract enforcement as follows:
 
@@
 
-The theorem states that if a data store correctly enforces $\scc$,
-$\ccc$, and $\ecc$ contracts in all well-formed executions, then the
-same store, extended with classification scheme, can enforce all
-well-formed \name contracts. The proof of the theorem has been
-relegated to our tech report~\cite{techrep} owing to space
-constraints \footnote{Tech report also contains operational semantics of our
-implementation of the data store, which concretizes the rewrite
+The theorem states that if a data store correctly enforces $\scc$, $\ccc$,
+and $\ecc$ contracts in all well-formed executions, then the same store,
+extended with classification scheme, can enforce all well-formed \name
+contracts. The proof of the theorem has been relegated to our tech report
+and is provided as part of the supplementary
+material~\cite{techrep}\footnote{The supplementary material and tech report
+  also defines an operational semantics that concretizes the rewrite
   relation ($\xrightarrow{}$), along with the proof that it correctly

### 2014-10-30 motivation.tex
edits

--- 2014-10-30 motivation.tex
@@
 eventually consistent data store, with the \emph{integrity} constraint that the
-balance should never drop below 0. We will begin by implementing the bank
+balance should never drop below 0. We begin by implementing a bank
 account replicated data type (RDT) in \name, and then describe the mechanisms
@@
 A key novelty in \name is that it allows the addition of new RDTs to the store,
-immensely improving the programmability of data stores. The programmer no
+greatly improving the programmability of data stores. The programmer no
 longer has to coerce the application logic to utilize the store provided data
 types. In addition, \name treats the convergence semantics of the data type
-separately from the consistency properties. The separation of concern allows
+separately from the consistency properties. This separation of concerns allows
 the programmer to \emph{operationally} reason about conflict resolution, while
@@
 where each operation is described as a reduction over the state of the object
-i.e., the list of effects. Let us assume that the bank account object provides
+in terms of a list of effects. Let us assume that the bank account object provides
 three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the
@@
 
-\noindent It takes a list of effects (the \emph{context} for the operation), an
-argument and returns a result along with an optional effect. Read-only
+\noindent It takes a list of effects (the \emph{context} for the operation),
+and an input argument, and returns a result along with an optional effect. Read-only
 operations do not return any effect. If an effect is returned, it is added to
@@
 other replicas. The implementation of the bank account operations in \name is
-given below:
+given in Figure~\ref{fig:ex}:
 
+\begin{figure}
 \begin{codehaskell}
@@
 \end{codehaskell}
+\label{Definition of a bank account expressed in Quelea.}
+\label{fig:ex}
+\end{figure}
 
@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} that axiomatically specify the set of allowed
-executions involving this operation.  In the case of the bank account example,
-any execution that does not exhibit the anomalies described in the previous
-section is a \emph{well-formed} execution on the bank account object.  By
-specifying the set of legal executions for each data type in terms of a trace
-of operation invocations on that type, \name\ ensures that all executions over
-that type are well-formed.
+constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}) that axiomatically
+specify the set of allowed executions involving this operation. In the case of
+the bank account, any execution that does not exhibit the anomalies
+described in the previous section is a \emph{well-formed} execution on the bank
+account object.  By specifying the set of legal executions for each data type in
+terms of a trace of operation invocations on that type, \name\ ensures that all
+executions over that type are well-formed.
 
@@
 \noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
-The syntax $a:\rcf{withdraw}$ states that $a$ is an effect which was emitted
+The syntax $a:\rcf{withdraw}$ states that $a$ is an effect  emitted
 by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
@@
 
-Our specification language (Fig~\ref{fig:contract-lang}) provides
-operators to compose relations. Thus, the syntax $(R_1 \cap R_2)(a,b)$
-is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
-above contract says that if a \cf{deposit} or a \cf{withdraw}
-operation precedes a \cf{getBalance} operation in session order, and
-is applied on the same object as the \cf{getBalance} operation, then
-it must be the case that the \cf{getBalance} operation witnesses the
-effects of the preceding operations.
-
-Since there are no restrictions on when and how a \cf{deposit}
-operation can execute, its contract is simply $\small \true$.
+Our specification language provides operators to compose relations.  Thus,
+the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge
+R_2(a,b)$. The last line of the above contract says that if a \cf{deposit}
+or a \cf{withdraw} operation precedes a \cf{getBalance} operation in session
+order, and is applied on the same object as the \cf{getBalance} operation,
+then it must be the case that the \cf{getBalance} operation witnesses the
+effects of the preceding operations. Since there are no restrictions on when
+or how a \cf{deposit} operation can execute, its contract is simply $\small
+\true$.
 
@@
 
-How might one go about ensuring that any execution adheres to the contract? The
-challenge is that a contract provides a declarative (axiomatic) specification
-of an execution, while what is required is is an operational procedure for
-\emph{enforcing} its implicit constraints. One strategy would be to execute
-operations speculatively.  Here, operations are tentatively applied as they are
-received from the client or other replicas. We can maintain a runtime
-manifestation of executions, and check well-formedness conditions at runtime,
-rolling back executions if they are ill-formed. However, the overhead of state
-maintenance and the complexity of user-defined contracts make this technique
-infeasible in practice.
+How might one go about ensuring that an execution adheres to a contract? The
+challenge is that a contract provides a declarative (axiomatic)
+specification of an execution, while what is required is is an operational
+procedure for \emph{enforcing} its implicit constraints. One strategy would
+be to execute operations speculatively.  Here, operations are tentatively
+applied as they are received from the client or other replicas. We can
+maintain a runtime manifestation of executions, and check well-formedness
+conditions at runtime, rolling back executions if they are
+ill-formed. However, the overhead of state maintenance and the complexity of
+user-defined contracts is likely to make this technique infeasible in
+practice.
 

### 2014-10-30 transactions.tex
edits

@@
 While contracts on individual operations offer the programmer object-level
-declarative reasoning, real-world scenarios often involve operations that span
-multiple objects. In order to address this problem, several recent
+declarative reasoning, real-world scenarios often involve operations that
+span multiple objects. In order to address this problem, several recent
 systems~\cite{COPS,BurckhardtESOP,BailisHAT} have proposed a variety of
@@
 serializability~\cite{} and snapshot isolation~\cite{} require inter-replica
-coordination, these systems espouse \emph{coordination-free transactions} that
-remain available under network partitions, but only provide weaker isolation
-guarantees. Coordination-free transactions have intricate consistency semantics
-and widely varying runtime overheads. As with operation-level consistency, the
-onus is on the programmer to pick the correct transaction kind, the choice of
-which is complicated by the consistency requirement of the operations it
-composes.
+coordination, these systems espouse \emph{coordination-free transactions}
+that remain available under network partitions, but only provide weaker
+isolation guarantees. Coordination-free transactions have intricate
+consistency semantics and widely varying runtime overheads. As with
+operation-level consistency, the onus is on the programmer to pick the
+correct transaction kind; this selection is complicated by operation
+consistency requirements.
 
@@
 consistency specification. We extend the contract language with a new term
-under quantifier-free propositions ${\small \txnZ}~S_1~S_2$, where $S_1$ and
+under quantifier-free propositions - ${\small \txnZ}~S_1~S_2$, where $S_1$ and
 $S_2$ are sets of effects, and introduce a new primitive equivalence relation
@@
 
-We assume that operations not part of any transaction to belong to their own
-unique transaction. While the transactions may have varying isolation
+We assume that operations not part of any transaction belong to their own
+unique transaction. While transactions may have varying isolation
 guarantees, we make the standard assumption that all transactions provide
@@
 In order to illustrate the utility of declarative reasoning for transactions,
-let us extend our running bank account example with use two accounts (objects)
+let us extend our running bank account example to use two accounts (objects)
 -- current ($c$) and savings ($s$). Each account provides operations
@@
 defined previously. We consider two transactions -- \cf{save(amt)}, which
-transfers \cf{amt} from current to savings, and \cf{totalBalance} returns the
+transfers \cf{amt} from current to savings, and \cf{totalBalance}, which returns the
 sum of the balances of individual accounts. The \name code for the transactions

### 2014-10-30 motivation.tex
Minor edits

@@
 \begin{codehaskell}
-data Acc = Deposit_ Int | Withdraw_ Int | GetBalance_
+data Acc = Deposit Int | Withdraw Int
+				 | GetBalance
 

### 2014-10-30 transactions.tex
minor edits

@@
 In order to illustrate the utility of declarative reasoning for transactions,
-let us extend our running bank account example to use two accounts (objects)
--- current ($c$) and savings ($s$). Each account provides operations
+let us extend our running bank account example to use two accounts (objects) --
+current ($c$) and savings ($s$). Each account provides operations
 \cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
 defined previously. We consider two transactions -- \cf{save(amt)}, which
-transfers \cf{amt} from current to savings, and \cf{totalBalance}, which returns the
-sum of the balances of individual accounts. The \name code for the transactions
-is given below:
+transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
+returns the sum of the balances of individual accounts. The \name code for the
+transactions is given below:
 \vspace{-1em}
@@
 \noindent $\cv_1$ and $\cv_2$ are the contracts on the corresponding
-transactions. Our goal is to ensure that \cf{totalBalance} returns the result
-obtained from a consistent snapshot of the object states.
+transactions. The function \cf{classify} assigns the contracts
+\emph{statically} to one of the transaction isolation levels offered by the
+store; \cf{\$()} is meta-programming syntax for splicing the result into the
+program. The \cf{atomically} construct invokes the enclosing operations at the
+given isolation level $x$, ensuring that the effects of the operations are made
+visible atomically. Our goal is to ensure that \cf{totalBalance} returns the
+result obtained from a consistent snapshot of the object states.
 

### 2014-10-31 contract-language.tex
edits

@@
 contracts in our contract language lets us compare and determine if
-contract ($\cv_{op}$) of an operation ($op$) is weak enough to be
+contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
 satisfied under a store consistency level identified by the contract
@@
 
-If the contract ($\cv_{op}$) of an operation ($op$) is \emph{weaker
+If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
 than} a store contract ($\cv_{st}$), then constraints expressed by the
@@
 $\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
-\cv_{op}$). Consequently, it is safe to execute operation $op$ under
+\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
 store consistency level captured by $\cv_{st}$.
@@
 interacting with the store. A session is modeled simply as a sequence
-of replicated data type operations ($op$), tagged with the consistency
+of replicated data type operations ($\mathit{op}$), tagged with the consistency
 class ($\tau$) of their contracts (as determined by the contract
@@
 \noindent on the system state. The relation captures the progress of
-the execution (from $\E$ to $\E'$)  due to the application of an
-operation ($op$) from one of the sessions in $\Sigma$ to a replica,
+the execution (from $\E$ to $\E'$)  due to the successful completion
+of a client operation ($\mathit{op}$) from one of the sessions in $\Sigma$,
 generating a new effect ($\eff$). If the resultant execution ($\E'$)
@@
 \label{lem:core-preservation}
-Let $\cv$ be a well-formed contract of a replicated data type operation $op$,
+Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
 and let $\tau$ denote the consistency class of $\cv$ as determined by

### 2014-10-31 transactions.tex
edits

@@
 guarantees, we make the standard assumption that all transactions provide
-atomicity. Hence, we include the following axiom in $\Delta$: $\small \forall
+atomicity. Hence, we include the following axiom in $\Delta$: \[\small \forall
 a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
-\vis{c}{a}$. The semantics of this contract is illustrated represented in
+\vis{c}{a}\] \noindent The semantics of this contract is illustrated in
 Figure~\ref{fig:txn_atomicity}.
@@
 In order to illustrate the utility of declarative reasoning for transactions,
-let us extend our running bank account example to use two accounts (objects) --
+consider an extension of our running example to use two accounts (objects) --
 current ($c$) and savings ($s$). Each account provides operations

### 2014-10-31 implementation.tex
edits

@@
 
+\subsection{Operation Consistency}
+
 Every effect generated as a result of an effectful operation on an object
@@
 columns.
+
+\subsection{Transactions}
+
+While cassandra offers all-or-nothing failure semantics for multiple writes
+through batching, readers may witness the initial write while the batch is in
+progress. \name implements atomic visibility by exploiting shim layer causality
+guarantee -- an effect is included only if all the effects if depends on are
+also included.
+
+For every transaction, we instantiate a special transaction marker effect $m$,
+but importantly, do not insert into the backing store. $m$ is included as a
+dependence to every effect (let $e$ be one such effect) generated in the
+transaction. Since $m$ has not yet been written to the store, no operation will
+witness $e$ while the transaction in progress. Let $S$ be the set of effects
+generated in the transaction. At the end of the transaction, we insert $m$ into
+the backing store with $S$ as its dependence. Now any replica which includes
+$e$, must include $m$, and transitively must include every effect from the
+transaction. This ensures atomicity and satisfies RC requirement.
+
+The above scheme prevents a transaction from witnessing its own effects. This
+might conflict with the causality requirement on the operations. Hence,
+transactions piggy-back the previous effects from the same transaction for each
+request. MAV semantics is implemented by keeping track of the set of
+transaction markers $M$ witnessed by the transaction, and before performing an
+operation at some replica, ensuring that $M$ is a subset of the transaction
+markers included at that replica. If not, the missing effects are synchronously
+fetched. RR semantics is realized by capturing a optimized snapshot of the
+state of some replica; each operation from an RR transaction is applied to this
+snapshot state. Any generated effects are added to this snapshot.
+
+\subsection{Summarization}
+
+The main challenge in realizing an efficient implementation of operation-based
+replicated data types is that the state of the object i.e., the set of effects
+grows with every effectful operation on the object. If left unchecked, the
+operations slow down over time, until the shim layer memory or backing store
+disk runs out of memory. Luckily, the state of the operation-based replicated
+data type can often be summarized to an \emph{observably equivalent} smaller
+state. For example,
+
+\begin{itemize}
+\setlength{\itemsep}{2pt}
+\item A last-writer-wins register with multiple updates where $v$ is the value
+of the last write is observably equivalent to a register with a single write
+$v$.
+
+\item A bank account with a series of deposits and withdraws with current
+balance $b$ is equivalent to a bank account with a single deposit of $b$.
+
+\item A set with collection of add and remove operations is equivalent to a set
+with a series of add operations of live elements from the original set.
+\end{itemize}
+
+Since the semantics of summarization depends on the semantics of the data type,
+we expect the programmer to provide a summarization function for each RDT with
+the following type:
+
+\begin{codehaskell}
+summarize :: Effect e => [e] -> [e]
+\end{codehaskell}
+
+\noindent with the intention that the length of the result is smaller that the
+length of the argument. We utilize the \cf{summarize} function to summarize the
+object state both in the shim layer node and the backing store, typically when
+the number of effects on an object crosses a tunable threshold. Shim layer
+summarization is straight-forward; a summarization thread takes the local lock
+on the object, and replaces its state with the summarized state. The shim layer
+node remains unavailable for that particular object during summarization
+(usually a few milliseconds).
+
+Compared to the shim layer, summarization in the backing store is more
+complicated. The main challenge is that unlike the shim layer, summarization
+cannot run as an atomic operation. Summarization in the backing store involves
+deleting previously inserted rows and inserting new rows, where each row
+corresponds to an effect. It is essential that concurrent client operations are
+permitted, but are not allowed to witness the intermediate state of the
+summarization process.
+
+To this end, we adopt a novel summarization strategy that builds on the
+causality property of the store. Suppose we have a set $O$ of effects on a
+object, which when summarized yields a new set of effects $N$. We first
+instantiate a summarization marker $s$, and similar to transaction marker, we
+do not insert it into the store immediately. We insert the new effects in $N$
+with strong consistency, including $s$ as a dependence. Since $s$ is not yet in
+the store, effects from $N$ are not made visible to the clients. Then we insert
+$s$ with strong consistency, including $O$ as dependence, and finally delete
+the rows corresponding to $O$. When the shim layer node witnesses the
+summarization marker $s$ on the object, it removes the effects from $O$ from
+its cache. But due to strong consistency of insertion of $N$ and $s$, the new
+effects $N$ are also observed. Thus, the clients witness the observably
+equivalent new set of effects $N$, and not the old effects $O$. This ensures
+atomicity of summarization in the backing store.

### 2014-10-31 implementation.tex
edits

@@
 
-Figure~\ref{fig:impl_mod} describes the overall system architecture. The main
-idea in our implementation of \name is to offload the majority of concerns
-related to data management such as replication, fault tolerance, availability
-and convergence to Cassandra, an off-the-shelf eventually consistent store
-(called \emph{backing store} in the sequel). Replicated data type
-implementation and the stronger consistency semantics are enforced in the
-\emph{shim layer}. Our implementation supports eventual, causal and strong
-consistency for operations, and RC, MAV and RR transaction semantics. All of
-this functionality is implemented on top of the standard interface exposed by
-Cassandra, and illustrates the portability of \name programming model. Complex
-business logic is implemented with the help of transactions on the server side,
-with clients interacting with the web-service over REST interface.
+\name is implemented as a shallow extension of Haskell on the Glasgow Haskell
+Compiler (GHC). We utilize Template Haskell functionality to implemement static
+contract classification. Our proof obligations are discharged automatically
+with the help of Z3~\cite{} SMT solver. Figure~\ref{fig:impl_mod} describes the
+overall system architecture. The main idea in our implementation of \name is to
+offload the majority of concerns related to data management such as
+replication, fault tolerance, availability and convergence to Cassandra, an
+off-the-shelf eventually consistent store (called \emph{backing store} in the
+sequel).
+
+Replicated data type implementation and the stronger consistency semantics are
+enforced in the \emph{shim layer}. Our implementation supports eventual, causal
+and strong consistency for operations, and RC, MAV and RR transaction
+semantics. All of this functionality is implemented on top of the standard
+interface exposed by Cassandra, and illustrates the portability of \name
+programming model. Complex business logic is implemented with the help of
+transactions on the server side, with clients interacting with the web-service
+over REST interface. From an engineering perspective, resorting to an
+off-the-shelf backing store meant that our entire implementation comprised of
+only around 2500 lines of Haskell code, which is packaged as a Haskell library.
 

### 2014-10-31 implementation.tex
edits

@@
 \name is implemented as a shallow extension of Haskell on the Glasgow Haskell
-Compiler (GHC). We utilize Template Haskell functionality to implemement static
+Compiler (GHC). We utilize Template Haskell functionality to implement static
 contract classification. Our proof obligations are discharged automatically
@@
 
-While cassandra offers all-or-nothing failure semantics for multiple writes
+While Cassandra offers all-or-nothing failure semantics for multiple writes
 through batching, readers may witness the initial write while the batch is in

### 2014-10-31 evaluation.tex
edits

\section{Evaluation}

In this section, we evaluate \name programs and report the contract profile as
well as illustrating the performance benefits of fine-grained consistency
classification on operations and transactions. We also report on the impact of
the summarization. We implemented the following applications, which includes
RDTs as well as larger applications composed of severals RDTs:

\begin{itemize}
\item \textbf{LWW register}: A last-write-wins register.
\item \textbf{DynamoDB register}: A integer register that allows eventual and
strong puts and gets, conditional puts, increment and decrement operations.
\item \textbf{Bank account}: Our running example, with savings and current
accounts.
\item \textbf{Shopping list}: Collaborative shopping list which allows adding
and deleting items.
\item \textbf{Online store}: Models an online store with shopping cart and
dynamically changing item prices. Checkout process verifies that the custormer
only pays the accepted price.
\item \textbf{RuBis}: An ebay-like auction site~\cite{}.
\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
Twissandra~\cite{}.
\end{itemize}

### 2014-10-31 implementation.tex
Minor edits to implementation section

@@
 programming model. Complex business logic is implemented with the help of
-transactions on the server side, with clients interacting with the web-service
-over REST interface. From an engineering perspective, resorting to an
-off-the-shelf backing store meant that our entire implementation comprised of
-only around 2500 lines of Haskell code, which is packaged as a Haskell library.
+transactions on the web-server side, with clients interacting with the
+web-service over REST interface. From an engineering perspective, resorting to
+an off-the-shelf backing store meant that our entire implementation comprised of
+only around 2500 lines of Haskell code, which is packaged as a Haskell
+library\footnote{url to the hackage}.
 
@@
 effects visible to this operation, $txn$ is an optional transaction identifier,
-and $val$ is the value associated with the effect. The shim layer periodically
-fetches updates from the backing store for those objects which were accessed in
-the previous time slice. Since causally consistent operations require an
-up-to-date view of the current session, the shim layer node synchronously
-fetches operations if the causally preceding operations in the current session
-are not available in the cache. Strongly consistent operations are performed
-after obtaining exclusive leases on objects. The lease mechanism is implemented
-with the help of Cassandra's support for conditional updates and expiring
-columns.
+and $val$ is the value associated with the effect (eg: \cf{Withdraw 50}). The
+shim layer periodically fetches updates from the backing store for those objects
+which were accessed since updates were last fetched. Since causally consistent
+operations require an up-to-date view of the current session, the shim layer
+node synchronously fetches operations if the causally preceding operations in
+the current session are not available in the cache.  Strongly consistent
+operations are performed after obtaining exclusive leases on objects. The lease
+mechanism is implemented with the help of Cassandra's support for conditional
+updates and expiring columns.
 
@@
 $e$, must include $m$, and transitively must include every effect from the
-transaction. This ensures atomicity and satisfies RC requirement.
+transaction. This ensures atomicity and satisfies the RC requirement.
 

### 2014-10-31 evaluation.tex
edits

@@
 \end{itemize}
+
+\begin{table}
+\setlength{\tabcolsep}{4pt}
+{\sffamily \small
+\begin{center}
+\begin{tabular} {|l|r|r|r|r|r|r|r|}
+\hline
+{\bf Benchmark} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
+\hline
+{LWW Reg} & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
+{DynamoDB} & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
+{Bank Account} & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
+{Shopping List} & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
+{Online store} & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
+{Rubis} & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
+{Microblog} & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
+\hline
+\end{tabular}
+\end{center}
+}
+\caption{The distribution of classfied contracts. \#T refers to the number of tables in the application.}
+\label{tab:ctrts}
+\end{table}
+
+The distribution of contracts in these applications is given in
+Table~\ref{tab:ctrts}. We see that majority of the operations are classified as
+eventually consistent. Typically, causally consistent operations tend to be
+reads that require to see operations in the same session or causally preceding
+effects. Transactions are useful for maintaining cross-object consistency and
+integrity constraints. For example, we maintain foregin key relationship
+between tables using RC isolation level on the insertion transaction, and MAV
+on the read transaction; if a MAV transaction witnesses the foreign key field
+on table, then it \emph{will} witness the corresponding primiary key row in the
+referenced table.
+
+RR transaction comes in handy for maintaining the integrity of follow user
+functionality in microblog. If user $A$ is followed by $B$, then we add a
+\emph{followedBy} edge from $A$ to $B$ and a \emph{following} edge in the other
+direction. The integrity constraint is that either both the edges are added, or
+none. RR transaction isolation level is used to ensure correctness when
+concurrently $A$ blocks $B$, and $B$ follows $A$. By operating on a snapshot of
+consistent state, the block user transaction will not be interfered by the
+concurrent follow user transaction.

### 2014-10-31 implementation.tex
edits

--- 2014-10-31 implementation.tex
@@
 permitted, but are not allowed to witness the intermediate state of the
-summarization process.
-
-To this end, we adopt a novel summarization strategy that builds on the
-causality property of the store. Suppose we have a set $O$ of effects on a
-object, which when summarized yields a new set of effects $N$. We first
-instantiate a summarization marker $s$, and similar to transaction marker, we
-do not insert it into the store immediately. We insert the new effects in $N$
-with strong consistency, including $s$ as a dependence. Since $s$ is not yet in
-the store, effects from $N$ are not made visible to the clients. Then we insert
-$s$ with strong consistency, including $O$ as dependence, and finally delete
-the rows corresponding to $O$. When the shim layer node witnesses the
-summarization marker $s$ on the object, it removes the effects from $O$ from
-its cache. But due to strong consistency of insertion of $N$ and $s$, the new
-effects $N$ are also observed. Thus, the clients witness the observably
-equivalent new set of effects $N$, and not the old effects $O$. This ensures
-atomicity of summarization in the backing store.
+summarization process. To this end, we adopt a summarization strategy that
+builds on the causality property of the store. The details of the summarization
+are beyond the scope of the paper, and is presented in detail in the tech
+report~\cite{}.

### 2014-10-31 motivation.tex
edits

@@
 deposit :: [Acc] -> Int -> ((), Maybe Acc)
-deposit _ amt = (amt, Just $ Deposit amt)
+deposit _ amt = (amt, Just dollar Deposit amt)
 
@@
 withdraw ctxt v =
-	if sel1 $ getBalance ctxt () >= v
-  then (True, Just $ Withdraw v)
+	if sel1 dollar getBalance ctxt () >= v
+  then (True, Just dollar Withdraw v)
 	else (False, Nothing)

### 2014-10-31 not_included.tex
edits

%% ----------------------------------------------------------------------------
%% Summarization details
%% ----------------------------------------------------------------------------

Suppose we have a set $O$ of effects on a object, which when summarized yields
a new set of effects $N$. We first instantiate a summarization marker $s$, and
similar to transaction marker, we do not insert it into the store immediately.
We insert the new effects in $N$ with strong consistency, including $s$ as a
dependence. Since $s$ is not yet in the store, effects from $N$ are not made
visible to the clients. Then we insert $s$ with strong consistency, including
$O$ as dependence, and finally delete the rows corresponding to $O$. When the
shim layer node witnesses the summarization marker $s$ on the object, it
removes the effects from $O$ from its cache. But due to strong consistency of
insertion of $N$ and $s$, the new effects $N$ are also observed. Thus, the
clients witness the observably equivalent new set of effects $N$, and not the
old effects $O$. This ensures atomicity of summarization in the backing store.

%% ----------------------------------------------------------------------------
%% Example for MAV
%% ----------------------------------------------------------------------------

\if{0}
RC transactions typically perform writes on multiple objects, and want to
ensure that the write is atomic. For example, consider the process of adding a
new user in the micrblog example. The user information is maintained in the
\cf{User} table, with \cf{userID} as the primary key and \cf{userName} and
\cf{password} as values (written \cf{User :: UserID -> (UserName,Password)}. In
order to lookup user id from user name, we need to maintain a secondary index
(\cf{UserIdx :: UserName -> UserID}). Adding a new user involves insertions
into both the tables, and we would like the insertions to be atomic:

\begin{codehaskell}
addUser userID userName password = do
	x <- dollar(classify $true$)
	atomically x dollar do
	  User.addUser userName userID
		UserIdx.addUserIdx userID (userName, password)
\end{codehaskell}

MAV transactions are typically used to preserve foreign key constraints, and
integrity of secondary indexes. For example, consider the code for getting
password for a user name:

\begin{codehaskell}
getPassword userName = do
  x <- dollar(classify $\psi$)
	atomically x dollar do
	  mUid <- UserIdx.getUserID userName
		case mUid of
		  Nothing -> error "User name does not exist"
			Just uid -> User.getPassword uid
\end{codehaskell}

If \cf{getUserID} successfully returns a user id, indicating that the user name
exists in \cf{UserIdx} table, we expect to see the user in the \cf{User} table
as well. For this purpose, we write the contract $\psi$ as:

\begin{smathpar}
\begin{array}{l}
\forall (a:{\sf GetUserID}), (b:{\sf GetPassword}), (c:{\sf AddUser}), \\
~~(d:{\sf AddUserIdx}). ~\txn{a,b}{c,d} ~\wedge~ \vis{d}{a} ~\wedge~ \sameobj{b}{c} \\
\qquad \qquad \qquad \qquad \wedge~ \so{a}{b} \Rightarrow \vis{c}{b}
\end{array}
\end{smathpar}

\noindent which assigns MAV transaction isolation level to \cf{getPassword}
transaction.
\fi

### 2014-10-31 transactions.tex
edits

@@
 save amt =
-  x <- $(classify psi1)
-  atomically x $ do
+  x <- dollar(classify $\psi_1$)
+  atomically x dollar do
     b <- withdraw c amt
-    when b $ deposit s amt
+    when b dollar deposit s amt
 \end{codehaskell}
@@
 totalBalance =
-  x <- $(classify psi2)
-  atomically x $ do
+  x <- dollar(classify $\psi_2$)
+  atomically x dollar do
     b1 <- getBalance s

### 2014-10-31 contract-language.tex
Demonstrated classfication for getBalance contract

@@
 corresponding consistency level as the \emph{consistency class of the
-contract}. The classification scheme is presented formally in
-Figure~\ref{sem:classify}.
+contract}. The classification scheme, presented formally in
+Figure~\ref{sem:classify}, defines rules to judge consistency class of
+a contact. For example, the scheme classifies \cf{getBalance} contract
+($\cv_{gb}$) from Sec.~\ref{sec:motivation} as a {\sf
+CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
+\Rightarrow \cv_{gb}$ is valid in first-order logic (therefore,
+$\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
+\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$).
+Due to the confinement of our contract language to a decidable subset
+of the logic, validity of such sequents can be decided mechanically
+allowing us to automate the classification scheme in \name.
 

### 2014-10-31 motivation.tex
Demonstrated classfication for getBalance contract

@@
 Section~\ref{sec:sysmod}, we express this application-level consistency
-requirement as a contract or specification over \cf{withdraw}:
+requirement as a contract or specification ($\cv_w$) over \cf{withdraw}:
 
@@
 
-\noindent For \cf{getBalance}, we construct the following contract:
+\noindent For \cf{getBalance}, we construct the following contract ($\cv_{gb}$):
 

### 2014-11-01 evaluation.tex
edits

@@
 
+\subsection{Contract Characteristics}
+
 \begin{table}
@@
 \begin{center}
-\begin{tabular} {|l|r|r|r|r|r|r|r|}
+\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
 \hline
-{\bf Benchmark} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
+{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
 \hline
-{LWW Reg} & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
-{DynamoDB} & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
-{Bank Account} & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
-{Shopping List} & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
-{Online store} & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
-{Rubis} & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
-{Microblog} & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
+{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
+{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
+{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
+{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
+{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
+{Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
+{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
 \hline
@@
 }
-\caption{The distribution of classfied contracts. \#T refers to the number of tables in the application.}
+\caption{The distribution of classfied contracts. \#T refers to the number of
+tables in the application. The columns 4-6 (7-9) represent operations
+(transactions) assigned to this consistency (isolation) level.}
 \label{tab:ctrts}
@@
 concurrent follow user transaction.
+
+The proof obligations associated with contract classfication is discarged
+through the Z3 SMT Solver. In particular, the contract classification process
+is completely performed compile time and has no overheads at runtime. Across
+our benchmarks, classifying a contract took 11.5 milliseconds on average.

### 2014-11-01 implementation.tex
implementation

@@
 
-In this section, we describe the instantiation of \name on top of off-the-shelf
-eventually consistent distributed store Cassandra. We particularly focus on the
-challenges in realizing an efficient implementation of our model.
+\name is implemented as a shallow extension of GHC Haskell and runs on top
+of Cassandra, an off-the-shelf eventually consistent distributed data (or
+backing) store responsible for all data management issues (i.e.,
+replication, fault tolerance, availability, and convergence).  Template
+Haskell is used implement static contract classification, and proof
+obligations are discharged with the help of the Z3~\cite{} SMT
+solver. Figure~\ref{fig:impl_mod} describes the overall system architecture.
 
@@
 
-\name is implemented as a shallow extension of Haskell on the Glasgow Haskell
-Compiler (GHC). We utilize Template Haskell functionality to implement static
-contract classification. Our proof obligations are discharged automatically
-with the help of Z3~\cite{} SMT solver. Figure~\ref{fig:impl_mod} describes the
-overall system architecture. The main idea in our implementation of \name is to
-offload the majority of concerns related to data management such as
-replication, fault tolerance, availability and convergence to Cassandra, an
-off-the-shelf eventually consistent store (called \emph{backing store} in the
-sequel).
+Replicated data types and the stronger consistency semantics are
+implemmented and enforced in the \emph{shim layer}. Our implementation
+supports eventual, causal, and strong consistency for data type operations,
+and RC, MAV, and RR semantics for transactions.  This functionality is
+implemented entirely on top of the standard interface exposed by Cassandra.
+From an engineering perspective, leveraging an off-the-shelf data store
+enables an implementation comprising roughly only 2500 lines of Haskell
+code, packaged as a library\footnote{url to the hackage}.
 
-Replicated data type implementation and the stronger consistency semantics are
-enforced in the \emph{shim layer}. Our implementation supports eventual, causal
-and strong consistency for operations, and RC, MAV and RR transaction
-semantics. All of this functionality is implemented on top of the standard
-interface exposed by Cassandra, and illustrates the portability of \name
-programming model. Complex business logic is implemented with the help of
-transactions on the web-server side, with clients interacting with the
-web-service over REST interface. From an engineering perspective, resorting to
-an off-the-shelf backing store meant that our entire implementation comprised of
-only around 2500 lines of Haskell code, which is packaged as a Haskell
-library\footnote{url to the hackage}.
+%% Complex business logic is implemented with the help of
+%% transactions on the web-server side, with clients interacting with the
+%% web-service over REST interface.
 
@@
 
-The shim layer maintains a causally consistent in-memory snapshot of a subset
-of objects in the backing store, by explicitly tracking dependencies introduced
-between the effects due to visibility, session and same transaction relations.
-The dependence tracking is similar to the techniques presented in~\cite{BoltOn}
-and~\cite{Eiger}, with the usual optimizations making use of transitivity
-properties for minimizing the number of dependencies. Shim layer performs the
-reductions associated with replicated datatype operations corresponding to
-client requests. As the backing store provides durability, convergence and
-fault tolerance, each shim layer node simply acts as a soft-state cache, and
-can safely be terminated at any instant. Similarly, more shim layer nodes can
-be spawned on demand.
+The shim layer maintains a causally consistent in-memory snapshot of a
+subset of objects in the backing store, by explicitly tracking dependencies
+introduced between effects due to visibility, session and same transaction
+relations.  Dependence tracking is similar to the techniques presented
+in~\cite{BoltOn} and~\cite{Eiger}. Because Cassandra provides durability,
+convergence and fault tolerance, each shim layer node simply acts as a
+soft-state cache, and can safely be terminated at any point. Similarly, new
+shim layer nodes can be spawned on demand.
 
@@
 
-Every effect generated as a result of an effectful operation on an object
-inserts a new row $(o,e,vis,txn,val)$ into the backing store, where $o$ and $e$
-are object and (unique) effect identifiers, $vis$ is the set of identifiers of
-effects visible to this operation, $txn$ is an optional transaction identifier,
-and $val$ is the value associated with the effect (eg: \cf{Withdraw 50}). The
-shim layer periodically fetches updates from the backing store for those objects
-which were accessed since updates were last fetched. Since causally consistent
-operations require an up-to-date view of the current session, the shim layer
-node synchronously fetches operations if the causally preceding operations in
-the current session are not available in the cache.  Strongly consistent
-operations are performed after obtaining exclusive leases on objects. The lease
-mechanism is implemented with the help of Cassandra's support for conditional
-updates and expiring columns.
+Every generated effect on an object inserts a new row
+$(o,e,\mathit{vis},\mathit{txn},\mathit{val})$ into the backing store, where
+$o$ and $e$ are object and (unique) effect identifiers, \emph{vis} is the
+set of effect identifiers visible to this operation, \emph{txn} is an
+optional transaction identifier, and \mathi{val is the value associated with the
+effect. (eg: \cf{Withdraw 50}).  Since causally consistent operations require an
+up-to-date view of the current session, the shim layer node synchronously
+fetches operations if the causally preceding operations in the current
+session are not available in its cache.  Strongly consistent operations are
+performed after obtaining exclusive leases on objects. The lease mechanism
+is implemented with the help of Cassandra's support for conditional updates
+and expiring columns.
 
@@
 While Cassandra offers all-or-nothing failure semantics for multiple writes
-through batching, readers may witness the initial write while the batch is in
-progress. \name implements atomic visibility by exploiting shim layer causality
-guarantee -- an effect is included only if all the effects if depends on are
-also included.
+through batching, readers may witness the initial write while the batch is
+in progress. \name implements atomic visibility by exploiting the shim layer
+causality guarantee -- an effect is included only if all the effects if
+depends on are also included.
 
-For every transaction, we instantiate a special transaction marker effect $m$,
-but importantly, do not insert into the backing store. $m$ is included as a
-dependence to every effect (let $e$ be one such effect) generated in the
-transaction. Since $m$ has not yet been written to the store, no operation will
-witness $e$ while the transaction in progress. Let $S$ be the set of effects
-generated in the transaction. At the end of the transaction, we insert $m$ into
-the backing store with $S$ as its dependence. Now any replica which includes
-$e$, must include $m$, and transitively must include every effect from the
-transaction. This ensures atomicity and satisfies the RC requirement.
+For every transaction, we instantiate a special transaction marker effect
+$m$, but importantly, do not insert this into the backing store. $m$ is
+included as a dependence to every effect (let $e$ be one such effect)
+generated in the transaction. Since $m$ has not yet been written to the
+store, no operation will witness $e$ while the transaction in progress. Let
+$S$ be the set of effects generated in the transaction. At the end of the
+transaction, we insert $m$ into the backing store with $S$ as its
+dependence. Now any replica which includes $e$, must include $m$, and
+transitively must include every effect from the transaction. This ensures
+atomicity and satisfies the RC requirement.
 
@@
 
-The main challenge in realizing an efficient implementation of operation-based
-replicated data types is that the state of the object i.e., the set of effects
-grows with every effectful operation on the object. If left unchecked, the
-operations slow down over time, until the shim layer memory or backing store
-disk runs out of memory. Luckily, the state of the operation-based replicated
-data type can often be summarized to an \emph{observably equivalent} smaller
-state. For example,
+The main challenge in realizing an efficient implementation of
+operation-based replicated data types is that the state of the object
+defined in terms of its set of effects, grows with every effectful
+operation. If left unchecked, performance slow down over time, until the
+shim layer memory or backing store disk runs out of memory. Luckily, the
+state of the operation-based replicated data type can often be summarized to
+an \emph{observably equivalent} smaller state. For example,
 
@@
 
-Since the semantics of summarization depends on the semantics of the data type,
-we expect the programmer to provide a summarization function for each RDT with
-the following type:
-
+\noindent Since the semantics of summarization depends on the semantics of the data
+type, our programming model provides a summarization function for each RDT:
 \begin{codehaskell}
@@
 \end{codehaskell}
-
 \noindent with the intention that the length of the result is smaller that the
@@
 
-Compared to the shim layer, summarization in the backing store is more
-complicated. The main challenge is that unlike the shim layer, summarization
-cannot run as an atomic operation. Summarization in the backing store involves
-deleting previously inserted rows and inserting new rows, where each row
-corresponds to an effect. It is essential that concurrent client operations are
-permitted, but are not allowed to witness the intermediate state of the
-summarization process. To this end, we adopt a summarization strategy that
-builds on the causality property of the store. The details of the summarization
-are beyond the scope of the paper, and is presented in detail in the tech
-report~\cite{}.
+Summarization in the backing store is more complicated. The main challenge
+is that unlike the shim layer, summarization cannot run as an atomic
+operation. Summarization in the backing store involves deleting previously
+inserted rows and inserting new rows, where each row corresponds to an
+effect. To ensure high availability, it is essential that concurrent client
+operations are permitted, but are not allowed to witness the intermediate
+state of the summarization process. To this end, we adopt a summarization
+strategy that builds on the causality property of the store. The details of
+the summarization are beyond the scope of the paper, and is presented in
+detail in the tech report~\cite{}.

### 2014-11-01 implementation.tex
implementation

@@
 set of effect identifiers visible to this operation, \emph{txn} is an
-optional transaction identifier, and \mathi{val is the value associated with the
+optional transaction identifier, and \emph{val} is the value associated with the
 effect. (eg: \cf{Withdraw 50}).  Since causally consistent operations require an

### 2014-11-01 contract-language.tex
edits

@@
 
-\item \textbf{Causal consistency}: Causally consistent operations are
-  required to see a causally consistent snapshot of the object state,
-  including the actions performed on the same session.  The latter
-  requirement entails that if two operations $o_1$ and $o_2$ from the same
-  session are applied to two different replicas $r_1$ and $r_2$, the second
-  operation cannot be discharged until the effect of $o_1$ is merged with
-  $o_2$ in both $r_1$ and $r_2$. The \cf{getBalance} operation requires
-  causal consistency, as it requires the operations from the same session to
-  be visible, which cannot be guaranteed under just eventual consistency. We
-  assume that causality is only tracked through operations on the same
-  object; two operations in the same session but on different objects are
-  considered causally unrelated under this definition. Stores typically
-  avoid tracking causality across objects to mitigate overheads when
-  causality tracking is unnecessary. The corresponding store semantics is
+\item \textbf{Causal consistency}: Causally consistent operations are required
+	to see a causally consistent snapshot of the object state, including the
+	actions performed on the same session.  The latter requirement entails that
+	if two operations $o_1$ and $o_2$ from the same session are applied to two
+	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
+	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
+	\cf{getBalance} operation requires causal consistency, as it requires the
+	operations from the same session to be visible, which cannot be guaranteed
+	under just eventual consistency. We assume that causality is only tracked
+	through operations on the same object. The corresponding store semantics is
 	captured by the contract $\ccc$ defined below:
@@
   \end{smathpar}
+
+  %%; two operations in the same session but on different objects are
+  %% considered causally unrelated under this definition. Stores typically
+  %%avoid tracking causality across objects to mitigate overheads when
+  %% causality tracking is unnecessary.
 

### 2014-11-01 evaluation.tex
edits

@@
 
-In this section, we evaluate \name programs and report the contract profile as
-well as illustrating the performance benefits of fine-grained consistency
+In this section, we evaluate \name programs and report their contract profile
+as well as illustrating the performance benefits of fine-grained consistency
 classification on operations and transactions. We also report on the impact of
@@
 \item \textbf{Bank account}: Our running example, with savings and current
-accounts.
-\item \textbf{Shopping list}: Collaborative shopping list which allows adding
-and deleting items.
-\item \textbf{Online store}: Models an online store with shopping cart and
-dynamically changing item prices. Checkout process verifies that the custormer
-only pays the accepted price.
-\item \textbf{RuBis}: An ebay-like auction site~\cite{}.
-\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
-Twissandra~\cite{}.
-\end{itemize}
+accounts. \item \textbf{Shopping list}: Collaborative shopping list which
+allows adding and deleting items. \item \textbf{Online store}: Models an online
+store with shopping cart and dynamically changing item prices. Checkout process
+verifies that the custormer only pays the accepted price. \item \textbf{RuBis}:
+An ebay-like auction site~\cite{}. \item \textbf{Microblog}: A twitter-like
+microblogging site, modelled after Twissandra~\cite{}. \end{itemize}
 
@@
 
-\begin{table}
-\setlength{\tabcolsep}{4pt}
-{\sffamily \small
-\begin{center}
-\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
-\hline
-{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
-\hline
-{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
-{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
-{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
-{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
-{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
-{Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
-{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
-\hline
-\end{tabular}
-\end{center}
-}
-\caption{The distribution of classfied contracts. \#T refers to the number of
-tables in the application. The columns 4-6 (7-9) represent operations
-(transactions) assigned to this consistency (isolation) level.}
-\label{tab:ctrts}
-\end{table}
+\begin{table} \setlength{\tabcolsep}{4pt} {\sffamily \small \begin{center}
+\begin{tabular} {|l|r|r|r|r|r|r|r|r|} \hline {\bf Benchmark} & {\bf LOC} & {\bf
+\#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
+\hline {LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\ {DynamoDB} & 126 & 1 & 3
+& 1 & 2 & 0 & 0 & 0 \\ {Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
+{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\ {Online store} & 340 & 4 &
+9 & 1 & 0 & 2 & 0 & 1 \\ {Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
+{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\ \hline \end{tabular}
+\end{center} } \caption{The distribution of classfied contracts. \#T refers to
+the number of tables in the application. The columns 4-6 (7-9) represent
+operations (transactions) assigned to this consistency (isolation) level.}
+\label{tab:ctrts} \end{table}
 
@@
 our benchmarks, classifying a contract took 11.5 milliseconds on average.
+
+\subsection{Performance}
+
+In this section, we study the performance benefit of fine-grained contract
+classification, and the impact of summarization optimization.
+
+\subsection{Configuration}
+
+We deploy \name applications in \emph{clusters} where each cluster is composed
+of 5 fully replicated Cassandra replicas within the same datacenter. We
+instantiate one shim layer node for every cassandra replica, an place it on the
+same VM as the cassandra replica. The clients are instantiated on the same data
+center as the store.

### 2014-11-01 implementation.tex
implementation
edits

--- 2014-10-31 implementation.tex
@@
 and strong consistency for operations, and RC, MAV and RR transaction
-semantics. All of this functionality is implemented on top of the standard
-interface exposed by Cassandra, and illustrates the portability of \name
-programming model. Complex business logic is implemented with the help of
-transactions on the web-server side, with clients interacting with the
-web-service over REST interface. From an engineering perspective, resorting to
-an off-the-shelf backing store meant that our entire implementation comprised of
-only around 2500 lines of Haskell code, which is packaged as a Haskell
-library\footnote{url to the hackage}.
+semantics. Complex business logic is implemented with the help of transactions
+on the server side, with clients interacting with the web-service over REST
+interface. From an engineering perspective, resorting to an off-the-shelf
+backing store meant that our entire implementation comprised of only around
+2500 lines of Haskell code, which is packaged as a Haskell library\footnote{url
+to the hackage}.
 
+% All of this functionality is implemented on top of the standard interface
+% exposed by Cassandra, and illustrates the portability of \name programming
+% model.
 \subsection{Shim Layer}
@@
 
+\KC{Perhaps the following two subsections can be elided/shortened?}
+
 \subsection{Operation Consistency}
@@
 and $val$ is the value associated with the effect (eg: \cf{Withdraw 50}). The
-shim layer periodically fetches updates from the backing store for those objects
-which were accessed since updates were last fetched. Since causally consistent
-operations require an up-to-date view of the current session, the shim layer
-node synchronously fetches operations if the causally preceding operations in
-the current session are not available in the cache.  Strongly consistent
-operations are performed after obtaining exclusive leases on objects. The lease
-mechanism is implemented with the help of Cassandra's support for conditional
-updates and expiring columns.
+shim layer periodically fetches updates from the backing store for those
+objects which were accessed since updates were last fetched. Since causally
+consistent operations require an up-to-date view of the current session, the
+shim layer node synchronously fetches operations if the causally preceding
+operations in the current session are not available in the cache.  Strongly
+consistent operations are performed after obtaining exclusive leases on
+objects. The lease mechanism is implemented with the help of Cassandra's
+support for conditional updates and expiring columns.
 

### 2014-11-01 implementation.tex

--- 2014-11-01 implementation.tex
@@
 
-Replicated data types and the stronger consistency semantics are
-implemmented and enforced in the \emph{shim layer}. Our implementation
-supports eventual, causal, and strong consistency for data type operations,
-and RC, MAV, and RR semantics for transactions.  This functionality is
-implemented entirely on top of the standard interface exposed by Cassandra.
-From an engineering perspective, leveraging an off-the-shelf data store
-enables an implementation comprising roughly only 2500 lines of Haskell
-code, packaged as a library\footnote{url to the hackage}.
+Replicated data types and the stronger consistency semantics are implemented
+and enforced in the \emph{shim layer}. Our implementation supports eventual,
+causal, and strong consistency for data type operations, and RC, MAV, and RR
+semantics for transactions.  This functionality is implemented entirely on top
+of the standard interface exposed by Cassandra. From an engineering
+perspective, leveraging an off-the-shelf data store enables an implementation
+comprising roughly only 2500 lines of Haskell code, packaged as a
+library\footnote{url to the hackage}.
 
@@
 
-The shim layer maintains a causally consistent in-memory snapshot of a
-subset of objects in the backing store, by explicitly tracking dependencies
-introduced between effects due to visibility, session and same transaction
-relations.  Dependence tracking is similar to the techniques presented
-in~\cite{BoltOn} and~\cite{Eiger}. Because Cassandra provides durability,
-convergence and fault tolerance, each shim layer node simply acts as a
-soft-state cache, and can safely be terminated at any point. Similarly, new
-shim layer nodes can be spawned on demand.
+The shim layer maintains a causally consistent in-memory snapshot of a subset
+of objects in the backing store, by explicitly tracking dependencies introduced
+between effects due to visibility, session and same transaction relations.
+Dependence tracking is similar to the techniques presented in~\cite{BoltOn}
+and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
+tolerance, each shim layer node simply acts as a soft-state cache, do not
+commicate directly with each other, and can safely be terminated at any point.
+Similarly, new shim layer nodes can be spawned on demand.
 
-\subsection{Operation Consistency}
+The shim layer nodes periodically fetch updates from the store for eventually
+consistent operations, and on-demand for causally consistent and stronly
+consistent operations. Strongly consistent operations are performed after
+obtaining exclusive leases on objects. The lease mechanism is implemented with
+the help of Cassandra's support for conditional updates and expiring columns.
 
-Every generated effect on an object inserts a new row
-$(o,e,\mathit{vis},\mathit{txn},\mathit{val})$ into the backing store, where
-$o$ and $e$ are object and (unique) effect identifiers, \emph{vis} is the
-set of effect identifiers visible to this operation, \emph{txn} is an
-optional transaction identifier, and \emph{val} is the value associated with the
-effect. (eg: \cf{Withdraw 50}).  Since causally consistent operations require an
-up-to-date view of the current session, the shim layer node synchronously
-fetches operations if the causally preceding operations in the current
-session are not available in its cache.  Strongly consistent operations are
-performed after obtaining exclusive leases on objects. The lease mechanism
-is implemented with the help of Cassandra's support for conditional updates
-and expiring columns.
-
-\subsection{Transactions}
-
-While Cassandra offers all-or-nothing failure semantics for multiple writes
-through batching, readers may witness the initial write while the batch is
-in progress. \name implements atomic visibility by exploiting the shim layer
-causality guarantee -- an effect is included only if all the effects if
-depends on are also included.
-
-For every transaction, we instantiate a special transaction marker effect
-$m$, but importantly, do not insert this into the backing store. $m$ is
-included as a dependence to every effect (let $e$ be one such effect)
-generated in the transaction. Since $m$ has not yet been written to the
-store, no operation will witness $e$ while the transaction in progress. Let
-$S$ be the set of effects generated in the transaction. At the end of the
-transaction, we insert $m$ into the backing store with $S$ as its
-dependence. Now any replica which includes $e$, must include $m$, and
-transitively must include every effect from the transaction. This ensures
-atomicity and satisfies the RC requirement.
-
-The above scheme prevents a transaction from witnessing its own effects. This
-might conflict with the causality requirement on the operations. Hence,
-transactions piggy-back the previous effects from the same transaction for each
-request. MAV semantics is implemented by keeping track of the set of
-transaction markers $M$ witnessed by the transaction, and before performing an
-operation at some replica, ensuring that $M$ is a subset of the transaction
-markers included at that replica. If not, the missing effects are synchronously
-fetched. RR semantics is realized by capturing a optimized snapshot of the
-state of some replica; each operation from an RR transaction is applied to this
-snapshot state. Any generated effects are added to this snapshot.
+Cassandra does not provide general-purpose transactions. Since the transaction
+guarantees provided by \name are coordination-free~\cite{BailisHAT}, we realize
+efficient implementations by explicitly tracking the dependencies between
+operations and transactions. Importantly, the weaker isolation semantics of
+transactions in \name permit transactions to be discharged if at least one shim
+layer node is reachable.
 
@@
 
-Summarization in the backing store is more complicated. The main challenge
-is that unlike the shim layer, summarization cannot run as an atomic
-operation. Summarization in the backing store involves deleting previously
-inserted rows and inserting new rows, where each row corresponds to an
-effect. To ensure high availability, it is essential that concurrent client
-operations are permitted, but are not allowed to witness the intermediate
-state of the summarization process. To this end, we adopt a summarization
-strategy that builds on the causality property of the store. The details of
-the summarization are beyond the scope of the paper, and is presented in
-detail in the tech report~\cite{}.
+Performing summarization in the backing store is more complicated; the whole
+process needs to be atomic from a clients perspective, but Cassadra does not
+provide multirow transactions. We have implemented a highly available
+implementation of summarization that permits concurrent client operations, but
+nonetheless prohibits these operations from witnessing intermediate states of
+the summarization process.

### 2014-11-01 implementation.tex
implementation spell-check

@@
 tolerance, each shim layer node simply acts as a soft-state cache, do not
-commicate directly with each other, and can safely be terminated at any point.
+communicate directly with each other, and can safely be terminated at any point.
 Similarly, new shim layer nodes can be spawned on demand.
@@
 The shim layer nodes periodically fetch updates from the store for eventually
-consistent operations, and on-demand for causally consistent and stronly
+consistent operations, and on-demand for causally consistent and strongly
 consistent operations. Strongly consistent operations are performed after
@@
 Performing summarization in the backing store is more complicated; the whole
-process needs to be atomic from a clients perspective, but Cassadra does not
-provide multirow transactions. We have implemented a highly available
+process needs to be atomic from a clients perspective, but Cassandra does not
+provide multi-row transactions. We have implemented a highly available
 implementation of summarization that permits concurrent client operations, but

### 2014-11-01 contract-language.tex
edits

@@
 as a formula in our contract language, and informally describe the kind
-of application-level consistency requirements that can met under the
+of application-level consistency requirements that can be met under the
 consistency level:
@@
 	to see a causally consistent snapshot of the object state, including the
-	actions performed on the same session.  The latter requirement entails that
+	actions performed on the same session.  The latter requirement implies that
 	if two operations $o_1$ and $o_2$ from the same session are applied to two
@@
 
-Our goal is to map application-level consistency constraints on
-operations to appropriate store-level consistency guarantees capable
-of satisfying these constraints.  The ability to express both of them as
-contracts in our contract language lets us compare and determine if
+Our goal is to map application-level consistency constraints on operations
+to appropriate store-level consistency guarantees capable of satisfying
+these constraints.  The ability to express both these kinds of constraints
+as contracts in our contract language lets us compare and determine if
 contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
 satisfied under a store consistency level identified by the contract
-$\cv_{st}$. Towards this end, we define a binary \emph{weaker than}
-relation for our contract language as following:
-
+$\cv_{st}$. Towards this end, we define a binary \emph{weaker than} relation
+for our contract language as following:
 \begin{definition}
@@
 \end{definition}
-
-\vspace{-1em}
+\vspace{-2em}
 \noindent The $\Delta$ referred in the above defintion is a
 conjunction of assumptions about the nature of primitive relations,
-such as $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
+e.g., $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
 relation $\hbZ$ is acyclic. A \emph{well-formed} axiomatic execution
@@
 than} a store contract ($\cv_{st}$), then constraints expressed by the
-former are implied by guarantees provided by the later. Due to the
-completeness of first-order logic\cite{completeness}, this means that
-any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
+former are implied by guarantees provided by the latter.  The
+completeness of first-order logic\cite{completeness} allows us to assert
+that any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
 $\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
 \cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
-store consistency level captured by $\cv_{st}$.
+a store consistency level captured by $\cv_{st}$.
 
@@
 contract}. The classification scheme, presented formally in
-Figure~\ref{sem:classify}, defines rules to judge consistency class of
-a contact. For example, the scheme classifies \cf{getBalance} contract
-($\cv_{gb}$) from Sec.~\ref{sec:motivation} as a {\sf
+Figure~\ref{sem:classify}, defines rules to judge the consistency class of
+a contact. For example, the scheme classifies the \cf{getBalance} contract
+($\cv_{gb}$) from Section~\ref{sec:motivation} as a {\sf
 CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
@@
 
-In this section, we present a meta-theoretic result that certifies the
-soundness of classification-based contract enforcement. To help us
-state the result, we present an axiomatic model of the our system
-described informally in Sec.~\ref{sec:sysmod}:
+We now present a meta-theoretic result that certifies the soundness of
+classification-based contract enforcement. To help us state the result, we
+define an axiomatic model of the our system described informally in
+Section~\ref{sec:sysmod}:
 
@@
 and let $\tau$ denote the consistency class of $\cv$ as determined by
-the contract classification scheme. Forall well-formed execution
+the contract classification scheme. For all well-formed execution
 states $\E$, $\E'$ such that
 $\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
- {\E';\langle \sigma \rangle \pll \Sigma}$
+ {\E';\langle \sigma \rangle \pll \Sigma}$,
 if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$,
-and $\ecc$ contracts in all well-formed executions, then the same store,
-extended with classification scheme, can enforce all well-formed \name
-contracts. The proof of the theorem has been relegated to our tech report
-and is provided as part of the supplementary
-material~\cite{techrep}\footnote{The supplementary material and tech report
-  also defines an operational semantics that concretizes the rewrite
-  relation ($\xrightarrow{}$), along with the proof that it correctly
-  enforces all \name contracts.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given in a tech report and is provided
+as part of the supplementary material~\cite{techrep}\footnote{The
+  supplementary material and tech report also defines an operational
+  semantics that concretizes the rewrite relation ($\xrightarrow{}$), along
+  with the proof that it correctly enforces all \name contracts.}.

### 2014-11-01 implementation.tex
edits

@@
 
-Performing summarization in the backing store is more complicated; the whole
-process needs to be atomic from a clients perspective, but Cassandra does not
+Performing summarization in the backing store is more complicated since the whole
+process needs to be atomic from a client's perspective, but Cassandra does not
 provide multi-row transactions. We have implemented a highly available

### 2014-11-01 evaluation.tex
Committing first experiments

@@
 
-\begin{itemize}
-\item \textbf{LWW register}: A last-write-wins register.
-\item \textbf{DynamoDB register}: A integer register that allows eventual and
-strong puts and gets, conditional puts, increment and decrement operations.
-\item \textbf{Bank account}: Our running example, with savings and current
-accounts. \item \textbf{Shopping list}: Collaborative shopping list which
-allows adding and deleting items. \item \textbf{Online store}: Models an online
-store with shopping cart and dynamically changing item prices. Checkout process
-verifies that the custormer only pays the accepted price. \item \textbf{RuBis}:
-An ebay-like auction site~\cite{}. \item \textbf{Microblog}: A twitter-like
+\begin{itemize} \item \textbf{LWW register}: A last-write-wins register. \item
+\textbf{DynamoDB register}: A integer register that allows eventual and strong
+puts and gets, conditional puts, increment and decrement operations. \item
+\textbf{Bank account}: Our running example, with savings and current accounts.
+\item \textbf{Shopping list}: Collaborative shopping list which allows adding
+and deleting items. \item \textbf{Online store}: Models an online store with
+shopping cart and dynamically changing item prices. Checkout process verifies
+that the custormer only pays the accepted price. \item \textbf{RuBis}: An
+ebay-like auction site~\cite{}. \item \textbf{Microblog}: A twitter-like
 microblogging site, modelled after Twissandra~\cite{}. \end{itemize}
@@
 same VM as the cassandra replica. The clients are instantiated on the same data
-center as the store.
+center as the store, and run the transactions. We deploy the each node in the
+cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,
+30GB memory, and 320GB of SSD storage. Our shim layer nodes are multi-threaded,
+and we allocate 8 CPUs for each shim layer node. The clients also run on
+\cf{c2.4xlarge} instances. We call this deployment \cf{1DC}. For our WAN
+experiments, we instantiate 2 clusters (each with 5 nodes), one in US-east (N.
+Virginia) and another in US-west (Oregon). The average latency between the
+regions was 85 ms. We call this deployment \cf{2DC}.
+
+\subsection{LWW Register}
+
+\begin{figure*}
+  \centering
+  \subfigure[Perf. Oper. 1DC]{\label{grf:LWW-perf-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC.pdf}}
+  \subfigure[Perf. Oper. 2DC]{\label{grf:LWW-perf-2DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-2DC.pdf}}
+  \subfigure[Perf. Txn. 1DC]{\label{grf:LWW-txn-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC-txn.pdf}}
+  \subfigure[Mutation Prof. 1DC]{\label{grf:LWW-txn-wp}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn-wp-vs-tp.pdf}}
+	\caption{Performance of LWW Register RDT.}
+  \label{grf:LWW_perf}
+\end{figure*}
+
+We analyze the performance of different consistency classifications in \name
+using the LWW register. Our client workload was generated using YCSB
+benchmark~\cite{}, with the default workload of 50\% reads and 50\% writes,
+with keys uniformly chosen from a set of 1000 keys. We varied the number of
+clients from 128 to 1024, with each workload running for 180 seconds.
+Last-writer-wins is the conflict resolution policy used by Cassandra. Hence, we
+are able to directly compare the performance of \name implementation against
+Cassandra.
+
+Figure~\ref{grf:LWW-perf-1DC} shows the throughput vs. latency graph for the
+different consistency levels. We also include Cassandra, which provides basic
+eventual consistency. Recall that \name's eventual consistency is stronger than
+the guarantees provided by Cassandra (Section~\ref{sec:store_sem}). With 512
+clients, the average latency of operations on Cassandra was 32.24 milliseconds.
+EC and CC under \name was 30\% and 57\% slower whereas strong consistency was
+241\% slower. EC(SC) was able to achieve 80\%(40\%) of throughput of vanilla
+Cassandra. The results show that while \name provides an expressive programming
+model with stronger consistency guarantees, its performance is comparable to
+vanilla Cassandra. However, SC has a huge performance hit due to coordination
+between all the replicas for obtaining object leases as well as reading and
+writing to all replicas.

### 2014-11-01 Data/LWW-tp-vs-lat.py

@@ -1,0 +1,73 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["LWW-tp-vs-lat-1DC.dat", "LWW-tp-vs-lat-2DC.dat", "LWW-tp-vs-lat-1DC-txn.dat"]
+	for f in files:
+		Cassandra = []
+		EC = []
+		CC = []
+		SC = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+
+			d = line.split ()
+			Cassandra.append ((float(d[0]),float (d[1])))
+			EC.append ((float(d[2]),float (d[3])))
+			CC.append ((float(d[4]),float (d[5])))
+			SC.append ((float(d[6]),float (d[7])))
+
+		(cx,cy) = zip(*Cassandra)
+		(ecx,ecy) = zip(*EC)
+		(ccx,ccy) = zip(*CC)
+		(scx,scy) = zip(*SC)
+
+
+		plt.xlabel ("Throughput (X 1000 ops/s)")
+		plt.ylabel ("Latency (ms)")
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		if (f == "LWW-tp-vs-lat-1DC-txn.dat"):
+			plt.plot (cx, cy, nodeKind [0], label="NoTxn", linewidth=2.0, ms=15)
+			plt.plot (ecx, ecy, nodeKind [1], label="RC", linewidth=2.0, ms=15)
+			plt.plot (ccx, ccy, nodeKind [2], label="MAV", linewidth=2.0, ms=15)
+			plt.plot (scx, scy, nodeKind [3], label="RR", linewidth=2.0, ms=15)
+		else:
[…]

### 2014-11-01 Data/LWW-txn-wp-vs-tp.py

@@ -1,0 +1,63 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["LWW-txn-wp-vs-tp.dat"]
+	for f in files:
+		Cassandra = []
+		EC = []
+		CC = []
+		SC = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+
+			d = line.split ()
+			Cassandra.append ((float(d[0]),float (d[1])))
+			EC.append ((float(d[0]),float (d[2])))
+			CC.append ((float(d[0]),float (d[3])))
+			SC.append ((float(d[0]),float (d[4])))
+
+		(cx,cy) = zip(*Cassandra)
+		(ecx,ecy) = zip(*EC)
+		(ccx,ccy) = zip(*CC)
+		(scx,scy) = zip(*SC)
+
+
+		plt.xlabel ("Writes proportion")
+		plt.ylabel ("Throughput (X1000 ops/s)")
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		plt.plot (cx, cy, nodeKind [0], label="NoTxn", linewidth=2.0, ms=15)
+		plt.plot (ecx, ecy, nodeKind [1], label="RC", linewidth=2.0, ms=15)
+		plt.plot (ccx, ccy, nodeKind [2], label="MAV", linewidth=2.0, ms=15)
+		plt.plot (scx, scy, nodeKind [3], label="RR", linewidth=2.0, ms=15)
+
+		plt.xlim(xmin = 0)
+		plt.legend ()
+		print ("saving fig " + f.replace("dat","pdf"))
+		plt.savefig (f.replace("dat","pdf"))
+		plt.close ()
+
+main ()

### 2014-11-02 evaluation.tex
edits

@@
   \centering
-  \subfigure[Perf. Oper. 1DC]{\label{grf:LWW-perf-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC.pdf}}
-  \subfigure[Perf. Oper. 2DC]{\label{grf:LWW-perf-2DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-2DC.pdf}}
-  \subfigure[Perf. Txn. 1DC]{\label{grf:LWW-txn-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC-txn.pdf}}
-  \subfigure[Mutation Prof. 1DC]{\label{grf:LWW-txn-wp}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn-wp-vs-tp.pdf}}
+  \subfigure[Operation performance 1DC]{\label{grf:LWW-perf-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC.pdf}}
+  \subfigure[Operation performance 2DC]{\label{grf:LWW-perf-2DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-2DC.pdf}}
+  \subfigure[Transaction performance 1DC]{\label{grf:LWW-txn-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC-txn.pdf}}
+  \subfigure[Mutations in transactions]{\label{grf:LWW-txn-wp}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn-wp-vs-tp.pdf}}
 	\caption{Performance of LWW Register RDT.}
@@
 benchmark~\cite{}, with the default workload of 50\% reads and 50\% writes,
-with keys uniformly chosen from a set of 1000 keys. We varied the number of
+with keys uniformly chosen from a set of 100,000 keys. We varied the number of
 clients from 128 to 1024, with each workload running for 180 seconds.
@@
 Figure~\ref{grf:LWW-perf-1DC} shows the throughput vs. latency graph for the
-different consistency levels. We also include Cassandra, which provides basic
-eventual consistency. Recall that \name's eventual consistency is stronger than
-the guarantees provided by Cassandra (Section~\ref{sec:store_sem}). With 512
-clients, the average latency of operations on Cassandra was 32.24 milliseconds.
-EC and CC under \name was 30\% and 57\% slower whereas strong consistency was
-241\% slower. EC(SC) was able to achieve 80\%(40\%) of throughput of vanilla
-Cassandra. The results show that while \name provides an expressive programming
-model with stronger consistency guarantees, its performance is comparable to
-vanilla Cassandra. However, SC has a huge performance hit due to coordination
-between all the replicas for obtaining object leases as well as reading and
-writing to all replicas.
+different consistency levels in \cf{1DC} configuration. We also include
+Cassandra, which provides basic eventual consistency. Recall that \name's
+eventual consistency is stronger than the guarantees provided by Cassandra
+(Section~\ref{sec:store_sem}). With 512 clients, the average latency of
+operations on Cassandra was 32.24 milliseconds. EC and CC under \name was 30\%
+and 57\% slower whereas strong consistency was 241\% slower. EC(SC) was able to
+achieve 80\%(40\%) of throughput of vanilla Cassandra. The results show that
+while \name provides an expressive programming model with stronger consistency
+guarantees, its performance is comparable to vanilla Cassandra. However, SC has
+a huge performance hit due to coordination between all the replicas for
+obtaining object leases as well as reading and writing to all replicas. This
+overhead worsens in a WAN setting. Figure~\ref{grf:LWW-perf-2DC} shows the
+throughput vs. latency graph under 2DC configuration. While the Cassandra, EC
+and CC numbers largely remain the same when compared to 1DC configuration, SC
+is 6$\times$ to 12$\times$ higher latency and lower throughput than than the
+1DC case.
+
+We compare the performance of different transaction isolation level choices in
+Figure~\ref{grf:LWW-txn-1DC}. The numbers were obtained under 1DC
+configuration. The YCSB workload was modified to issue 10 operations per
+transaction. Each operation is assumed to have eventual consistency. \cf{NoTxn}
+corresponds to a configuration that does not use transactions. Compared to this
+RC is only 12\% shower in terms of latency with 512 clients, where as RR is
+2.3X slower. The difference between RC and \cf{NoTxn} is due to the meta-data
+overhead of recording transaction information in the object state. For RR
+transaction, The cost of capturing and maintaining the snapshot in an RR
+transaction is the biggest source of overhead.
+
+Moreover, the size of the snapshot is proportional to the the number of
+different objects accessed by the session as well as the number of updates.
+Hence, with increasing proportion of writes, the overhead of RR transaction
+increases as shown in Figure~\ref{grf:LWW-txn-wp}. The numbers were obtained
+with 128 clients. Compared to an RC transaction whose throughput decreasess
+45\% as the write proportion increases from 0\% to 50\%, RR transaction's
+throughput drops 78\% in the same interval.

### 2014-11-02 evaluation.tex
edits

@@
 
-\subsection{Configuration}
+\subsubsection{Configuration}
 
@@
 
-\subsection{LWW Register}
+\subsubsection{LWW Register}
 
@@
 throughput drops 78\% in the same interval.
+
+\subsection{Bank Account}
+
+\begin{figure}
+  \centering
+	\includegraphics[width=0.5\columnwidth]{Figures/BA-legend.pdf}
+  \subfigure[Latency 1DC]{\label{grf:BA-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-lat.pdf}}
+  \subfigure[Throughput 1DC]{\label{grf:BA-tp}\includegraphics[width=0.23\textwidth]{graphs/BA-tp.pdf}}
+	\caption{Performance of Bank Account RDT.}
+  \label{grf:BA_perf}
+\end{figure}
+
+Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
+of operations in bank account example as we increase the number of clients. The
+benchmark uniformly chose from 100,000 keys, where the operation spread was
+25\% withdraw, 25\% deposit and 50\% getBalance. Here, the lines marked EC and
+CC correspond to all operations being assigned EC and CC consistency levels.
+These levels compromise correctness as withdraw has to be and SC operation. The
+line SC corresponds to a configuration where all operations are strongly
+consistent; this ensures application correctness. \name corresponds to our
+implementation, which classifies operations based on their contracts. At 512
+clients, \name implementation was within 30\% of latency and 12\% of throughput
+of EC, while providing correctness, whereas SC operations had 162\% higher
+latency and 52\% lower throughput than EC operations. This illustrates the
+benefit of fine-grained contract classification in \name.

### 2014-11-02 Data/BA.py

@@ -1,0 +1,72 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter, MultipleLocator
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["BA-lat.dat","BA-tp.dat"]
+	for f in files:
+		Cassandra = []
+		EC = []
+		CC = []
+		SC = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+
+			d = line.split ()
+			Cassandra.append ((float(d[0]),float (d[1])))
+			EC.append ((float(d[0]),float (d[2])))
+			CC.append ((float(d[0]),float (d[3])))
+			SC.append ((float(d[0]),float (d[4])))
+
+		(cx,cy) = zip(*Cassandra)
+		(ecx,ecy) = zip(*EC)
+		(ccx,ccy) = zip(*CC)
+		(scx,scy) = zip(*SC)
+
+
+		plt.xlabel ("Clients")
+		xmajorLocator = MultipleLocator(128)
+		plt.axes().xaxis.set_major_locator (xmajorLocator)
+		if (f=="BA-lat.dat"):
+			plt.ylabel ("Latency (ms)")
+			ymajorLocator = MultipleLocator(40)
+			plt.axes().yaxis.set_major_locator (ymajorLocator)
+			plt.axes().set_aspect(2.5)
+		else:
+			plt.ylabel ("Throughput (X1000 ops/s)")
+			plt.axes().set_aspect(50)
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		plt.plot (cx, cy, nodeKind [0], label="Quelea", linewidth=2.0, ms=15)
[…]

### 2014-11-02 evaluation.tex
edits

@@
 Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
-of operations in bank account example as we increase the number of clients. The
-benchmark uniformly chose from 100,000 keys, where the operation spread was
-25\% withdraw, 25\% deposit and 50\% getBalance. Here, the lines marked EC and
-CC correspond to all operations being assigned EC and CC consistency levels.
-These levels compromise correctness as withdraw has to be and SC operation. The
-line SC corresponds to a configuration where all operations are strongly
-consistent; this ensures application correctness. \name corresponds to our
-implementation, which classifies operations based on their contracts. At 512
-clients, \name implementation was within 30\% of latency and 12\% of throughput
-of EC, while providing correctness, whereas SC operations had 162\% higher
-latency and 52\% lower throughput than EC operations. This illustrates the
-benefit of fine-grained contract classification in \name.
+of operations in bank account example as we increase the number of clients in
+\cf{1DC} configuration. The benchmark uniformly chose from 100,000 keys, where
+the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance. Here,
+the lines marked EC and CC correspond to all operations being assigned EC and
+CC consistency levels. These levels compromise correctness as withdraw has to
+be and SC operation. The line SC corresponds to a configuration where all
+operations are strongly consistent; this ensures application correctness. \name
+corresponds to our implementation, which classifies operations based on their
+contracts. At 512 clients, \name implementation was within 30\% of latency and
+12\% of throughput of EC, while providing correctness, whereas SC operations
+had 162\% higher latency and 52\% lower throughput than EC operations. This
+illustrates the benefit of fine-grained contract classification in \name.

### 2014-11-02 evaluation.tex
edits

@@
 
-\begin{itemize} \item \textbf{LWW register}: A last-write-wins register. \item
-\textbf{DynamoDB register}: A integer register that allows eventual and strong
-puts and gets, conditional puts, increment and decrement operations. \item
-\textbf{Bank account}: Our running example, with savings and current accounts.
+\begin{itemize}
+
+\item \textbf{LWW register}: A last-write-wins register.
+
+\item \textbf{DynamoDB register}: A integer register that allows eventual and
+strong puts and gets, conditional puts, increment and decrement operations.
+
+\item \textbf{Bank account}: Our running example, with savings and current
+accounts.
+
 \item \textbf{Shopping list}: Collaborative shopping list which allows adding
-and deleting items. \item \textbf{Online store}: Models an online store with
-shopping cart and dynamically changing item prices. Checkout process verifies
-that the custormer only pays the accepted price. \item \textbf{RuBis}: An
-ebay-like auction site~\cite{}. \item \textbf{Microblog}: A twitter-like
-microblogging site, modelled after Twissandra~\cite{}. \end{itemize}
+and deleting items.
+
+\item \textbf{Online store}: Models an online store with shopping cart and
+dynamically changing item prices. Checkout process verifies that the custormer
+only pays the accepted price. \item \textbf{RuBis}: An ebay-like auction
+site~\cite{}.
+
+\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
+Twissandra~\cite{}.
+
+\end{itemize}
 
@@
 
-\begin{table} \setlength{\tabcolsep}{4pt} {\sffamily \small \begin{center}
-\begin{tabular} {|l|r|r|r|r|r|r|r|r|} \hline {\bf Benchmark} & {\bf LOC} & {\bf
-\#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
-\hline {LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\ {DynamoDB} & 126 & 1 & 3
-& 1 & 2 & 0 & 0 & 0 \\ {Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
-{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\ {Online store} & 340 & 4 &
-9 & 1 & 0 & 2 & 0 & 1 \\ {Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
-{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\ \hline \end{tabular}
-\end{center} } \caption{The distribution of classfied contracts. \#T refers to
-the number of tables in the application. The columns 4-6 (7-9) represent
-operations (transactions) assigned to this consistency (isolation) level.}
-\label{tab:ctrts} \end{table}
+\begin{table}
+\setlength{\tabcolsep}{4pt}
+{\sffamily \small
+\begin{center}
+\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
+\hline
+{\bf Benchmark} & {\bf LOC} & {\bf \#T} & {\bf EC} & {\bf CC} & {\bf SC} & {\bf RC} & {\bf MAV} & {\bf RR} \\
+\hline
+{LWW Reg} & 108 & 1 & 2 & 2 & 2 & 0 & 0 & 0 \\
+{DynamoDB} & 126 & 1 & 3 & 1 & 2 & 0 & 0 & 0 \\
+{Bank Account} & 155 & 1 & 1 & 1 & 1 & 1 & 0 & 1 \\
+{Shopping List} & 140 & 1 & 2 & 1 & 1 & 0 & 0 & 0 \\
+{Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
+{Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
+{Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
+\hline
+\end{tabular}
+\end{center} }
+\caption{The distribution of classfied contracts. \#T refers to the number of
+tables in the application. The columns 4-6 (7-9) represent operations
+(transactions) assigned to this consistency (isolation) level.}
+\label{tab:ctrts}
+\end{table}
 

### 2014-11-03 evaluation.tex
Major addition to implementation

@@
 
-\item \textbf{LWW register}: A last-write-wins register.
+\item \textbf{LWW register}: A last-write-wins register that provides read and
+write operations. Each write is associated with a timestamp, which is used to
+resolve conflicting concurrent writes -- newer write wins.
 
@@
 dynamically changing item prices. Checkout process verifies that the custormer
-only pays the accepted price. \item \textbf{RuBis}: An ebay-like auction
-site~\cite{}.
+only pays the accepted price.
+
+\item \textbf{RuBis}: An ebay-like auction site~\cite{}. The application allows
+users to browse items, bid for items on sale, and pay for items from a wallet
+modelled after a bank account.
 
 \item \textbf{Microblog}: A twitter-like microblogging site, modelled after
-Twissandra~\cite{}.
+Twissandra~\cite{}. The application allows adding a new user, adding and
+replying to tweets, following, unfollowing and blocking users, and fetching a
+user's timeline, userline, followers and following.
 
 \end{itemize}
-
-\subsection{Contract Characteristics}
 
@@
 The distribution of contracts in these applications is given in
-Table~\ref{tab:ctrts}. We see that majority of the operations are classified as
-eventually consistent. Typically, causally consistent operations tend to be
-reads that require to see operations in the same session or causally preceding
-effects. Transactions are useful for maintaining cross-object consistency and
-integrity constraints. For example, we maintain foregin key relationship
-between tables using RC isolation level on the insertion transaction, and MAV
-on the read transaction; if a MAV transaction witnesses the foreign key field
-on table, then it \emph{will} witness the corresponding primiary key row in the
-referenced table.
-
-RR transaction comes in handy for maintaining the integrity of follow user
-functionality in microblog. If user $A$ is followed by $B$, then we add a
-\emph{followedBy} edge from $A$ to $B$ and a \emph{following} edge in the other
-direction. The integrity constraint is that either both the edges are added, or
-none. RR transaction isolation level is used to ensure correctness when
-concurrently $A$ blocks $B$, and $B$ follows $A$. By operating on a snapshot of
-consistent state, the block user transaction will not be interfered by the
-concurrent follow user transaction.
-
-The proof obligations associated with contract classfication is discarged
-through the Z3 SMT Solver. In particular, the contract classification process
-is completely performed compile time and has no overheads at runtime. Across
-our benchmarks, classifying a contract took 11.5 milliseconds on average.
+Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
+are classified as eventually consistent and RC, respectively. Operation
+contracts are used to enforce integrity and visibility constraints on
+individual fields in the tables. Transactions are mainly used to consistently
+modify and access related fields across tables. The proof obligations
+associated with contract classfication is discarged through the Z3 SMT Solver.
+In \name, the contract classification process is completely performed compile
+time and has no overheads at runtime. Across our benchmarks, classifying a
+contract took 11.5 milliseconds on average.
 
@@
 classification, and the impact of summarization optimization.
-
-\subsubsection{Configuration}
 
@@
 and we allocate 8 CPUs for each shim layer node. The clients also run on
-\cf{c2.4xlarge} instances. We call this deployment \cf{1DC}. For our WAN
-experiments, we instantiate 2 clusters (each with 5 nodes), one in US-east (N.
-Virginia) and another in US-west (Oregon). The average latency between the
-regions was 85 ms. We call this deployment \cf{2DC}.
-
-\subsubsection{LWW Register}
+\cf{c2.4xlarge} instances. We call this \cf{1DC} configuration. For our
+geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with 5
+nodes, and place the clusters on US-east (Virginia) and US-west (Oregon). The
+average inter-region latency was 85ms.
 
@@
   \centering
-  \subfigure[Operation performance 1DC]{\label{grf:LWW-perf-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC.pdf}}
-  \subfigure[Operation performance 2DC]{\label{grf:LWW-perf-2DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-2DC.pdf}}
-  \subfigure[Transaction performance 1DC]{\label{grf:LWW-txn-1DC}\includegraphics[width=0.24\textwidth]{graphs/LWW-tp-vs-lat-1DC-txn.pdf}}
-  \subfigure[Mutations in transactions]{\label{grf:LWW-txn-wp}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn-wp-vs-tp.pdf}}
-	\caption{Performance of LWW Register RDT.}
+  \subfigure[Bank account latency]{\label{grf:BA-lat}\includegraphics[width=0.252\textwidth]{graphs/BA-lat.pdf}}
+  \subfigure[Bank account throughput]{\label{grf:BA-tp}\includegraphics[width=0.24\textwidth]{graphs/BA-tp.pdf}}
+  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
+  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
+	\caption{Quelea Performance.}
   \label{grf:LWW_perf}
@@
 
-We analyze the performance of different consistency classifications in \name
-using the LWW register. Our client workload was generated using YCSB
-benchmark~\cite{}, with the default workload of 50\% reads and 50\% writes,
-with keys uniformly chosen from a set of 100,000 keys. We varied the number of
-clients from 128 to 1024, with each workload running for 180 seconds.
-Last-writer-wins is the conflict resolution policy used by Cassandra. Hence, we
-are able to directly compare the performance of \name implementation against
-Cassandra.
+Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
+of operations in bank account example as we increase the number of clients in
+\cf{1DC} configuration. Our client workload was generated using YCSB
+benchmark~\cite{}. The benchmark uniformly chose from 100,000 keys, where the
+operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
+corresponds to the default 50:50 read:write mix in YCSB. Here, the lines marked
+EC and CC correspond to all operations being assigned EC and CC consistency
+levels. These levels compromise correctness as withdraw has to be and SC
+operation. The line SC corresponds to a configuration where all operations are
+strongly consistent; this ensures application correctness. \name corresponds to
+our implementation, which classifies operations based on their contracts.
 
-Figure~\ref{grf:LWW-perf-1DC} shows the throughput vs. latency graph for the
-different consistency levels in \cf{1DC} configuration. We also include
-Cassandra, which provides basic eventual consistency. Recall that \name's
-eventual consistency is stronger than the guarantees provided by Cassandra
-(Section~\ref{sec:store_sem}). With 512 clients, the average latency of
-operations on Cassandra was 32.24 milliseconds. EC and CC under \name was 30\%
-and 57\% slower whereas strong consistency was 241\% slower. EC(SC) was able to
-achieve 80\%(40\%) of throughput of vanilla Cassandra. The results show that
-while \name provides an expressive programming model with stronger consistency
-guarantees, its performance is comparable to vanilla Cassandra. However, SC has
-a huge performance hit due to coordination between all the replicas for
-obtaining object leases as well as reading and writing to all replicas. This
-overhead worsens in a WAN setting. Figure~\ref{grf:LWW-perf-2DC} shows the
-throughput vs. latency graph under 2DC configuration. While the Cassandra, EC
-and CC numbers largely remain the same when compared to 1DC configuration, SC
-is 6$\times$ to 12$\times$ higher latency and lower throughput than than the
-1DC case.
+Both \name and SC ensure correctness. However, with 512 clients, \name
+implementation was within 30\% of latency and 12\% of throughput of EC, whereas
+SC operations had 162\% higher latency and 52\% lower throughput than EC
+operations. In \cf{2DC} configuration (not-shown), the average latency of SC
+operations with 512 clients increased by 9.4$\times$ due to the cost of
+geo-distributed coordination, whereas \name operations were only 2.2$\times$
+slower, mainly due to the increased cost of \cf{withdraw} operations.
+Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
+the same. This illustrates the benefit of fine-grained contract classification
+in \name.
 
 We compare the performance of different transaction isolation level choices in
-Figure~\ref{grf:LWW-txn-1DC}. The numbers were obtained under 1DC
-configuration. The YCSB workload was modified to issue 10 operations per
-transaction. Each operation is assumed to have eventual consistency. \cf{NoTxn}
-corresponds to a configuration that does not use transactions. Compared to this
-RC is only 12\% shower in terms of latency with 512 clients, where as RR is
-2.3X slower. The difference between RC and \cf{NoTxn} is due to the meta-data
-overhead of recording transaction information in the object state. For RR
-transaction, The cost of capturing and maintaining the snapshot in an RR
-transaction is the biggest source of overhead.
+Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
+under 1DC configuration. The YCSB workload was modified to issue 10 operations
+per transaction, with the default 50:50 read:write mix. Each operation is
+assumed to have eventual consistency. \cf{NoTxn} corresponds to a configuration
+that does not use transactions. Compared to this RC is only 12\% shower in
+terms of latency with 512 clients, where as RR is 2.3X slower. The difference
+between RC and \cf{NoTxn} is due to the meta-data overhead of recording
+transaction information in the object state. For RR transaction, the cost of
+capturing and maintaining the snapshot in an RR transaction is the biggest
+source of overhead.
 
-Moreover, the size of the snapshot is proportional to the the number of
-different objects accessed by the session as well as the number of updates.
-Hence, with increasing proportion of writes, the overhead of RR transaction
-increases as shown in Figure~\ref{grf:LWW-txn-wp}. The numbers were obtained
-with 128 clients. Compared to an RC transaction whose throughput decreasess
-45\% as the write proportion increases from 0\% to 50\%, RR transaction's
-throughput drops 78\% in the same interval.
+We also compared the performance of EC LWW operations directly against
+Cassandra (our backing store), which uses last-writer-wins as the only
+convergence semantics. While Cassandra provides no stronger-than-eventual
+consistency properties, \name was within 30\%(20\%) of latency(throughput) of
+Cassandra with 512 clients, illustrating that the programmers only have to pay
+a minimal overhead for the expressive and stronger \name programming model.
 
-\subsection{Bank Account}
+Finally, we study the impact of summarization in
+Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
+replica, with all the clients operating on the \emph{same} LWW register to
+stress test the summarization mechanism. We utilize 50:50 read:write mix, and
+33:33:33 EC:CC:SC mix. The shim layer cache (mem) of operations is summarized
+every 64 updates, while the updates in the backing store (disk) are summarized
+every 4096 updates. Each point in the graph represents the average latency of
+the previous 1000 operations. Each experiment is run for 60s.
 
-\begin{figure}
-  \centering
-	\includegraphics[width=0.5\columnwidth]{Figures/BA-legend.pdf}
-  \subfigure[Latency 1DC]{\label{grf:BA-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-lat.pdf}}
-  \subfigure[Throughput 1DC]{\label{grf:BA-tp}\includegraphics[width=0.23\textwidth]{graphs/BA-tp.pdf}}
-	\caption{Performance of Bank Account RDT.}
-  \label{grf:BA_perf}
-\end{figure}
-
-Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
-of operations in bank account example as we increase the number of clients in
-\cf{1DC} configuration. The benchmark uniformly chose from 100,000 keys, where
-the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance. Here,
-the lines marked EC and CC correspond to all operations being assigned EC and
-CC consistency levels. These levels compromise correctness as withdraw has to
-be and SC operation. The line SC corresponds to a configuration where all
-operations are strongly consistent; this ensures application correctness. \name
-corresponds to our implementation, which classifies operations based on their
-contracts. At 512 clients, \name implementation was within 30\% of latency and
-12\% of throughput of EC, while providing correctness, whereas SC operations
-had 162\% higher latency and 52\% lower throughput than EC operations. This
-illustrates the benefit of fine-grained contract classification in \name.
+The results show that without summarization, the average latency of operations
+increase exponentially to almost 1 second, and only 13K operations were
+performed in a minute. Since every operation has to reduce over the set of all
+previous operations, with a ever growing set, the operations take increasingly
+more time to complete. With summarization only in memory, the performance still
+degrades due to the cost of fetching all previous updates from the backing
+store into the shim layer. Fetching the latest updates is essential for SC
+operations. With both summarizations enabled, we see that the latency does not
+increase over time, and we were able to perform 67K operations. This graph
+illustrates the importance and effectiveness of summarization in \name.

### 2014-11-03 Data/BA.py

@@ -49,10 +49,8 @@ def main ():
 			plt.ylabel ("Latency (ms)")
 			ymajorLocator = MultipleLocator(40)
 			plt.axes().yaxis.set_major_locator (ymajorLocator)
-			plt.axes().set_aspect(2.5)
 		else:
 			plt.ylabel ("Throughput (X1000 ops/s)")
-			plt.axes().set_aspect(50)
 		plt.grid (True)
 		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
 		matplotlib.rc('font', **font)
@@ -64,7 +62,10 @@ def main ():
 		plt.plot (scx, scy, nodeKind [3], label="SC", linewidth=2.0, ms=15)
 
 		plt.xlim(xmin = 0)
-		#plt.legend(bbox_to_anchor=(-0.1, 1.02, 1.2, .102), loc=3, ncol=4, mode="expand", borderaxespad=0.)
+		if (f=="BA-lat.dat"):
+			plt.legend(loc=2,ncol=2)
+		else:
+			plt.legend(loc=2, framealpha=0.5)
 		print ("saving fig " + f.replace("dat","pdf"))
 		plt.savefig (f.replace("dat","pdf"))
 		plt.close ()

### 2014-11-03 Data/GC.py

@@ -1,0 +1,65 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["GC.dat"]
+	for f in files:
+		Cassandra = []
+		EC = []
+		CC = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+
+			d = line.split ()
+			if (len(d) == 6):
+				Cassandra.append ((float(d[0])/10000,float (d[1])/100))
+				EC.append ((float(d[2])/10000,float (d[3])/100))
+				CC.append ((float(d[4])/10000,float (d[5])/100))
+			elif (len(d) == 4):
+				EC.append ((float(d[0])/10000,float (d[1])/100))
+				CC.append ((float(d[2])/10000,float (d[3])/100))
+			elif (len(d) == 2):
+				CC.append ((float(d[0])/10000,float (d[1])/100))
+
+
+		(cx,cy) = zip(*Cassandra)
+		(ecx,ecy) = zip(*EC)
+		(ccx,ccy) = zip(*CC)
+
+
+		plt.xlabel ("Operations (X 10,000)")
+		plt.ylabel ("Latency (X 100 ms)")
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		plt.plot (cx, cy, nodeKind [0], label="No Sum", linewidth=2.0, ms=10)
+		plt.plot (ecx, ecy, nodeKind [1], label="Mem Only", linewidth=2.0, ms=10)
+		plt.plot (ccx, ccy, nodeKind [2], label="Mem & Disk", linewidth=2.0, ms=10)
+		plt.xlim(xmin = 0)
+		plt.legend (bbox_to_anchor=(1, 0.75))
[…]

### 2014-11-03 evaluation.tex
edits

@@
 Both \name and SC ensure correctness. However, with 512 clients, \name
-implementation was within 30\% of latency and 12\% of throughput of EC, whereas
+implementation was within 41\% of latency and 18\% of throughput of EC, whereas
 SC operations had 162\% higher latency and 52\% lower throughput than EC

### 2014-11-04 evaluation.tex
Added RUBiS results

@@
 \begin{itemize}
-
+\setlength{\itemsep}{2pt}
 \item \textbf{LWW register}: A last-write-wins register that provides read and
@@
 
-\item \textbf{RuBis}: An ebay-like auction site~\cite{}. The application allows
+\item \textbf{RUBiS}: An ebay-like auction site~\cite{}. The application allows
 users to browse items, bid for items on sale, and pay for items from a wallet
@@
 user's timeline, userline, followers and following.
-
 \end{itemize}
@@
 {Online store} & 340 & 4 & 9 & 1 & 0 & 2 & 0 & 1 \\
-{Rubis} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
+{RUBiS} & 640 & 6 & 14 & 2 & 1 & 4 & 2 & 0 \\
 {Microblog} & 659 & 5 & 13 & 6 & 1 & 6 & 3 & 1 \\
@@
   \centering
-  \subfigure[Bank account latency]{\label{grf:BA-lat}\includegraphics[width=0.252\textwidth]{graphs/BA-lat.pdf}}
-  \subfigure[Bank account throughput]{\label{grf:BA-tp}\includegraphics[width=0.24\textwidth]{graphs/BA-tp.pdf}}
+  \subfigure[Bank account]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-tp-vs-lat.pdf}}
   \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
+  \subfigure[RUBiS bidding mix]{\label{grf:rubis}\includegraphics[width=0.24\textwidth]{graphs/Rubis.pdf}}
   \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
@@
 
-Figure~\ref{grf:BA-lat} and~\ref{grf:BA-tp} analyzes the latency and throughput
-of operations in bank account example as we increase the number of clients in
-\cf{1DC} configuration. Our client workload was generated using YCSB
-benchmark~\cite{}. The benchmark uniformly chose from 100,000 keys, where the
-operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
-corresponds to the default 50:50 read:write mix in YCSB. Here, the lines marked
-EC and CC correspond to all operations being assigned EC and CC consistency
-levels. These levels compromise correctness as withdraw has to be and SC
-operation. The line SC corresponds to a configuration where all operations are
-strongly consistent; this ensures application correctness. \name corresponds to
-our implementation, which classifies operations based on their contracts.
+Figure~\ref{grf:BA-tp-vs-lat} shows the throughtput vs. latency of operations
+in bank account example as we increase the number of clients in \cf{1DC}
+configuration. Our client workload was generated using YCSB benchmark~\cite{}.
+The benchmark uniformly chose from 100,000 keys, where the operation spread was
+25\% withdraw, 25\% deposit and 50\% getBalance, which corresponds to the
+default 50:50 read:write mix in YCSB. We increased the number of clients from
+128 to 1024, and each experiment ran for 180 seconds.
 
-Both \name and SC ensure correctness. However, with 512 clients, \name
-implementation was within 41\% of latency and 18\% of throughput of EC, whereas
-SC operations had 162\% higher latency and 52\% lower throughput than EC
-operations. In \cf{2DC} configuration (not-shown), the average latency of SC
-operations with 512 clients increased by 9.4$\times$ due to the cost of
-geo-distributed coordination, whereas \name operations were only 2.2$\times$
-slower, mainly due to the increased cost of \cf{withdraw} operations.
-Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
-the same. This illustrates the benefit of fine-grained contract classification
-in \name.
+The lines marked EC and CC correspond to all operations being assigned EC and
+CC consistency levels. These levels compromise correctness as withdraw has to
+be and SC operation. The line SC corresponds to a configuration where all
+operations are strongly consistent; this ensures application correctness. \name
+corresponds to our implementation, which classifies operations based on their
+contracts. Both \name and SC ensure correctness. However, with 512 clients,
+\name implementation was within 41\% of latency and 18\% of throughput of EC,
+whereas SC operations had 162\% higher latency and 52\% lower throughput than
+EC operations. Observe that there is a point in each case after which the
+latency increases while the throughput decreases. This indicates a point after
+which the store becomes saturated with client requests.
+
+In \cf{2DC} configuration (not-shown), the average latency of SC operations
+with 512 clients increased by 9.4$\times$ due to the cost of geo-distributed
+coordination, whereas \name operations were only 2.2$\times$ slower, mainly due
+to the increased cost of \cf{withdraw} operations. Importantly, the latency of
+\cf{getBalance} and \cf{deposit} remained almost the same. This illustrates the
+benefit of fine-grained contract classification in \name.
 
@@
 
-We also compared the performance of EC LWW operations directly against
-Cassandra (our backing store), which uses last-writer-wins as the only
+We also compared (not shown) the performance of EC LWW operations directly
+against Cassandra (our backing store), which uses last-writer-wins as the only
 convergence semantics. While Cassandra provides no stronger-than-eventual
@@
 a minimal overhead for the expressive and stronger \name programming model.
+
+Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
+configuration against a single replica (NoRep) and strongly replicated
+(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
+15\% read-write interactions, which is representative of the auction workload.
+Without replication, NoRep trivially provides strong consistency. However, this
+deployment does not scale beyond 1750 operations per second. Strong replication
+offers better throughput at the cost of greater latency due to inter-replica
+coordination. \name deployment offers the benefit of replication, while only
+paying the cost of coordination when necessary.
 

### 2014-11-04 Data/BA-tp-vs-lat.py

@@ -1,0 +1,62 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["BA-tp-vs-lat.dat"]
+	for f in files:
+		EC = []
+		CC = []
+		SC = []
+		Q = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+			d = line.split ()
+			EC.append ((float(d[0]),float (d[1])))
+			CC.append ((float(d[2]),float (d[3])))
+			SC.append ((float(d[4]),float (d[5])))
+			Q.append ((float(d[6]),float (d[7])))
+
+		(ecx,ecy) = zip(*EC)
+		(ccx,ccy) = zip(*CC)
+		(scx,scy) = zip(*SC)
+		(qx,qy) = zip(*Q)
+
+
+		plt.xlabel ("Throughput (X 1000 ops/s)")
+		plt.ylabel ("Latency (ms)")
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		plt.plot (ecx, ecy, nodeKind [0], label="EC", linewidth=2.0, ms=15)
+		plt.plot (ccx, ccy, nodeKind [1], label="CC", linewidth=2.0, ms=15)
+		plt.plot (scx, scy, nodeKind [2], label="SC", linewidth=2.0, ms=15)
+		plt.plot (qx, qy, nodeKind [3], label="Q", linewidth=2.0, ms=15)
+
+		plt.xlim(xmin = 0)
+		plt.legend (framealpha=0.5)
+		print ("saving fig " + f.replace("dat","pdf"))
+		plt.savefig (f.replace("dat","pdf"))
+		plt.close ()
+
+main ()

### 2014-11-04 Data/Rubis.py

@@ -1,0 +1,58 @@
+import subprocess
+import shlex
+import re
+from optparse import OptionParser
+import numpy
+import matplotlib
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from matplotlib.ticker import FormatStrFormatter
+import sys
+import os
+import fnmatch
+import locale
+
+nodeKind = ['o-', 's--', 'D-.', '^:', 'X-', 'V--', '>-.', '<:']
+
+matplotlib.rcParams['ps.fonttype'] = 42
+matplotlib.rcParams['pdf.fonttype'] = 42
+
+def main ():
+	files = ["Rubis.dat"]
+	for f in files:
+		Q = []
+		NoRep = []
+		StrRep = []
+
+		infile = open (f, "r")
+		for line in infile.readlines ():
+			if line == "\n" or line.startswith ("#"):
+				continue
+			d = line.split ()
+			Q.append ((float(d[0]),float (d[1])))
+			NoRep.append ((float(d[2]),float (d[3])))
+			StrRep.append ((float(d[4]),float (d[5])))
+
+		(qx,qy) = zip(*Q)
+		(nrx,nry) = zip(*NoRep)
+		(strx,stry) = zip(*StrRep)
+
+
+		plt.xlabel ("Throughput (X 1000 ops/s)")
+		plt.ylabel ("Latency (ms)")
+		plt.grid (True)
+		font = {'family' : 'normal', 'weight' : 'normal', 'size' : 20}
+		matplotlib.rc('font', **font)
+
+		plt.subplots_adjust (bottom=0.12)
+		plt.plot (qx, qy, nodeKind [0], label="Quelea", linewidth=2.0, ms=15)
+		plt.plot (nrx, nry, nodeKind [1], label="NoRep", linewidth=2.0, ms=15)
+		plt.plot (strx, stry, nodeKind [2], label="StrongRep", linewidth=2.0, ms=15)
+
+		plt.xlim(xmin = 0)
+		plt.legend ()
+		print ("saving fig " + f.replace("dat","pdf"))
+		plt.savefig (f.replace("dat","pdf"))
+		plt.close ()
+
+main ()

### 2014-11-05 contract-language.tex
Committing first experiments
Major addition to implementation
Edits throughout

@@
 \section{Contract Language}
-\label{sec:contract-lang}
+\label{sec:lang}
 
@@
 \subsection{Capturing Store Semantics}
+\label{sec:store_sem}
 
 An important aspect of our contract language is its ability to capture
-store-level consistency guarantees, along with application-level
-consistency requirements. Similar to~\cite{Burckhardt2014}, we can
-rigorously define a wide variety of store semantics including those
-that combine any subset of session and causality guarantees, and
-multiple consistency levels.  However, for our purposes, we identify
-three particular consistency levels -- eventual, causal, and strong,
-commonly offered by many distributed stores with tunable consistency,
-with increasing overhead in terms of latency and availability. For
-each of the these three consistency levels, we capture the semantics
-as a formula in our contract language, and informally describe the kind
-of application-level consistency requirements that can be met under the
+store-level consistency guarantees, along with application-level consistency
+requirements. Similar to~\cite{Burckhardt2014}, we can rigorously define a wide
+variety of store semantics including those that combine any subset of session
+and causality guarantees, and multiple consistency levels.  However, for our
+purposes, we identify three particular consistency levels -- eventual, causal,
+and strong, commonly offered by many distributed stores with tunable
+consistency, with increasing overhead in terms of latency and availability. For
+each of the these three consistency levels, we capture the semantics as a
+formula in our contract language, and informally describe the kind of
+application-level consistency requirements that can be met under the
 consistency level:
@@
 	does not require its action to manifest on all replicas before other
-  operations in its session are allowed to proceed. While eventually
-  consistent data stores typically offer \emph{basic} eventual consistency
-  with all possible anomalies, we assume that our store provides stronger
-  semantics that remain highly-available~\cite{BailisHAT,COPS}; the store
-  always exposes a causal cut of the updates. This semantics can be formally
-  captured in terms of the following contract definition:
+	operations in its session are allowed to proceed. While eventually consistent
+	data stores typically offer \emph{basic} eventual consistency with all
+	possible anomalies, we assume that our store provides stronger semantics that
+	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
+	causal cut of the updates. This semantics can be formally captured in terms
+	of the following contract definition:
 
@@
 	operations from the same session to be visible, which cannot be guaranteed
-	under just eventual consistency. We assume that causality is only tracked
-	through operations on the same object. The corresponding store semantics is
+	under just eventual consistency. The corresponding store semantics is
 	captured by the contract $\ccc$ defined below:
@@
   \begin{smathpar}
-  \ccc = \forall a.~(\hboZ \cap \sameobjZ) (a,\cureff) \Rightarrow \vis{a}{\cureff}
+  \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
   \end{smathpar}
@@
 
-% SJ: not sure this is necessary, or appropriate here.
-%% \noindent Observe that out contract language does not incorporate real (i.e.,
-%% wall-clock) time. Hence, it cannot describe store
-%% semantics such as recency or bounded-staleness
-%% guarantees offered by certain stores~\cite{Pileus}.
-
 % Contract classification rules
@@
 
-Our goal is to map application-level consistency constraints on operations
-to appropriate store-level consistency guarantees capable of satisfying
-these constraints.  The ability to express both these kinds of constraints
-as contracts in our contract language lets us compare and determine if
-contract ($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be
-satisfied under a store consistency level identified by the contract
-$\cv_{st}$. Towards this end, we define a binary \emph{weaker than} relation
-for our contract language as following:
+Our goal is to map application-level consistency constraints on operations to
+appropriate store-level consistency guarantees capable of satisfying these
+constraints.  The ability to express both these kinds of constraints as
+contracts in our contract language lets us compare and determine if contract
+($\cv_{op}$) of an operation ($\mathit{op}$) is weak enough to be satisfied
+under a store consistency level identified by the contract $\cv_{st}$. Towards
+this end, we define a binary \emph{weaker than} relation for our contract
+language as following:
 \begin{definition}
@@
 If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
-than} a store contract ($\cv_{st}$), then constraints expressed by the
-former are implied by guarantees provided by the latter.  The
-completeness of first-order logic\cite{completeness} allows us to assert
-that any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
-$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
-\cv_{op}$). Consequently, it is safe to execute operation $\mathit{op}$ under
-a store consistency level captured by $\cv_{st}$.
+than} a store contract ($\cv_{st}$), then constraints expressed by the former
+are implied by guarantees provided by the latter. The completeness of
+first-order logic allows us to assert that any well-formed execution ($\E$)
+that satisfies $\cv_{st}$ (i.e., $\E\models \cv_{st}$) also satisfies
+$\cv_{op}$ (i.e., $\E \models \cv_{op}$). Consequently, it is safe to execute
+operation $\mathit{op}$ under a store consistency level captured by $\cv_{st}$.
 
-Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves
-totally ordered with respect to the $\le$ relation: $\ecc \le \ccc \le
-\scc$.  This concurs with the intuition that any contract satisfiable
-under $\ecc$ or $\ccc$ is satisfiable under $\scc$, and any contract
-that is satisfiable under $\ecc$ is satisfiable under $\ccc$. However,
-we are interested in the \emph{weakest} guarantee (among $\ecc$,
-$\ccc$, and $\scc$) required to satisfy the contract. We define the
-corresponding consistency level as the \emph{consistency class of the
-contract}. The classification scheme, presented formally in
-Figure~\ref{sem:classify}, defines rules to judge the consistency class of
-a contact. For example, the scheme classifies the \cf{getBalance} contract
-($\cv_{gb}$) from Section~\ref{sec:motivation} as a {\sf
-CausallyConsistent} contract, because the sequent $\Delta \vdash \ccc
-\Rightarrow \cv_{gb}$ is valid in first-order logic (therefore,
-$\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
-\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$).
-Due to the confinement of our contract language to a decidable subset
-of the logic, validity of such sequents can be decided mechanically
-allowing us to automate the classification scheme in \name.
+Observe that the contracts $\scc$, $\ccc$ and $\ecc$ are themselves totally
+ordered with respect to the $\le$ relation: $\ecc \le \ccc \le \scc$.  This
+concurs with the intuition that any contract satisfiable under $\ecc$ or $\ccc$
+is satisfiable under $\scc$, and any contract that is satisfiable under $\ecc$
+is satisfiable under $\ccc$. However, we are interested in the \emph{weakest}
+guarantee (among $\ecc$, $\ccc$, and $\scc$) required to satisfy the contract.
+We define the corresponding consistency level as the \emph{consistency class of
+the contract}. The classification scheme, presented formally in
+Figure~\ref{sem:classify}, defines rules to judge the consistency class of a
+contact. For example, the scheme classifies the \cf{getBalance} contract
+($\cv_{gb}$) from \S~\ref{sec:motivation} as a {\sf CausallyConsistent}
+contract, because the sequent $\Delta \vdash \ccc \Rightarrow \cv_{gb}$ is
+valid in first-order logic (therefore, $\cv_{gb} \le \ccc$), whereas the
+sequent $\Delta \vdash \ecc \Rightarrow \cv_{gb}$ is invalid (therefore,
+$\cv_{gb} \not\le \ecc$). Due to the confinement of our contract language to a
+decidable subset of the logic, validity of such sequents can be decided
+mechanically allowing us to automate the classification scheme in \name.
 
-Along with three straightforward rules that classify contracts into
-consistency classes, the classification scheme also presents a rule
-that judges well-formedness of a contract. A contract is well-formed
-if and only if it is satisfiable under $\scc$ - the strongest possible
-consistency guarantee any store can provide. Otherwise, it is
-considered ill-formed, and rejected statically.
+Along with three straightforward rules that classify contracts into consistency
+classes, the classification scheme also presents a rule that judges
+well-formedness of a contract. A contract is well-formed if and only if it is
+satisfiable under $\scc$ - the strongest possible consistency guarantee any
+store can provide. Otherwise, it is considered ill-formed, and rejected
+statically.
 
@@
 define an axiomatic model of the our system described informally in
-Section~\ref{sec:sysmod}:
+\S~\ref{sec:sysmod}:
 
@@
 Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
-proof of the theorem is given in a tech report and is provided
-as part of the supplementary material~\cite{techrep}\footnote{The
-  supplementary material and tech report also defines an operational
-  semantics that concretizes the rewrite relation ($\xrightarrow{}$), along
-  with the proof that it correctly enforces all \name contracts.}.
+proof of the theorem is given in a tech report and is provided as part of the
+supplementary material~\cite{techrep}\footnote{The supplementary material and
+tech report also defines an operational semantics that concretizes the rewrite
+relation ($\xrightarrow{}$), along with the proof that it correctly enforces
+all \name contracts.}.

### 2014-11-05 evaluation.tex
Edits throughout

@@
 \section{Evaluation}
+\label{sec:results}
 
@@
 
-\item \textbf{RUBiS}: An ebay-like auction site~\cite{}. The application allows
-users to browse items, bid for items on sale, and pay for items from a wallet
-modelled after a bank account.
+\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The application
+allows users to browse items, bid for items on sale, and pay for items from a
+wallet modelled after a bank account.
 
 \item \textbf{Microblog}: A twitter-like microblogging site, modelled after
-Twissandra~\cite{}. The application allows adding a new user, adding and
-replying to tweets, following, unfollowing and blocking users, and fetching a
-user's timeline, userline, followers and following.
+Twissandra~\cite{Twissandra}. The application allows adding a new user, adding
+and replying to tweets, following, unfollowing and blocking users, and fetching
+a user's timeline, userline, followers and following.
 \end{itemize}
@@
 in bank account example as we increase the number of clients in \cf{1DC}
-configuration. Our client workload was generated using YCSB benchmark~\cite{}.
-The benchmark uniformly chose from 100,000 keys, where the operation spread was
-25\% withdraw, 25\% deposit and 50\% getBalance, which corresponds to the
-default 50:50 read:write mix in YCSB. We increased the number of clients from
-128 to 1024, and each experiment ran for 180 seconds.
+configuration. Our client workload was generated using YCSB
+benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
+the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
+corresponds to the default 50:50 read:write mix in YCSB. We increased the
+number of clients from 128 to 1024, and each experiment ran for 180 seconds.
 
@@
 replica, with all the clients operating on the \emph{same} LWW register to
-stress test the summarization mechanism. We utilize 50:50 read:write mix, and
-33:33:33 EC:CC:SC mix. The shim layer cache (mem) of operations is summarized
-every 64 updates, while the updates in the backing store (disk) are summarized
-every 4096 updates. Each point in the graph represents the average latency of
-the previous 1000 operations. Each experiment is run for 60s.
+stress test the summarization mechanism. The shim layer cache (mem) of
+operations is summarized every 64 updates, while the updates in the backing
+store (disk) are summarized every 4096 updates. Each point in the graph
+represents the average latency of the previous 1000 operations. Each experiment
+is run for 60s.
 

### 2014-11-05 implementation.tex
Major addition to implementation
Edits throughout

@@
 \section{Implementation}
+\label{sec:impl}
 
-\name is implemented as a shallow extension of GHC Haskell and runs on top
-of Cassandra, an off-the-shelf eventually consistent distributed data (or
-backing) store responsible for all data management issues (i.e.,
-replication, fault tolerance, availability, and convergence).  Template
-Haskell is used implement static contract classification, and proof
-obligations are discharged with the help of the Z3~\cite{} SMT
-solver. Figure~\ref{fig:impl_mod} describes the overall system architecture.
+\name is implemented as a shallow extension of GHC Haskell and runs on top of
+Cassandra, an off-the-shelf eventually consistent distributed data (or backing)
+store responsible for all data management issues (i.e., replication, fault
+tolerance, availability, and convergence).  Template Haskell is used implement
+static contract classification, and proof obligations are discharged with the
+help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod} illustrates the
+overall system architecture.
 
@@
 \begin{center}
-\includegraphics[width=0.9\columnwidth]{Figures/ImplModel}
+\includegraphics[width=0.7\columnwidth]{Figures/ImplModel}
 \end{center}

### 2014-11-05 intro.tex
edits
Edits throughout

@@
 non-uniform latency (i.e., availability) and network
-partitions~\cite{Brewer2000,Gilbert}. Indeed, modern web services, which aim
+partitions~\cite{Brewer2000,Gilbert2002}. Indeed, modern web services, which aim
 to provide an "always on" experience, overwhelmingly favour availability and
@@
 conflicting updates, and has to pay careful attention to avoid unwanted
-inconsistencies (e.g., negative balances in a bank account, or having an
-item appear in a shopping cart after it has been
-removed~\cite{Dynamo}). Oftentimes, the inconsistency leaks from the
-application and is witnessed by the user.  Ultimately, the developer must
-decide the consistency level appropriate for a particular operation; this is
-understandably an error-prone process requiring intricate knowledge of both
-the application as well as the semantics and implementation of the
-underlying data store, which typically have only informal
-descriptions~\cite{}.  Nonetheless, picking the correct consistency level is
-critical not only for correctness but also for scalability of the
-application. While choosing a weaker consistency level than required may
-introduce program errors and anomalies, choosing a stronger one than
-necessary can negatively impact program scalability.
-
-%--KC--
-% The choice of correct consistency level is critical to the correctness and
-% scalability of the application. Choosing a weaker consistency level than
-% required introduces anomalies, while choosing a stronger than necessary level
-% affects program scalability.
-%\KC{Do we include the LWW-Register number here to add emphasis to the fact that
-%choosing the correct consistency level is essential for program scalability}.
+inconsistencies (e.g., negative balances in a bank account, or having an item
+appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
+the inconsistency leaks from the application and is witnessed by the user.
+Ultimately, the developer must decide the consistency level appropriate for a
+particular operation; this is understandably an error-prone process requiring
+intricate knowledge of both the application as well as the semantics and
+implementation of the underlying data store, which typically have only informal
+descriptions. Nonetheless, picking the correct consistency level is critical
+not only for correctness but also for scalability of the application. While
+choosing a weaker consistency level than required may introduce program errors
+and anomalies, choosing a stronger one than necessary can negatively impact
+program scalability.
 
@@
 
-The rest of the paper is organized as follows. The next section describes
-the system model. Section~\ref{sec:motivation} describes the challenges in
-programming with eventually consistent data stores, and introduces
-\name\ contracts as a proposed solution to overcome these issues.
-Section~\ref{sec:lang} provides more details the contract language.
-Section~\ref{sec:opsem} presents an operational semantics.
-Section~\ref{sec:txns} introduces transaction contracts and classification.
-Section~\ref{sec:impl} describes the implementation and provides details 
-about the optimizations needed to make the system practical.
-Section~\ref{sec:results} discusses experimental evaluation.
-Section~\ref{sec:related} and~\ref{sec:concl} present related work and
-conclusions.
+The rest of the paper is organized as follows. The next section describes the
+system model.  We describe the challenges in programming with eventually
+consistent data stores, and introduces \name\ contracts as a proposed solution
+to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
+more details the contract language, and its mapping to the store consistency
+levels, along with metatheory for certifying the correctness of the mapping.
+\S~\ref{sec:txns} introduces transaction contracts and classification.
+\S~\ref{sec:impl} describes the implementation and provides details about the
+optimizations needed to make the system practical. \S~\ref{sec:results}
+discusses experimental evaluation. \S~\ref{sec:related} and~\ref{sec:concl}
+present related work and conclusions.

### 2014-11-05 related.tex
Edits throughout

\section{Related Work}
\label{sec:related}

Operation-based RDTs have been widely studied in terms of their algorithmic
properties~\cite{SSS,Burckhardt2014}, and several systems utilize this model to
construct distributed data structures~\cite{Cassandra,Bayou,Tango}. These
systems typically propose to implement the datatypes directly over a cluster of
nodes, and only focus on basic eventual consistency. Hence, these systems
implement custom solutions for durability and fault-tolerance. \name realizes
RDTs stronger consistency models on top of off-the-shelf eventually consistent
distributed stores. In this respect, \name is similar to~\cite{BoltOn} where
causal consistency is achieved through a shim layer on Cassandra, which
explicitly tracks and enforces depedencies between updates.
However,~\cite{BoltOn} does not support user-defined RDTs, automatic contract
classification and transactions.

Since eventual consistency alone is insufficient to build correct applications,
several systems~\cite{Bayou,Pileus,RedBlue} propose a lattice of stronger
consistency levels. Similarly, traditional database processing
systems~\cite{Berenson95} and their replicated variants~\cite{BailisHAT}
propose weaker isolation levels for performance. In these systems, the onus is
on the developer to choose the correct consistency(isolation) level for
operations(transactions). \name relieves the developer of this burden, and
instead expects contracts expressing declarative visibility requirements.

Our contract language is inspired by the axiomatic decription of RDT semantics
proposed by~\cite{Burckhardt2014}. While they use axioms for formal
verification of correctness of an RDT implementation, we utilize them as a
means for the user to express the desired consistency guarantees in the
application. Similar to their work, our contract language does not incorporate
real (i.e., wall-clock) time. Hence, it cannot describe store semantics such as
recency or bounded-staleness guarantees offered by certain
stores~\cite{Pileus}.

Program analysis has been proposed to relieve the annotation burden associated
with consistency classification.~\cite{Calm} identifies code regions that
require injection of coordination.~\cite{Sieve} automatically classifies
operations into strong and eventual consistency.~\cite{IConfluence} defines the
conditions under which an operation requires coordination. These works focus of
coarse-grained classification (eventual or strong), and do not focus on
transaction consistency. Indeed, we do see these approaches as a next step in
\name to automatically infer the contracts.

### 2014-11-05 transactions.tex
Edits throughout

@@
 \section{Transaction Contracts}
+\label{sec:txns}
 
 While contracts on individual operations offer the programmer object-level
-declarative reasoning, real-world scenarios often involve operations that
-span multiple objects. In order to address this problem, several recent
-systems~\cite{COPS,BurckhardtESOP,BailisHAT} have proposed a variety of
+declarative reasoning, real-world scenarios often involve operations that span
+multiple objects. In order to address this problem, several recent
+systems~\cite{Walter,Burckhardt2012,BailisHAT} have proposed a variety of
 transactions, with varying semantics, in order to compose operations on
 multiple objects. However, given that classical transaction models such as
-serializability~\cite{} and snapshot isolation~\cite{} require inter-replica
-coordination, these systems espouse \emph{coordination-free transactions}
-that remain available under network partitions, but only provide weaker
-isolation guarantees. Coordination-free transactions have intricate
-consistency semantics and widely varying runtime overheads. As with
-operation-level consistency, the onus is on the programmer to pick the
-correct transaction kind; this selection is complicated by operation
-consistency requirements.
+serializability and snapshot isolation require inter-replica coordination,
+these systems espouse \emph{coordination-free transactions} that remain
+available under network partitions, but only provide weaker isolation
+guarantees. Coordination-free transactions have intricate consistency semantics
+and widely varying runtime overheads. As with operation-level consistency, the
+onus is on the programmer to pick the correct transaction kind; this selection
+is complicated by operation consistency requirements.
 
@@
 
-\noindent \begin{minipage}[t]{0.5\columnwidth}
+\noindent \begin{minipage}[t]{0.53\columnwidth}
 \begin{codehaskell}
@@
 \end{minipage}
-\begin{minipage}[t]{0.5\columnwidth}
+\begin{minipage}[t]{0.47\columnwidth}
 \begin{codehaskell}
@@
 identify three well-understood coordination-free transaction semantics -- Read
-Committed (RC)~\cite{}, Monotonic Atomic View (MAV)~\cite{} and Repeatable Read
-(RR)~\cite{}, and illustrate the classification strategy. Our technique can
-indeed be applied to a different isolation level lattice.
+Committed (RC)~\cite{Berenson95}, Monotonic Atomic View (MAV)~\cite{BailisHAT}
+and Repeatable Read (RR)~\cite{Berenson95}, and illustrate the classification
+strategy. Our technique can indeed be applied to a different isolation level
+lattice.
 

### 2014-11-05 concl.tex
edits

\section{Conclusions}
\label{sec:concl}

In this paper, we have presented \name a shallow Haskell extension for
declarative programming over eventually consistent data stores. The key idea of
\name is the automatic classification of fine-grained consistency contracts on
operations and distributed transactions with respect to the consistency and
isolation levels offered by the store. Our contract language is carefully
crafted from a decidable subset of first-order logic, enabling the use of
automatic theorem prover to discharge the proof obligations associated with
contract classification. We provide meta-theory that certifies the correctness
of contract classification. We realize an instantiation of \name on top of
off-the-shelf distributed store, Cassandra, and illustrate the benefit of
fine-grained contract classification by implementing and evaluating several
scalable applications.

### 2014-11-05 related.tex
edits

@@
 causal consistency is achieved through a shim layer on Cassandra, which
-explicitly tracks and enforces depedencies between updates.
+explicitly tracks and enforces dependencies between updates.
 However,~\cite{BoltOn} does not support user-defined RDTs, automatic contract
@@
 
-Our contract language is inspired by the axiomatic decription of RDT semantics
+Our contract language is inspired by the axiomatic description of RDT semantics
 proposed by~\cite{Burckhardt2014}. While they use axioms for formal

### 2014-11-05 intro.tex
edits

@@
 
-Many real-world web services --- such as those built and maintained by
-Amazon, Facebook, Google, Twitter, etc. --- replicate application state and
-logic across multiple \emph{replicas} within and across data
-centers. Replication is intended not only to improve application throughput
-and reduce user-perceived latency, but also to tolerate partial failures
-without compromising overall service availability. Traditionally programmers
-have relied on \emph{strong consistency} guarantees such as
-linerarizability~\cite{Herlihy1990} or
-serializability~\cite{Serializability} in order to build correct
-applications.  While strong consistency is an easily stated property, it
-masks the reality underlying large-scale distributed systems with respect to
-non-uniform latency (i.e., availability) and network
-partitions~\cite{Brewer2000,Gilbert2002}. Indeed, modern web services, which aim
-to provide an "always on" experience, overwhelmingly favour availability and
-partition tolerance over strong consistency. To this end, several \emph{weak
-  consistency} models such as eventual consistency, causal consistency,
-session guarantees, and timeline consistency have been proposed.
+Many real-world web services --- such as those built and maintained by Amazon,
+Facebook, Google, Twitter, etc. --- replicate application state and logic
+across multiple \emph{replicas} within and across data centers. Replication is
+intended not only to improve application throughput and reduce user-perceived
+latency, but also to tolerate partial failures without compromising overall
+service availability. Traditionally programmers have relied on \emph{strong
+consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
+serializability~\cite{Serializability} in order to build correct applications.
+While strong consistency is an easily stated property, it masks the reality
+underlying large-scale distributed systems with respect to non-uniform latency,
+availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
+modern web services, which aim to provide an "always on" experience,
+overwhelmingly favour availability and partition tolerance over strong
+consistency. To this end, several \emph{weak consistency} models such as
+eventual consistency, causal consistency, session guarantees, and timeline
+consistency have been proposed.
 
@@
 \setlength{\itemsep}{2pt}
-\item We introduce \name, a shallow extension of Haskell that supports the description
-  and validation of replicated data types found in programs intended to
-  execute on eventually consistent stores. Contracts are used to specify
-  fine-grained application-level consistency properties, and are analyzed to
-  assign the most efficient and sound store consistency level to the
-  corresponding operation.
+\item We introduce \name, a shallow extension of Haskell that supports the
+	description and validation of replicated data types found on eventually
+	consistent stores. Contracts are used to specify fine-grained
+	application-level consistency properties, and are analyzed to assign the most
+	efficient and sound store consistency level to the corresponding operation.
 \item \name supports coordination-free transactions over arbitrary datatypes.
@@
 	guarantees, and utilize the contract enforcement system to automatically
-	assign the correct and most efficient isolation level for a transaction.
+	assign the correct isolation level for a transaction.
 \item We provide metatheory that certifies the soundness of our contract
@@
 \item An implementation of \name as a transparent shim layer over
-	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.  An
-	extensive experimental evaluation over a collection of real-world
-	applications, including a Twitter-like micro-blogging site and an eBay-like
-	auction site illustrates the practicality of our approach.
+	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
+	Experimental evaluation over a set of real-world applications, including a
+	Twitter-like micro-blogging site and an eBay-like auction site illustrates
+	the practicality of our approach.
 \end{itemize}

### 2014-11-05 model.tex
edits

@@
 \includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
-\caption{Quelea system model.}
+\caption{\name system model.}
 \label{fig:sysmod}
@@
 an operation is applied to; the data store may choose to route the operation to
-the closest replica for minimizing latency, or route it to a different replica
-in order to balance the load in the system. For example, the operations
-\emph{foo} and \emph{bar} invoked by the same session on the same object, might
-end up being applied to different replicas because replica 1 (to which
-\emph{foo} was applied) might be unreachable when the client invokes
+any replica in order to minimize latency, balance load, etc. For example, the
+operations \emph{foo} and \emph{bar} invoked by the same session on the same
+object, might end up being applied to different replicas because replica 1 (to
+which \emph{foo} was applied) might be unreachable when the client invokes
 \emph{bar}.
@@
 over a \emph{snapshot} of the state of the object on which it is invoked. In
-this case, we say that the effect $w^x_4$ \emph{witnesses} $w^x_1$ and $w^x_2$.
-Alternatively, $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$, written
-logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where $\visZ$ is
-the visibility relation between effects. Visibility is an irreflexive (an
-operation cannot see its own effect) and asymmetric relation, and only relates
-effects produced by operations on the same object. Executing a read-only
-operation is similar except that no new effects are produced.
+this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
+written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
+$\visZ$ is the visibility relation between effects. Visibility is an
+irreflexive and asymmetric relation, and only relates effects produced by
+operations on the same object. Executing a read-only operation is similar
+except that no new effects are produced.
 
 The effect added to a particular replica is asynchronously sent to other
-replicas. We assume that \emph{eventually} an effect applied at one replica
-will be merged into all other replicas. Two effects $w^x_4$ and $w^x_5$ that
-arise from the same session are said to be in \emph{session order} (written
-logically as $\so{w^x_4}{w^x_5}$). Session order is an irreflexive, transitive
-relation. The effects $w^x_4$ and $w^x_5$ arising from operations applied to
-the same object $x$ are said to be under the \emph{same object} relation,
-written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate every effect with
-the operation that generated the effect with the help of a relation $\operZ$.
-In the current example, $\oper{w^x_4}{foo}$ and $\oper{w^x_5}{bar}$ hold. For
-simplicity, we assume all operation names across all object types are distinct.
+replicas, and eventually merged into all other replicas. Two effects $w^x_4$
+and $w^x_5$ that arise from the same session are said to be in \emph{session
+order} (written logically as $\so{w^x_4}{w^x_5}$). Session order is an
+irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
+operations applied to the same object $x$ are said to be under the \emph{same
+object} relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate
+every effect with the operation that generated the effect with the help of a
+relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
+$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names across
+all object types are distinct.
 

### 2014-11-05 evaluation.tex
Grammar
Edits to related work.

@@
 associated with contract classfication is discarged through the Z3 SMT Solver.
-In \name, the contract classification process is completely performed compile
-time and has no overheads at runtime. Across our benchmarks, classifying a
-contract took 11.5 milliseconds on average.
+In \name, the contract classification process is completely performed
+at compile time and has no overheads at runtime. Across our
+benchmarks, classifying a contract took 11.5 milliseconds on average.
 
@@
 of 5 fully replicated Cassandra replicas within the same datacenter. We
-instantiate one shim layer node for every cassandra replica, an place it on the
-same VM as the cassandra replica. The clients are instantiated on the same data
-center as the store, and run the transactions. We deploy the each node in the
+instantiate one shim layer node for every cassandra replica, and place it on the
+same VM as the cassandra replica. Clients are instantiated on the same data
+center as the store, and run transactions. We deploy the each node in the
 cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,

### 2014-11-05 related.tex
Edits to related work.

@@
 
-Program analysis has been proposed to relieve the annotation burden associated
-with consistency classification.~\cite{Calm} identifies code regions that
-require injection of coordination.~\cite{Sieve} automatically classifies
-operations into strong and eventual consistency.~\cite{IConfluence} defines the
-conditions under which an operation requires coordination. These works focus of
-coarse-grained classification (eventual or strong), and do not focus on
-transaction consistency. Indeed, we do see these approaches as a next step in
-\name to automatically infer the contracts.
+Several conditions have been proposed to judge whether an operation on a
+replicated data object needs coordination or not. ~\cite{Calm} defines
+\emph{logical monotonicity} as a sufficient condition for coordination freedom,
+and proposes a consistency analysis that marks code regions performing
+non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as potential
+coordination points.  ~\cite{IConfluence} and ~\cite{Sieve} define
+\emph{invariant confluence} and \emph{invariant safety}, respectively, as
+conditions for safely executing an operation without coordination. ~\cite{Sieve}
+also proposes a program analysis that conservatively marks operations as
+\emph{blue} (coordination not required), while marking the remaining as
+\emph{red} (coordination required). Unlike \name, these works focus on a
+coarse-grained classification of consistency as eventual or strong, and do not
+focus on transaction isolation levels. However, program analyses they propose
+relieve programmers of the burden to tag operations with consistency levels.
+Indeed, we do consider automatic inference of consistency contracts from
+application-specific integrity constraints as the next step for \name.

### 2014-11-05 motivation.tex
Major addition to implementation
Edits throughout

@@
 eventually consistent data store, with the \emph{integrity} constraint that the
-balance should never drop below 0. We begin by implementing a bank
-account replicated data type (RDT) in \name, and then describe the mechanisms
-to obtain the desired correctness guarantees.
+balance must be non-negative. We begin by implementing a bank account
+replicated data type (RDT) in \name, and then describe the mechanisms to obtain
+the desired correctness guarantees.
 
@@
 A key novelty in \name is that it allows the addition of new RDTs to the store,
-greatly improving the programmability of data stores. The programmer no
-longer has to coerce the application logic to utilize the store provided data
-types. In addition, \name treats the convergence semantics of the data type
-separately from the consistency properties. This separation of concerns allows
-the programmer to \emph{operationally} reason about conflict resolution, while
-permitting \emph{declarative} reasoning for consistency. As mentioned in the
-previous section, RDTs are described in terms of operations on a RDT object,
-where each operation is described as a reduction over the state of the object
-in terms of a list of effects. Let us assume that the bank account object provides
-three operations: \cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the
-assumption that the withdraw fails if the account has insufficient balance.
-Every operation in \name is of the following type, written in Haskell syntax:
+which obviates the need for coercing the application logic to utilize the store
+provided data types. In addition, \name treats the convergence semantics of the
+data type separately from its consistency properties. This separation of
+concerns permits \emph{operational} reasoning for conflict resolution, and
+\emph{declarative} reasoning for consistency. The combination of these
+techniques enhances the programmability of the store.
+
+Let us assume that the bank account object provides three operations:
+\cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the assumption that the
+withdraw fails if the account has insufficient balance. Every operation in
+\name is of the following type, written in Haskell syntax:
 
 \begin{codehaskell}
-type Operation = (Effect e, Arg a, Res r)
-               => [e] -> a -> (r, Maybe e)
+type Operation = [e] -> a -> (r, Maybe e)
 \end{codehaskell}
@@
 \noindent It takes a list of effects (the \emph{context} for the operation),
-and an input argument, and returns a result along with an optional effect. Read-only
-operations do not return any effect. If an effect is returned, it is added to
-the state of the object at the current replica, and asynchronously sent to
-other replicas. The implementation of the bank account operations in \name is
-given in Figure~\ref{fig:ex}:
+and an input argument, and returns a result along with an optional effect
+(read-only operations return \cf{Nothing}). The new effect (if emitted) is
+added to the state of the object at the current replica, and asynchronously
+sent to other replicas. The implementation of the bank account operations in
+\name is given in Figure~\ref{fig:ex}:
 
@@
 \end{codehaskell}
-\label{Definition of a bank account expressed in Quelea.}
+\caption{Definition of a bank account expressed in Quelea.}
 \label{fig:ex}
@@
 Our goal is to choose the correct consistency level for each of the bank
-account operations such that (1) the balance never drops below zero and (2) the
-\cf{getBalance} operation never incorrectly returns a negative balance. To be
-able to do this, we need to consider the anomalies that could arise under
-eventual consistency.
+account operations such that (1) the balance remains non-negative and (2) the
+\cf{getBalance} operation never incorrectly returns a negative balance. Let us
+first consider the anomalies that could arise under eventual consistency.
 
@@
 
-It is easy to see that the only way to ensure the integrity constraint on
-\cf{balance} is to prevent concurrent withdrawals on the same bank account
-object. This can be done by insisting that \cf{withdraw} be executed with
-strong consistency properties. Although this prevents the integrity constraint
-from being violated, a strongly consistent \cf{withdraw} itself is not
-sufficient to prevent \cf{getBalance} from ever reporting a negative balance to
-the end user.
+It is easy to see that preventing concurrent \cf{withdraw} operations
+eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
+executed as a strongly consistent operation. Despite this strengthening,
+\cf{getBalance} operation may incorrectly report a negative balance to the end
+user. Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
+which consists of three concurrent sessions performing a \cf{deposit}, a
+\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
+account object. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in
+session 2, witnesses the effects of \cf{deposit(100)} from session 1, concludes
+that there is sufficient balance, and completes successfully. However, the
+\cf{getBalance} operation may only witness this successful withdraw, but not
+the causally preceding \cf{dopsit}, and reports the balance of negative 50 to
+the user.
 
-Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly}, which
-consists of three concurrent sessions performing a \cf{deposit}, a
-\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
-account object. As before, assume that the initial balance in the account is
-zero. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in session 2,
-can witnesses the effects of \cf{deposit(100)} from session 1, leading it to
-rightfully conclude that the current balance is 50. Consequently,
-\cf{withdraw(50)} is performed safely. The effect of this \cf{withdraw}
-operation subsequently becomes visible to the \cf{getBalance} in Session 3.
-However, because \cf{getBalance} only witnesses the effect of the \cf{withdraw}
-operation from session 2, but not the \cf{deposit} from session 1, reports the
-balance of negative 50 to the user.
-
-Producing a negative balance is not the only inconsistency that the user is
-exposed to under an eventually consistent bank account. Figure
-~\ref{fig:missingUpdateAnomaly} shows an execution where the \cf{getBalance}
-operation in a session does not witness the effects of an earlier \cf{withdraw}
-operation performed in the same session, possibly because it was served by a
-replica that has not yet merged the \cf{withdraw} effect, served by a different
- replica. This anomaly leads the user to incorrectly conclude that the
-\cf{withdraw} operation failed to go through.
+Under eventual consistency, the users may also be exposed to other forms of
+inconsistencies. Figure~\ref{fig:missingUpdateAnomaly} shows an execution where
+the \cf{getBalance} operation in a session does not witness the effects of an
+earlier \cf{withdraw} operation performed in the same session, possibly because
+it was served by a replica that has not yet merged the \cf{withdraw} effect.
+This anomaly leads the user to incorrectly conclude that the \cf{withdraw}
+operation failed to go through.
 
 Although it is easy to understand the reasons behind the occurrence of the
-aforementioned negative balance and missing update anomalies, finding the
-appropriate fixes is not readily apparent. Making \cf{getBalance} a strongly
-consistent operation is definitely sufficient to avert anomalies, but is it
-really necessary? Given the cost of enforcing strong
-consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid it unless
-there are no viable alternatives.  Exploring the space of these alternatives
-requires understanding the subtle differences in semantics of various kinds
-of weak consistency alternatives.
+aforementioned anomalies, finding the appropriate fixes is not readily
+apparent. Making \cf{getBalance} a strongly consistent operation is definitely
+sufficient to avert anomalies, but is it really necessary? Given the cost of
+enforcing strong consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid
+it unless there are no viable alternatives. Exploring the space of these
+alternatives requires understanding the subtle differences in semantics of
+various kinds of weak consistency alternatives.
 
-\subsection{Specifications and Contracts}
+\subsection{Contracts}
 
@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}) that axiomatically
-specify the set of allowed executions involving this operation. In the case of
-the bank account, any execution that does not exhibit the anomalies
-described in the previous section is a \emph{well-formed} execution on the bank
-account object.  By specifying the set of legal executions for each data type in
-terms of a trace of operation invocations on that type, \name\ ensures that all
-executions over that type are well-formed.
+constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
+exposes the contract construction language as a Haskell library}) that
+axiomatically specify the set of allowed executions involving this operation.
+In the case of the bank account, any execution that does not exhibit the
+anomalies described in the previous section is a \emph{well-formed} execution
+on the bank account object. By specifying the set of legal executions for each
+data type in terms of a trace of operation invocations on that type, \name\
+ensures that all executions over that type are well-formed.
 
@@
 given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. Utilizing the primitive relations described in
-Section~\ref{sec:sysmod}, we express this application-level consistency
-requirement as a contract or specification ($\cv_w$) over \cf{withdraw}:
+vice versa. We express this application-level consistency requirement as a
+contract ($\cv_w$) over \cf{withdraw}:
 
@@
 
-\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$ is an
-effect that was emitted either by a \cf{deposit} or a \cf{withdraw} operation.
-If a \cf{withdraw} $b$ is visible to \cf{getBalance}, then all \cf{deposit}
-operations $a$ visible to the \cf{withdraw} should also be visible to
-\cf{getBalance}. This prevents negative balance anomalies.
+\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
+is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
+operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
+all \cf{deposit} operations $a$ visible to $b$ should also be visible to
+$\cureff$. This prevents negative balance anomalies. Our contract language
+provides operators to compose relations. The syntax $(R_1 \cap R_2)(a,b)$ is
+equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the above contract
+says that if a \cf{deposit} or a \cf{withdraw} operation precedes a
+\cf{getBalance} operation in session order, and is applied on the same object
+as the \cf{getBalance} operation, then it must be the case that the
+\cf{getBalance} operation witnesses the effects of the preceding operations.
 
-Our specification language provides operators to compose relations.  Thus,
-the syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge
-R_2(a,b)$. The last line of the above contract says that if a \cf{deposit}
-or a \cf{withdraw} operation precedes a \cf{getBalance} operation in session
-order, and is applied on the same object as the \cf{getBalance} operation,
-then it must be the case that the \cf{getBalance} operation witnesses the
-effects of the preceding operations. Since there are no restrictions on when
-or how a \cf{deposit} operation can execute, its contract is simply $\small
-\true$.
+Finally, since there are no restrictions on when or how a \cf{deposit}
+operation can execute, its contract is simply $\small \true$.
 
@@
 The mapping of application-level consistency requirements to appropriate
-store-level guarantees is done automatically behind-the-scene.
+store-level guarantees is done automatically behind-the-scene. How might one go
+about ensuring that an execution adheres to a contract? The challenge is that a
+contract provides a declarative (axiomatic) specification of an execution,
+while what is required is an operational procedure for \emph{enforcing} its
+implicit constraints.
 
-How might one go about ensuring that an execution adheres to a contract? The
-challenge is that a contract provides a declarative (axiomatic)
-specification of an execution, while what is required is is an operational
-procedure for \emph{enforcing} its implicit constraints. One strategy would
-be to execute operations speculatively.  Here, operations are tentatively
-applied as they are received from the client or other replicas. We can
-maintain a runtime manifestation of executions, and check well-formedness
-conditions at runtime, rolling back executions if they are
+One strategy would be to execute operations speculatively.  Here, operations
+are tentatively applied as they are received from the client or other replicas.
+We can maintain a runtime manifestation of executions, and check
+well-formedness conditions at runtime, rolling back executions if they are
 ill-formed. However, the overhead of state maintenance and the complexity of
-user-defined contracts is likely to make this technique infeasible in
-practice.
+user-defined contracts is likely to make this technique infeasible in practice.
 

### 2014-11-05 contract-language.tex
edits

@@
 
-% Contract Language Syntax
-% ------------------------
 \input{contract-syntax.tex}
-
-In this section, we formalize \name's contract language, and describe our
-contract classification scheme, which analyzes a contract and maps it to the
-weakest store consistency level sufficient to satisfy its consistency
-requirements.
-
-% classifies contracts on the basis of the weakest
-% store-level consistency guarantee
 
@@
 \EffSoup & \in & \mathtt{EffSoup}	  & \coloneqq & \set{\eff} \\
-\visZ, \soZ, &	\in & \mathtt{Relations} & \coloneqq &
-  \set{\eff}\times\set{\eff} \\
-\sameobjZ		&     &  & \\
+\visZ, \soZ, \sameobjZ &	\in & \mathtt{Relations} & \coloneqq & \EffSoup \times \EffSoup \\
+%\sameobjZ		&     &  & \\
 {\E} 		& \in & \mathtt{ExecState}  & \coloneqq & \Exec \\
@@
 \name contracts are constraints over axiomatic definitions of program
-executions.Fig.~\ref{sem:contracts} summarizes artifacts relevant to
-define an axiomatic execution. We formalize an axiomatic execution as
-a tuple $\Exec$, where $\EffSoup$, called the \emph{effect soup}, is
-the set of all effects generated during the program execution, and
-$\visZ,\soZ,\sameobjZ \subseteq \EffSoup \times \EffSoup$ are
-\emph{visibility}, \emph{session order}, and \emph{same object}
-relations, respectively, witnessed over generated effects at run-time.
+executions.Fig.~\ref{sem:contracts} summarizes artifacts relevant to define an
+axiomatic execution. We formalize an axiomatic execution as a tuple $\Exec$,
+where $\EffSoup$, called the \emph{effect soup}, is the set of all effects
+generated during the program execution, and $\visZ,\soZ,\sameobjZ \subseteq
+\EffSoup \times \EffSoup$ are \emph{visibility}, \emph{session order}, and
+\emph{same object} relations, respectively, witnessed over generated effects at
+run-time.
 
@@
 and strong, commonly offered by many distributed stores with tunable
-consistency, with increasing overhead in terms of latency and availability. For
-each of the these three consistency levels, we capture the semantics as a
-formula in our contract language, and informally describe the kind of
-application-level consistency requirements that can be met under the
-consistency level:
-
-% Indeed, the techniques presented here can be
-% extended to other consistency stratifications. We assign each
-% application-level contract into one of these following classes:
+consistency, with increasing overhead in terms of latency and availability.
 
@@
 \item \textbf{Eventually consistency}: Eventually consistent operations can
-	be satisfied as long as the client can reach at least one replica. For
-	example, \cf{deposit} is an eventually consistent operation; its semantics
-	does not require its action to manifest on all replicas before other
-	operations in its session are allowed to proceed. While eventually consistent
-	data stores typically offer \emph{basic} eventual consistency with all
-	possible anomalies, we assume that our store provides stronger semantics that
-	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
-	causal cut of the updates. This semantics can be formally captured in terms
-	of the following contract definition:
+	be satisfied as long as the client can reach at least one replica. In the
+	bank account example, \cf{deposit} is an eventually consistent operation.
+	While eventually consistent data stores typically offer \emph{basic} eventual
+	consistency with all possible anomalies, we assume that our store provides
+	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
+	store always exposes a \emph{causal cut} of the updates. This semantics can
+	be formally captured in terms of the following contract definition:
 
@@
   \end{smathpar}
-  \noindent where $\small \hboZ = ((\soZ \cap \sameobjZ) \cup \visZ)^+$.
 
@@
 	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
-	until the effect of $o_1$ is merged with $o_2$ in both $r_1$ and $r_2$. The
-	\cf{getBalance} operation requires causal consistency, as it requires the
-	operations from the same session to be visible, which cannot be guaranteed
-	under just eventual consistency. The corresponding store semantics is
-	captured by the contract $\ccc$ defined below:
+	until the effect of $o_1$ is included in $r_2$. The \cf{getBalance} operation
+	requires causal consistency, as it requires the operations from the same
+	session to be visible, which cannot be guaranteed under eventual consistency.
+	The corresponding store semantics is captured by the contract $\ccc$ defined
+	below:
 
@@
   \end{smathpar}
-
-  %%; two operations in the same session but on different objects are
-  %% considered causally unrelated under this definition. Stores typically
-  %%avoid tracking causality across objects to mitigate overheads when
-  %% causality tracking is unnecessary.
 
@@
 \vspace{-2em}
-\noindent The $\Delta$ referred in the above defintion is a
-conjunction of assumptions about the nature of primitive relations,
-e.g., $\Rvis$ and $\Rso$ are irreflexive, and the happens-before
-relation $\hbZ$ is acyclic. A \emph{well-formed} axiomatic execution
-($\E$) is expected to satisfy these assumptions (i.e., $\E \models
-\Delta$).
+\noindent The $\Delta$ referred in the above defintion is a conjunction of
+assumptions about the nature of primitive relations, e.g., $\Rvis$ and $\Rso$
+are irreflexive, and the happens-before relation $\hbZ$ is acyclic. A
+\emph{well-formed} axiomatic execution ($\E$) is expected to satisfy these
+assumptions (i.e., $\E \models \Delta$).
 
@@
 is satisfiable under $\scc$, and any contract that is satisfiable under $\ecc$
-is satisfiable under $\ccc$. However, we are interested in the \emph{weakest}
-guarantee (among $\ecc$, $\ccc$, and $\scc$) required to satisfy the contract.
-We define the corresponding consistency level as the \emph{consistency class of
-the contract}. The classification scheme, presented formally in
-Figure~\ref{sem:classify}, defines rules to judge the consistency class of a
-contact. For example, the scheme classifies the \cf{getBalance} contract
-($\cv_{gb}$) from \S~\ref{sec:motivation} as a {\sf CausallyConsistent}
-contract, because the sequent $\Delta \vdash \ccc \Rightarrow \cv_{gb}$ is
-valid in first-order logic (therefore, $\cv_{gb} \le \ccc$), whereas the
-sequent $\Delta \vdash \ecc \Rightarrow \cv_{gb}$ is invalid (therefore,
-$\cv_{gb} \not\le \ecc$). Due to the confinement of our contract language to a
-decidable subset of the logic, validity of such sequents can be decided
-mechanically allowing us to automate the classification scheme in \name.
+is satisfiable under $\ccc$. We are interested in the \emph{weakest} guarantee
+(among $\ecc$, $\ccc$, and $\scc$) required to satisfy the contract. We define
+the corresponding consistency level as the \emph{consistency class} of the
+contract.
+
+The classification scheme, presented formally in Figure~\ref{sem:classify},
+defines rules to judge the consistency class of a contact. For example, the
+scheme classifies the \cf{getBalance} contract ($\cv_{gb}$) from
+\S~\ref{sec:motivation} as a {\sf\small CausallyConsistent} contract, because
+the sequent $\Delta \vdash \ccc \Rightarrow \cv_{gb}$ is valid in first-order
+logic (therefore, $\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
+\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$). Since we
+confine of our contract language to a decidable subset of the logic, validity
+of such sequents can be decided mechanically allowing us to automate the
+classification scheme in \name.
 
@@
 well-formedness of a contract. A contract is well-formed if and only if it is
-satisfiable under $\scc$ - the strongest possible consistency guarantee any
-store can provide. Otherwise, it is considered ill-formed, and rejected
+satisfiable under $\scc$ - the strongest possible consistency guarantee that
+the store can provide. Otherwise, it is considered ill-formed, and rejected
 statically.
@@
 
+\vspace{-1em}
 \begin{smathpar}
 \begin{array}{lclcl}
-\multicolumn{5}{l}{
+\multicolumn{5}{c}{
   {op} \in \mathtt{Operation} \qquad
@@
   \langle s,{\sigma} \rangle \pll \Sigma \ALT \emptyset \\
-				&			&	\mathtt{Store}		  & \coloneqq & \E;\Sigma \\
+				&			&	\mathtt{Config}		  & \coloneqq & \E;\Sigma \\
 \end{array}
@@
 
-We model the system as a tuple $\E;\Sigma$, where the execution $\E$
-captures the data store's current state of the execution, and
-$\Sigma$, the session soup, is the set of concurrent client sessions
-interacting with the store. A session is modeled simply as a sequence
-of replicated data type operations ($\mathit{op}$), tagged with the consistency
-class ($\tau$) of their contracts (as determined by the contract
-classification scheme). We assume a rewrite relation of form:
+We model the system as a tuple $\E;\Sigma$, where the axiomatic execution $\E$
+captures the data store's current state, and session soup $\Sigma$ is the
+set of concurrent client sessions interacting with the store. A session is
+modeled simply as a sequence of replicated data type operations
+($\mathit{op}$), tagged with the consistency class ($\tau$) of their contracts
+(as determined by the contract classification scheme). We assume a rewrite
+relation of form:
 

### 2014-11-05 contract-language.tex
edits

@@
 
-The syntax of our core contract language is shown in Fig.
+The syntax of our core contract language is shown in Figure
 ~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
@@
 \name contracts are constraints over axiomatic definitions of program
-executions.Fig.~\ref{sem:contracts} summarizes artifacts relevant to define an
-axiomatic execution. We formalize an axiomatic execution as a tuple $\Exec$,
+executions. Figure~\ref{sem:contracts} summarizes artifacts relevant to define
+an axiomatic execution. We formalize an axiomatic execution as a tuple $\Exec$,
 where $\EffSoup$, called the \emph{effect soup}, is the set of all effects
@@
 \begin{array}{lclcl}
-\multicolumn{5}{c}{
-  {op} \in \mathtt{Operation} \qquad
-  \tau \in \mathtt{Consistency\; Class}
-}\\
-\cv(\tau) & \in & \mathtt{Store\; Contract\; of \; \tau} & \coloneqq & \scc,
-  \ccc, \ecc\\
-{\sigma} & \in & \mathtt{Session} & \coloneqq & \cdot \ALT (op,\tau); \sigma \\
-\Sigma 	& \in & \mathtt{Session\;Soup}   & \coloneqq &
-  \langle s,{\sigma} \rangle \pll \Sigma \ALT \emptyset \\
-				&			&	\mathtt{Config}		  & \coloneqq & \E;\Sigma \\
+{\it op} 	& \in & \mathtt{Operation} \\
+{\tau}		& \in & \mathtt{Consistency Class} 	& \coloneqq & {\sf ec},{\sf cc},{\sf sc} \\
+{\sigma} 	& \in & \mathtt{Session} 					 	& \coloneqq & \cdot \ALT \langle op,\tau \rangle; \sigma \\
+\Sigma 		& \in & \mathtt{Session\;Soup}   	 	& \coloneqq & \sigma \pll \Sigma \ALT \emptyset \\
+					&			&	\mathtt{Config}		  			 	& \coloneqq & \E,\Sigma \\
 \end{array}
@@
 
-We model the system as a tuple $\E;\Sigma$, where the axiomatic execution $\E$
-captures the data store's current state, and session soup $\Sigma$ is the
-set of concurrent client sessions interacting with the store. A session is
-modeled simply as a sequence of replicated data type operations
-($\mathit{op}$), tagged with the consistency class ($\tau$) of their contracts
-(as determined by the contract classification scheme). We assume a rewrite
-relation of form:
+We model the system as a tuple $\E,\Sigma$, where the axiomatic execution $\E$
+captures the data store's current state, and session soup $\Sigma$ is the set
+of concurrent client sessions interacting with the store. A session $\sigma$ is
+a sequence of pairs composed of replicated data type operations $\mathit{op}$,
+tagged with the consistency class $\tau$ of their contracts (as determined by
+the contract classification scheme). We assume a rewrite relation of form:
 
@@
 \begin{smathpar}
-  \auxred{} {\E;\langle (op,\tau);\sigma \rangle \pll \Sigma} {\eff}
-    {\E';\langle \sigma \rangle \pll \Sigma'}
+  \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E',\sigma \pll \Sigma}
 \end{smathpar}
 
-\noindent on the system state. The relation captures the progress of
-the execution (from $\E$ to $\E'$)  due to the successful completion
-of a client operation ($\mathit{op}$) from one of the sessions in $\Sigma$,
-generating a new effect ($\eff$). If the resultant execution ($\E'$)
-satisfies the store contract ($\cv(\tau)$) of $\tau$ (i.e., $\E
-\models \cv(\tau)$), then we say that the store has \emph{enforced} the
-contract $\cv(\tau)$ in the execution $\E'$.
-
-With help of the axiomatic model, we now state the soundness of
-contract enforcement as follows:
+\noindent on the system state. The relation captures the progress of the
+execution (from $\E$ to $\E'$)  due to the successful completion of a client
+operation $\mathit{op}$ from one of the sessions in $\Sigma$, generating a new
+effect $\eff$. If the resultant execution $\E'$ satisfies the store contract
+$\cv_\tau$ (i.e., $\E \models \cv_\tau$), then we say that the store has
+\emph{enforced} the contract $\cv_\tau$ in the execution $\E'$. With help of
+the axiomatic model, we now state the soundness of contract enforcement as
+follows:
 
@@
 \label{lem:core-preservation}
-Let $\cv$ be a well-formed contract of a replicated data type operation $\mathit{op}$,
-and let $\tau$ denote the consistency class of $\cv$ as determined by
-the contract classification scheme. For all well-formed execution
+Let $\cv$ be a well-formed contract of a replicated data type operation
+$\mathit{op}$, and let $\tau$ denote the consistency class of $\cv$ as
+determined by the contract classification scheme. For all well-formed execution
 states $\E$, $\E'$ such that
-$\auxred{} {\E;\langle (op,\tau);\,\sigma \rangle \pll \Sigma} {\eff}
- {\E';\langle \sigma \rangle \pll \Sigma}$,
-if $\E' \models [\eff/\cureff]\,\cv(\tau)$, then $\E' \models [\eff/\cureff]\,\cv$
+$\auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff} {\E', \sigma
+\pll \Sigma}$, if $\E' \models \cv_\tau[\eff/\cureff]$, then $\E' \models
+\cv[\eff/\cureff]$
 \end{theorem}
@@
 Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
-proof of the theorem is given in a tech report and is provided as part of the
-supplementary material~\cite{techrep}\footnote{The supplementary material and
-tech report also defines an operational semantics that concretizes the rewrite
-relation ($\xrightarrow{}$), along with the proof that it correctly enforces
-all \name contracts.}.
+proof of the theorem is given in the supplementary material\footnote{The
+supplementary material also defines an operational semantics that concretizes
+the rewrite relation ($\xrightarrow{}$), along with the proof that it correctly
+enforces all \name contracts.}.

### 2014-11-05 core-lang.tex
Edits
edits

@@
 
-Fig.~\ref{sem:oper} summarizes the artifacts needed. We formalize an
-abstract model of a replicated data store as a tuple $\E;\Sigma$, where $\E$
-denotes the execution state, and $\Sigma$, the session soup, is the set of
-concurrent client sessions interacting with the store. The execution state
-($\E$) of a data store is modeled as a tuple $\Exec$ where $\EffSoup$,
-called the \emph{effect soup}, is a set of effects generated during the
-execution, and $\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are
-visibility and session order relations, respectively, witnessed over
-generated effects. A session is modeled simply as a sequence of replicated
-data type operations ($op$), tagged with the consistency class ($\tau$) of
-their contracts (as determined by the contract classification scheme). Our
-contract language and classification scheme is independent of store
-semantics. Accordingly, we avoid making any specific assumptions about the
-semantics of the store. However, we do assume the existence of a reduction
-relation of the form:
+Figure~\ref{sem:oper} summarizes the artifacts needed. We formalize an abstract
+model of a replicated data store as a tuple $\E;\Sigma$, where $\E$ denotes the
+execution state, and $\Sigma$, the session soup, is the set of concurrent
+client sessions interacting with the store. The execution state ($\E$) of a
+data store is modeled as a tuple $\Exec$ where $\EffSoup$, called the
+\emph{effect soup}, is a set of effects generated during the execution, and
+$\visZ,\soZ \subseteq \EffSoup \times \EffSoup$ are visibility and session
+order relations, respectively, witnessed over generated effects. A session is
+modeled simply as a sequence of replicated data type operations ($op$), tagged
+with the consistency class ($\tau$) of their contracts (as determined by the
+contract classification scheme). Our contract language and classification
+scheme is independent of store semantics. Accordingly, we avoid making any
+specific assumptions about the semantics of the store. However, we do assume
+the existence of a reduction relation of the form:
 

### 2014-11-05 motivation.tex
edits

@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
+constraints as \emph{contracts} (Figure~\ref{fig:contract-lang}\footnote{\name
 exposes the contract construction language as a Haskell library}) that

### 2014-11-05 implementation.tex
edits

@@
 perspective, leveraging an off-the-shelf data store enables an implementation
-comprising roughly only 2500 lines of Haskell code, packaged as a
-library\footnote{url to the hackage}.
-
-%% Complex business logic is implemented with the help of
-%% transactions on the web-server side, with clients interacting with the
-%% web-service over REST interface.
+comprising roughly only 2500 lines of Haskell code, which is packaged as a
+library.
 
@@
 
-\noindent Since the semantics of summarization depends on the semantics of the data
-type, our programming model provides a summarization function for each RDT:
-\begin{codehaskell}
-summarize :: Effect e => [e] -> [e]
-\end{codehaskell}
-\noindent with the intention that the length of the result is smaller that the
-length of the argument. We utilize the \cf{summarize} function to summarize the
-object state both in the shim layer node and the backing store, typically when
-the number of effects on an object crosses a tunable threshold. Shim layer
-summarization is straight-forward; a summarization thread takes the local lock
-on the object, and replaces its state with the summarized state. The shim layer
-node remains unavailable for that particular object during summarization
-(usually a few milliseconds).
+Since the semantics of summarization depends on the semantics of the data type,
+our programming model provides a summarization function (\cf{summarize}) for
+each RDT of type \cf{[e] -> [e]} with the intention that the length of the
+result is smaller that the length of the argument. We utilize the
+\cf{summarize} function to summarize the object state both in the shim layer
+node and the backing store, typically when the number of effects on an object
+crosses a tunable threshold. Shim layer summarization is straight-forward; a
+summarization thread takes the local lock on the object, and replaces its state
+with the summarized state. The shim layer node remains unavailable for that
+particular object during summarization (usually a few milliseconds).
 

### 2014-11-05 transactions.tex
edits

@@
 multiple objects. In order to address this problem, several recent
-systems~\cite{Walter,Burckhardt2012,BailisHAT} have proposed a variety of
-transactions, with varying semantics, in order to compose operations on
-multiple objects. However, given that classical transaction models such as
-serializability and snapshot isolation require inter-replica coordination,
-these systems espouse \emph{coordination-free transactions} that remain
-available under network partitions, but only provide weaker isolation
-guarantees. Coordination-free transactions have intricate consistency semantics
-and widely varying runtime overheads. As with operation-level consistency, the
-onus is on the programmer to pick the correct transaction kind; this selection
-is complicated by operation consistency requirements.
+systems~\cite{Walter,Burckhardt2012,BailisHAT} have proposed eventually
+consistent transactions in order to compose operations on multiple objects.
+However, given that classical transaction models such as serializability and
+snapshot isolation require inter-replica coordination, these systems espouse
+\emph{coordination-free transactions} that remain available under network
+partitions, but only provide weaker isolation guarantees. Coordination-free
+transactions have intricate consistency semantics and widely varying runtime
+overheads. As with operation-level consistency, the onus is on the programmer
+to pick the correct transaction kind. This choice is further is complicated by
+consistency semantics of individual operations.
 
@@
 We assume that operations not part of any transaction belong to their own
-unique transaction. While transactions may have varying isolation
-guarantees, we make the standard assumption that all transactions provide
-atomicity. Hence, we include the following axiom in $\Delta$: \[\small \forall
-a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a} \Rightarrow
-\vis{c}{a}\] \noindent The semantics of this contract is illustrated in
+unique transaction. While transactions may have varying isolation guarantees,
+we make the standard assumption that all transactions provide atomicity. Hence,
+we include the following axiom in $\Delta$:
+
+\vspace{-1.5em}
+\begin{smathpar}
+\forall a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a}
+\Rightarrow \vis{c}{a}
+\end{smathpar}
+
+\noindent The semantics of this contract is illustrated in
 Figure~\ref{fig:txn_atomicity}.
@@
 save amt =
-  x <- dollar(classify $\psi_1$)
+  x <- dollar(classify $\cv_{sv}$)
   atomically x dollar do
@@
 totalBalance =
-  x <- dollar(classify $\psi_2$)
+  x <- dollar(classify $\cv_{tb}$)
   atomically x dollar do
@@
 
-\noindent $\cv_1$ and $\cv_2$ are the contracts on the corresponding
+\noindent $\cv_{sv}$ and $\cv_{tb}$ are the contracts on the corresponding
 transactions. The function \cf{classify} assigns the contracts
@@
 While making both transactions serializable would ensure correctness,
-distributed stores rarely offer serializable transactions. Moreover,
-serializability is unavailable and hinders scalability. As we will see, these
+distributed stores rarely offer serializable transactions since it is
+unavailable and hinders scalability~\cite{BailisHAT}. As we will see, these
 transactions can be satisfied with much weaker isolation guarantees. Despite
@@
 operations to witness the same set of \cf{save} transactions, we define the
-constraint on \cf{totalBalance} transaction $\psi_2$ as:
+constraint on \cf{totalBalance} transaction $\psi_{tb}$ as:
 
-\vspace{-1em}
+\vspace{-1.2em}
 \begin{smathpar}
 \begin{array}{l}
-\cv_{2} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, \\
-\qquad (c:\rcf{withdraw} \vee \rcf{deposit}), (c:\rcf{withdraw} \vee \rcf{deposit}). \\
+\cv_{tb} = \forall a:\rcf{getBalance}, b:\rcf{getBalance}, \\
+\qquad (c:\rcf{withdraw} \vee \rcf{deposit}), (d:\rcf{withdraw} \vee \rcf{deposit}). \\
 \qquad \txn{a,b}{c,d} ~\wedge~ \vis{c}{a} ~\wedge~ \sameobj{d}{b} \Rightarrow \vis{d}{b}
@@
 \noindent The key idea in the above definition is that the $\txnZ$ primitive
-allows us to relate operations on different objects.
-
-The \cf{save} transaction only needs to ensure that the two writes it performs
-are made visible atomically. Since this is ensured by combining them in a
-transaction, \cf{save} does not require any additional constraints, and
-$\psi_1$ is simply $\small \true$.
+allows us to relate operations on different objects. The \cf{save} transaction
+only needs to ensure that the two writes it performs are made visible
+atomically. Since this is ensured by combining them in a transaction, \cf{save}
+does not require any additional constraints, and $\psi_1$ is simply $\small
+\true$.
 
@@
 buffered updates are made visible. Otherwise, the buffered updates are
-discarded. In particular, a store implementing RC does not require
-inter-replica coordination. RC does not entail any further isolation
-guarantees. We can express RC as follows:
+discarded. RC does not entail any further isolation guarantees. Hence, a store
+implementing RC does not require inter-replica coordination. We can express RC
+as follows:
 
@@
 \begin{smathpar}
-\rcc = \forall a,b,c.~\txn{a}{b,c} \wedge \sameobj{b}{c} \wedge \vis{b}{a} \Rightarrow \vis{c}{a}
+\begin{array}{l}
+\rcc = \forall a,b,c.~\txn{a}{b,c} \wedge \sameobj{b}{c} \\
+\qquad \qquad \qquad \wedge~ \vis{b}{a} \Rightarrow \vis{c}{a}
+\end{array}
 \end{smathpar}
@@
 
-RR semantics requires that the transaction witness a snapshot of the data store
-state. Importantly, this snapshot can be obtained from any replica, and hence
-RR is coordination-free. An example for such a transaction is the
+ANSI RR semantics requires that the transaction witness a snapshot of the data
+store state. Importantly, this snapshot can be obtained from any replica, and
+hence RR is coordination-free. An example for such a transaction is the
 \cf{totalBalance} transaction. The semantics of RR is captured by the following

### 2014-11-06 evaluation.tex
edits

@@
 
-In this section, we evaluate \name programs and report their contract profile
-as well as illustrating the performance benefits of fine-grained consistency
-classification on operations and transactions. We also report on the impact of
-the summarization. We implemented the following applications, which includes
-RDTs as well as larger applications composed of severals RDTs:
+In this section, we evaluate \name programs, report their contract profile and
+illustrate the performance benefits of fine-grained consistency classification
+on operations and transactions. We also evaluate on the impact of the
+summarization. We implemented the following applications, which includes
+individual RDTs as well as larger applications composed of several RDTs:
 
@@
 \item \textbf{Online store}: Models an online store with shopping cart and
-dynamically changing item prices. Checkout process verifies that the custormer
+dynamically changing item prices. Checkout process verifies that the customer
 only pays the accepted price.
@@
 \end{center} }
-\caption{The distribution of classfied contracts. \#T refers to the number of
+\caption{The distribution of classified contracts. \#T refers to the number of
 tables in the application. The columns 4-6 (7-9) represent operations
@@
 individual fields in the tables. Transactions are mainly used to consistently
-modify and access related fields across tables. The proof obligations
-associated with contract classfication is discarged through the Z3 SMT Solver.
-In \name, the contract classification process is completely performed
-at compile time and has no overheads at runtime. Across our
-benchmarks, classifying a contract took 11.5 milliseconds on average.
+modify and access related fields across tables. In \name, the contract
+classification process is completely performed at compile time and has no
+overheads at runtime. The proof obligations associated with contract
+classification is discharged through the Z3 SMT Solver. Across our benchmarks,
+classifying a contract took 11.5 milliseconds on average.
 
-\subsection{Performance}
-
-In this section, we study the performance benefit of fine-grained contract
-classification, and the impact of summarization optimization.
-
-We deploy \name applications in \emph{clusters} where each cluster is composed
-of 5 fully replicated Cassandra replicas within the same datacenter. We
-instantiate one shim layer node for every cassandra replica, and place it on the
-same VM as the cassandra replica. Clients are instantiated on the same data
-center as the store, and run transactions. We deploy the each node in the
-cluster cluster on \cf{c3.4xlarge} Amazon EC2 instances with 16 virtual CPUs,
-30GB memory, and 320GB of SSD storage. Our shim layer nodes are multi-threaded,
-and we allocate 8 CPUs for each shim layer node. The clients also run on
-\cf{c2.4xlarge} instances. We call this \cf{1DC} configuration. For our
-geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with 5
-nodes, and place the clusters on US-east (Virginia) and US-west (Oregon). The
-average inter-region latency was 85ms.
+For our performance evaluation, we deploy \name applications in \emph{clusters},
+where each cluster is composed of 5 fully replicated Cassandra replicas within
+the same datacenter. We instantiate one shim layer node for every Cassandra
+replica, and place it on the same VM as the Cassandra replica. Clients are
+instantiated on the same data center as the store, and run transactions. We
+deploy the each node in the cluster on \cf{c3.4xlarge} Amazon EC2 instances.
+Our shim layer nodes are multi-threaded, and we allocate 8 CPUs (out of 16
+available) for each shim layer node. The clients also run on \cf{c2.4xlarge}
+instances. We call this \cf{1DC} configuration. For our geo-distributed
+experiments (\cf{2DC}), we instantiate 2 clusters, each with 5 nodes, and place
+the clusters on US-east (Virginia) and US-west (Oregon). The average
+inter-region latency was 85ms.
 
@@
 
-Figure~\ref{grf:BA-tp-vs-lat} shows the throughtput vs. latency of operations
+Figure~\ref{grf:BA-tp-vs-lat} shows the throughput vs. latency of operations
 in bank account example as we increase the number of clients in \cf{1DC}

### 2014-11-06 implementation.tex
edits

@@
 
-The shim layer maintains a causally consistent in-memory snapshot of a subset
-of objects in the backing store, by explicitly tracking dependencies introduced
-between effects due to visibility, session and same transaction relations.
-Dependence tracking is similar to the techniques presented in~\cite{BoltOn}
+Each new effect in \name is realized as a row insertion in Cassandra, and the
+state of an object is the set of all corresponding rows. The shim layer
+maintains a causally consistent in-memory snapshot of a subset of objects in
+the backing store, by explicitly tracking dependencies introduced between
+effects due to visibility, session and same transaction relations. Dependence
+tracking is similar to the techniques presented in~\cite{BoltOn}
 and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
 tolerance, each shim layer node simply acts as a soft-state cache, do not
-communicate directly with each other, and can safely be terminated at any point.
-Similarly, new shim layer nodes can be spawned on demand.
+communicate directly with each other, and can safely be terminated at any
+point. Similarly, new shim layer nodes can be spawned on demand.
 
-The shim layer nodes periodically fetch updates from the store for eventually
-consistent operations, and on-demand for causally consistent and strongly
-consistent operations. Strongly consistent operations are performed after
-obtaining exclusive leases on objects. The lease mechanism is implemented with
-the help of Cassandra's support for conditional updates and expiring columns.
-
-Cassandra does not provide general-purpose transactions. Since the transaction
-guarantees provided by \name are coordination-free~\cite{BailisHAT}, we realize
-efficient implementations by explicitly tracking the dependencies between
-operations and transactions. Importantly, the weaker isolation semantics of
-transactions in \name permit transactions to be discharged if at least one shim
-layer node is reachable.
+The shim layer nodes periodically fetch updates from the backing store for
+eventually consistent operations, and on-demand for causally consistent and
+strongly consistent operations. Strongly consistent operations are performed
+after obtaining exclusive leases on objects. The lease mechanism is implemented
+with the help of Cassandra's support for conditional updates and expiring
+columns. Cassandra does not provide general-purpose transactions. Since the
+transaction guarantees provided by \name are
+coordination-free~\cite{BailisHAT}, we realize efficient implementations by
+explicitly tracking the dependencies between operations and transactions.
+Importantly, the weaker isolation semantics of transactions in \name permit
+transactions to be discharged if at least one shim layer node is reachable.
 
@@
 
-The main challenge in realizing an efficient implementation of
-operation-based replicated data types is that the state of the object
-defined in terms of its set of effects, grows with every effectful
-operation. If left unchecked, performance slow down over time, until the
-shim layer memory or backing store disk runs out of memory. Luckily, the
-state of the operation-based replicated data type can often be summarized to
-an \emph{observably equivalent} smaller state. For example,
+The main challenge in realizing an efficient implementation of operation-based
+replicated data types is that the state of the object defined in terms of its
+set of effects grows with every effectful operation. If left unchecked,
+performance slow down over time, until the shim layer memory or backing store
+disk runs out of memory. Luckily, the state of the operation-based replicated
+data type can often be summarized to an \emph{observably equivalent} smaller
+state. For example, a last-writer-wins register with multiple updates can be
+summarized to an equivalent register with only the last write. Similarly, a
+bank account with a series of deposits and updates with balance $b$ is
+equivalent to a bank account with a single deposit of $b$. A set with a
+collection of add and remove operations is equivalent to a set with a series of
+addition of live elements from the original set.
 
-\begin{itemize}
-\setlength{\itemsep}{2pt}
-\item A last-writer-wins register with multiple updates where $v$ is the value
-of the last write is observably equivalent to a register with a single write
-$v$.
+Since the semantics of summarization is specific to each data type, \name
+expects a summarization function (\cf{summarize}) of type \cf{[e]->[e]} for
+each RDT, where the result is smaller than the argument but is observably
+equivalent. We utilize the \cf{summarize} function to summarize the object
+state both in the shim layer node and the backing store, typically when the
+number of effects on an object crosses a tunable threshold. Shim layer
+summarization is straight-forward; a summarization thread takes the local lock
+on the object, and replaces its state with the summarized state. The shim layer
+node remains unavailable for that particular object during summarization
+(usually a few milliseconds).
 
-\item A bank account with a series of deposits and withdraws with current
-balance $b$ is equivalent to a bank account with a single deposit of $b$.
-
-\item A set with collection of add and remove operations is equivalent to a set
-with a series of add operations of live elements from the original set.
-\end{itemize}
-
-Since the semantics of summarization depends on the semantics of the data type,
-our programming model provides a summarization function (\cf{summarize}) for
-each RDT of type \cf{[e] -> [e]} with the intention that the length of the
-result is smaller that the length of the argument. We utilize the
-\cf{summarize} function to summarize the object state both in the shim layer
-node and the backing store, typically when the number of effects on an object
-crosses a tunable threshold. Shim layer summarization is straight-forward; a
-summarization thread takes the local lock on the object, and replaces its state
-with the summarized state. The shim layer node remains unavailable for that
-particular object during summarization (usually a few milliseconds).
-
-Performing summarization in the backing store is more complicated since the whole
+Performing summarization in the backing store is more complicated; the whole
 process needs to be atomic from a client's perspective, but Cassandra does not
-provide multi-row transactions. We have implemented a highly available
-implementation of summarization that permits concurrent client operations, but
-nonetheless prohibits these operations from witnessing intermediate states of
-the summarization process.
+provide multi-row transactions. We have engineered and implemented a scalable
+summarization mechanism for the backing that permits concurrent client
+operations, but nonetheless prohibits these operations from witnessing
+intermediate states of the summarization process.

### 2014-11-06 appendix.tex
Commiting appendix

\section{Appendix}
\label{sec:appendix}

\subsection{Soundness of Contract Classification}

We restate theorem ~\ref{thm:classification-sound} and provide its proof
below:

\begin{theorem}[Soundness of Contract Enforcement]
\label{lem:classification-sound}
Let $\cv$ be a well-formed contract of a replicated data type
operation $\mathit{op}$, and let $\tau$ denote the consistency class
of $\cv$ as determined by the contract classification scheme. For all
well-formed execution states $\E$, $\E'$ such that $\auxred{}
{\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff} {\E', \sigma
\pll \Sigma}$, if $\E' \models \cv_\tau[\eff/\cureff]$, then $\E'
\models \cv[\eff/\cureff]$
\end{theorem}
\begin{proof}
  Hypothesis:
  \begin{smathpar}
  \begin{array}{cr}
    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
    {\E', \sigma \pll \Sigma} & H\npp\\
    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
  \end{array}
  \end{smathpar}
  Since $\tau$ is the contract class of $\cv$, by inversion, we have
  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
  \end{array}
  \end{smathpar}
   Since $\eff$ denotes new effect, it is a fresh variable that does
   not occur free in $\Delta$. From $H2$, after instantiating bound
   $\cureff$ with $\eff$, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff] 
      & H\npp\\
  \end{array}
  \end{smathpar}
  Due to the soundness of natural deduction for first-order logic,
  $H3$ implies that for all models $\mathcal{M}$ such that
  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
  \begin{smathpar}
  \begin{array}{cr}
    \E' \models \Delta & H\npp\\
  \end{array}
  \end{smathpar}
  Proof follows from $H1$ and $H4$.
  \qed
\end{proof}

\subsection{Axiomatic Semantics}
\renewcommand{\auxred}[4]{#1 \vdash #2 \;\xhookrightarrow{#3}\; #4 }

We now describe axiomatic semantics of a data store that implements
strong, causal and eventual consistency guarantees. The semantics also
serves as an elegant representation of our implementation of \name.

Figure ~\ref{sem:oper} presents our semantics as concretization of the
rewrite relation ($\xrightarrow{}$) over execution state. Since we now have a
concrete store, we extend our system model with $\Theta$, a representation of
the store as a map from replicas to their local states. The local state of a
replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
visible at $r$.  An operation $op$ performed at replica $r$ can only witness
the set of effects ($\Theta(r)$) visible at $r$. For the sake of clarity, we
only consider a single replicated object of well-defined type (for eg: a
replicated object of type \cf{BankAccount}) in our formalization.  Our
semantics are parametric over the specification of this replicated data type.
Figure~\ref{sem:oper} formalizes replicated data type (RDT) specification as
tuple $(\delta,\Ops,\Ctrts)$, where $\delta$ is the data type, $\Ops$ maps
labels ($op$) of operations on $\delta$ to their definitions, while $\Ctrts$
maps them to their consistency contracts ($\cv$). The definition of an
operation is expected to be a lambda expression, although we do not enforce
this in our formalization. For technical reasons, we tag each session with a
session identifier ($s$) and the sequence number ($i$) of the next operation
in the session.

The state of an axiomatic execution ($\E$) is a tuple $\Exec$ where $\EffSoup$
is a set of effects, and $\visZ,\soZ, \sameobjZ \subseteq \EffSoup \times
\EffSoup$ are \emph{visibility}, \emph{session order}, and \emph{same object}
relations over effects, respectively. We concretize an effect ($\eff$) as a
tuple $(s,i,op,v)$, which records the fact that $i^{th}$ action in session
with {\sf SessID} $s$, which is an operation $op$ on the replicated object,
has been successfully executed on some replica yielding a return value $v$.
Note that the combination of $s$ and $i$ uniquely identifies the effect.
Session order relation ($\soZ$) relates effects generated by the same session.
An effect $\eff = (s,i,op,v)$ is said to precede another effect $\eff' =
(s',i',op',v')$ in session order if and only if $s'=s$ and $i'\ge i$. Since we
only consider one replicated object in our formalization, the $\sameobjZ$
relation relates every pair of effects in the effect soup ($\EffSoup$). An
effect generated at a replica becomes visible at rest of the replicas
eventually.  If we denote the effect generated by the operation $op$ as
$\eff_{op}$, then $\Theta(r) \times \{\eff_{op}\} ~\subseteq~ \visZ$. Often,
in our formalization, we use $\visZ$ and $\soZ$ binary relations to obtain a
set of effects visible to a given effect $\eff$, or set of effects that
precede a given effect $\eff$ in the session order. As a syntactic
convenience, whenever $R$ is a binary relation, we write $R(\eff)$ to denote
the set of all $\eff'$ such that $(\eff,\eff') \in R$.  Conversely, we write
$R^{-1}(\eff)$ to denote the set of all $\eff'$ such that $(\eff',\eff) \in
R$.

Basic gurantee provided by the store is causal visibility, which is
captured by the rule $\rulelabel{EffVis}$ as a condition for an effect
to be visible at a replica. The rule makes an effect ($\eff$) visible
at a replica $r$ only after all the effects that causally precede
$\eff$ are made visible at $r$.  It is important to note that that
enforcing causal visibility does not require any inter-replica
coordination. Any eventually consistent store can provide causal
visibility while being eventually consistent.  Therefore, we do not lose
any generality by assuming that the store provides causal visiblity.

% Semantics
% ---------
\input{core-opsem}

Rule $\rulelabel{Oper}$ is an auxiliary reduction of the
form:\vspace{-1.7mm}
\begin{smathpar}
\auxred{\Theta} {(\E,\langle s,i,op \rangle)} {r} {(\E',\eff)}
\vspace{-1.7mm}
\end{smathpar}
\noindent Under the store configuration $\Theta$, the rule captures
the progress in execution (from $\E$ to $\E'$) due to the application
of operation $op$ to replica $r$ resulting in a new effect $\eff$.
The rule first constructs a \emph{context} for the application from
the local state ($\Theta(r)$) of the replica, by
projecting\footnote{{\textsf{ctxt*}} is auxiliary function
\textsf{ctxt} extended straightforwardly to set of effects} relevant
information from effects in $\Theta(r)$. It then substitutes the
definition ($\Ops(op)$) of the operation for its label ($op$), and
relies on the reduction relation ($\rdtredsto$) of the server-side
language to reduce the application $\Ops(op)(ctxt)$ to a value
$v'$.  Subsequently, the the attributes of execution state, namely
$\EffSoup$, $\visZ$, $\soZ$, and $\sameobjZ$ are extended to
incorporate the new effect ($\eff$).

If the operation $op$ is {\sf EventuallyConsistent}, we simply apply
the operation to any replica $r$. Since the store provides causal
visibility, eventually consistent operations are satisfiable under any
replica. If the operation is {\sf CausallyConsistent}, the operation
can be applied to a replica $r$ only if it already contains the
effects of all the previous operations from the same session. This
guarantee can be satisfied by applying all operations from the same
session to the same logical copy of the database.  If such a logical
copy is untenable, then the operation might block. Since the store is
assumed to converge eventually, the blocked causally consistent
operation guaranteed to unblock eventually.

A {\sf StronglyConsistent} operation expects sequential consistency.
That is, universe of all effects ($\EffSoup$) in an execution ($\E$)
must be partitionable into a set of effects that \emph{happened
before} $\eff$ and another set that \emph{happened after} $\eff$,
where $\eff$ is the effect generated by an strongly consistent
operation. The rule $\rulelabel{UA}$ enforces this sequencing in two
steps; firstly, it insists that the the strongly consistent operation
($op$) witness effects of all operations executed so far by requiring
the global set of effects $\EffSoup$ to be a subset of local state
($\Theta(r)$) of the replica ($r$) executing $op$. Secondly, the rule
requires the effect ($\eff$) generated by $op$ to be added to the
local state of every other replica in the store, so that further
operations on these replicas can witness the effect of $op$. Since
both these steps require global coordination among replicas, the store
is \emph{unavailable} during the time it is executing $op$. 
% However, it must be noted that executing strongly consistent
% operation does not entail the convergence of local states of all
% replicas; it simply means that there is one replica ($r$) whose
% local state is the upper bound of local states of all replicas, and
% there exists a set containing atleast one effect ($\eff$), which is
% their lower bound.

\subsection{Soundness of Axiomatic Semantics}

Our main meta-theoretic result guarantees the soundness of our
axiomatic semantics in enforcing $\ecc$, $\ccc$, and $\scc$
consistency guarantess by ensuring that every reduction step preserves
correctness of axiomatic execution.

\begin{theorem}[Soundness preservation]
\label{lem:core-preservation}
Let $\E$ be a well-formed axiomatic execution state, and $\tau$ denote
a contract class.  If:
\begin{smathpar}
\E; \Theta; \langle s,i,\langle op,\tau \rangle;
\sigma \rangle \pll \Sigma \xhookrightarrow{\eff} \E'; \Theta';
\langle s,i,\sigma \rangle \Sigma'
\end{smathpar}
then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
$\E' \models \cv_{\tau}[\eff/\cureff]$ 
\end{theorem}

### 2014-11-06 contract-language.tex
edits
Commiting appendix

@@
 
-  \vspace{-1em}
+  \vspace{-1.5em}
   \begin{smathpar}
@@
 \begin{theorem}[Soundness of Contract Enforcement]
-\label{lem:core-preservation}
+\label{thm:classification-sound}
 Let $\cv$ be a well-formed contract of a replicated data type operation

### 2014-11-07 appendix.tex
Techreport is a separate file

@@
 
-\subsection{Axiomatic Semantics}
+\subsection{Operational Semantics}
 \renewcommand{\auxred}[4]{#1 \vdash #2 \;\xhookrightarrow{#3}\; #4 }
 
-We now describe axiomatic semantics of a data store that implements
-strong, causal and eventual consistency guarantees. The semantics also
-serves as an elegant representation of our implementation of \name.
+We now describe operational semantics of a data store that implements strong,
+causal and eventual consistency guarantees. The semantics also serves as an
+elegant representation of our implementation of \name.
 
-Figure ~\ref{sem:oper} presents our semantics as concretization of the
-rewrite relation ($\xrightarrow{}$) over execution state. Since we now have a
-concrete store, we extend our system model with $\Theta$, a representation of
-the store as a map from replicas to their local states. The local state of a
-replica $r$ (i.e., $\Theta(r)$) is a set of effects that are currently
-visible at $r$.  An operation $op$ performed at replica $r$ can only witness
-the set of effects ($\Theta(r)$) visible at $r$. For the sake of clarity, we
-only consider a single replicated object of well-defined type (for eg: a
-replicated object of type \cf{BankAccount}) in our formalization.  Our
-semantics are parametric over the specification of this replicated data type.
-Figure~\ref{sem:oper} formalizes replicated data type (RDT) specification as
-tuple $(\delta,\Ops,\Ctrts)$, where $\delta$ is the data type, $\Ops$ maps
-labels ($op$) of operations on $\delta$ to their definitions, while $\Ctrts$
-maps them to their consistency contracts ($\cv$). The definition of an
-operation is expected to be a lambda expression, although we do not enforce
-this in our formalization. For technical reasons, we tag each session with a
-session identifier ($s$) and the sequence number ($i$) of the next operation
-in the session.
+Figure ~\ref{sem:oper} presents our semantics as concretization of the rewrite
+relation ($\xrightarrow{}$) over execution state. Since we now have a concrete
+store, we extend our system model with $\Theta$, a representation of the store
+as a map from replicas to their local states. The local state of a replica $r$
+(i.e., $\Theta(r)$) is a set of effects that are currently visible at $r$.  An
+operation $op$ performed at replica $r$ can only witness the set of effects
+($\Theta(r)$) visible at $r$. For the sake of clarity, we only consider a
+single replicated object of well-defined type (for eg: a replicated object of
+type \cf{BankAccount}) in our formalization.  Our semantics are parametric over
+the specification of this replicated data type. Figure~\ref{sem:oper}
+formalizes replicated data type (RDT) specification as tuple
+$(\delta,\Ops,\Ctrts)$, where $\delta$ is the data type, $\Ops$ maps labels
+($op$) of operations on $\delta$ to their definitions, while $\Ctrts$ maps them
+to their consistency contracts ($\cv$). The definition of an operation is
+expected to be a lambda expression, although we do not enforce this in our
+formalization. For technical reasons, we tag each session with a session
+identifier ($s$) and the sequence number ($i$) of the next operation in the
+session.
 
-The state of an axiomatic execution ($\E$) is a tuple $\Exec$ where $\EffSoup$
-is a set of effects, and $\visZ,\soZ, \sameobjZ \subseteq \EffSoup \times
-\EffSoup$ are \emph{visibility}, \emph{session order}, and \emph{same object}
-relations over effects, respectively. We concretize an effect ($\eff$) as a
-tuple $(s,i,op,v)$, which records the fact that $i^{th}$ action in session
-with {\sf SessID} $s$, which is an operation $op$ on the replicated object,
-has been successfully executed on some replica yielding a return value $v$.
-Note that the combination of $s$ and $i$ uniquely identifies the effect.
-Session order relation ($\soZ$) relates effects generated by the same session.
-An effect $\eff = (s,i,op,v)$ is said to precede another effect $\eff' =
+The state of an operational execution ($\E$) is a tuple $\Exec$ where
+$\EffSoup$ is a set of effects, and $\visZ,\soZ, \sameobjZ \subseteq \EffSoup
+\times \EffSoup$ are \emph{visibility}, \emph{session order}, and \emph{same
+object} relations over effects, respectively. We concretize an effect ($\eff$)
+as a tuple $(s,i,op,v)$, which records the fact that $i^{th}$ action in session
+with {\sf SessID} $s$, which is an operation $op$ on the replicated object, has
+been successfully executed on some replica yielding a return value $v$. Note
+that the combination of $s$ and $i$ uniquely identifies the effect. Session
+order relation ($\soZ$) relates effects generated by the same session. An
+effect $\eff = (s,i,op,v)$ is said to precede another effect $\eff' =
 (s',i',op',v')$ in session order if and only if $s'=s$ and $i'\ge i$. Since we
@@
 eventually.  If we denote the effect generated by the operation $op$ as
-$\eff_{op}$, then $\Theta(r) \times \{\eff_{op}\} ~\subseteq~ \visZ$. Often,
-in our formalization, we use $\visZ$ and $\soZ$ binary relations to obtain a
-set of effects visible to a given effect $\eff$, or set of effects that
-precede a given effect $\eff$ in the session order. As a syntactic
-convenience, whenever $R$ is a binary relation, we write $R(\eff)$ to denote
-the set of all $\eff'$ such that $(\eff,\eff') \in R$.  Conversely, we write
-$R^{-1}(\eff)$ to denote the set of all $\eff'$ such that $(\eff',\eff) \in
-R$.
+$\eff_{op}$, then $\Theta(r) \times \{\eff_{op}\} ~\subseteq~ \visZ$. Often, in
+our formalization, we use $\visZ$ and $\soZ$ binary relations to obtain a set
+of effects visible to a given effect $\eff$, or set of effects that precede a
+given effect $\eff$ in the session order. As a syntactic convenience, whenever
+$R$ is a binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
+such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$ to denote
+the set of all $\eff'$ such that $(\eff',\eff) \in R$.
 
@@
 
-\subsection{Soundness of Axiomatic Semantics}
+\subsection{Soundness of Operational Semantics}
 
-Our main meta-theoretic result guarantees the soundness of our
-axiomatic semantics in enforcing $\ecc$, $\ccc$, and $\scc$
-consistency guarantess by ensuring that every reduction step preserves
-correctness of axiomatic execution.
+Our main meta-theoretic result guarantees the soundness of our operational
+semantics in enforcing $\ecc$, $\ccc$, and $\scc$ consistency guarantess by
+ensuring that every reduction step preserves correctness of operational
+execution.
 
@@
 \label{lem:core-preservation}
-Let $\E$ be a well-formed axiomatic execution state, and $\tau$ denote
-a contract class.  If:
+Let $\E$ be a well-formed operational execution state, and $\tau$ denote a
+contract class.  If:
 \begin{smathpar}

### 2014-11-07 intro.tex
edits

@@
 
-To address these issues, we describe \name, a declarative programming model
-and implementation for eventually consistent geo-distributed data
-stores. The key novelty of \name is an expressive \emph{contract language}
-to declare and verify fine-grained application-level consistency
-properties. The programmer uses the contract language to axiomatically
-specify the set of legal executions allowed over the replicated data
-type. Contracts are constructed using primitive consistency relations such
-as \emph{visibility} and \emph{session order} along with standard logical
-and relational operators. A \emph{contract enforcement system} automatically
-maps operations over the datatype to a particular consistency level
-available on the store, and provably validates the correctness of the
-mapping.  The paper makes the following contributions:
+To address these issues, we describe \name, a declarative programming model and
+implementation for eventually consistent geo-distributed data stores. The key
+novelty of \name is an expressive \emph{contract language} to declare and
+verify fine-grained application-level consistency properties. The programmer
+uses the contract language to axiomatically specify the set of legal executions
+allowed over the replicated data type. Contracts are constructed using
+primitive consistency relations such as \emph{visibility} and \emph{session
+order} along with standard logical and relational operators. A \emph{contract
+enforcement system} statically maps operations over the datatype to a
+particular consistency level available on the store, and provably validates the
+correctness of the mapping. The paper makes the following contributions:
 
@@
 	consistent stores. Contracts are used to specify fine-grained
-	application-level consistency properties, and are analyzed to assign the most
-	efficient and sound store consistency level to the corresponding operation.
+	application-level consistency properties, and are statically analyzed to
+	assign the most efficient and sound store consistency level to the
+	corresponding operation.
 \item \name supports coordination-free transactions over arbitrary datatypes.

### 2014-11-07 contract-language.tex
edits
Edits
Added operational semantics to the techrep

@@
 
-\input{contract-syntax.tex}
+\begin{figure}
+\begin{smathpar}
+\begin{array}{rclcl}
+\multicolumn{5}{l}{
+  {x,y,z} \in \mathtt{EffVar} \qquad
+  {\cureff} \in \mathtt{CurEff} \qquad
+  {\sf Op} \in \mathtt{OperName}
+}\\
+\cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
+        \ALT \pi \\
+\tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
+        \ALT \tau \vee \tau \\
+\pi			&	\in & \mathtt{Prop} & \coloneqq & \true \ALT R(x,y)
+        \ALT \pi \vee \pi \\
+			  & 		&	 &  \ALT & \pi \wedge \pi \ALT \pi \Rightarrow \pi \\
+R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
+        \ALT \sameobjZ \ALT R^+ \\
+				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
+\end{array}
+\end{smathpar}
+\caption{Contract language.}
+\label{fig:contract-lang}
+\end{figure}
 
@@
 
-\caption{An Axiomatic Execution}
+\caption{Axiomatic execution.}
 \label{sem:contracts}
@@
 Note that the axiomatic definition of an execution ($\E$) provides
-interpretations for primitive relations (eg: $\visZ$) that occur free
-in contract formulas, and also fixes the domain of quantification to
-set of all effects ($\EffSoup$) observed during the program execution.
-As such, $\E$ is a potential model for any first-order formula ($\cv$)
-expressible in our contract language. If $\E$ is indeed a valid model
-for $\cv$ (expressed using the model-theoretic consequence relation as $\E
-\models \cv$), we say that the execution $\E$ satisfied the contract
-$\cv$:
+interpretations for primitive relations (eg: $\visZ$) that occur free in
+contract formulas, and also fixes the domain of quantification to set of all
+effects ($\EffSoup$) observed during the program execution. As such, $\E$ is a
+potential model for any first-order formula ($\cv$) expressible in our contract
+language. If $\E$ is indeed a valid model for $\cv$ (expressed using the
+model-theoretic consequence relation as $\E \models \cv$), we say that the
+execution $\E$ satisfied the contract $\cv$:
 \begin{definition}
@@
   \end{smathpar}
-
 \end{itemize}
-
-% Contract classification rules
-% ------------------------------
-\input{contract-classification}
+\vspace{-1em}
 
 \subsection{Contract Comparison and Classification}
+
+\newcommand{\DDe}[1]{#1}
+\begin{figure}
+\begin{smathpar}
+\begin{array}{c}
+\vspace{3mm}
+\RuleTwo
+{\DDe{\cv} \le \DDe{\scc}}
+{{\sf WellFormed}(\cv)}  \qquad
+
+\RuleTwo
+{\DDe{\cv} \le \DDe{\ecc}}
+{{\sf EventuallyConsistent}(\cv)} \\
+
+\vspace{3mm}
+\RuleTwo
+{\DDe{\cv} \not\le \DDe{\ecc}
+\quad \DDe{\cv} \le \DDe{\ccc}}
+{{\sf CausallyConsistent}(\cv)} \qquad
+
+\RuleTwo
+{\DDe{\cv} \not\le \DDe{\ccc}
+\quad \DDe{\cv} \le \DDe{\scc}}
+{{\sf StronglyConsistent}(\cv)}
+\end{array}
+\end{smathpar}
+\vspace{-5mm}
+\caption{Contract classification.}
+\label{sem:classify}
+\end{figure}
 
@@
 language as following:
+
 \begin{definition}
@@
 \le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
-\begin{center}
-\end{center}
 \end{definition}
-\vspace{-2em}
+
 \noindent The $\Delta$ referred in the above defintion is a conjunction of
@@
 classification-based contract enforcement. To help us state the result, we
-define an axiomatic model of the our system described informally in
+define an operational semantics of the our system described informally in
 \S~\ref{sec:sysmod}:
@@
 tagged with the consistency class $\tau$ of their contracts (as determined by
-the contract classification scheme). We assume a rewrite relation of form:
+the contract classification scheme). We assume a reduction relation of form:
 
@@
 \emph{enforced} the contract $\cv_\tau$ in the execution $\E'$. With help of
-the axiomatic model, we now state the soundness of contract enforcement as
-follows:
+the operational semantics, we now state the soundness of contract enforcement
+as follows:
 
@@
 Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
-proof of the theorem is given in the supplementary material\footnote{The
-supplementary material also defines an operational semantics that concretizes
-the rewrite relation ($\xrightarrow{}$), along with the proof that it correctly
-enforces all \name contracts.}.
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H2$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H3$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$ and $H4$.
+  \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2014-11-07 store-opsem.tex
Added operational semantics to the techrep

--- 2014-11-07 appendix.tex
@@
-\section{Appendix}
-\label{sec:appendix}
+\section{Operational Semantics}
+\label{sec:store-opsem}
 
-\subsection{Soundness of Contract Classification}
-
-We restate theorem ~\ref{thm:classification-sound} and provide its proof
-below:
-
-\begin{theorem}[Soundness of Contract Enforcement]
-\label{lem:classification-sound}
-Let $\cv$ be a well-formed contract of a replicated data type
-operation $\mathit{op}$, and let $\tau$ denote the consistency class
-of $\cv$ as determined by the contract classification scheme. For all
-well-formed execution states $\E$, $\E'$ such that $\auxred{}
-{\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff} {\E', \sigma
-\pll \Sigma}$, if $\E' \models \cv_\tau[\eff/\cureff]$, then $\E'
-\models \cv[\eff/\cureff]$
-\end{theorem}
-\begin{proof}
-  Hypothesis:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
-    {\E', \sigma \pll \Sigma} & H\npp\\
-    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Since $\tau$ is the contract class of $\cv$, by inversion, we have
-  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
-  \end{array}
-  \end{smathpar}
-   Since $\eff$ denotes new effect, it is a fresh variable that does
-   not occur free in $\Delta$. From $H2$, after instantiating bound
-   $\cureff$ with $\eff$, we have:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
-      & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Due to the soundness of natural deduction for first-order logic,
-  $H3$ implies that for all models $\mathcal{M}$ such that
-  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
-  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
-  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \E' \models \Delta & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Proof follows from $H1$ and $H4$.
-  \qed
-\end{proof}
-
-\subsection{Operational Semantics}
 \renewcommand{\auxred}[4]{#1 \vdash #2 \;\xhookrightarrow{#3}\; #4 }
 
-We now describe operational semantics of a data store that implements strong,
-causal and eventual consistency guarantees. The semantics also serves as an
-elegant representation of our implementation of \name.
+We now describe operational semantics of a data store that implements
+strong, causal and eventual consistency guarantees. The semantics also
+serves as a high-level description of our implementation of the store
+underlying \name.
 
-Figure ~\ref{sem:oper} presents our semantics as concretization of the rewrite
-relation ($\xrightarrow{}$) over execution state. Since we now have a concrete
-store, we extend our system model with $\Theta$, a representation of the store
-as a map from replicas to their local states. The local state of a replica $r$
-(i.e., $\Theta(r)$) is a set of effects that are currently visible at $r$.  An
-operation $op$ performed at replica $r$ can only witness the set of effects
-($\Theta(r)$) visible at $r$. For the sake of clarity, we only consider a
-single replicated object of well-defined type (for eg: a replicated object of
-type \cf{BankAccount}) in our formalization.  Our semantics are parametric over
-the specification of this replicated data type. Figure~\ref{sem:oper}
-formalizes replicated data type (RDT) specification as tuple
-$(\delta,\Ops,\Ctrts)$, where $\delta$ is the data type, $\Ops$ maps labels
-($op$) of operations on $\delta$ to their definitions, while $\Ctrts$ maps them
-to their consistency contracts ($\cv$). The definition of an operation is
-expected to be a lambda expression, although we do not enforce this in our
-formalization. For technical reasons, we tag each session with a session
-identifier ($s$) and the sequence number ($i$) of the next operation in the
-session.
+Figure ~\ref{sem:oper} presents operational semantics as rules
+defining the reduction relation ($\xrightarrow{}$) over the execution
+state.  Since we now have a concrete store, we extend our system model
+with $\Theta$, a representation of the store as a map from replicas to
+their local states. The local state of a replica $r$ (i.e.,
+$\Theta(r)$) is a set of effects that are currently visible at $r$.
+An operation $op$ performed at replica $r$ can only witness the set of
+effects ($\Theta(r)$) visible at $r$. For the sake of clarity, we only
+consider a single replicated object of well-defined type (for eg: a
+replicated object of type \cf{BankAccount}) in our formalization.  Our
+semantics are parametric over the specification of this replicated
+data type. Figure~\ref{sem:oper} formalizes replicated data type (RDT)
+specification as tuple $(\delta,\Ops,\Ctrts)$, where $\delta$ is the
+data type, $\Ops$ maps labels ($op$) of operations on $\delta$ to
+their definitions, while $\Ctrts$ maps them to their consistency
+contracts ($\cv$). The definition of an operation is expected to be a
+lambda expression, although we do not enforce this in our
+formalization. For technical reasons, we tag each session with a
+session identifier ($s$) and the sequence number ($i$) of the next
+operation in the session.
 
 The state of an operational execution ($\E$) is a tuple $\Exec$ where
-$\EffSoup$ is a set of effects, and $\visZ,\soZ, \sameobjZ \subseteq \EffSoup
-\times \EffSoup$ are \emph{visibility}, \emph{session order}, and \emph{same
-object} relations over effects, respectively. We concretize an effect ($\eff$)
-as a tuple $(s,i,op,v)$, which records the fact that $i^{th}$ action in session
-with {\sf SessID} $s$, which is an operation $op$ on the replicated object, has
-been successfully executed on some replica yielding a return value $v$. Note
-that the combination of $s$ and $i$ uniquely identifies the effect. Session
-order relation ($\soZ$) relates effects generated by the same session. An
-effect $\eff = (s,i,op,v)$ is said to precede another effect $\eff' =
-(s',i',op',v')$ in session order if and only if $s'=s$ and $i'\ge i$. Since we
-only consider one replicated object in our formalization, the $\sameobjZ$
-relation relates every pair of effects in the effect soup ($\EffSoup$). An
+$\EffSoup$ is a set of effects, and $\visZ$, $\soZ$, $\sameobjZ$
+$\subseteq$ $\EffSoup \times \EffSoup$ are \emph{visibility},
+\emph{session order}, and \emph{same object} relations over effects,
+respectively. We concretize an effect ($\eff$) as a tuple
+$(s,i,op,v)$, which records the fact that $i^{th}$ action in session
+with {\sf SessID} $s$, which is an operation $op$ on the replicated
+object, has been successfully executed on some replica yielding a
+return value $v$. Note that the combination of $s$ and $i$ uniquely
+identifies the effect. Session order relation ($\soZ$) relates effects
+generated by the same session.  An effect $\eff = (s,i,op,v)$ is said
+to precede another effect $\eff' = (s',i',op',v')$ in session order if
+and only if $s'=s$ and $i'\ge i$. Since we only consider one
+replicated object in our formalization, the $\sameobjZ$ relation
+relates every pair of effects in the effect soup ($\EffSoup$). An
 effect generated at a replica becomes visible at rest of the replicas
-eventually.  If we denote the effect generated by the operation $op$ as
-$\eff_{op}$, then $\Theta(r) \times \{\eff_{op}\} ~\subseteq~ \visZ$. Often, in
-our formalization, we use $\visZ$ and $\soZ$ binary relations to obtain a set
-of effects visible to a given effect $\eff$, or set of effects that precede a
-given effect $\eff$ in the session order. As a syntactic convenience, whenever
-$R$ is a binary relation, we write $R(\eff)$ to denote the set of all $\eff'$
-such that $(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$ to denote
+eventually.  If we denote the effect generated by the operation $op$
+as $\eff_{op}$, then $\Theta(r) \times \{\eff_{op}\} ~\subseteq~
+\visZ$. Often, in our formalization, we use $\visZ$ and $\soZ$ binary
+relations to obtain a set of effects visible to a given effect $\eff$,
+or set of effects that precede a given effect $\eff$ in the session
+order. As a syntactic convenience, whenever $R$ is a binary relation,
+we write $R(\eff)$ to denote the set of all $\eff'$ such that
+$(\eff,\eff') \in R$.  Conversely, we write $R^{-1}(\eff)$ to denote
 the set of all $\eff'$ such that $(\eff',\eff) \in R$.
@@
 coordination. Any eventually consistent store can provide causal
-visibility while being eventually consistent.  Therefore, we do not lose
-any generality by assuming that the store provides causal visiblity.
+visibility while being eventually consistent.  Therefore, we do not
+lose any generality by assuming that the store provides causal
+visiblity.
 
@@
 % ---------
-\input{core-opsem}
+\input{store-opsem-rules}
 
@@
 relies on the reduction relation ($\rdtredsto$) of the server-side
-language to reduce the application $\Ops(op)(ctxt)$ to a value
-$v'$.  Subsequently, the the attributes of execution state, namely
+language to reduce the application $\Ops(op)(ctxt)$ to a value $v'$.
+Subsequently, the the attributes of execution state, namely
 $\EffSoup$, $\visZ$, $\soZ$, and $\sameobjZ$ are extended to
@@
 where $\eff$ is the effect generated by an strongly consistent
-operation. The rule $\rulelabel{UA}$ enforces this sequencing in two
+operation. The rule $\rulelabel{SC}$ enforces this sequencing in two
 steps; firstly, it insists that the the strongly consistent operation
@@
 
-Our main meta-theoretic result guarantees the soundness of our operational
-semantics in enforcing $\ecc$, $\ccc$, and $\scc$ consistency guarantess by
-ensuring that every reduction step preserves correctness of operational
-execution.
+We now prove a meta-theoretic property that establishes the soundness
+of our operational semantics in enforcing $\ecc$, $\ccc$, and $\scc$
+consistency guarantees at every reduction step. As a corollary of this
+result, and Theorem~\ref{thm:classification-sound}, we have the
+assurance that \name correctly enforces all well-formed consistency
+contracts.
 
-\begin{theorem}[Soundness preservation]
-\label{lem:core-preservation}
-Let $\E$ be a well-formed operational execution state, and $\tau$ denote a
-contract class.  If:
+\begin{theorem}[Soundness Modulo Classification]
+\label{thm:weak-soundness}
+For every well-formed execution state $\E$, and contract class $\tau 
+\in \{{\sf ec},{\sf cc},{\sf sc}\}$, If:
 \begin{smathpar}
-\E; \Theta; \langle s,i,\langle op,\tau \rangle;
-\sigma \rangle \pll \Sigma \xhookrightarrow{\eff} \E'; \Theta';
-\langle s,i,\sigma \rangle \Sigma'
+(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
+\langle s,i,\sigma \rangle \Sigma')
 \end{smathpar}
@@
 \end{theorem}
+\begin{proof}
+  By case analysis on $(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+  \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
+  \langle s,i,\sigma \rangle \Sigma')$. 
+  \GK{Lets first finalize all notations, rules, and also agree on
+  all theorem statements. This way we don't have to change the proof
+  again.} 
+\end{proof}
+
+\begin{corollary}[Soundness]
+\label{thm:soundness}
+For every well-formed execution state $\E$, and contract class $\tau
+\in \{{\sf ec},{\sf cc},{\sf sc}\}$, and for every consistency
+contract $\cv$ in the contract class $\tau$, If:
+\begin{smathpar}
+(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
+\langle s,i,\sigma \rangle \Sigma')
+\end{smathpar}
+then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
+$\E' \models \cv[\eff/\cureff]$
+\end{corollary}
+\begin{proof}
+Follows from Theorems~\ref{thm:classification-sound} and
+\ref{thm:weak-soundness}
+\qed
+\end{proof}

### 2014-11-07 store-opsem.tex
Fixed typos

@@
 \emph{session order}, and \emph{same object} relations over effects,
-respectively. We concretize an effect ($\eff$) as a tuple
+respectively. We define an effect ($\eff$) as a tuple
 $(s,i,op,v)$, which records the fact that $i^{th}$ action in session
@@
 \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-\langle s,i,\sigma \rangle \Sigma')
+\langle s,i,\sigma \rangle \pll \Sigma')
 \end{smathpar}
@@
   \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-  \langle s,i,\sigma \rangle \Sigma')$. 
+  \langle s,i,\sigma \rangle \pll \Sigma')$. 
   \GK{Lets first finalize all notations, rules, and also agree on

### 2014-11-07 concl.tex
edits

@@
 
-In this paper, we have presented \name a shallow Haskell extension for
-declarative programming over eventually consistent data stores. The key idea of
-\name is the automatic classification of fine-grained consistency contracts on
-operations and distributed transactions with respect to the consistency and
-isolation levels offered by the store. Our contract language is carefully
-crafted from a decidable subset of first-order logic, enabling the use of
-automatic theorem prover to discharge the proof obligations associated with
-contract classification. We provide meta-theory that certifies the correctness
-of contract classification. We realize an instantiation of \name on top of
-off-the-shelf distributed store, Cassandra, and illustrate the benefit of
-fine-grained contract classification by implementing and evaluating several
-scalable applications.
+This paper presents \name, a shallow Haskell extension for declarative
+programming over eventually consistent data stores. The key idea underlying
+\name's design is the automatic classification of fine-grained consistency
+contracts on operations and distributed transactions with respect to the
+consistency and isolation levels offered by the store. Our contract language
+is carefully crafted from a decidable subset of first-order logic, enabling
+the use of automated verification tools to discharge the proof obligations
+associated with contract classification. We realize an instantiation of
+\name on top of off-the-shelf distributed store, Cassandra, and illustrate
+the benefit of fine-grained contract classification by implementing and
+evaluating several scalable applications.

### 2014-11-07 evaluation.tex
edits

@@
 
-In this section, we evaluate \name programs, report their contract profile and
-illustrate the performance benefits of fine-grained consistency classification
-on operations and transactions. We also evaluate on the impact of the
-summarization. We implemented the following applications, which includes
-individual RDTs as well as larger applications composed of several RDTs:
+We present an evaluation study of our implementation, report contract
+profiles of benchmark programs, and illustrate the performance benefits of
+fine-grained consistency classification on operations and transactions. We
+also evaluate the impact of the summarization. We have implemented the
+following applications, which includes individual RDTs as well as larger
+applications composed of several RDTs:
 
@@
 \setlength{\itemsep}{2pt}
-\item \textbf{LWW register}: A last-write-wins register that provides read and
-write operations. Each write is associated with a timestamp, which is used to
-resolve conflicting concurrent writes -- newer write wins.
+\item \textbf{LWW register}: A last-write-wins register that provides read
+  and write operations. Each write is associated with a timestamp, which is
+  used to resolve conflicting concurrent writes -- a newer write always
+  wins.
 
-\item \textbf{DynamoDB register}: A integer register that allows eventual and
-strong puts and gets, conditional puts, increment and decrement operations.
+\item \textbf{DynamoDB register}: An integer register that allows eventual
+  and strong puts and gets, conditional puts, increment and decrement
+  operations.
 
-\item \textbf{Bank account}: Our running example, with savings and current
+\item \textbf{Bank account}: Our running example, with savings and checking
   accounts.
 
-\item \textbf{Shopping list}: Collaborative shopping list which allows adding
-and deleting items.
+\item \textbf{Shopping list}: A collaborative shopping list that allows
+  adding and deleting items.
 
-\item \textbf{Online store}: Models an online store with shopping cart and
-dynamically changing item prices. Checkout process verifies that the customer
-only pays the accepted price.
+\item \textbf{Online store}: An online store with shopping cart
+  functionality and dynamically changing item prices.  The checkout process
+  verifies that the customer only pays the accepted price.
 
-\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The application
-allows users to browse items, bid for items on sale, and pay for items from a
-wallet modelled after a bank account.
+\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The
+  application allows users to browse items, bid for items on sale, and pay
+  for items from a wallet modelled after a bank account.
 
 \item \textbf{Microblog}: A twitter-like microblogging site, modelled after
-Twissandra~\cite{Twissandra}. The application allows adding a new user, adding
-and replying to tweets, following, unfollowing and blocking users, and fetching
-a user's timeline, userline, followers and following.
+  Twissandra~\cite{Twissandra}. The application allows adding new users,
+  adding and replying to tweets, following, unfollowing and blocking users,
+  and fetching a user's timeline, userline, followers and following.
 \end{itemize}
@@
 The distribution of contracts in these applications is given in
-Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
-are classified as eventually consistent and RC, respectively. Operation
-contracts are used to enforce integrity and visibility constraints on
-individual fields in the tables. Transactions are mainly used to consistently
-modify and access related fields across tables. In \name, the contract
-classification process is completely performed at compile time and has no
-overheads at runtime. The proof obligations associated with contract
-classification is discharged through the Z3 SMT Solver. Across our benchmarks,
-classifying a contract took 11.5 milliseconds on average.
+Table~\ref{tab:ctrts}. We see that majority of the operations and
+transactions are classified as eventually consistent and RC,
+respectively. Operation contracts are used to enforce integrity and
+visibility constraints on individual fields in the tables. Transactions are
+mainly used to consistently modify and access related fields across
+tables. In \name, the contract classification process is completely
+performed at compile time and has no overheads at runtime. The proof
+obligations associated with contract classification is discharged through
+the Z3 SMT Solver. Across our benchmarks, classifying a contract took 11.5
+milliseconds on average.
 
-For our performance evaluation, we deploy \name applications in \emph{clusters},
-where each cluster is composed of 5 fully replicated Cassandra replicas within
-the same datacenter. We instantiate one shim layer node for every Cassandra
-replica, and place it on the same VM as the Cassandra replica. Clients are
-instantiated on the same data center as the store, and run transactions. We
-deploy the each node in the cluster on \cf{c3.4xlarge} Amazon EC2 instances.
-Our shim layer nodes are multi-threaded, and we allocate 8 CPUs (out of 16
-available) for each shim layer node. The clients also run on \cf{c2.4xlarge}
-instances. We call this \cf{1DC} configuration. For our geo-distributed
-experiments (\cf{2DC}), we instantiate 2 clusters, each with 5 nodes, and place
-the clusters on US-east (Virginia) and US-west (Oregon). The average
-inter-region latency was 85ms.
+For our performance evaluation, we deploy \name applications in
+\emph{clusters}, where each cluster is composed of five fully replicated
+Cassandra replicas within the same datacenter. We instantiate one shim layer
+node co-located on the same VM as a Cassandra replica. Clients are
+instantiated within the same data center as the store, and run
+transactions. We deploy each node in the cluster on \cf{c3.4xlarge} Amazon
+EC2 instances.  Our shim layer nodes are multi-threaded, and we allocate 8
+CPUs (out of 16 available) for each shim layer node. The clients also run on
+\cf{c2.4xlarge} instances. We call this a \cf{1DC} configuration. For our
+geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with
+five nodes, and place the clusters on US-east (Virginia) and US-west
+(Oregon) locations. The average inter-region latency was 85ms.
 
@@
 
-Figure~\ref{grf:BA-tp-vs-lat} shows the throughput vs. latency of operations
-in bank account example as we increase the number of clients in \cf{1DC}
-configuration. Our client workload was generated using YCSB
-benchmark~\cite{YCSB}. The benchmark uniformly chose from 100,000 keys, where
+Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations
+in the bank account example as we increase the number of clients in a \cf{1DC}
+configuration. Our client workload was generated using the YCSB
+benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
 the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
@@
 
-The lines marked EC and CC correspond to all operations being assigned EC and
-CC consistency levels. These levels compromise correctness as withdraw has to
-be and SC operation. The line SC corresponds to a configuration where all
-operations are strongly consistent; this ensures application correctness. \name
-corresponds to our implementation, which classifies operations based on their
-contracts. Both \name and SC ensure correctness. However, with 512 clients,
-\name implementation was within 41\% of latency and 18\% of throughput of EC,
-whereas SC operations had 162\% higher latency and 52\% lower throughput than
-EC operations. Observe that there is a point in each case after which the
-latency increases while the throughput decreases. This indicates a point after
-which the store becomes saturated with client requests.
+The lines marked EC and CC correspond to all operations (including
+\cf{withdraw}) being assigned EC and CC consistency levels. These levels
+compromise correctness as \cf{withdraw} has to be an SC operation. The SC
+line corresponds to a configuration where all operations are strongly
+consistent; this ensures application correctness, at the cost of
+performance. \name corresponds to our implementation, which classifies
+operations based on their contract specifications.  Both \name and SC ensure
+correctness. With 512 clients, the \name implementation was within 41\% of
+the latency and 18\% of the throughput of EC, whereas SC operations had
+162\% higher latency and 52\% lower throughput than EC operations. Observe
+that there is a point in each case after which the latency increases while
+the throughput decreases; these points correspond to client request
+saturation points in the data store.
 
-In \cf{2DC} configuration (not-shown), the average latency of SC operations
-with 512 clients increased by 9.4$\times$ due to the cost of geo-distributed
-coordination, whereas \name operations were only 2.2$\times$ slower, mainly due
-to the increased cost of \cf{withdraw} operations. Importantly, the latency of
-\cf{getBalance} and \cf{deposit} remained almost the same. This illustrates the
-benefit of fine-grained contract classification in \name.
+In a \cf{2DC} configuration (not shown here), the average latency of SC
+operations with 512 clients increased by 9.4$\times$ due to the cost of
+geo-distributed coordination, whereas \name operations were only 2.2$\times$
+slower, mainly due to the increased cost of \cf{withdraw}
+operations. Importantly, the latency of \cf{getBalance} and \cf{deposit}
+remained almost the same, illustrating the benefit of fine-grained
+contract classification.
 
-We compare the performance of different transaction isolation level choices in
-Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
-under 1DC configuration. The YCSB workload was modified to issue 10 operations
-per transaction, with the default 50:50 read:write mix. Each operation is
-assumed to have eventual consistency. \cf{NoTxn} corresponds to a configuration
-that does not use transactions. Compared to this RC is only 12\% shower in
-terms of latency with 512 clients, where as RR is 2.3X slower. The difference
-between RC and \cf{NoTxn} is due to the meta-data overhead of recording
-transaction information in the object state. For RR transaction, the cost of
-capturing and maintaining the snapshot in an RR transaction is the biggest
-source of overhead.
+We compare the performance of different transaction isolation level choices
+in Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were
+obtained under a 1DC configuration. The YCSB workload was modified to issue
+10 operations per transaction, with a default 50:50 read:write mix. Each
+operation is assumed to be eventual consistent. \cf{NoTxn} corresponds to a
+configuration that does not use transactions. Compared to this, RC is only
+12\% shower in terms of latency with 512 clients, whereas RR is 2.3X
+slower. The difference between RC and \cf{NoTxn} is due to the meta-data
+overhead of recording transaction information in the object state. For RR
+transactions, the cost of capturing and maintaining a snapshot is the
+biggest source of overhead.
 
 We also compared (not shown) the performance of EC LWW operations directly
-against Cassandra (our backing store), which uses last-writer-wins as the only
-convergence semantics. While Cassandra provides no stronger-than-eventual
-consistency properties, \name was within 30\%(20\%) of latency(throughput) of
-Cassandra with 512 clients, illustrating that the programmers only have to pay
-a minimal overhead for the expressive and stronger \name programming model.
+against Cassandra, which uses last-writer-wins as the only convergence
+semantics. While Cassandra provides no stronger-than-eventual consistency
+properties, \name was within 30\%(20\%) of latency(throughput) of Cassandra
+with 512 clients, supporting our thesis that programmers only have to incur
+relatively low overhead for a more expressive programming model which
+provides stronger provable consistency guarantees.
 
-Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
-configuration against a single replica (NoRep) and strongly replicated
-(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
-15\% read-write interactions, which is representative of the auction workload.
-Without replication, NoRep trivially provides strong consistency. However, this
-deployment does not scale beyond 1750 operations per second. Strong replication
-offers better throughput at the cost of greater latency due to inter-replica
-coordination. \name deployment offers the benefit of replication, while only
-paying the cost of coordination when necessary.
+Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a
+\cf{1DC} configuration against a single replica ({\sf NoRep}) and a strongly
+replicated ({\sf StrongRep}) \cf{1DC} deployment. The benchmark uses the
+default RUBiS bidding mix, which has 15\% read-write interactions, which is
+representative of the auction workload.  Without replication, {\sf NoRep}
+trivially provides strong consistency. However, this deployment does not
+scale beyond 1750 operations per second. Strong replication offers better
+throughput at the cost of greater latency due to inter-replica
+coordination. The \name deployment offers the benefit of replication, while
+only paying the cost of coordination when necessary.
 
 Finally, we study the impact of summarization in
-Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
-replica, with all the clients operating on the \emph{same} LWW register to
-stress test the summarization mechanism. The shim layer cache (mem) of
-operations is summarized every 64 updates, while the updates in the backing
-store (disk) are summarized every 4096 updates. Each point in the graph
-represents the average latency of the previous 1000 operations. Each experiment
-is run for 60s.
-
-The results show that without summarization, the average latency of operations
-increase exponentially to almost 1 second, and only 13K operations were
-performed in a minute. Since every operation has to reduce over the set of all
-previous operations, with a ever growing set, the operations take increasingly
-more time to complete. With summarization only in memory, the performance still
-degrades due to the cost of fetching all previous updates from the backing
-store into the shim layer. Fetching the latest updates is essential for SC
-operations. With both summarizations enabled, we see that the latency does not
-increase over time, and we were able to perform 67K operations. This graph
-illustrates the importance and effectiveness of summarization in \name.
+Figure~\ref{grf:summarization}. We use 128 clients and a single \name
+replica, with all clients operating on the \emph{same} LWW register to
+stress test the summarization mechanism. The shim layer cache (memory) is
+summarized every 64 updates, while the updates in the backing store (disk)
+are summarized every 4096 updates. Each point in the graph represents the
+average latency of the previous 1000 operations. Each experiment is run for
+one minute.  Without summarization, the average latency of operations
+increases exponentially to almost one second, and only 13K operations were
+performed in a minute. Since every operation has to reduce over the set of
+all previous operations, operations take increasingly more time to complete
+since they must contend with an ever growing set.  With summarization only
+in memory, performance still degrades due to the cost of fetching all
+previous updates from the backing store into the shim layer. Fetching the
+latest updates is essential for SC operations. With summarization enabled on
+both disk and memory, latency does not increase over time, and the
+implementation realizes throughput of 67K operations/minute.

### 2014-11-07 implementation.tex
edits

@@
 
-Replicated data types and the stronger consistency semantics are implemented
-and enforced in the \emph{shim layer}. Our implementation supports eventual,
+Replicated data types and various consistency semantics are implemented and
+enforced in the \emph{shim layer}. Our implementation supports eventual,
 causal, and strong consistency for data type operations, and RC, MAV, and RR
-semantics for transactions.  This functionality is implemented entirely on top
-of the standard interface exposed by Cassandra. From an engineering
-perspective, leveraging an off-the-shelf data store enables an implementation
-comprising roughly only 2500 lines of Haskell code, which is packaged as a
-library.
+semantics for transactions.  This functionality is implemented entirely on
+top of the standard interface exposed by Cassandra. From an engineering
+perspective, leveraging an off-the-shelf data store enables an
+implementation comprising roughly only 2500 lines of Haskell code, which is
+packaged as a library.
 
@@
 
-Each new effect in \name is realized as a row insertion in Cassandra, and the
-state of an object is the set of all corresponding rows. The shim layer
+Each new effect in \name is realized as a row insertion in Cassandra, and
+the state of an object is the set of all corresponding rows. The shim layer
 maintains a causally consistent in-memory snapshot of a subset of objects in
 the backing store, by explicitly tracking dependencies introduced between
-effects due to visibility, session and same transaction relations. Dependence
-tracking is similar to the techniques presented in~\cite{BoltOn}
-and~\cite{Eiger}. Because Cassandra provides durability, convergence and fault
-tolerance, each shim layer node simply acts as a soft-state cache, do not
-communicate directly with each other, and can safely be terminated at any
-point. Similarly, new shim layer nodes can be spawned on demand.
+effects due to visibility, session and same transaction
+relations. Dependence tracking is similar to the techniques presented
+in~\cite{BoltOn} and~\cite{Eiger}. Because Cassandra provides durability,
+convergence, and fault tolerance, each shim layer node simply acts as a
+soft-state cache, with no inter-node communication, and can safely be
+terminated at any point. Similarly, new shim layer nodes can be spawned on
+demand.
 
@@
 strongly consistent operations. Strongly consistent operations are performed
-after obtaining exclusive leases on objects. The lease mechanism is implemented
-with the help of Cassandra's support for conditional updates and expiring
-columns. Cassandra does not provide general-purpose transactions. Since the
-transaction guarantees provided by \name are
+after obtaining exclusive leases on objects. The lease mechanism is
+implemented with the help of Cassandra's support for conditional updates and
+expiring columns. Cassandra does not provide general-purpose
+transactions. Since the transaction guarantees provided by \name are
 coordination-free~\cite{BailisHAT}, we realize efficient implementations by
-explicitly tracking the dependencies between operations and transactions.
+explicitly tracking dependencies between operations and transactions.
 Importantly, the weaker isolation semantics of transactions in \name permit
@@
 set of effects grows with every effectful operation. If left unchecked,
-performance slow down over time, until the shim layer memory or backing store
+performance slows down over time, until the shim layer memory or backing store
 disk runs out of memory. Luckily, the state of the operation-based replicated
@@
 
-Performing summarization in the backing store is more complicated; the whole
-process needs to be atomic from a client's perspective, but Cassandra does not
-provide multi-row transactions. We have engineered and implemented a scalable
-summarization mechanism for the backing that permits concurrent client
-operations, but nonetheless prohibits these operations from witnessing
-intermediate states of the summarization process.
+Performing summarization in the backing store is more complicated since the
+whole process needs to be atomic from a client's perspective, but Cassandra
+does not provide multi-row transactions. We have engineered and implemented
+a scalable summarization mechanism for the backing store that permits
+concurrent client operations, but nonetheless prohibits these operations
+from witnessing intermediate states of the summarization process.

### 2014-11-07 evaluation.tex
Minor edit to evaluation

@@
 Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a
-\cf{1DC} configuration against a single replica ({\sf NoRep}) and a strongly
-replicated ({\sf StrongRep}) \cf{1DC} deployment. The benchmark uses the
+\cf{1DC} configuration against a single replica ({\sf\small NoRep}) and a strongly
+replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
 default RUBiS bidding mix, which has 15\% read-write interactions, which is
-representative of the auction workload.  Without replication, {\sf NoRep}
+representative of the auction workload.  Without replication, {\sf\small NoRep}
 trivially provides strong consistency. However, this deployment does not

### 2014-11-08 contract-language.tex
Definition of weaker-than corrected.

@@
 A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
-\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
+\le \cv_{st}$ ) if and only if $\Delta \vdash \forall \cureff. \cv_{st} \Rightarrow \cv_{op}$.
 \end{definition}
@@
 Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
-proof of the theorem is given below:
-
-\begin{proof}
-  Hypothesis:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
-    {\E', \sigma \pll \Sigma} & H\npp\\
-    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Since $\tau$ is the contract class of $\cv$, by inversion, we have
-  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
-  \end{array}
-  \end{smathpar}
-   Since $\eff$ denotes new effect, it is a fresh variable that does
-   not occur free in $\Delta$. From $H2$, after instantiating bound
-   $\cureff$ with $\eff$, we have:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
-      & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Due to the soundness of natural deduction for first-order logic,
-  $H3$ implies that for all models $\mathcal{M}$ such that
-  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
-  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
-  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
-  \begin{smathpar}
-  \begin{array}{cr}
-    \E' \models \Delta & H\npp\\
-  \end{array}
-  \end{smathpar}
-  Proof follows from $H1$ and $H4$.
-  \qed
-\end{proof}
-
-It is important to note that Theorem~\ref{thm:classification-sound}
-does not ascribe any semantics to the reduction relation
-($\xrightarrow{}$). As such, it makes no assumptions about how the
-store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
-guarantees. The specific implementation strategy is determined by the
-operational semantics of the store, which \emph{defines} the reduction
-relation for that particular store. The following section describes
-operational semantics of the store used by the \name implementation.
+proof of the theorem is given in the supplementary material\footnote{The
+supplementary material also provides the concrete reduction rules for enforcing
+the consistency classes.}.

### 2014-11-08 contract-language.tex
Added Delta axioms

--- 2014-11-07 contract-language.tex
@@
 
-\noindent The $\Delta$ referred in the above defintion is a conjunction of
-assumptions about the nature of primitive relations, e.g., $\Rvis$ and $\Rso$
-are irreflexive, and the happens-before relation $\hbZ$ is acyclic. A
-\emph{well-formed} axiomatic execution ($\E$) is expected to satisfy these
-assumptions (i.e., $\E \models \Delta$).
+\noindent The $\Delta$ referred in the above definition is a conjunction of
+following assumptions about the nature of primitive relations:
 
-If the contract ($\cv_{op}$) of an operation ($\mathit{op}$) is \emph{weaker
-than} a store contract ($\cv_{st}$), then constraints expressed by the former
-are implied by guarantees provided by the latter. The completeness of
-first-order logic allows us to assert that any well-formed execution ($\E$)
-that satisfies $\cv_{st}$ (i.e., $\E\models \cv_{st}$) also satisfies
-$\cv_{op}$ (i.e., $\E \models \cv_{op}$). Consequently, it is safe to execute
-operation $\mathit{op}$ under a store consistency level captured by $\cv_{st}$.
+\begin{itemize}
+\item $\visZ$ relates actions on the same object:
+  \begin{itemize}
+    \item $\forall a,b. \visZ(a,b) \Rightarrow \sameobjZ(a,b)$
+  \end{itemize}
+\item $\soZ$ is a transitive relation:
+  \begin{itemize}
+    \item $\forall a,b,c. \soZ(a,b) \conj \soZ(b,c) \Rightarrow \soZ(a,c)$
+  \end{itemize}
+\item $\sameobjZ$ is an equivalence relation:
+  \begin{itemize}
+    \item $\forall a. \sameobjZ(a,a)$
+    \item $\forall a,b. \sameobjZ(a,b) \Rightarrow \sameobjZ(b,a)$
+    \item $\forall a,b,c. \soZ(a,b) \conj \soZ(b,c) \Rightarrow \soZ(a,c)$
+  \end{itemize}
+\end{itemize}
+
+\noindent A \emph{well-formed} axiomatic execution ($\E$) is expected
+to satisfy these assumptions:
+
+\begin{definition}
+An axiomatic execution $\E$ is said to be well-formed if and only if
+$\E \models \Delta$.
+\end{definition}
+
+\noindent If the contract ($\cv_{op}$) of an operation ($\mathit{op}$)
+is \emph{weaker than} a store contract ($\cv_{st}$), then constraints
+expressed by the former are implied by guarantees provided by the
+latter. The completeness of first-order logic allows us to assert that
+any well-formed execution ($\E$) that satisfies $\cv_{st}$ (i.e.,
+$\E\models \cv_{st}$) also satisfies $\cv_{op}$ (i.e., $\E \models
+\cv_{op}$). Consequently, it is safe to execute operation
+$\mathit{op}$ under a store consistency level captured by $\cv_{st}$.
 
@@
   \begin{array}{cr}
+    \Delta \vdash \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\cureff$ occurs free in $\cv$ and $\cv_\tau$, but not in
+  $\Delta$, we introduce quantifier to bind $\cureff$:
+  \begin{smathpar}
+  \begin{array}{cr}
     \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
@@
    Since $\eff$ denotes new effect, it is a fresh variable that does
-   not occur free in $\Delta$. From $H2$, after instantiating bound
+   not occur free in $\Delta$. From $H3$, after instantiating bound
    $\cureff$ with $\eff$, we have:
@@
   Due to the soundness of natural deduction for first-order logic,
-  $H3$ implies that for all models $\mathcal{M}$ such that
+  $H4$ implies that for all models $\mathcal{M}$ such that
   $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
@@
   \end{smathpar}
-  Proof follows from $H1$ and $H4$.
+  Proof follows from $H1$, $H5$, and $H4$.
   \qed

### 2014-11-08 store-opsem.tex
Added proof for Soundness Modulo Classification

@@
 An operation $op$ performed at replica $r$ can only witness the set of
-effects ($\Theta(r)$) visible at $r$. For the sake of clarity, we only
+effects ($\Theta(r)$) visible at $r$. To avoid too many paranthesis in
+our formal presentation, we represent operation $op$, whose contract
+is in consistency class $\tau$, as $op_\tau$ instead of the usual
+$\langle op,\tau \rangle$. For the sake of clarity, we only
 consider a single replicated object of well-defined type (for eg: a
@@
 
+% First, we prove few useful lemmas:
+
+% \begin{lemma}[]
+% \end{lemma}
+
 \begin{theorem}[Soundness Modulo Classification]
@@
 \begin{smathpar}
-(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+(\E, \Theta, \langle s,i,op_\tau ::
 \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-\langle s,i,\sigma \rangle \pll \Sigma')
+\langle s,i+1,\sigma \rangle \pll \Sigma)
 \end{smathpar}
@@
 \begin{proof}
-  By case analysis on $(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+  By case analysis on the derivation:
+  \begin{smathpar}
+    (\E, \Theta, \langle s,i,op_\tau ::
     \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-  \langle s,i,\sigma \rangle \pll \Sigma')$. 
-  \GK{Lets first finalize all notations, rules, and also agree on
-  all theorem statements. This way we don't have to change the proof
-  again.} 
+    \langle s,i+1,\sigma \rangle \pll \Sigma)
+  \end{smathpar}
+  Cases:
+  \begin{itemize}
+    \item Case \rulelabel{EC}: Hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+      \tau = {\sf EventuallyConsistent} & H\npp \\
+      \auxred{\Theta}{(\E,\langle s,i,op \rangle)} {r} 
+        {(\E',\eff)} & H\npp\\
+      \end{array}
+      \end{smathpar}
+      Goal (\romannumeral 1) follows from $H1$ and
+      Lemma~\ref{lem:auxred-wf}. Goal (\romannumeral 2) is the
+      following: 
+      \begin{smathpar}
+      \begin{array}{cr}
+        \E' \models \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\eff}
+          \Rightarrow \vis{a}{\eff} & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
+      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$.  By inversion
+      on $H1$, we get the following hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
+        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
+        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
+        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
+      \end{array}
+      \end{smathpar}
+       Since $\E'$ defines $\EffSoup'$ as the universe of values.
+       Therefore, the goal can be rewritten:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \forall (a,b\in\EffSoup').
+        \msentails{E'}{\De{\hbo{a}{b} \wedge
+        \vis{b}{\eff}} \Rightarrow 
+          \De{\vis{a}{\eff}}} & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      New hypotheses after \emph{intros}:
+      \begin{smathpar}
+      \begin{array}{cr}
+        a \in \EffSoup' & H\npp\\
+        b \in \EffSoup' & H\npp\\
+      \end{array}
+      \end{smathpar}
+      And new goal:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \msentails{E'}{\De{\hbo{a}{b} \wedge
+        \vis{b}{\eff}} \Rightarrow 
+          \De{\vis{a}{\eff}}} & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      Since $(\mathcal{M} \models A \Rightarrow B) \Leftrightarrow
+      (\mathcal{M} \models A \Rightarrow \mathcal{M} \models B)$, we
+      prove $G0$ by proving: 
+      \begin{smathpar}
+      \begin{array}{cr}
+        (\E' \models ~\hbo{a}{b} \wedge \vis{b}{\eff})
+          \Rightarrow (\E' \models \vis{a}{\eff}) & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      After intros:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \E' \models ~\hbo{a}{b} \wedge \vis{b}{\eff} &
+        H\npp\\
+      \end{array}
+      \end{smathpar}
+      Since $\E'$ defines $\hboZ'$ and $\Rvis'$ as interpretations for
+      $\hboZ$ and $\Rvis$ respectively, we have:
+      \begin{smathpar}
+      \begin{array}{cr}
+          {{\sf hbo'}(a,b)} \wedge \Rvis'(b,\eff) & H\npp\\
+      \end{array}
+      \end{smathpar}
+      And the goal is:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \Rvis'(a,\eff) & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      Inversion on $H9$:
+      \begin{smathpar}
+      \begin{array}{cr}
+          {{\sf hbo'}(a,b)} & H\npp\\
+          \Rvis'(b,\eff) & H\npp\\
+      \end{array}
+      \end{smathpar}
+      Since $\eff$ is unique, from $H3$ and $H11$ we have the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+            b\in\Theta(r)& H\npp\\
+      \end{array}
+      \end{smathpar}
+      Since local state (i.e., $\Theta(r)$) is causally consistent
+      (lemma~\ref{lem:local-state-is-cc}) using $H10$ and $H12$ we
+      derive the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        a \in \Theta(r) & H\npp\\
+      \end{array}
+      \end{smathpar}
+      Now, from $H3$ and $H13$, we deduce:
+      \begin{smathpar}
+      \begin{array}{cr}
+        (a,\eff) \in \Rvis'& \\
+      \end{array}
+      \end{smathpar}
+      which is what needs to be proven ($G4$).
+
+    \item Case $\rulelabel{CC}$: Hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \tau = {\sf CausallyConsistent} & H\npp \\
+        \auxred{\Theta}{(\E,\langle s,i,op \rangle)} {r} 
+          {(\E',\eff)} & H\npp\\
+         \E'.\Rso(\eff) \subseteq \Theta(r)& H\npp\\
+      \end{array}
+      \end{smathpar}
+     Goal (\romannumeral 1) follows from $H15$ and
+     Lemma~\ref{lem:auxred-wf}. We now prove Goal (\romannumeral 2).
+     Expanding the definition of $\ccc$, goal is the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \ccc = \forall a.~\hbo{a}{\eff} \Rightarrow
+        \vis{a}{\eff} & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
+      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$. By inversion
+      on $H15$, we get the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
+        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
+        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
+        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
+      \end{array}
+      \end{smathpar}
+%     From $H16$ and $H19$, we have:
+%     \begin{smathpar}
+%     \begin{array}{cr}
+%       \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i} \in \Theta(r) &
+%       H\npp \\
+%     \end{array}
+%     \end{smathpar}
+      By similar reasoning as with the previous case (\rulelabel{EC}),
+      we get a context with following hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+        a \in \EffSoup' & H\npp\\
+        {\hboZ'(a,\eff)}  & H\npp\\
+      \end{array}
+      \end{smathpar}
+      And the goal is the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \Rvis'(a,\eff) & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      Since \emph{happens-before} is transitive, by inversion on
+      $H23$, we get two cases:
+      \begin{itemize}
+        \item SCase 1: Hypotheses:
+        \begin{smathpar}
+        \begin{array}{cr}
+          (\Rso' \cup \Rvis')(a,\eff) & H\npp\\
+        \end{array}
+        \end{smathpar}
+        Inversion on $H24$ leads to two subcases. In one case, we
+        assume $\Rvis'(a,\eff)$ and try to prove the goal $G6$. The
+        proof for this case mimics the proof for Case \rulelabel{EC}.
+        Alternatively, in second case, we assume:
+        \begin{smathpar}
+        \begin{array}{cr}
+          \Rso'(a,\eff) & H\npp\\
+        \end{array}
+        \end{smathpar}
+        and prove $G6$. From $H25$ and $H16$, we infer:
+        \begin{smathpar}
+        \begin{array}{cr}
+          a \in \Theta(r) & H\npp\\
+        \end{array}
+        \end{smathpar}
+        Now, from $H26$ and $H18$ we know:
+        \begin{smathpar}
+        \begin{array}{cr}
+          (a,\eff) \in \Rvis' & \\
+        \end{array}
+        \end{smathpar}
+        which is the goal ($G6$).
+
+        \item SCase 2: Hypotheses (after abbreviating the occurance
+        of$(\Rso' \cup \Rvis')^{+}$ as {\sf hbo'}):
+        \begin{smathpar}
+        \begin{array}{cr}
+          \exists(c\in\EffSoup').{\sf hbo'}(a,c) \wedge (\Rso' \cup
+          \Rvis')(c,\eff) & H\npp\\
+        \end{array}
+        \end{smathpar}
+        Inverting $H27$, followed by expanding $\Rso' \cup \Rvis'$:
+        \begin{smathpar}
+        \begin{array}{cr}
+          c\in\EffSoup' & H\npp\\
+           {\sf hbo'}(a,c) \wedge (\Rso'(c,\eff) \vee \Rvis'(c,\eff))& H\npp\\
+        \end{array}
+        \end{smathpar}
+        Inverting the disjunction in $H29$, we get two cases:
+        \begin{itemize}
+          \item SSCase R: Hypothesis is
+          \begin{smathpar}
+          \begin{array}{cr}
+             {\sf hbo'}(a,c) \wedge \Rvis'(c,\eff)& H\npp\\
+          \end{array}
+          \end{smathpar}
+          Observe that hypothesis $H30$ and current goal ($G6$) are
+          same as hypothesis $H9$ and goal ($G4$) in Case
+          \rulelabel{EC}. The proof for this SSCase is also the same.
+
+          \item SSCase L: Hypothesis is
+          \begin{smathpar}
+          \begin{array}{cr}
+             {\sf hbo'}(a,c) \wedge \Rso'(c,\eff)& H\npp\\
+          \end{array}
+          \end{smathpar}
+          Inverting $H31$:
+          \begin{smathpar}
+          \begin{array}{cr}
+             {\sf hbo'}(a,c)& H\npp\\
+             \Rso'(c,\eff)& H\npp\\
+          \end{array}
+          \end{smathpar}
+          From $H33$ and $H16$, we infer:
+          \begin{smathpar}
+          \begin{array}{cr}
+            c \in \Theta(r) & H\npp \\
+          \end{array}
+          \end{smathpar}
+          Since local state is causally consistent
+          (lemma~\ref{lem:local-state-is-cc}), from $H32$ and $H34$,
+          we have:
+          \begin{smathpar}
+          \begin{array}{cr}
+            a \in \Theta(r) & H\npp\\
+          \end{array}
+          \end{smathpar}
+           Proof follows from $H35$ and $H18$.
+        % End of SSCase
+        \end{itemize}
+      % End of SCases
+      \end{itemize}
+
+    \item Case \rulelabel{SC}: Hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \tau = {\sf StronglyConsistent} & H\npp \\
+        \auxred{\Theta} {(\E,\langle s,i,op \rangle)} {r}
+        {(\E',\eff)} & H\npp\\ 
+        \E.A \subseteq \Theta(r) & H\npp\\
+        \dom(\Theta') = \dom(\Theta) & H\npp\\
+        \forall r'\in \dom(\Theta'). \Theta'(r') = \Theta(r') \cup
+          \{\eff\} & H\npp\\
+      \end{array}
+      \end{smathpar}
+      Let $\EffSoup' = \E'.\EffSoup$, $\Rvis' = \E'.\Rvis$, $\Rso' =
+      \E'.\Rso$, and $\hboZ' =(\Rso' \cup \Rvis')^{+}$. Inversion on
+      $H37$ gives: 
+      \begin{smathpar}
+      \begin{array}{cr}
+        \EffSoup' = \EffSoup \cup \{\eff\} & H\npp\\
+        \visZ' = \Theta(r)\times\eff ~\cup~ \visZ & H\npp\\
+        \Rso' = \EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times\eff ~\cup~ \Rso & H\npp\\
+        \sameobjZ' = \EffSoup' \times \EffSoup'& H\npp\\
+      \end{array}
+      \end{smathpar}
+       Goal (\romannumeral 1) follows from $H36$ and
+       Lemma~\ref{lem:auxred-wf}. Expanding the definition of $\scc$,
+       goal is the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \forall a.~\sameobj{a}{\eff} \Rightarrow \vis{a}{\eff} 
+          ~\vee~ \vis{\eff}{a} ~\vee~ a = \eff & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      By similar reasoning as with the \rulelabel{EC} case,
+      we get a context with following hypotheses:
+      \begin{smathpar}
+      \begin{array}{cr}
+        a \in \EffSoup' & H\npp\\
+        {\sameobjZ'(a,\eff)}  & H\npp\\
+      \end{array}
+      \end{smathpar}
+      And the goal is the following:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \vis{a}{\eff} ~\vee~ \vis{\eff}{a} ~\vee~ a = \eff & G\mpp\\
+      \end{array}
+      \end{smathpar}
+      From $H40$ and $H44$, we know that either $a \in \EffSoup$ or $a
+      = \eff$. If $a = \eff$, then $G8$ follows trivially. On the
+      other hand, when $a \in \EffSoup$, from $H37$ we know that $a
+      \in \Theta(r)$. From $H41$, we derive:
+      \begin{smathpar}
+      \begin{array}{cr}
+        a \in \visZ'(\eff)& \\
+      \end{array}
+      \end{smathpar}
+      Which means first disjunct of $G8$ is valid, proving the
+      theorem.
+  % End of all cases    
+  \end{itemize}
+
 \end{proof}
@@
 \begin{smathpar}
-(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+(\E, \Theta, \langle s,i,op_\tau :: 
 \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-\langle s,i,\sigma \rangle \pll \Sigma')
+\langle s,i+1,\sigma \rangle \pll \Sigma)
 \end{smathpar}

### 2014-11-09 store-opsem.tex
All proofs added

@@
 
-% First, we prove few useful lemmas:
+First, we prove a useful lemma:
 
-% \begin{lemma}[]
-% \end{lemma}
+\begin{lemma}[Auxiliary Reduction Preserves Well-Formedness]
+For every well-formed execution state $\E$, if
+$\auxred{\Theta}{(\E,\langle s,i,op \rangle)} {r} {(\E',\eff)}$, then
+$\E'$ is well-formed.
+\end{lemma}
+\begin{proof}
+  Let us denote $\E.\EffSoup$, $\E.\Rvis$, $\E.\Rso$, and $\E.\sameobjZ$
+  as $\EffSoup$, $\Rvis$, $\Rso$, and $\sameobjZ$ respectively.
+  Likewise, let us denote $\E'.\EffSoup$, $\E'.\Rvis$, $\E'.\Rso$, and
+  $\E'.\sameobjZ$ as $\EffSoup$, $\Rvis$, $\Rso$, and $\sameobjZ$
+  respectively. By inversion on $\auxred{\Theta}{(\E,\langle s,i,op
+  \rangle)} {r} {(\E',\eff)}$, we have following hypotheses:
+  \begin{smathpar}
+  \begin{array}{cr}
+    r \in \dom(\Theta) & H\npp\\
+    ctxt = {\ctxtFn}^{*}(\Theta(r)) & H\npp\\
+    \Ops(op)(ctxt) {\rdtredsto}^{*} v & H\npp\\
+    \eff = (s,i,op,v) & H\npp\\
+    \{\eff'\} = \EffSoup_{({\sf SessID}=s,\,{\sf SeqNo}=i-1)}\\
+    \EffSoup' = \{\eff\} \cup \EffSoup & H\npp\\
+    \visZ' = \Theta(r)\times\{\eff\} \cup \visZ & H\npp\\
+    \soZ' = (\soZ^{-1}(\eff') \cup \eff') \times\{\eff\} \cup \soZ
+      & H\npp \\
+    \sameobjZ' = \EffSoup' \times \EffSoup' & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\E$ is well-formed, from the definition of
+  well-formedness and \emph{models} relation, we have the following:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a \in \EffSoup). \neg\hboZ(a,a) & H\npp\\
+    \forall (a,b \in \EffSoup). \visZ(a,b) \Rightarrow 
+      \sameobjZ(a,b) & H\npp\\
+    \forall (a,b,c \in \EffSoup). \soZ(a,b) \conj \soZ(b,c) \Rightarrow
+      \soZ(a,c) & H\npp\\
+    \forall (a \in \EffSoup). \sameobjZ(a,a) & H\npp\\
+    \forall (a,b \in \EffSoup). \sameobjZ(a,b) \Rightarrow 
+      \sameobjZ(b,a) & H\npp\\
+    \forall (a,b,c \in \EffSoup). \soZ(a,b) \conj \soZ(b,c) \Rightarrow
+      \soZ(a,c) & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\hboZ = (\Rvis \cup \Rso)^{+}$, $H8$ is equivalent to
+  conjunction of following assertions:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a \in \EffSoup). \neg\visZ(a,a) & H\npp\\
+    \forall (a \in \EffSoup). \neg\soZ(a,a) & H\npp\\
+    \forall (a,b \in \EffSoup). \neg (\visZ(a,b) \conj \soZ(b,a)) & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\eff$ is fresh, $\eff \notin \Theta(r)$. From $H4$, $H5$, and
+  $H14$, we have the acyclicity property for $\visZ'$:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a \in \EffSoup'). \neg\visZ'(a,a) & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Also, $\eff \notin \EffSoup$, and from $H4$, $H6$ and $H15$, we have
+  acyclicity for $\soZ'$: 
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a \in \EffSoup'). \neg\soZ'(a,a) & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Similarly, from the uniqueness of $\eff$, and $H5$, $H6$, and $H16$,
+  we have the following:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a,b \in \EffSoup'). \neg (\visZ'(a,b) \conj \soZ'(b,a)) & H\npp\\
+  \end{array}
+  \end{smathpar}
+  From $H17-19$, we prove the acyclicity of $\hboZ'$: 
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a \in \EffSoup').\neg\hboZ(a,a) & G\mpp\\
+  \end{array}
+  \end{smathpar}
+  The $\sameobjZ'$ relation is simply the cross product
+  $\EffSoup' \times \EffSoup'$. Hence following trivially hold:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a,b \in \EffSoup'). \visZ'(a,b) \Rightarrow 
+      \sameobjZ'(a,b) & G\mpp\\
+    \forall (a \in \EffSoup'). \sameobjZ'(a,a) & G\mpp\\
+    \forall (a,b \in \EffSoup'). \sameobjZ'(a,b) \Rightarrow 
+      \sameobjZ'(b,a) & G\mpp\\
+  \end{array}
+  \end{smathpar}
+  Finally, from $H4$, $H6$ and $H10$, we have transitivity for $\soZ'$:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \forall (a,b,c \in \EffSoup'). \soZ'(a,b) \conj \soZ'(b,c) \Rightarrow
+      \soZ'(a,c) & G\mpp\\
+  \end{array}
+  \end{smathpar}
+  Well-formedness of $\E'$ follows from $G0-4$.
+\end{proof}
+
+We now define causal consistency property of the store formally:
+
+\begin{definition}
+  A store $\Theta$ is said to be causally consistent under an execution
+  $\E=\Exec$ if and only if:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \hspace*{-0.5in}\forall (r\in{\sf dom}(\Theta)). \forall (\eff \in
+      \Theta(r)). & \\
+    \hspace*{0.3in}\forall (a\in \EffSoup). \hbo{a}{\eff} \Rightarrow a 
+      \in \Theta(r) & H\npp \\
+  \end{array}
+  \end{smathpar}
+  Where, $\hboZ = (\Rvis \cup \Rso)^{+}$
+\end{definition}
 
@@
 \label{thm:weak-soundness}
-For every well-formed execution state $\E$, and contract class $\tau 
-\in \{{\sf ec},{\sf cc},{\sf sc}\}$, If:
+For every well-formed execution state $\E$, for every store $\Theta$
+that is causally consistent under $\E$, and for every contract class
+$\tau \in \{{\sf ec},{\sf cc},{\sf sc}\}$, if:
 \begin{smathpar}
@@
       Goal (\romannumeral 1) follows from $H1$ and
-      Lemma~\ref{lem:auxred-wf}. Goal (\romannumeral 2) is the
+      lemma~\ref{lem:auxred-wf}. Goal (\romannumeral 2) is the
       following: 
@@
       \end{smathpar}
-      Since local state (i.e., $\Theta(r)$) is causally consistent
-      (lemma~\ref{lem:local-state-is-cc}) using $H10$ and $H12$ we
-      derive the following:
+      Since $a,b \neq \eff$, we have that $\hboZ'(a,b) \Rightarrow
+      \hboZ(a,b)$.  Since $\Theta(r)$ is causally consistent under
+      $\E$, using $H10$ and $H12$ we derive the following:
       \begin{smathpar}
@@
           \end{smathpar}
-          Since local state is causally consistent
-          (lemma~\ref{lem:local-state-is-cc}), from $H32$ and $H34$,
-          we have:
+          Since $a,c \neq \eff$, we know that $\hboZ'(a,c) \Rightarrow
+          \hboZ(a,c)$.  Since $\Theta(r)$ is causally consistent under
+          $\E$, using $H32$ and $H34$ we derive the following:
           \begin{smathpar}
@@
   \end{itemize}
-
 \end{proof}
@@
   \label{thm:soundness}
-For every well-formed execution state $\E$, and contract class $\tau
+  For every well-formed execution state $\E$, for every store $\Theta$
+  that is causally consistent under $\E$, for every contract class $\tau
   \in \{{\sf ec},{\sf cc},{\sf sc}\}$, and for every consistency
@@
 \end{proof}
+
+We now show that every configuration of the store that is reachable
+via the reduction relation ($\xrightarrow{}$) is causally consistent.
+
+\begin{lemma}[Causal Consistency Preservation]
+  \label{lem:local-state-is-cc}
+  For every well-formed execution state $\E$, and a store $\Theta$
+  that is causally consistent under $\E$, if: 
+  \begin{smathpar}
+    (\E, \Theta, \langle s,i,op_\tau ::
+    \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
+    \langle s,i+1,\sigma \rangle \pll \Sigma)
+  \end{smathpar}
+  then $\Theta'$ is causally consistent under $\E'$.
+\end{lemma}
+\begin{proof}
+  By case analysis on on:
+  \begin{smathpar}
+    (\E, \Theta, \langle s,i,op_\tau ::
+    \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
+    \langle s,i+1,\sigma \rangle \pll \Sigma)
+  \end{smathpar}
+  Cases:
+  \begin{itemize}
+    \item Case \rulelabel{EffVis}: Final execution state is same as
+    the initial (i.e., $\E' = \E$). Therefore, we need to prove that
+    new store configuration $\Theta$ is causally consistent under
+    $\E=\Exec$. Hypotheses:
+    \begin{smathpar}
+    \begin{array}{cr}
+      r \in {\sf ReplID} & H\npp \\
+      \eff \in A & H\npp \\
+      \eff \notin \Theta(r)& H\npp \\
+      \E.\visZ^{-1}(\eff) \cup \E.\soZ^{-1}(\eff) \subseteq \Theta(r)
+      & H\npp \\
+      \Theta' = \Theta \cup [r \mapsto \{\eff\} \cup \Theta(r)] & H\npp \\
+      \hspace*{-0.5in}\forall (r\in{\sf dom}(\Theta)). \forall (\eff \in
+        \Theta(r)). & \\
+      \hspace*{0.3in}\forall (a\in\EffSoup). \hbo{a}{\eff} \Rightarrow a \in \Theta(r) & H\npp \\
+    \end{array}
+    \end{smathpar}
+    From $H4$ and $H5$, it suffices to prove:
+    \begin{smathpar}
+    \begin{array}{cr}
+      \forall (a\in \EffSoup). \hbo{a}{\eff} 
+        \Rightarrow a \in \Theta'(r) & G\mpp\\
+    \end{array}
+    \end{smathpar}
+    After \emph{intros}, hypotheses:
+    \begin{smathpar}
+    \begin{array}{cr}
+      a\in\EffSoup & H\npp\\
+      \hbo{a}{\eff} & H\npp \\
+    \end{array}
+    \end{smathpar}
+    Goal:
+    \begin{smathpar}
+    \begin{array}{cr}
+      a \in \Theta'(r) & G\mpp \\
+    \end{array}
+    \end{smathpar}
+    Inversion on $H7$ leads to two cases:
+    \begin{itemize}
+      \item SCase \emph{$a$ directly precedes $\eff$ }: Hypothesis:
+      \begin{smathpar}
+      \begin{array}{cr}
+        (\Rvis \cup \Rso)(a,\eff) & H\npp\\
+      \end{array}
+      \end{smathpar}
+      From $H3$ and $H8$, we conclude that $a \in \Theta'(r)$.
+
+      \item SCase \emph{$a$ transitively precedes $\eff$}: Hypothesis:
+      \begin{smathpar}
+      \begin{array}{cr}
+        \exists (c \in \EffSoup). \hbo{a}{c} \wedge (\Rvis \cup
+        \Rso)(c,\eff) & H\npp \\
+      \end{array}
+      \end{smathpar}
+      Inverting $H9$:
+      \begin{smathpar}
+      \begin{array}{cr}
+        c \in A & H\npp \\
+        \hbo{a}{c} & H\npp \\
+        (\Rvis \cup \Rso)(c,\eff) & H\npp \\
+      \end{array}
+      \end{smathpar}
+      From $H3$ and $H12$, we have:
+      \begin{smathpar}
+      \begin{array}{cr}
+        c \in \Theta'(r) & H\npp\\
+      \end{array}
+      \end{smathpar}
+      From $H5$, $H13$ and $H11$, we conclude that $a \in \Theta'(r)$.
+    \end{itemize}
+
+    \item Cases \rulelabel{EC} and \rulelabel{CC}: Store configuration
+    ($\Theta$) remains unchanged. Proof trivial.
+
+    \item Case \rulelabel{SC}: $\Sigma = \tuplee{s,\sigma} \pll
+    \Sigma_1$. Hypotheses:
+    \begin{smathpar}
+    \begin{array}{cr}
+      r \in {\sf ReplID} & H\npp \\
+      {\sf Unavailable}(\cv) & H\npp \\
+      \E_1;\Theta_1;\tuplee{s,op(v); \sigma} 
+        \;\xhookrightarrow{\eff,r}\; 
+      \E;\Theta_2;\tuplee{s,\sigma} & H\npp\\
+      \E_1.A \subseteq \Theta_1(r) & H\npp\\
+      \dom(\Theta) = \dom(\Theta_2) & H\npp\\
+      \forall r'\in \dom(\Theta). \Theta(r') = \Theta_2(r') \cup
+      \{\eff\} & H\npp\\
+    \end{array}
+    \end{smathpar}
+    Inductive hypothesis:
+    \begin{smathpar}
+    \begin{array}{cr}
+      \hspace*{-0.5in}\forall (r'\in{\sf dom}(\Theta_1)). \forall (\eff' \in
+        \Theta_1(r')). & \\
+      \hspace*{0.3in}\forall (a\in\E_1.A). \hbo{a}{\eff'} \Rightarrow a
+        \in \Theta_1(r') & IH\npp \\
+    \end{array}
+    \end{smathpar}
+    The goal:
+    \begin{smathpar}
+    \begin{array}{cr}
+      \hspace*{-0.5in}\forall (r'\in{\sf dom}(\Theta)). \forall (\eff' \in
+        \Theta(r')). & \\
+      \hspace*{0.3in}\forall (a\in\E.A). \hbo{a}{\eff'} \Rightarrow a
+        \in \Theta(r') & G\mpp \\
+    \end{array}
+    \end{smathpar}
+    Inverting $H16$:
+    \begin{smathpar}
+    \begin{array}{cr}
+        \E.\EffSoup = \E_1.\EffSoup \cup \{\eff\} & H\npp\\
+        \E.\visZ = \Theta_1(r)\times \{\eff\} ~\cup~ \E_1.\visZ & H\npp\\
+        \E.\Rso = \E_1.\EffSoup_{{\sf SessID}=s,\,{\sf SeqNo}<i}\times
+          \{\eff\} ~\cup~ \E_1.\Rso & H\npp\\
+        \Theta_2 = \Theta_1 & H\npp\\
+    \end{array}
+    \end{smathpar}
+    Using $H24$ to rewrite $H19$:
+    \begin{smathpar}
+    \begin{array}{cr}
+      \forall r'\in \dom(\Theta). \Theta(r') = \Theta_1(r') \cup
+      \{\eff\} & H\npp\\
+    \end{array}
+    \end{smathpar}
+    Using $H25$ and $IH20$, we can reduce the goal ($G2$) to:
+    \begin{smathpar}
+    \begin{array}{cr}
+      \forall (r'\in{\sf dom}(\Theta)). 
+      \forall (a\in\E.\EffSoup). \hbo{a}{\eff} \Rightarrow a
+        \in \Theta(r') & G\mpp \\
+    \end{array}
+    \end{smathpar}
+    After \emph{intros}, hypotheses:
+    \begin{smathpar}
+    \begin{array}{cr}
+      r' \in \dom(\Theta) & H\npp \\
+      a \in \E.\EffSoup & H\npp\\
+      \hbo{a}{\eff} & H\npp\\
+    \end{array}
+    \end{smathpar}
+    Goal:
+    \begin{smathpar}
+    \begin{array}{cr}
+      a \in \Theta(r) & G\mpp \\
+    \end{array}
+    \end{smathpar}
+    From $H27$ and $H21$, we know that either $a \in \E_1.\EffSoup$ or
+    $a = \eff$.
+    \begin{itemize}
+      \item If $a \in \E_1.\EffSoup$, then from $H17$, we know that $a
+      \in \Theta_1(r)$. However, from $H25$ we know that $\Theta_1(r)
+      \subset \Theta(r)$, which lets us conclude that $a \in
+      \Theta(r)$.
+
+      \item If $a = \eta$, then $H28$ is $\hbo{\eff}{\eff}$, which is
+      $false$ since {\sf hbo} is irreflexive.
+      \emph{ex\_falso\_quodlibet}.
+    \end{itemize}
+  % End of main cases  
+  \end{itemize}
+\qed
+\end{proof}

### 2014-11-09 implementation.tex
Integrating Gustavo\'s edits

@@
 store responsible for all data management issues (i.e., replication, fault
-tolerance, availability, and convergence).  Template Haskell is used implement
-static contract classification, and proof obligations are discharged with the
-help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod} illustrates the
-overall system architecture.
+tolerance, availability, and convergence).  Template Haskell is used to
+implement static contract classification, and proof obligations are discharged
+with the help of the Z3~\cite{Z3} SMT solver. Figure~\ref{fig:impl_mod}
+illustrates the overall system architecture.
 

### 2014-11-09 intro.tex
Integrating Gustavo\'s edits

@@
 availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
-modern web services, which aim to provide an "always on" experience,
+modern web services, which aim to provide an "always-on" experience,
 overwhelmingly favour availability and partition tolerance over strong

### 2014-11-09 motivation.tex
edits
Integrating Gustavo\'s edits

--- 2014-11-05 motivation.tex
@@
 \caption{Anomalies possible under eventual consistency for the get balance operation.}
-\label{fig:cleanliness_examples}
+\label{fig:ba_anomalies}
 \end{figure}
@@
 executed as a strongly consistent operation. Despite this strengthening,
-\cf{getBalance} operation may incorrectly report a negative balance to the end
-user. Consider the execution shown in fig.~\ref{fig:negativeBalanceAnomaly},
-which consists of three concurrent sessions performing a \cf{deposit}, a
-\cf{withdraw}, and a \cf{getBalance} operation, respectively, on the same bank
-account object. As the \cf{vis} edge indicates, operation \cf{withdraw(50)} in
-session 2, witnesses the effects of \cf{deposit(100)} from session 1, concludes
-that there is sufficient balance, and completes successfully. However, the
-\cf{getBalance} operation may only witness this successful withdraw, but not
-the causally preceding \cf{dopsit}, and reports the balance of negative 50 to
-the user.
+\cf{getBalance} operation may still incorrectly report a negative balance to
+the end user. Consider the execution shown in
+fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
+sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
+operation, respectively, on the same bank account object. As the \cf{vis} edge
+indicates, operation \cf{withdraw(50)} in session 2, witnesses the effects of
+\cf{deposit(100)} from session 1, concludes that there is sufficient balance,
+and completes successfully. However, the \cf{getBalance} operation may only
+witness this successful withdraw, but not the causally preceding \cf{dopsit},
+and reports the balance of negative 50 to the user.
 
@@
 levels by letting the programmer declare application-level consistency
-constraints as \emph{contracts} (Fig~\ref{fig:contract-lang}\footnote{\name
-exposes the contract construction language as a Haskell library}) that
-axiomatically specify the set of allowed executions involving this operation.
-In the case of the bank account, any execution that does not exhibit the
-anomalies described in the previous section is a \emph{well-formed} execution
-on the bank account object. By specifying the set of legal executions for each
-data type in terms of a trace of operation invocations on that type, \name\
-ensures that all executions over that type are well-formed.
+constraints as \emph{contracts}\footnote{\name exposes the contract
+construction language as a Haskell library} (Figure~\ref{fig:contract-lang})
+that axiomatically specify the set of allowed executions involving this
+operation. In the case of the bank account, any execution that does not exhibit
+the anomalies described in the previous section is a \emph{well-formed}
+execution on the bank account object. By specifying the set of legal executions
+for each data type in terms of a trace of operation invocations on that type,
+\name\ ensures that all executions over that type are well-formed.
 
@@
 \forall (a:\rcf{deposit}), (b:\rcf{withdraw}), (c: \rcf{deposit} \vee \rcf{withdraw}). \\
-\qquad \vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff} \\
-\qquad \wedge~ (\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff}
+\qquad (\vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}) \\
+\qquad \wedge~ ((\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff})
 \end{array}

### 2014-11-09 store-opsem.tex
Cleaned up proofs

--- 2014-11-07 store-opsem.tex
@@
 
-\begin{theorem}[Soundness Modulo Classification]
-\label{thm:weak-soundness}
-For every well-formed execution state $\E$, and contract class $\tau 
-\in \{{\sf ec},{\sf cc},{\sf sc}\}$, If:
+First, we prove a useful lemma:
+
+\input{auxred-wf-proof}
+
+We now define causal consistency property of the store formally:
+
+\begin{definition}
+  A store $\Theta$ is said to be causally consistent under an execution
+  $\E=\Exec$ if and only if:
   \begin{smathpar}
-(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
-\sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-\langle s,i,\sigma \rangle \pll \Sigma')
+  \begin{array}{cr}
+    \hspace*{-0.5in}\forall (r\in{\sf dom}(\Theta)). \forall (\eff \in
+      \Theta(r)). & \\
+    \hspace*{0.3in}\forall (a\in \EffSoup). \hbo{a}{\eff} \Rightarrow a 
+      \in \Theta(r) & H\npp \\
+  \end{array}
   \end{smathpar}
-then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
-$\E' \models \cv_{\tau}[\eff/\cureff]$
-\end{theorem}
-\begin{proof}
-  By case analysis on $(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
-  \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-  \langle s,i,\sigma \rangle \pll \Sigma')$. 
-  \GK{Lets first finalize all notations, rules, and also agree on
-  all theorem statements. This way we don't have to change the proof
-  again.} 
-\end{proof}
+  Where, $\hboZ = (\Rvis \cup \Rso)^{+}$
+\end{definition}
+
+The following theorem proves that our operational semantics correctly
+enforce $\ecc$, $\ccc$, and $\scc$ guarantees:
+
+% Main Theorem
+% ------------
+\input{soundness-theorem-proof}
+
+We now show that every configuration of the store that is reachable
+via the reduction relation ($\xrightarrow{}$) is causally consistent.
+
+\input{causal-consistency-proof}
 
@@
   \begin{smathpar}
-(\E, \Theta, \langle s,i,\langle op,\tau \rangle;
+  (\E, \Theta, \langle s,i,op_\tau :: 
   \sigma \rangle \pll \Sigma) \xrightarrow{\eff} (\E', \Theta',
-\langle s,i,\sigma \rangle \pll \Sigma')
+  \langle s,i+1,\sigma \rangle \pll \Sigma)
   \end{smathpar}
-then (\romannumeral 1) $\E'$ is well-formed, and (\romannumeral 2)
+  then (\romannumeral 1) $\E'$ is well-formed, (\romannumeral 2)
+  $\Theta'$ is causally consistent under $\E'$, and (\romannumeral 3)
   $\E' \models \cv[\eff/\cureff]$
@@
 \begin{proof}
-Follows from Theorems~\ref{thm:classification-sound} and
-\ref{thm:weak-soundness}
+  Follows from Theorems \ref{thm:classification-sound},
+  \ref{thm:local-state-is-cc}, and \ref{thm:weak-soundness}.
   \qed

### 2014-11-10 contract-language.tex
Reverted the weaker-than definition
All proofs added
Integrating Gustavo\'s edits
Added basic axioms

--- 2014-11-08 contract-language.tex
@@
 \begin{array}{rclcl}
-\multicolumn{5}{l}{
-  {x,y,z} \in \mathtt{EffVar} \qquad
-  {\cureff} \in \mathtt{CurEff} \qquad
+\multicolumn{5}{c}{
+  {x,y,\cureff} \in \mathtt{EffVar} \qquad \qquad
   {\sf Op} \in \mathtt{OperName}
@@
 primitive relations.  Commonly used derived relations are the \emph{same object
-session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), and the \emph{same
-object happens-before order} ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
+session order} ($\small \sooZ = \soZ ~\cap~ \sameobjZ$), \emph{happens-before
+order} ($\small \hbZ = (\soZ ~\cup~ \visZ)^+$) and the \emph{same object
+happens-before order} ($\small \hboZ = (\sooZ ~\cup~ \visZ)^+$).
 
@@
 A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
-\le \cv_{st}$ ) if and only if $\Delta \vdash \forall \cureff. \cv_{st} \Rightarrow \cv_{op}$.
+\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
 \end{definition}
 
-\noindent The $\Delta$ referred in the above defintion is a conjunction of
-assumptions about the nature of primitive relations, e.g., $\Rvis$ and $\Rso$
-are irreflexive, and the happens-before relation $\hbZ$ is acyclic. A
-\emph{well-formed} axiomatic execution ($\E$) is expected to satisfy these
-assumptions (i.e., $\E \models \Delta$).
+The $\Delta$ referred in the above defintion is a conjunction of assumptions
+about the nature of primitive relations. A \emph{well-formed} axiomatic
+execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
+\Delta$).
+
+\begin{definition}
+An axiomatic executions $\small \E = \Exec$ is said to be well-formed if the
+following axioms ($\Delta$) hold:
+
+\begin{itemize}
+\setlength{\itemsep}{2pt}
+\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
+\item Visibility only relates actions on the same object:
+	\begin{itemize}
+	\item $\small \forall a,b.~\vis{a}{b} \Rightarrow \sameobj{a}{b}$.
+	\end{itemize}
+\item Session order is a transitive relation:
+	\begin{itemize}
+	\item $\small \forall a,b,c.~\so{a}{b} ~\wedge~ \so{b}{c} \Rightarrow \so{a}{c}$.
+	\end{itemize}
+\item Same object is an equivalence relation:
+	\begin{itemize}
+	\item $\small \forall a.~\sameobj{a}{a}$.
+	\item $\small \forall a,b.~\sameobj{a}{b} \Rightarrow \sameobj{b}{a}$.
+	\item $\small \forall a,b,c.~\sameobj{a}{b} ~\wedge~ \sameobj{b}{c} \Rightarrow \sameobj{a}{c}$.
+	\end{itemize}
+\end{itemize}
+\end{definition}
 
@@
 
-\noindent The theorem states that if a data store correctly enforces $\scc$,
-$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
-store, extended with the classification scheme shown in
-Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
-proof of the theorem is given in the supplementary material\footnote{The
-supplementary material also provides the concrete reduction rules for enforcing
-the consistency classes.}.
+The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
+$\ecc$ contracts in all well-formed executions, then the same store, extended
+with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
+all well-formed \name contracts. The proof of the theorem is given in the
+supplementary material\footnote{The supplementary material also provides the
+concrete reduction rules for enforcing the consistency classes.}.

### 2014-11-10 evaluation.tex
Added basic axioms

@@
 The distribution of contracts in these applications is given in
-Table~\ref{tab:ctrts}. We see that majority of the operations and
-transactions are classified as eventually consistent and RC,
-respectively. Operation contracts are used to enforce integrity and
-visibility constraints on individual fields in the tables. Transactions are
-mainly used to consistently modify and access related fields across
-tables. In \name, the contract classification process is completely
-performed at compile time and has no overheads at runtime. The proof
-obligations associated with contract classification is discharged through
-the Z3 SMT Solver. Across our benchmarks, classifying a contract took 11.5
-milliseconds on average.
+Table~\ref{tab:ctrts}. We see that majority of the operations and transactions
+are classified as eventually consistent and RC, respectively. Operation
+contracts are used to enforce integrity and visibility constraints on
+individual fields in the tables. Transactions are mainly used to consistently
+modify and access related fields across tables. In \name, the contract
+classification process is completely performed at compile time and has no
+overheads at runtime. The proof obligations associated with contract
+classification is discharged through the Z3 SMT Solver. Across our benchmarks,
+classifying a contract took 11.5 milliseconds on average.
 
@@
 Cassandra replicas within the same datacenter. We instantiate one shim layer
-node co-located on the same VM as a Cassandra replica. Clients are
-instantiated within the same data center as the store, and run
-transactions. We deploy each node in the cluster on \cf{c3.4xlarge} Amazon
-EC2 instances.  Our shim layer nodes are multi-threaded, and we allocate 8
-CPUs (out of 16 available) for each shim layer node. The clients also run on
-\cf{c2.4xlarge} instances. We call this a \cf{1DC} configuration. For our
-geo-distributed experiments (\cf{2DC}), we instantiate 2 clusters, each with
-five nodes, and place the clusters on US-east (Virginia) and US-west
-(Oregon) locations. The average inter-region latency was 85ms.
+node co-located on the same VM as a Cassandra replica. Clients are instantiated
+within the same data center as the store, and run transactions. We deploy each
+cluster node on an \cf{c3.4xlarge} Amazon EC2 instance.  Our shim layer nodes
+are multi-threaded, and we allocate 8 CPUs (out of 16 available) for each shim
+layer node. The clients also run on \cf{c2.4xlarge} instances. We call this a
+\cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
+instantiate 2 clusters, each with five nodes, and place the clusters on US-east
+(Virginia) and US-west (Oregon) locations. The average inter-region latency was
+85ms.
 
 \begin{figure*}
+  \vspace{-5pt}
   \centering
@@
 \cf{withdraw}) being assigned EC and CC consistency levels. These levels
-compromise correctness as \cf{withdraw} has to be an SC operation. The SC
-line corresponds to a configuration where all operations are strongly
-consistent; this ensures application correctness, at the cost of
-performance. \name corresponds to our implementation, which classifies
-operations based on their contract specifications.  Both \name and SC ensure
-correctness. With 512 clients, the \name implementation was within 41\% of
-the latency and 18\% of the throughput of EC, whereas SC operations had
-162\% higher latency and 52\% lower throughput than EC operations. Observe
-that there is a point in each case after which the latency increases while
-the throughput decreases; these points correspond to client request
-saturation points in the data store.
+compromise correctness as \cf{withdraw} has to be an SC operation. The SC line
+corresponds to a configuration where all operations are strongly consistent;
+this ensures application correctness, at the cost of performance. \name
+corresponds to our implementation, which classifies operations based on their
+contract specifications.  Both \name and SC ensure correctness. With 512
+clients, the \name implementation was within 41\% of the latency and 18\% of
+the throughput of EC, whereas SC operations had 162\% higher latency and 52\%
+lower throughput than EC operations. Observe that there is a point in each case
+after which the latency increases while the throughput decreases; these
+correspond to points where the store becomes saturated with client requests.
 
@@
 geo-distributed coordination, whereas \name operations were only 2.2$\times$
-slower, mainly due to the increased cost of \cf{withdraw}
-operations. Importantly, the latency of \cf{getBalance} and \cf{deposit}
-remained almost the same, illustrating the benefit of fine-grained
-contract classification.
+slower, mainly due to the increased cost of \cf{withdraw} operations.
+Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
+the same, illustrating the benefit of fine-grained contract classification.
 
@@
 with 512 clients, supporting our thesis that programmers only have to incur
-relatively low overhead for a more expressive programming model which
-provides stronger provable consistency guarantees.
+relatively low overhead for a more expressive programming model which provides
+stronger provable consistency guarantees.
 
-Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a
-\cf{1DC} configuration against a single replica ({\sf\small NoRep}) and a strongly
+Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
+configuration against a single replica ({\sf\small NoRep}) and a strongly
 replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
@@
 representative of the auction workload.  Without replication, {\sf\small NoRep}
-trivially provides strong consistency. However, this deployment does not
-scale beyond 1750 operations per second. Strong replication offers better
-throughput at the cost of greater latency due to inter-replica
-coordination. The \name deployment offers the benefit of replication, while
-only paying the cost of coordination when necessary.
+trivially provides strong consistency. However, this deployment does not scale
+beyond 1750 operations per second. Strong replication offers better throughput
+at the cost of greater latency due to inter-replica coordination. The \name
+deployment offers the benefit of replication, while only paying the cost of
+coordination when necessary.
 
 Finally, we study the impact of summarization in
-Figure~\ref{grf:summarization}. We use 128 clients and a single \name
-replica, with all clients operating on the \emph{same} LWW register to
-stress test the summarization mechanism. The shim layer cache (memory) is
-summarized every 64 updates, while the updates in the backing store (disk)
-are summarized every 4096 updates. Each point in the graph represents the
-average latency of the previous 1000 operations. Each experiment is run for
-one minute.  Without summarization, the average latency of operations
-increases exponentially to almost one second, and only 13K operations were
-performed in a minute. Since every operation has to reduce over the set of
-all previous operations, operations take increasingly more time to complete
-since they must contend with an ever growing set.  With summarization only
-in memory, performance still degrades due to the cost of fetching all
-previous updates from the backing store into the shim layer. Fetching the
-latest updates is essential for SC operations. With summarization enabled on
-both disk and memory, latency does not increase over time, and the
-implementation realizes throughput of 67K operations/minute.
+Figure~\ref{grf:summarization}. We use 128 clients and a single \name replica,
+with all clients operating on the \emph{same} LWW register to stress test the
+summarization mechanism. The shim layer cache (memory) is summarized every 64
+updates, while the updates in the backing store (disk) are summarized every
+4096 updates. Each point in the graph represents the average latency of the
+previous 1000 operations. Each experiment is run for one minute.  Without
+summarization, the average latency of operations increases exponentially to
+almost one second, and only 13K operations were performed in a minute. Since
+every operation has to reduce over the set of all previous operations,
+operations take increasingly more time to complete since they must contend with
+an ever growing set.  With summarization only in memory, performance still
+degrades due to the cost of fetching all previous updates from the backing
+store into the shim layer. Fetching the latest updates is essential for SC
+operations. With summarization enabled on both disk and memory, latency does
+not increase over time, and the implementation realizes throughput of 67K
+operations/minute.

### 2014-11-10 intro.tex
Added basic axioms

@@
 
-Weak consistency also hinders compositional reasoning about programs.  While
-an application might be naturally expressed in terms of well-understood and
-expressive data types such as maps, trees, queues, or graphs,
-geo-distributed stores typically only provide a minimal set of data types
-with in-built conflict resolution strategies such as last-writer-wins (LWW)
-registers, counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while
-traditional database systems enable composability through transactions,
-geo-distributed stores typically lack unrestricted transactional access to
-the data.  Working in this environment thus requires application state to be
+Weak consistency also hinders compositional reasoning about programs.  While an
+application might be naturally expressed in terms of well-understood and
+expressive data types such as maps, trees, queues, or graphs, geo-distributed
+stores typically only provide a minimal set of data types with in-built
+conflict resolution strategies such as last-writer-wins (LWW) registers,
+counters, and sets~\cite{Cassandra,DynamoDB}.  Furthermore, while traditional
+database systems enable composability through transactions, geo-distributed
+stores typically lack unrestricted serializable transactional access to the
+data. Working in this environment thus requires application state to be
 suitably coerced to function using only the capabilities of the store.

### 2014-11-10 transactions.tex
Integrating Gustavo\'s edits
Added basic axioms

@@
 overheads. As with operation-level consistency, the onus is on the programmer
-to pick the correct transaction kind. This choice is further is complicated by
+to pick the correct transaction kind. This choice is further complicated by
 consistency semantics of individual operations.
@@
 
-\noindent $\cv_{sv}$ and $\cv_{tb}$ are the contracts on the corresponding
+$\cv_{sv}$ and $\cv_{tb}$ are the contracts on the corresponding
 transactions. The function \cf{classify} assigns the contracts
@@
 
-\noindent The key idea in the above definition is that the $\txnZ$ primitive
+\noindent The key idea in the above definition is that the $\small \txnZ$ primitive
 allows us to relate operations on different objects. The \cf{save} transaction
@@
 \noindent Notice that the above definition is the same as the atomicity
-guarantee of transaction described in Section~\ref{sec:syn_sem_ext}. The
-\cf{save} is an example for RC transaction.
+guarantee of transaction described in \S~\ref{sec:syn_sem_ext}. The \cf{save}
+is an example for RC transaction.
 
@@
 
-\vspace{-1em}
+\vspace{-1.2em}
 \begin{smathpar}

### 2014-11-10 contract-language.tex
Fixed Typo
Ran spellcheck.

@@
 an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
-the effect. We admit disjuntion in types to let an effect variable range over
+the effect. We admit disjunction in types to let an effect variable range over
 multiple operation names. The contract $\small \forall (a : \tau_1 \vee
@@
 
-\item \textbf{Eventually consistency}: Eventually consistent operations can
+\item \textbf{Eventual consistency}: Eventually consistent operations can
 	be satisfied as long as the client can reach at least one replica. In the
@@
 
-The $\Delta$ referred in the above defintion is a conjunction of assumptions
+The $\Delta$ referred in the above definition is a conjunction of assumptions
 about the nature of primitive relations. A \emph{well-formed} axiomatic

### 2014-11-10 intro.tex
Ran spellcheck.

@@
 service availability. Traditionally programmers have relied on \emph{strong
-consistency} guarantees such as linerarizability~\cite{Herlihy1990} or
+consistency} guarantees such as linearizability~\cite{Herlihy1990} or
 serializability~\cite{Serializability} in order to build correct applications.
@@
 modern web services, which aim to provide an "always-on" experience,
-overwhelmingly favour availability and partition tolerance over strong
+overwhelmingly favor availability and partition tolerance over strong
 consistency. To this end, several \emph{weak consistency} models such as
@@
 	assign the correct isolation level for a transaction.
-\item We provide metatheory that certifies the soundness of our contract
+\item We provide meta-theory that certifies the soundness of our contract
 	enforcement system, and ensures that an operation is only executed if the
@@
 more details the contract language, and its mapping to the store consistency
-levels, along with metatheory for certifying the correctness of the mapping.
+levels, along with meta-theory for certifying the correctness of the mapping.
 \S~\ref{sec:txns} introduces transaction contracts and classification.

### 2014-11-10 motivation.tex
Added basic axioms
Ran spellcheck.

@@
 \begin{codehaskell}
-type Operation = [e] -> a -> (r, Maybe e)
+type Operation = forall e a r. [e] -> a -> (r, Maybe e)
 \end{codehaskell}
@@
 sent to other replicas. The implementation of the bank account operations in
-\name is given in Figure~\ref{fig:ex}:
+\name is given in Figure~\ref{fig:ex}.
 
@@
 It is easy to see that preventing concurrent \cf{withdraw} operations
-eliminates this anamoly. This can be done by insisting that \cf{withdraw} be
+eliminates this anomaly. This can be done by insisting that \cf{withdraw} be
 executed as a strongly consistent operation. Despite this strengthening,
@@
 and completes successfully. However, the \cf{getBalance} operation may only
-witness this successful withdraw, but not the causally preceding \cf{dopsit},
+witness this successful withdraw, but not the causally preceding \cf{deposit},
 and reports the balance of negative 50 to the user.
@@
 given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice versa. We express this application-level consistency requirement as a
+vice-versa. We express this application-level consistency requirement as a
 contract ($\cv_w$) over \cf{withdraw}:
@@
 procedure \emph{contract classification}. Given the variety and complexity of
-store level consistency properties, the idea is that the system implementor
+store level consistency properties, the idea is that the system implementer
 parameterizes the classification procedure by describing the store semantics in

### 2014-11-10 evaluation.tex

@@
 
-\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The
+\item \textbf{RUBiS}: An eBay-like auction site~\cite{RUBiS}. The
   application allows users to browse items, bid for items on sale, and pay
-  for items from a wallet modelled after a bank account.
+  for items from a wallet modeled after a bank account.
 
-\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
+\item \textbf{Microblog}: A twitter-like microblogging site, modeled after
   Twissandra~\cite{Twissandra}. The application allows adding new users,

### 2014-11-10 contract-language.tex
minor edit
Applied changes from paper to techrep

@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
-$\ecc$ contracts in all well-formed executions, then the same store, extended
-with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
-all well-formed \name contracts. The proof of the theorem is given in the
-supplementary material\footnote{The supplementary material also provides the
-concrete reduction rules for enforcing the consistency classes.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\cureff$ occurs free in $\cv$ and $\cv_\tau$, but not in
+  $\Delta$, we introduce quantifier to bind $\cureff$:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff.\,\cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H3$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H4$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$, $H5$, and $H4$.
+  \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2014-11-10 contract-language.tex
Fixed grammar.

--- 2014-11-10 contract-language.tex
@@
 \begin{figure}
+\vspace{-7pt}
 \begin{smathpar}
@@
 \Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$). Since we
-confine of our contract language to a decidable subset of the logic, validity
+confine our contract language to a decidable subset of the logic, validity
 of such sequents can be decided mechanically allowing us to automate the

### 2014-11-10 intro.tex
Fixed grammar.

@@
 system model.  We describe the challenges in programming with eventually
-consistent data stores, and introduces \name\ contracts as a proposed solution
+consistent data stores, and introduce \name\ contracts as a proposed solution
 to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides

### 2014-11-10 intro.tex
edits

--- 2014-11-10 intro.tex
@@
 appear in a shopping cart after it has been removed~\cite{Dynamo}). Oftentimes,
-the inconsistency leaks from the application and is witnessed by the user.
+these inconsistencies leak from the application and is witnessed by the user.
 Ultimately, the developer must decide the consistency level appropriate for a
@@
 and anomalies, choosing a stronger one than necessary can negatively impact
-program scalability.
+program scalability and performance.
 
-Weak consistency also hinders compositional reasoning about programs.  While an
+Weak consistency also hinders compositional reasoning about programs.  Although an
 application might be naturally expressed in terms of well-understood and
@@
 correctness of the mapping. The paper makes the following contributions:
-
+\vspace*{-.4em}
 \begin{itemize}
@@
 
-The rest of the paper is organized as follows. The next section describes the
+\noindent The rest of the paper is organized as follows. The next section describes the
 system model.  We describe the challenges in programming with eventually
@@
 to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
-more details the contract language, and its mapping to the store consistency
+more details on the contract language, and its mapping to the store consistency
 levels, along with meta-theory for certifying the correctness of the mapping.
-\S~\ref{sec:txns} introduces transaction contracts and classification.
+\S~\ref{sec:txns} introduces transaction contracts and their classification.
 \S~\ref{sec:impl} describes the implementation and provides details about the

### 2014-11-10 motivation.tex
edits

@@
 account operations such that (1) the balance remains non-negative and (2) the
-\cf{getBalance} operation never incorrectly returns a negative balance. Let us
-first consider the anomalies that could arise under eventual consistency.
+\cf{getBalance} operation never incorrectly returns a negative balance.
 
@@
 executed as a strongly consistent operation. Despite this strengthening,
-\cf{getBalance} operation may still incorrectly report a negative balance to
-the end user. Consider the execution shown in
+the \cf{getBalance} operation may still incorrectly report a negative balance to
+the user. Consider the execution shown in
 fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
 sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
-operation, respectively, on the same bank account object. As the \cf{vis} edge
-indicates, operation \cf{withdraw(50)} in session 2, witnesses the effects of
+operation, respectively, on the same bank account object. As the {\sf vis} edge
+indicates, operation \cf{withdraw(50)} in session 2 witnesses the effects of
 \cf{deposit(100)} from session 1, concludes that there is sufficient balance,

### 2014-11-11 motivation.tex
edits

@@
 \begin{codehaskell}
-type Operation = forall e a r. [e] -> a -> (r, Maybe e)
+type Operation e a r = [e] -> a -> (r, Maybe e)
 \end{codehaskell}
@@
 
-getBalance :: [Acc] -> () -> (Int, Maybe Acc)
+getBalance :: Operation Acc () Int
 getBalance ctxt _ =
@@
 
-deposit :: [Acc] -> Int -> ((), Maybe Acc)
+deposit :: Operation Acc Int ()
 deposit _ amt = (amt, Just dollar Deposit amt)
 
-withdraw :: [Acc] -> Int -> (Bool, Maybe Acc)
+withdraw :: Operation Acc Int Bool
 withdraw ctxt v =

### 2014-11-11 contract-language.tex
Typo
Added untyped effvar quantification to contracts

@@
 \cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
-        \ALT \pi \\
+        \ALT \forall x.\cv \ALT \pi \\
 \tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
@@
 ~\ref{fig:contract-lang}. The language is based on first-order logic (FOL), and
-admits prenex universal quantification over typed effect variables. We use a
-special effect variable ($\cureff$) to denote the effect of \emph{current
-operation} - the operation for which a contract is being written. The type of
-an effect is simply the name of the operation (eg: \cf{withdraw}) that induced
-the effect. We admit disjunction in types to let an effect variable range over
-multiple operation names. The contract $\small \forall (a : \tau_1 \vee
-\tau_2).~\psi$ is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1}
-\vee \oper{a}{\tau_2}) \Rightarrow \psi$.
+admits prenex universal quantification over typed and untyped effect variables.
+We use a special effect variable ($\cureff$) to denote the effect of
+\emph{current operation} - the operation for which a contract is being written.
+The type of an effect is simply the name of the operation (eg: \cf{withdraw})
+that induced the effect. We admit disjunction in types to let an effect variable
+range over multiple operation names. The contract $\small \forall (a : \tau_1
+\vee \tau_2).~\psi$ is just syntactic sugar for $\small \forall a.
+(\oper{a}{\tau_1} \vee \oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect
+variable ranges over all operation names.
 
@@
 classification-based contract enforcement. To help us state the result, we
-define an operational semantics of the our system described informally in
+define an operational semantics of the system described informally in
 \S~\ref{sec:sysmod}:

### 2014-11-11 contract-language.tex
Quantifying cureff in weaker-than relation

@@
 A contract $\cv_{op}$ is said to be weaker than $\cv_{st}$ (written $\cv_{op}
-\le \cv_{st}$ ) if and only if $\Delta \vdash \cv_{st} \Rightarrow \cv_{op}$.
+\le \cv_{st}$ ) if and only if $\Delta \vdash \forall \cureff.
+\cv_{st} \Rightarrow \cv_{op}$.  
 \end{definition}
 
-The $\Delta$ referred in the above definition is a conjunction of assumptions
+The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$ and
+$\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of assumptions
 about the nature of primitive relations. A \emph{well-formed} axiomatic

### 2014-11-11 contract-language.tex
Merged changes from contract-language of paper

@@
 \begin{figure}
-\vspace{-7pt}
 \begin{smathpar}
@@
 \cv 		& \in & \mathtt{Contract} 	& \coloneqq & \forall (x : \tau).\cv
-        \ALT \forall x.\cv \ALT \pi \\
+        \ALT \pi \\
 \tau		& \in	& \mathtt{EffType}	& \coloneqq &  {\sf Op}
@@
 \Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$). Since we
-confine our contract language to a decidable subset of the logic, validity
+confine of our contract language to a decidable subset of the logic, validity
 of such sequents can be decided mechanically allowing us to automate the
@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
-$\ecc$ contracts in all well-formed executions, then the same store, extended
-with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
-all well-formed \name contracts. The proof of the theorem is given in the
-supplementary material\footnote{The supplementary material also provides the
-concrete reduction rules for enforcing the consistency classes.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H3$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H4$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$, $H5$, and $H4$.
+  \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2014-11-11 contract-language.tex
edits

--- 2014-11-11 contract-language.tex
@@
 \emph{current operation} - the operation for which a contract is being written.
-The type of an effect is simply the name of the operation (eg: \cf{withdraw})
-that induced the effect. We admit disjunction in types to let an effect variable
-range over multiple operation names. The contract $\small \forall (a : \tau_1
-\vee \tau_2).~\psi$ is just syntactic sugar for $\small \forall a.
-(\oper{a}{\tau_1} \vee \oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect
-variable ranges over all operation names.
+Notice that $\cureff$ occurs free in the contract. We will fix its scope when
+classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
+simply the name of the operation (eg: \cf{withdraw}) that induced the effect.
+We admit disjunction in types to let an effect variable range over multiple
+operation names. The contract $\small \forall (a : \tau_1 \vee \tau_2).~\psi$
+is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1} \vee
+\oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect variable ranges over all
+operation names.
 
@@
 
-\subsection{Contract Comparison and Classification}
+\subsection{Contract Classification}
+\label{sec:classify}
 
@@
 
-The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$ and
-$\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of assumptions
-about the nature of primitive relations. A \emph{well-formed} axiomatic
-execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
-\Delta$).
+The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
+and $\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of
+assumptions about the nature of primitive relations. A \emph{well-formed}
+axiomatic execution ($\E$) is expected to satisfy these assumptions (i.e., $\E
+\models \Delta$).
 

### 2014-11-11 model.tex
edits

@@
 
-Figure~\ref{fig:sysmod} provides a schematic diagram of our system model. The
-distributed store is composed of a collection of \emph{replicas}, each of which
-stores a set of \emph{objects} ($x,y,\ldots$). We assume that every object is
-replicated at every replica in the store. The state of an object at any replica
-is the set of all updates (\emph{effects}) performed on the object. For
-example, the state of $x$ at replica 1 is the set composed of effects $w^x_1$
-and $w^x_2$.
+In this section, we describe the system model and introduce the primitive
+relations that our contract language is seeded with. Figure~\ref{fig:sysmod}
+provides a schematic diagram of our system model. The distributed store is
+composed of a collection of \emph{replicas}, each of which stores a set of
+\emph{objects} ($x,y,\ldots$). We assume that every object is replicated at
+every replica in the store. The state of an object at any replica is the set of
+all updates (\emph{effects}) performed on the object. For example, the state of
+$x$ at replica 1 is the ordered set composed of effects $w^x_1$ and $w^x_2$.
 
@@
 this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
-written logically as $\vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
-$\visZ$ is the visibility relation between effects. Visibility is an
+written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
+$\small \visZ$ is the visibility relation between effects. Visibility is an
 irreflexive and asymmetric relation, and only relates effects produced by
@@
 and $w^x_5$ that arise from the same session are said to be in \emph{session
-order} (written logically as $\so{w^x_4}{w^x_5}$). Session order is an
+order} (written logically as $\small \so{w^x_4}{w^x_5}$). Session order is an
 irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
 operations applied to the same object $x$ are said to be under the \emph{same
-object} relation, written $\sameobj{w^x_4}{w^x_5}$. Finally, we can associate
-every effect with the operation that generated the effect with the help of a
-relation $\operZ$. In the current example, $\oper{w^x_4}{foo}$ and
-$\oper{w^x_5}{bar}$ hold. For simplicity, we assume all operation names across
-all object types are distinct.
+object} relation, written $\small \sameobj{w^x_4}{w^x_5}$. Finally, we can
+associate every effect with the operation that generated the effect with the
+help of a relation $\small \operZ$. In the current example, $\small
+\oper{w^x_4}{foo}$ and $\small \oper{w^x_5}{bar}$ hold. For simplicity, we
+assume all operation names across all object are distinct.
 
@@
 store. Subsequently, we will illustrate how our contract and specification
-language, armed with the primitive relations $\visZ$, $\soZ$, $\sameobjZ$ and
-$\operZ$, mitigates these challenges.
+language, armed with the primitive relations $\small \visZ$, $\small \soZ$,
+$\small \sameobjZ$ and $\small \operZ$, mitigates these challenges.

### 2014-11-11 motivation.tex
edits

@@
 which obviates the need for coercing the application logic to utilize the store
-provided data types. In addition, \name treats the convergence semantics of the
-data type separately from its consistency properties. This separation of
-concerns permits \emph{operational} reasoning for conflict resolution, and
-\emph{declarative} reasoning for consistency. The combination of these
-techniques enhances the programmability of the store.
+provided data types. In addition, \name treats the convergence semantics (i.e.,
+\emph{how} conflicting updates are resolved) of the data type separately from
+its consistency properties (i.e., \emph{when} updates become visible). This
+separation of concerns permits \emph{operational} reasoning for conflict
+resolution, and \emph{declarative} reasoning for consistency. The combination
+of these techniques enhances the programmability of the store.
 
@@
 
+\subsubsection{Summarization}
+\label{sec:summarize}
+
+Observe that the definition of \cf{getBalance} reduces over the \emph{entire
+history} of updates to an account. If we are to realize an efficient
+implementation of this bank account RDT, we need a \emph{summary} of the
+account history. Intuitively, the current account balance summarizes the state
+of an account. A bank account with the history \cf{[Deposit 10, Withdraw 5]} is
+\emph{observably equivalent} to a bank account with a single deposit operation
+\cf{[Deposit 5]}; we can replace the earlier history with the latter and a
+client of the store would not able to tell the difference between the two.
+
+This notion of observable equivalence can be generalized to other RDTs as well.
+For example, a last-writer-wins register with multiple updates is equivalent to
+a register with only the last write. Similarly, a set with a collection of add
+and remove operations is equivalent to a set with a series of additions of live
+elements from the original set. Since the notion of observable equivalence is
+specific to each RDT, we expect the programmer to provide a summarization
+function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT. We utilize this
+function to ensure that the state of an object remains summarized in each
+replica. The summarization function for the bank account is {\footnotesize
+\cf{summarize ctxt = [Deposit \$ sel1 \$ getBalance ctxt ()]}}.
+
 \subsection{Anomalies under Eventual Consistency}
@@
 
-\begin{figure}
+\begin{figure}[ht]
 \centering
@@
 by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
-witnesses the deposit and succeeds\footnote{Although visibility and session
+witnesses\footnote{Although visibility and session
 order relations relate effects, we have abused the notation in these examples
 to relate operations, with the idea that the relations relate the effect
-emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
+emitted by those operations} the deposit and succeeds. Subsequently, session 2 perform a \cf{withdraw}
 operation, but importantly, due to eventual consistency, only witnesses the
@@
 all \cf{deposit} operations $a$ visible to $b$ should also be visible to
-$\cureff$. This prevents negative balance anomalies. Our contract language
-provides operators to compose relations. The syntax $(R_1 \cap R_2)(a,b)$ is
-equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the above contract
-says that if a \cf{deposit} or a \cf{withdraw} operation precedes a
-\cf{getBalance} operation in session order, and is applied on the same object
-as the \cf{getBalance} operation, then it must be the case that the
+$\cureff$. Intuitively, this prevents negative balance anomalies. Our contract
+language provides operators to compose relations. The syntax $(R_1 \cap
+R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
+above contract says that if a \cf{deposit} or a \cf{withdraw} operation
+precedes a \cf{getBalance} operation in session order, and is applied on the
+same object as the \cf{getBalance} operation, then it must be the case that the
 \cf{getBalance} operation witnesses the effects of the preceding operations.

### 2014-11-11 model.tex
edits

@@
 this case, the effects $w^x_1$ and $w^x_2$ are \emph{visible} to $w^x_4$,
-written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$, where
-$\small \visZ$ is the visibility relation between effects. Visibility is an
-irreflexive and asymmetric relation, and only relates effects produced by
+written logically as $\small \vis{w^x_1}{w^x_4} \wedge \vis{w^x_2}{w^x_4}$,
+where $\small \visZ$ is the visibility relation between effects. Visibility is
+an irreflexive and asymmetric relation, and only relates effects produced by
 operations on the same object. Executing a read-only operation is similar
-except that no new effects are produced.
+except that no new effects are produced. The effect added to a particular
+replica is asynchronously sent to other replicas, and eventually merged into
+all other replicas. Observe that this model does not assume a particular
+resolution strategy for concurrent conflicting updates, and instead preserves
+\emph{every} update. Conflicts are resolved when an operation reduces over the
+set of effects on an object at a particular replica.
 
-The effect added to a particular replica is asynchronously sent to other
-replicas, and eventually merged into all other replicas. Two effects $w^x_4$
-and $w^x_5$ that arise from the same session are said to be in \emph{session
-order} (written logically as $\small \so{w^x_4}{w^x_5}$). Session order is an
-irreflexive, transitive relation. The effects $w^x_4$ and $w^x_5$ arising from
-operations applied to the same object $x$ are said to be under the \emph{same
-object} relation, written $\small \sameobj{w^x_4}{w^x_5}$. Finally, we can
-associate every effect with the operation that generated the effect with the
-help of a relation $\small \operZ$. In the current example, $\small
-\oper{w^x_4}{foo}$ and $\small \oper{w^x_5}{bar}$ hold. For simplicity, we
-assume all operation names across all object are distinct.
+Two effects $w^x_4$ and $w^x_5$ that arise from the same session are said to be
+in \emph{session order} (written logically as $\small \so{w^x_4}{w^x_5}$).
+Session order is an irreflexive, transitive relation. The effects $w^x_4$ and
+$w^x_5$ arising from operations applied to the same object $x$ are said to be
+under the \emph{same object} relation, written $\small \sameobj{w^x_4}{w^x_5}$.
+Finally, we can associate every effect with the operation that generated the
+effect with the help of a relation $\small \operZ$. In the current example,
+$\small \oper{w^x_4}{foo}$ and $\small \oper{w^x_5}{bar}$ hold. For simplicity,
+we assume all operation names across all object are distinct.
 

### 2014-11-12 evaluation.tex
Incorporated new edits

@@
 \item \textbf{LWW register}: A last-write-wins register that provides read
-  and write operations. Each write is associated with a timestamp, which is
-  used to resolve conflicting concurrent writes -- a newer write always
-  wins.
+	and write operations, where the read returns the value of the latest write.
 
@@
 \item \textbf{Shopping list}: A collaborative shopping list that allows
-  adding and deleting items.
+  concurrent addition and deletion of items.
 
@@
 
-We compare the performance of different transaction isolation level choices
-in Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were
-obtained under a 1DC configuration. The YCSB workload was modified to issue
-10 operations per transaction, with a default 50:50 read:write mix. Each
-operation is assumed to be eventual consistent. \cf{NoTxn} corresponds to a
-configuration that does not use transactions. Compared to this, RC is only
-12\% shower in terms of latency with 512 clients, whereas RR is 2.3X
-slower. The difference between RC and \cf{NoTxn} is due to the meta-data
-overhead of recording transaction information in the object state. For RR
-transactions, the cost of capturing and maintaining a snapshot is the
-biggest source of overhead.
+We compare the performance of different transaction isolation level choices in
+Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
+under a 1DC configuration. The YCSB workload was modified to issue 10
+operations per transaction, with a default 50:50 read:write mix. Each operation
+is assumed to be eventually consistent. \cf{NoTxn} corresponds to a
+configuration that does not use transactions. Compared to this, RC is only 12\%
+shower in terms of latency with 512 clients, whereas RR is 2.3X slower. The
+difference between RC and \cf{NoTxn} is due to the meta-data overhead of
+recording transaction information in the object state. For RR transactions, the
+cost of capturing and maintaining a snapshot is the biggest source of overhead.
 

### 2014-11-12 motivation.tex
Incorporated new edits

@@
 
-\noindent It takes a list of effects (the \emph{context} for the operation),
-and an input argument, and returns a result along with an optional effect
-(read-only operations return \cf{Nothing}). The new effect (if emitted) is
-added to the state of the object at the current replica, and asynchronously
+\noindent It takes a list of effects (the \emph{history} of updates to that
+object), and an input argument, and returns a result along with an optional
+effect (read-only operations return \cf{Nothing}). The new effect (if emitted)
+is added to the state of the object at the current replica, and asynchronously
 sent to other replicas. The implementation of the bank account operations in
@@
 \begin{figure}
+\vspace{-2.3mm}
 \begin{codehaskell}
-data Acc = Deposit Int | Withdraw Int
-				 | GetBalance
+data Acc = Deposit Int | Withdraw Int | GetBal
 
-getBalance :: Operation Acc () Int
-getBalance ctxt _ =
-  let res = sum [x | Deposit x <- ctxt]
-						- sum [x | Withdraw x <- ctxt]
+getBalance :: [Acc] -> () -> (Int, Maybe Acc)
+getBalance hist _ =
+  let res = sum [x | Deposit x <- hist]
+						- sum [x | Withdraw x <- hist]
 	in (res, Nothing)
 
-deposit :: Operation Acc Int ()
-deposit _ amt = (amt, Just dollar Deposit amt)
+deposit :: [Acc] -> Int -> ((), Maybe Acc)
+deposit hist amt = ((), Just dollar Deposit amt)
 
-withdraw :: Operation Acc Int Bool
-withdraw ctxt v =
-	if sel1 dollar getBalance ctxt () >= v
+withdraw :: [Acc] -> Int -> (Bool, Maybe Acc)
+withdraw hist v =
+	if sel1 dollar getBalance hist () >= v
   then (True, Just dollar Withdraw v)
@@
 The datatype \cf{Acc} represents the effect type for the bank account. The
-context of the operations is a snapshot of the state of the object at some
-replica. In this sense, every operation on the RDT is atomic, and thus
-permitting sequential reasoning for implementing eventually consistent data
-types. We have implemented a large corpus of RDTs for realistic benchmarks
-including shopping carts, auction and micro-blogging sites in few tens of lines
-of code.
+function \cf{sum} returns the sum of elements in the list, and \cf{sel1}
+returns the first element of a tuple. For each operation, \cf{hist} is a
+\emph{snapshot} of the state of the object at some replica. In this sense,
+every operation on the RDT is atomic, and thus permitting sequential reasoning
+for implementing eventually consistent data types. Here, \cf{getBalance} is a
+read-only operation, \cf{deposit} always emits an effect, and \cf{withdraw}
+only emits an effect if there is sufficient balance in the account. We have
+implemented a large corpus of RDTs for realistic benchmarks including shopping
+carts, auction and micro-blogging sites in few tens of lines of code.
 
@@
 specific to each RDT, we expect the programmer to provide a summarization
-function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT. We utilize this
-function to ensure that the state of an object remains summarized in each
-replica. The summarization function for the bank account is {\footnotesize
-\cf{summarize ctxt = [Deposit \$ sel1 \$ getBalance ctxt ()]}}.
+function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT, as a part of
+the RDT specification.
+
+The summarization function for the bank account is \cf{summarize hist =
+[Deposit \$ sel1 \$ getBalance hist ()]}. Given a bank account history
+\cf{hist}, the \cf{summarize} function returns a new history with a single
+deposit of the current account balance. We utilize the summarization function
+to ensure that the state of an object remains summarized in each replica.
 
@@
 \begin{figure}[ht]
+\vspace{-2mm}
 \centering

### 2014-11-12 contract-language.tex
Incorporated new edits
Back to 11 pages

@@
 
-\begin{itemize}
-\setlength{\itemsep}{2pt}
+\begin{itemize}[noitemsep]
 
@@
 	bank account example, \cf{deposit} is an eventually consistent operation.
-	While eventually consistent data stores typically offer \emph{basic} eventual
-	consistency with all possible anomalies, we assume that our store provides
-	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
-	store always exposes a \emph{causal cut} of the updates. This semantics can
-	be formally captured in terms of the following contract definition:
+	While an \ecds typically offer \emph{basic} eventual consistency with all
+	possible anomalies, we assume that our store provides stronger semantics that
+	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
+	\emph{causal cut} of the updates. This semantics can be formally captured in
+	terms of the following contract definition:
 
@@
 \begin{figure}
+\vspace{-10pt}
 \begin{smathpar}
-\begin{array}{c}
-\vspace{3mm}
+\begin{array}{cc}
 \RuleTwo
 {\DDe{\cv} \le \DDe{\scc}}
-{{\sf WellFormed}(\cv)}  \qquad
+{{\sf WellFormed}(\cv)}  &
 
@@
 {\DDe{\cv} \le \DDe{\ecc}}
-{{\sf EventuallyConsistent}(\cv)} \\
+{{\sf EventuallyConsistent}(\cv)} \\ \\
 
-\vspace{3mm}
 \RuleTwo
@@
 \quad \DDe{\cv} \le \DDe{\ccc}}
-{{\sf CausallyConsistent}(\cv)} \qquad
+{{\sf CausallyConsistent}(\cv)} &
 
@@
 \end{smathpar}
-\vspace{-5mm}
+\vspace{-3mm}
 \caption{Contract classification.}
@@
 
-\begin{itemize}
-\setlength{\itemsep}{2pt}
+\begin{itemize}[noitemsep]
 \item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.

### 2014-11-12 evaluation.tex
Back to 11 pages

@@
 
-\begin{itemize}
-\setlength{\itemsep}{2pt}
+\begin{itemize}[noitemsep]
 \item \textbf{LWW register}: A last-write-wins register that provides read
@@
 
-\item \textbf{Bank account}: Our running example, with savings and checking
-  accounts.
+\item \textbf{Bank account}: Our running example.
 
@@
 within the same data center as the store, and run transactions. We deploy each
-cluster node on an \cf{c3.4xlarge} Amazon EC2 instance.  Our shim layer nodes
-are multi-threaded, and we allocate 8 CPUs (out of 16 available) for each shim
-layer node. The clients also run on \cf{c2.4xlarge} instances. We call this a
-\cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
+cluster and client node on an \cf{c3.4xlarge} Amazon EC2 instance. We call this
+a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
 instantiate 2 clusters, each with five nodes, and place the clusters on US-east
@@
 corresponds to our implementation, which classifies operations based on their
-contract specifications.  Both \name and SC ensure correctness. With 512
-clients, the \name implementation was within 41\% of the latency and 18\% of
-the throughput of EC, whereas SC operations had 162\% higher latency and 52\%
-lower throughput than EC operations. Observe that there is a point in each case
-after which the latency increases while the throughput decreases; these
-correspond to points where the store becomes saturated with client requests.
-
-In a \cf{2DC} configuration (not shown here), the average latency of SC
-operations with 512 clients increased by 9.4$\times$ due to the cost of
-geo-distributed coordination, whereas \name operations were only 2.2$\times$
-slower, mainly due to the increased cost of \cf{withdraw} operations.
-Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
-the same, illustrating the benefit of fine-grained contract classification.
+contract specifications. With 512 clients, the \name implementation was within
+41\% of the latency and 18\% of the throughput of EC, whereas SC operations had
+162\% higher latency and 52\% lower throughput than EC operations. Observe that
+there is a point in each case after which the latency increases while the
+throughput decreases; these correspond to points where the store becomes
+saturated with client requests. In a \cf{2DC} configuration (not shown here),
+the average latency of SC operations with 512 clients increased by 9.4$\times$
+due to the cost of geo-distributed coordination, whereas \name operations were
+only 2.2$\times$ slower, mainly due to the increased cost of \cf{withdraw}
+operations. Importantly, the latency of \cf{getBalance} and \cf{deposit}
+remained almost the same, illustrating the benefit of fine-grained contract
+classification.
 

### 2014-11-12 model.tex
Incorporated new edits
Back to 11 pages

@@
 \centering
-\includegraphics[width=0.75\columnwidth]{Figures/SystemModel}
+\includegraphics[width=0.7\columnwidth]{Figures/SystemModel}
 \caption{\name system model.}
@@
 resolution strategy for concurrent conflicting updates, and instead preserves
-\emph{every} update. Conflicts are resolved when an operation reduces over the
-set of effects on an object at a particular replica.
+\emph{every} update. Update conflicts are resolved when an operation reduces
+over the set of effects on an object at a particular replica.
 
@@
 section, we will concretely describe the challenges associated with
-constructing a consistent bank account on top of an eventually consistent data
-store. Subsequently, we will illustrate how our contract and specification
-language, armed with the primitive relations $\small \visZ$, $\small \soZ$,
-$\small \sameobjZ$ and $\small \operZ$, mitigates these challenges.
+constructing a consistent bank account on top of an \ecds. Subsequently, we
+will illustrate how our contract and specification language, armed with the
+primitive relations $\small \visZ$, $\small \soZ$, $\small \sameobjZ$ and
+$\small \operZ$, mitigates these challenges.

### 2014-11-12 motivation.tex
Back to 11 pages

@@
 Consider how we might implement a highly available bank account on top of an
-eventually consistent data store, with the \emph{integrity} constraint that the
-balance must be non-negative. We begin by implementing a bank account
-replicated data type (RDT) in \name, and then describe the mechanisms to obtain
-the desired correctness guarantees.
+\ecds, with the \emph{integrity} constraint that the balance must be
+non-negative. We begin by implementing a bank account replicated data type
+(RDT) in \name, and then describe the mechanisms to obtain the desired
+correctness guarantees.
 
@@
 every operation on the RDT is atomic, and thus permitting sequential reasoning
-for implementing eventually consistent data types. Here, \cf{getBalance} is a
-read-only operation, \cf{deposit} always emits an effect, and \cf{withdraw}
-only emits an effect if there is sufficient balance in the account. We have
-implemented a large corpus of RDTs for realistic benchmarks including shopping
-carts, auction and micro-blogging sites in few tens of lines of code.
+for implementing the RDTs. Here, \cf{getBalance} is a read-only operation,
+\cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
+there is sufficient balance in the account. We have implemented a large corpus
+of RDTs for realistic benchmarks including shopping carts, auction and
+micro-blogging sites in few tens of lines of code.
 

### 2014-11-12 transactions.tex
Applied changes from paper to techrep
Back to 11 pages

@@
 \begin{figure}
+\vspace{-2.3mm}
 \centering
@@
 A transaction with ANSI RC semantics only witnesses committed operations. Let
-us assume that the store buffers updates until all the updates from the
-transaction are available at a replica. If the transaction commits, the
-buffered updates are made visible. Otherwise, the buffered updates are
-discarded. RC does not entail any further isolation guarantees. Hence, a store
+us assume that a replica will buffer transactional updates until all the
+updates from the same transaction are available at that replica. Once all the
+updates from a transaction are available, the buffered updates are made visible
+to subsequent client requests. This ensure atomicity of transactions.
+Importantly, RC does not entail any other guarantees. As a result, a store
 implementing RC does not require inter-replica coordination. We can express RC

### 2014-11-12 contract-language.tex
edits

@@
 \begin{definition}
-An axiomatic execution $\E$ is said to satisfy a contract $\cv$ if and
+An axiomatic execution $\E$  satisfies a contract $\cv$ if and
 only if $E \models \cv$.
@@
 \begin{definition}
-An axiomatic executions $\small \E = \Exec$ is said to be well-formed if the
+An axiomatic execution $\small \E = \Exec$ is well-formed if the
 following axioms ($\Delta$) hold:

### 2014-11-12 evaluation.tex
Added detailed results

--- 2014-11-10 evaluation.tex
@@
 
-We present an evaluation study of our implementation, report contract
-profiles of benchmark programs, and illustrate the performance benefits of
-fine-grained consistency classification on operations and transactions. We
-also evaluate the impact of the summarization. We have implemented the
-following applications, which includes individual RDTs as well as larger
-applications composed of several RDTs:
+In this section, we evaluate \name programs, report their contract profile and
+illustrate the performance benefits of fine-grained consistency classification
+on operations and transactions. We also evaluate on the impact of the
+summarization. We implemented the following applications, which includes
+individual RDTs as well as larger applications composed of several RDTs:
 
@@
 \setlength{\itemsep}{2pt}
-\item \textbf{LWW register}: A last-write-wins register that provides read
-  and write operations. Each write is associated with a timestamp, which is
-  used to resolve conflicting concurrent writes -- a newer write always
-  wins.
+\item \textbf{LWW register}: A last-write-wins register that provides read and
+write operations. Each write is associated with a timestamp, which is used to
+resolve conflicting concurrent writes -- newer write wins.
 
-\item \textbf{DynamoDB register}: An integer register that allows eventual
-  and strong puts and gets, conditional puts, increment and decrement
-  operations.
+\item \textbf{DynamoDB register}: A integer register that allows eventual and
+strong puts and gets, conditional puts, increment and decrement operations.
 
-\item \textbf{Bank account}: Our running example, with savings and checking
+\item \textbf{Bank account}: Our running example, with savings and current
 accounts.
 
-\item \textbf{Shopping list}: A collaborative shopping list that allows
-  adding and deleting items.
+\item \textbf{Shopping list}: Collaborative shopping list which allows adding
+and deleting items.
 
-\item \textbf{Online store}: An online store with shopping cart
-  functionality and dynamically changing item prices.  The checkout process
-  verifies that the customer only pays the accepted price.
+\item \textbf{Online store}: Models an online store with shopping cart and
+dynamically changing item prices. Checkout process verifies that the customer
+only pays the accepted price.
 
-\item \textbf{RUBiS}: An eBay-like auction site~\cite{RUBiS}. The
-  application allows users to browse items, bid for items on sale, and pay
-  for items from a wallet modeled after a bank account.
+\item \textbf{RUBiS}: An ebay-like auction site~\cite{RUBiS}. The application
+allows users to browse items, bid for items on sale, and pay for items from a
+wallet modelled after a bank account.
 
-\item \textbf{Microblog}: A twitter-like microblogging site, modeled after
-  Twissandra~\cite{Twissandra}. The application allows adding new users,
-  adding and replying to tweets, following, unfollowing and blocking users,
-  and fetching a user's timeline, userline, followers and following.
+\item \textbf{Microblog}: A twitter-like microblogging site, modelled after
+Twissandra~\cite{Twissandra}. The application allows adding a new user, adding
+and replying to tweets, following, unfollowing and blocking users, and fetching
+a user's timeline, userline, followers and following.
 \end{itemize}
@@
 
-For our performance evaluation, we deploy \name applications in
-\emph{clusters}, where each cluster is composed of five fully replicated
-Cassandra replicas within the same datacenter. We instantiate one shim layer
-node co-located on the same VM as a Cassandra replica. Clients are instantiated
-within the same data center as the store, and run transactions. We deploy each
-cluster node on an \cf{c3.4xlarge} Amazon EC2 instance.  Our shim layer nodes
-are multi-threaded, and we allocate 8 CPUs (out of 16 available) for each shim
-layer node. The clients also run on \cf{c2.4xlarge} instances. We call this a
-\cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we
-instantiate 2 clusters, each with five nodes, and place the clusters on US-east
-(Virginia) and US-west (Oregon) locations. The average inter-region latency was
-85ms.
+For our performance evaluation, we deploy \name applications in \emph{clusters},
+where each cluster is composed of 5 fully replicated Cassandra replicas within
+the same datacenter. We instantiate one shim layer node for every Cassandra
+replica, and place it on the same VM as the Cassandra replica. Clients are
+instantiated on the same data center as the store, and run transactions. We
+deploy the each node in the cluster on \cf{c3.4xlarge} Amazon EC2 instances.
+Our shim layer nodes are multi-threaded, and we allocate 8 CPUs (out of 16
+available) for each shim layer node. The clients also run on \cf{c2.4xlarge}
+instances. We call this \cf{1DC} configuration. For our geo-distributed
+experiments (\cf{2DC}), we instantiate 2 clusters, each with 5 nodes, and place
+the clusters on US-east (Virginia) and US-west (Oregon). The average
+inter-region latency was 85ms.
 
-\begin{figure*}
-  \vspace{-5pt}
+\begin{figure*}[t]
   \centering
-  \subfigure[Bank account]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-tp-vs-lat.pdf}}
-  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
-  \subfigure[RUBiS bidding mix]{\label{grf:rubis}\includegraphics[width=0.24\textwidth]{graphs/Rubis.pdf}}
-  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
-	\caption{Quelea Performance.}
-  \label{grf:LWW_perf}
+  \subfigure[Latency]{\label{grf:BA-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:BA-tp}\includegraphics[width=0.32\textwidth]{graphs/BA-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-tp-vs-lat.pdf}}
+	\caption{Bank account performance.}
+  \label{grf:BA}
 \end{figure*}
 
-Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations
-in the bank account example as we increase the number of clients in a \cf{1DC}
-configuration. Our client workload was generated using the YCSB
-benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
-the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
-corresponds to the default 50:50 read:write mix in YCSB. We increased the
-number of clients from 128 to 1024, and each experiment ran for 180 seconds.
+Figure~\ref{grf:BA} shows the performance of operations in bank account example
+as we increase the number of clients in \cf{1DC} configuration.. Our client
+workload was generated using YCSB benchmark~\cite{YCSB}. The benchmark
+uniformly chose from 100,000 keys, where the operation spread was 25\%
+withdraw, 25\% deposit and 50\% getBalance, which corresponds to the default
+50:50 read:write mix in YCSB. We increased the number of clients from 128 to
+1024, and each experiment ran for 180 seconds.
 
-The lines marked EC and CC correspond to all operations (including
-\cf{withdraw}) being assigned EC and CC consistency levels. These levels
-compromise correctness as \cf{withdraw} has to be an SC operation. The SC line
-corresponds to a configuration where all operations are strongly consistent;
-this ensures application correctness, at the cost of performance. \name
+The lines marked EC and CC correspond to all operations being assigned EC and
+CC consistency levels. These levels compromise correctness as withdraw has to
+be and SC operation. The line SC corresponds to a configuration where all
+operations are strongly consistent; this ensures application correctness. \name
 corresponds to our implementation, which classifies operations based on their
-contract specifications.  Both \name and SC ensure correctness. With 512
-clients, the \name implementation was within 41\% of the latency and 18\% of
-the throughput of EC, whereas SC operations had 162\% higher latency and 52\%
-lower throughput than EC operations. Observe that there is a point in each case
-after which the latency increases while the throughput decreases; these
-correspond to points where the store becomes saturated with client requests.
+contracts. Both \name and SC ensure correctness. However, with 512 clients,
+\name implementation was within 41\% of latency and 18\% of throughput of EC,
+whereas SC operations had 162\% higher latency and 52\% lower throughput than
+EC operations. Observe that there is a point in each case after which the
+latency increases while the throughput decreases. This indicates a point after
+which the store becomes saturated with client requests.
 
-In a \cf{2DC} configuration (not shown here), the average latency of SC
-operations with 512 clients increased by 9.4$\times$ due to the cost of
-geo-distributed coordination, whereas \name operations were only 2.2$\times$
-slower, mainly due to the increased cost of \cf{withdraw} operations.
-Importantly, the latency of \cf{getBalance} and \cf{deposit} remained almost
-the same, illustrating the benefit of fine-grained contract classification.
+In \cf{2DC} configuration (not-shown), the average latency of SC operations
+with 512 clients increased by 9.4$\times$ due to the cost of geo-distributed
+coordination, whereas \name operations were only 2.2$\times$ slower, mainly due
+to the increased cost of \cf{withdraw} operations. Importantly, the latency of
+\cf{getBalance} and \cf{deposit} remained almost the same. This illustrates the
+benefit of fine-grained contract classification in \name.
 
-We compare the performance of different transaction isolation level choices
-in Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were
-obtained under a 1DC configuration. The YCSB workload was modified to issue
-10 operations per transaction, with a default 50:50 read:write mix. Each
-operation is assumed to be eventual consistent. \cf{NoTxn} corresponds to a
-configuration that does not use transactions. Compared to this, RC is only
-12\% shower in terms of latency with 512 clients, whereas RR is 2.3X
-slower. The difference between RC and \cf{NoTxn} is due to the meta-data
-overhead of recording transaction information in the object state. For RR
-transactions, the cost of capturing and maintaining a snapshot is the
-biggest source of overhead.
+\begin{figure*}[t]
+  \centering
+  \subfigure[Latency]{\label{grf:LWW-txn-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:LWW-txn-tp}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:LWW-txn-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp-vs-lat.pdf}}
+	\caption{LWW register transaction performance.}
+  \label{grf:LWW-txn}
+\end{figure*}
+
+We compare the performance of different transaction isolation level choices in
+Figure~\ref{grf:LWW-txn} using the LWW register. The numbers were obtained
+under 1DC configuration. The YCSB workload was modified to issue 10 operations
+per transaction, with the default 50:50 read:write mix. Each operation is
+assumed to have eventual consistency. \cf{NoTxn} corresponds to a configuration
+that does not use transactions. Compared to this RC is only 12\% shower in
+terms of latency with 512 clients, where as RR is 2.3X slower. The difference
+between RC and \cf{NoTxn} is due to the meta-data overhead of recording
+transaction information in the object state. For RR transaction, the cost of
+capturing and maintaining the snapshot in an RR transaction is the biggest
+source of overhead.
 
 We also compared (not shown) the performance of EC LWW operations directly
-against Cassandra, which uses last-writer-wins as the only convergence
-semantics. While Cassandra provides no stronger-than-eventual consistency
-properties, \name was within 30\%(20\%) of latency(throughput) of Cassandra
-with 512 clients, supporting our thesis that programmers only have to incur
-relatively low overhead for a more expressive programming model which provides
-stronger provable consistency guarantees.
+against Cassandra (our backing store), which uses last-writer-wins as the only
+convergence semantics. While Cassandra provides no stronger-than-eventual
+consistency properties, \name was within 30\%(20\%) of latency(throughput) of
+Cassandra with 512 clients, illustrating that the programmers only have to pay
+a minimal overhead for the expressive and stronger \name programming model.
 
-Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
-configuration against a single replica ({\sf\small NoRep}) and a strongly
-replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
-default RUBiS bidding mix, which has 15\% read-write interactions, which is
-representative of the auction workload.  Without replication, {\sf\small NoRep}
-trivially provides strong consistency. However, this deployment does not scale
-beyond 1750 operations per second. Strong replication offers better throughput
-at the cost of greater latency due to inter-replica coordination. The \name
-deployment offers the benefit of replication, while only paying the cost of
-coordination when necessary.
+\begin{figure}[t]
+  \centering
+  \subfigure[Latency]{\label{grf:rubis-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:rubis-tp}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:Rubis-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp-vs-lat.pdf}}
+	\caption{Rubis bidding mix performance.}
+  \label{grf:rubis}
+\end{figure}
+
+Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
+configuration against a single replica (NoRep) and strongly replicated
+(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
+15\% read-write interactions, which is representative of the auction workload.
+Without replication, NoRep trivially provides strong consistency. However, this
+deployment does not scale beyond 1750 operations per second. Strong replication
+offers better throughput at the cost of greater latency due to inter-replica
+coordination. \name deployment offers the benefit of replication, while only
+paying the cost of coordination when necessary.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.45\textwidth]{graphs/summarization.pdf}
+\caption{Impact of summarization.}
+\label{grf:summarization}
+\end{figure}
 
 Finally, we study the impact of summarization in
-Figure~\ref{grf:summarization}. We use 128 clients and a single \name replica,
-with all clients operating on the \emph{same} LWW register to stress test the
-summarization mechanism. The shim layer cache (memory) is summarized every 64
-updates, while the updates in the backing store (disk) are summarized every
-4096 updates. Each point in the graph represents the average latency of the
-previous 1000 operations. Each experiment is run for one minute.  Without
-summarization, the average latency of operations increases exponentially to
-almost one second, and only 13K operations were performed in a minute. Since
-every operation has to reduce over the set of all previous operations,
-operations take increasingly more time to complete since they must contend with
-an ever growing set.  With summarization only in memory, performance still
+Figure~\ref{grf:summarization}. We utilize 128 clients and a single \name
+replica, with all the clients operating on the \emph{same} LWW register to
+stress test the summarization mechanism. The shim layer cache (mem) of
+operations is summarized every 64 updates, while the updates in the backing
+store (disk) are summarized every 4096 updates. Each point in the graph
+represents the average latency of the previous 1000 operations. Each experiment
+is run for 60s.
+
+The results show that without summarization, the average latency of operations
+increase exponentially to almost 1 second, and only 13K operations were
+performed in a minute. Since every operation has to reduce over the set of all
+previous operations, with a ever growing set, the operations take increasingly
+more time to complete. With summarization only in memory, the performance still
 degrades due to the cost of fetching all previous updates from the backing
 store into the shim layer. Fetching the latest updates is essential for SC
-operations. With summarization enabled on both disk and memory, latency does
-not increase over time, and the implementation realizes throughput of 67K
-operations/minute.
+operations. With both summarizations enabled, we see that the latency does not
+increase over time, and we were able to perform 67K operations. This graph
+illustrates the importance and effectiveness of summarization in \name.

### 2014-11-12 contract-language.tex
hfill before qed

--- 2014-11-11 contract-language.tex
@@
 \emph{current operation} - the operation for which a contract is being written.
-Notice that $\cureff$ occurs free in the contract. We will fix its scope when
-classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
-simply the name of the operation (eg: \cf{withdraw}) that induced the effect.
-We admit disjunction in types to let an effect variable range over multiple
-operation names. The contract $\small \forall (a : \tau_1 \vee \tau_2).~\psi$
-is just syntactic sugar for $\small \forall a. (\oper{a}{\tau_1} \vee
-\oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect variable ranges over all
-operation names.
+The type of an effect is simply the name of the operation (eg: \cf{withdraw})
+that induced the effect. We admit disjunction in types to let an effect variable
+range over multiple operation names. The contract $\small \forall (a : \tau_1
+\vee \tau_2).~\psi$ is just syntactic sugar for $\small \forall a.
+(\oper{a}{\tau_1} \vee \oper{a}{\tau_2}) \Rightarrow \psi$. An untyped effect
+variable ranges over all operation names.
 
@@
 
-\subsection{Contract Classification}
-\label{sec:classify}
+\subsection{Contract Comparison and Classification}
 
@@
 
-The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
-and $\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of
-assumptions about the nature of primitive relations. A \emph{well-formed}
-axiomatic execution ($\E$) is expected to satisfy these assumptions (i.e., $\E
-\models \Delta$).
+The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$ and
+$\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of assumptions
+about the nature of primitive relations. A \emph{well-formed} axiomatic
+execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
+\Delta$).
 
@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
-$\ecc$ contracts in all well-formed executions, then the same store, extended
-with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
-all well-formed \name contracts. The proof of the theorem is given in the
-supplementary material\footnote{The supplementary material also provides the
-concrete reduction rules for enforcing the consistency classes.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H3$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H4$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$, $H5$, and $H4$.
+  \hfill \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2014-11-12 motivation.tex
edits to motivation

@@
 
-\noindent It takes a list of effects (the \emph{history} of updates to that
+\noindent An operation takes a list of effects (the \emph{history} of updates to that
 object), and an input argument, and returns a result along with an optional
@@
 \emph{snapshot} of the state of the object at some replica. In this sense,
-every operation on the RDT is atomic, and thus permitting sequential reasoning
-for implementing the RDTs. Here, \cf{getBalance} is a read-only operation,
+every operation on the RDT is atomic, and thus amenable to sequential reasoning.
+Here, \cf{getBalance} is a read-only operation,
 \cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
@@
 of RDTs for realistic benchmarks including shopping carts, auction and
-micro-blogging sites in few tens of lines of code.
+micro-blogging sites, etc. in a few tens of lines of code, expressed in this style.
 
@@
 elements from the original set. Since the notion of observable equivalence is
-specific to each RDT, we expect the programmer to provide a summarization
-function (\cf{summarize}) of type \cf{[e] -> [e]} for each RDT, as a part of
-the RDT specification.
+specific to each RDT, programmers can provide a summarization
+function - (\cf{summarize}) of type \cf{[e] -> [e]} -  as a part of
+the RDT specification. The summarization function for the bank account is:
 
-The summarization function for the bank account is \cf{summarize hist =
-[Deposit \$ sel1 \$ getBalance hist ()]}. Given a bank account history
-\cf{hist}, the \cf{summarize} function returns a new history with a single
-deposit of the current account balance. We utilize the summarization function
-to ensure that the state of an object remains summarized in each replica.
+{\small
+\begin{verbatim}
+ summarize hist = 
+   [Deposit $ sel1 $ getBalance hist ()]
+\end{verbatim}}
+
+\noindent Given a bank account history \cf{hist}, the \cf{summarize}
+function returns a new history with a single deposit of the current
+account balance. Our implementation invokes the summarization function
+associated with an RDT to reduce the size of the effect sets
+maintained by replicas.
 

### 2014-11-13 motivation.tex
Minor edit to where footnote is placed

--- 2014-11-12 motivation.tex
@@
 by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
-witnesses\footnote{Although visibility and session
+witnesses the deposit and succeeds\footnote{Although visibility and session
 order relations relate effects, we have abused the notation in these examples
 to relate operations, with the idea that the relations relate the effect
-emitted by those operations} the deposit and succeeds. Subsequently, session 2 perform a \cf{withdraw}
+emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
 operation, but importantly, due to eventual consistency, only witnesses the

### 2014-11-13 contract-language.tex
Minor edit

--- 2014-11-12 contract-language.tex
@@
 potential model for any first-order formula ($\cv$) expressible in our contract
-language. If $\E$ is indeed a valid model for $\cv$ (expressed using the
-model-theoretic consequence relation as $\E \models \cv$), we say that the
+language. If $\E$ is indeed a valid model for $\cv$ (written as $\E \models \cv$), we say that the
 execution $\E$ satisfied the contract $\cv$:

### 2014-11-13 model.tex
Fixes

@@
 relations that our contract language is seeded with. Figure~\ref{fig:sysmod}
-provides a schematic diagram of our system model. The distributed store is
+presents a schematic diagram of our system model. The distributed store is
 composed of a collection of \emph{replicas}, each of which stores a set of
@@
 all updates (\emph{effects}) performed on the object. For example, the state of
-$x$ at replica 1 is the ordered set composed of effects $w^x_1$ and $w^x_2$.
+$x$ at replica 1 is the set composed of effects $w^x_1$ and $w^x_2$.
 

### 2014-11-13 contract-language.tex
minor edit
Fixes

@@
 potential model for any first-order formula ($\cv$) expressible in our contract
-language. If $\E$ is indeed a valid model for $\cv$ (written as $\E \models \cv$), we say that the
-execution $\E$ satisfied the contract $\cv$:
+language. If $\E$ is indeed a valid model for $\cv$ (written as $\E \models
+\cv$), we say that the execution $\E$ satisfied the contract $\cv$:
 \begin{definition}

### 2014-11-13 evaluation.tex
hfill before qed

--- 2014-11-12 evaluation.tex
@@
 \begin{figure*}
-  \vspace{-5pt}
   \centering
-  \subfigure[Bank account]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.24\textwidth]{graphs/BA-tp-vs-lat.pdf}}
-  \subfigure[LWW transactions]{\label{grf:LWW-txn}\includegraphics[width=0.24\textwidth]{graphs/LWW-txn.pdf}}
-  \subfigure[RUBiS bidding mix]{\label{grf:rubis}\includegraphics[width=0.24\textwidth]{graphs/Rubis.pdf}}
-  \subfigure[Impact of summarization]{\label{grf:summarization}\includegraphics[width=0.235\textwidth]{graphs/summarization.pdf}}
-	\caption{Quelea Performance.}
-  \label{grf:LWW_perf}
+  \subfigure[Latency]{\label{grf:BA-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:BA-tp}\includegraphics[width=0.32\textwidth]{graphs/BA-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:BA-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/BA-tp-vs-lat.pdf}}
+	\caption{Bank account performance.}
+  \label{grf:BA}
 \end{figure*}
 
-Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations
-in the bank account example as we increase the number of clients in a \cf{1DC}
-configuration. Our client workload was generated using the YCSB
-benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
-the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
-corresponds to the default 50:50 read:write mix in YCSB. We increased the
-number of clients from 128 to 1024, and each experiment ran for 180 seconds.
+Figure~\ref{grf:BA} shows the performance of operations in bank account example
+as we increase the number of clients in \cf{1DC} configuration.. Our client
+workload was generated using YCSB benchmark~\cite{YCSB}. The benchmark
+uniformly chose from 100,000 keys, where the operation spread was 25\%
+withdraw, 25\% deposit and 50\% getBalance, which corresponds to the default
+50:50 read:write mix in YCSB. We increased the number of clients from 128 to
+1024, and each experiment ran for 180 seconds.
 
@@
 
+\begin{figure*}
+  \centering
+  \subfigure[Latency]{\label{grf:LWW-txn-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:LWW-txn-tp}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:LWW-txn-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/LWW-txn-tp-vs-lat.pdf}}
+	\caption{LWW register transaction performance.}
+  \label{grf:LWW-txn}
+\end{figure*}
+
 We compare the performance of different transaction isolation level choices in
@@
 We also compared (not shown) the performance of EC LWW operations directly
-against Cassandra, which uses last-writer-wins as the only convergence
-semantics. While Cassandra provides no stronger-than-eventual consistency
-properties, \name was within 30\%(20\%) of latency(throughput) of Cassandra
-with 512 clients, supporting our thesis that programmers only have to incur
-relatively low overhead for a more expressive programming model which provides
-stronger provable consistency guarantees.
+against Cassandra (our backing store), which uses last-writer-wins as the only
+convergence semantics. While Cassandra provides no stronger-than-eventual
+consistency properties, \name was within 30\%(20\%) of latency(throughput) of
+Cassandra with 512 clients, illustrating that the programmers only have to pay
+a minimal overhead for the expressive and stronger \name programming model.
 
-Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
-configuration against a single replica ({\sf\small NoRep}) and a strongly
-replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
-default RUBiS bidding mix, which has 15\% read-write interactions, which is
-representative of the auction workload.  Without replication, {\sf\small NoRep}
-trivially provides strong consistency. However, this deployment does not scale
-beyond 1750 operations per second. Strong replication offers better throughput
-at the cost of greater latency due to inter-replica coordination. The \name
-deployment offers the benefit of replication, while only paying the cost of
-coordination when necessary.
+\begin{figure*}
+  \centering
+  \subfigure[Latency]{\label{grf:rubis-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-lat.pdf}}
+  \subfigure[Throughput]{\label{grf:rubis-tp}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp.pdf}}
+  \subfigure[Throughput vs. Latency]{\label{grf:Rubis-tp-vs-lat}\includegraphics[width=0.32\textwidth]{graphs/Rubis-tp-vs-lat.pdf}}
+	\caption{Rubis bidding mix performance.}
+  \label{grf:rubis}
+\end{figure*}
+
+Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in \cf{1DC}
+configuration against a single replica (NoRep) and strongly replicated
+(StrongRep) \cf{1DC} deployment. The benchmark was RUBiS bidding mix, which has
+15\% read-write interactions, which is representative of the auction workload.
+Without replication, NoRep trivially provides strong consistency. However, this
+deployment does not scale beyond 1750 operations per second. Strong replication
+offers better throughput at the cost of greater latency due to inter-replica
+coordination. \name deployment offers the benefit of replication, while only
+paying the cost of coordination when necessary.
+
+\begin{figure}
+\centering
+\includegraphics[width=0.75\columnwidth]{graphs/summarization.pdf}
+\caption{Impact of summarization.}
+\label{grf:summarization}
+\end{figure}
 

### 2014-11-13 implementation.tex
Added basic axioms

@@
 \begin{center}
-\includegraphics[width=0.7\columnwidth]{Figures/ImplModel}
+\includegraphics[width=\columnwidth]{Figures/ImplModel}
 \end{center}
@@
 packaged as a library.
-
-\subsection{Shim Layer}
 
@@
 
-\subsection{Summarization}
-
-The main challenge in realizing an efficient implementation of operation-based
-replicated data types is that the state of the object defined in terms of its
-set of effects grows with every effectful operation. If left unchecked,
-performance slows down over time, until the shim layer memory or backing store
-disk runs out of memory. Luckily, the state of the operation-based replicated
-data type can often be summarized to an \emph{observably equivalent} smaller
-state. For example, a last-writer-wins register with multiple updates can be
-summarized to an equivalent register with only the last write. Similarly, a
-bank account with a series of deposits and updates with balance $b$ is
-equivalent to a bank account with a single deposit of $b$. A set with a
-collection of add and remove operations is equivalent to a set with a series of
-addition of live elements from the original set.
-
-Since the semantics of summarization is specific to each data type, \name
-expects a summarization function (\cf{summarize}) of type \cf{[e]->[e]} for
-each RDT, where the result is smaller than the argument but is observably
-equivalent. We utilize the \cf{summarize} function to summarize the object
-state both in the shim layer node and the backing store, typically when the
-number of effects on an object crosses a tunable threshold. Shim layer
+We utilize the \cf{summarize} function (\S~\ref{sec:summarize}) to summarize
+the object state both in the shim layer node and the backing store, typically
+when the number of effects on an object crosses a tunable threshold. Shim layer
 summarization is straight-forward; a summarization thread takes the local lock
-on the object, and replaces its state with the summarized state. The shim layer
-node remains unavailable for that particular object during summarization
-(usually a few milliseconds).
-
-Performing summarization in the backing store is more complicated since the
-whole process needs to be atomic from a client's perspective, but Cassandra
-does not provide multi-row transactions. We have engineered and implemented
-a scalable summarization mechanism for the backing store that permits
-concurrent client operations, but nonetheless prohibits these operations
-from witnessing intermediate states of the summarization process.
+on the cached object, and replaces its state with the summarized state. The
+shim layer node only remains unavailable for that particular object during
+summarization (usually a few milliseconds). Performing summarization in the
+backing store is more complicated since the whole process needs to be atomic
+from a client's perspective, but Cassandra does not provide multi-row
+transactions. We have engineered and implemented a scalable summarization
+mechanism for the backing store that permits concurrent client operations, but
+nonetheless prohibits these operations from witnessing intermediate states of
+the summarization process.

### 2014-11-13 related.tex

--- 2014-11-05 related.tex
@@
 
-Program analysis has been proposed to relieve the annotation burden associated
-with consistency classification.~\cite{Calm} identifies code regions that
-require injection of coordination.~\cite{Sieve} automatically classifies
-operations into strong and eventual consistency.~\cite{IConfluence} defines the
-conditions under which an operation requires coordination. These works focus of
-coarse-grained classification (eventual or strong), and do not focus on
-transaction consistency. Indeed, we do see these approaches as a next step in
-\name to automatically infer the contracts.
+Several conditions have been proposed to judge whether an operation on
+a replicated data object needs coordination or not. ~\cite{Calm}
+defines \emph{logical monotonicity} as a sufficient condition for
+coordination freedom, and proposes a consistency analysis that marks
+code regions performing non-monotonic reasoning (eg: aggregations,
+such as \cf{COUNT}) as potential coordination points.
+~\cite{IConfluence} and ~\cite{Sieve} define \emph{invariant
+confluence} and \emph{invariant safety}, respectively, as conditions
+for safely executing an operation without coordination. ~\cite{Sieve}
+also proposes a program analysis that conservatively marks certain
+operations as \emph{blue} (coordination not required), while marking
+the remaining as \emph{red} (coordination required). Unlike \name,
+these works focus on a coarse-grained classification of consistency as
+eventual or strong, and do not focus on finer-grained transaction
+isolation levels. However, the analyses they propose relieve
+programmers of the burden to tag operations with consistency levels.
+Indeed, we do consider automatic inference of consistency contracts
+from application-specific integrity constraints as the next step for
+\name.

### 2014-11-13 evaluation.tex
edits

--- 2014-11-12 evaluation.tex
@@
 
-Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations
-in the bank account example as we increase the number of clients in a \cf{1DC}
+Figure~\ref{grf:BA-tp-vs-lat} shows throughput vs. latency of operations in the
+bank account example as we increase the number of clients in a \cf{1DC}
 configuration. Our client workload was generated using the YCSB
@@
 operations per transaction, with a default 50:50 read:write mix. Each operation
-is assumed to be eventually consistent. \cf{NoTxn} corresponds to a
-configuration that does not use transactions. Compared to this, RC is only 12\%
-shower in terms of latency with 512 clients, whereas RR is 2.3X slower. The
-difference between RC and \cf{NoTxn} is due to the meta-data overhead of
-recording transaction information in the object state. For RR transactions, the
-cost of capturing and maintaining a snapshot is the biggest source of overhead.
+is assumed to be eventually consistent. NoTxn corresponds to a configuration
+that does not use transactions. Compared to this, RC is only 12\% shower in
+terms of latency with 512 clients, whereas RR is 2.3X slower. The difference
+between RC and NoTxn is due to the meta-data overhead of recording transaction
+information in the object state. For RR transactions, the cost of capturing and
+maintaining a snapshot is the biggest source of overhead.
 
@@
 Figure~\ref{grf:rubis} compares the \name implementation of RUBiS in a \cf{1DC}
-configuration against a single replica ({\sf\small NoRep}) and a strongly
-replicated ({\sf\small StrongRep}) \cf{1DC} deployment. The benchmark uses the
-default RUBiS bidding mix, which has 15\% read-write interactions, which is
-representative of the auction workload.  Without replication, {\sf\small NoRep}
-trivially provides strong consistency. However, this deployment does not scale
-beyond 1750 operations per second. Strong replication offers better throughput
-at the cost of greater latency due to inter-replica coordination. The \name
-deployment offers the benefit of replication, while only paying the cost of
-coordination when necessary.
+configuration against a single replica (NoRep) and a strongly replicated
+(StrongRep) \cf{1DC} deployment. The benchmark uses the default RUBiS bidding
+mix, which has 15\% read-write interactions, which is representative of the
+auction workload.  Without replication, NoRep trivially provides strong
+consistency. However, this deployment does not scale beyond 1750 operations per
+second. Strong replication offers better throughput at the cost of greater
+latency due to inter-replica coordination. The \name deployment offers the
+benefit of replication, while only paying the cost of coordination when
+necessary.
 
@@
 degrades due to the cost of fetching all previous updates from the backing
-store into the shim layer. Fetching the latest updates is essential for SC
-operations. With summarization enabled on both disk and memory, latency does
-not increase over time, and the implementation realizes throughput of 67K
-operations/minute.
+store into the shim layer. Fetching the latest updates from the backing store
+is essential for SC operations. With summarization enabled on both disk and
+memory, latency does not increase over time, and the implementation realizes
+throughput of 67K operations/minute.

### 2014-11-13 transactions.tex
edits

@@
 
-\subsection{Syntax and Semantics Extension}
+\subsection{Syntax and Semantics Extensions}
 \label{sec:syn_sem_ext}
@@
 transfers \cf{amt} from current to savings, and \cf{totalBalance}, which
-returns the sum of the balances of individual accounts. The \name code for the
-transactions is given below:
+returns the sum of the balances of individual accounts. Our goal is to ensure
+that \cf{totalBalance} returns the result obtained from a consistent snapshot
+of the object states. The \name code for the transactions is given below:
 \vspace{-1em}
@@
   atomically x dollar do
-    b1 <- getBalance s
+    b1 <- getBalance c
     b2 <- getBalance s
@@
 given isolation level $x$, ensuring that the effects of the operations are made
-visible atomically. Our goal is to ensure that \cf{totalBalance} returns the
-result obtained from a consistent snapshot of the object states.
+visible atomically.
 
@@
 \noindent The key idea in the above definition is that the $\small \txnZ$ primitive
-allows us to relate operations on different objects. The \cf{save} transaction
-only needs to ensure that the two writes it performs are made visible
-atomically. Since this is ensured by combining them in a transaction, \cf{save}
-does not require any additional constraints, and $\psi_1$ is simply $\small
-\true$.
+allows us to relate operations on different objects.
+
+The \cf{save} transaction only needs to ensure that the two writes it performs
+are made visible atomically. Since this is ensured by combining them in a
+transaction, \cf{save} does not require any additional constraints, and
+$\psi_1$ is simply $\small \true$.
 
@@
 the effects of another transaction $T_2$, then subsequent operations in $T_1$
-will also witness the effects of $T_2$. MAV semantics is useful for
-maintaining the integrity of foreign key constraints, materialized views and
-secondary updates. In order to implement MAV, a store only needs to keep track
-of the set of transactions $S_t$ witnessed by the running transaction, and
-before performing an operation at some replica, ensure that the replica
+will also witness the effects of $T_2$. MAV semantics is useful for maintaining
+the integrity of foreign key constraints, materialized views and secondary
+updates~\cite{BailisHAT}. In order to implement MAV, a store only needs to keep
+track of the set of transactions $S_t$ witnessed by the running transaction,
+and before performing an operation at some replica, ensure that the replica
 includes all the transactions in $S_t$. Hence, MAV is coordination-free. MAV

### 2014-11-13 evaluation.tex

--- 2014-11-13 evaluation.tex
@@
 
-Figure~\ref{grf:BA} shows the performance of operations in bank account example
-as we increase the number of clients in \cf{1DC} configuration.. Our client
-workload was generated using YCSB benchmark~\cite{YCSB}. The benchmark
-uniformly chose from 100,000 keys, where the operation spread was 25\%
-withdraw, 25\% deposit and 50\% getBalance, which corresponds to the default
-50:50 read:write mix in YCSB. We increased the number of clients from 128 to
-1024, and each experiment ran for 180 seconds.
+Figure~\ref{grf:BA} shows throughput vs. latency of operations in the bank
+account example as we increase the number of clients in a \cf{1DC}
+configuration. Our client workload was generated using the YCSB
+benchmark~\cite{YCSB}. The benchmark uniformly chooses from 100,000 keys, where
+the operation spread was 25\% withdraw, 25\% deposit and 50\% getBalance, which
+corresponds to the default 50:50 read:write mix in YCSB. We increased the
+number of clients from 128 to 1024, and each experiment ran for 180 seconds.
 

### 2015-01-20 PLDI15_Response/response.txt
edits

We would like to thank the reviewers for their insightful comments. We first
address the concerns raised my multiple reviewers first followed by individual
concerns.

Complexity of contracts
-----------------------
We tend to disagree with the idea that the difficulty of expressing the
contracts is the same as choosing the correct store consistency level. While
the programmer does need to translate the integrity constraints in the
application to consistency constraints, the advantage is that the programmers
write contracts for the abstract system model. This results in generic RDTs
that can be realized on different data stores (as our implementation
illustrates) with varied consistency lattices.

The Quelea contract language itself is built upon three primitive,
well-understood relations -- vis, so and sameobj, along with the usual logical
and relational operators. Moreover, in the presence of coordination-free
transactions with weaker isolation levels, the choice of correct consistency
level is non-trivial. Quelea incorporates transactions with a single additional
primitive relation sametxn. We believe the apparent complexity of contracts can
be further reduced using syntactic sugar, similar to our relational operators.

Comparison to previous work
---------------------------
Our work varies significantly from the previous works Bailis2013b and
Burckhardt2014. While Quelea provides a generic framework to describe new RDTs
and verifiably compile them to existing data stores, Bailis2013b is concerned
with implementing stronger consistency levels on top of existing stores.
Bailis2013b is not concerned with language abstractions, but rather an
implementation strategy. Our implementation, while sharing similarities with
Bailis2013b, illustrates that the Quelea programming model can be efficiently
realized on existing data stores. Moreover, Bailis2013b did not incorporate
transactions, which Quelea does.

Burckhardt2014 describes a formal framework for expressing, verifying and
comparing different forms of eventual consistency, and is not meant for
practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
Quelea focuses on automatic verification of contracts using SMT solver, which
is essential for practicality. Quelea additionally incorporates support for
programming with coordination-free transactions.

Compile Time
------------
Our contract classification through the solver took on average took 11.5 ms
across the benchmarks. We expect this SMT solver overhead to remain low as the
application scales. This is because we do not analyze the RDT/transaction
implementation to obtain the verification conditions but rather simply analyze
the user provided contracts. The contracts are encoded as SMT formulas. In
particular, each of our four primitive relations is encoded as an uninterpreted
function. Hence, the state space for the validity check remains low. Moreover,
our language is defined carefully such that the validity check is decidable for
which efficient decision procedures exist.

The time to classify a contract is proportional to the number of consistency
levels in the partially ordered lattice. For example, Doug Terry et al.'s
Session Guarantees for Weakly Consistent Replicated Data introduced 4
incomparable session guarantees, and the consistency level for an operation is
any combination of the 4 session guarantees. The classification here, at worst,
requires 4! = 24 comparisons (SMT validity tests). Since each comparison is
very cheap, the solution remains scalable.

Review A
--------
The challenge with a dynamic strategy for enforcing contracts is the
translation of contracts, which axiomatically describe correct execution to a
validity check over the dynamic execution state. Such a translation is
non-obvious.

We used the same number of shim layer nodes for the results in Figure 9 a,b,c
-- 5. Shim Layer nodes are soft-state entities and can be elastically scaled on
demand. All the benchmarks except Twissandra and RUBiS were written from
scratch.

Review B
--------
Bounded staleness could be accommodated into the contract language, but the
contract classification becomes undecidable since the theory of integer
arithmetic is undecidable.

RDT specification language does to some extent guards against
malicious/incorrect code. In particular, RDT operations are side-effect free.
While it is possible that the reduction can end up not terminating, it cannot
do any more harm.

A crashed operation does not affect the durable log. An operation is added to
the log only if it is successful. An operation is considered successful if it
produces a result.

R+ could be a total relation, but the R+ is still useful because it is not
required to be a total relation. For example, consider causal
consistency (CC) contract which is defined as:

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of  transitive closure of (soo U vis). Nevertheless, all we can
assume about hbo is that it contains the  transitive closure of (soo U vis);
there is no proof of membership for other elements in hbo.  Consequently, by
asserting hbo \subseteq vis, the strongest guarantee that causal consistency 
contract is offering is that the transitive closure of (soo U vis) is contained
in vis. An application  contract (\psi) which requires more than the transitive
closure of (soo U vis) to be contained in vis  cannot be satisfied under CC
because the following implication is simply not valid even under our lenient
definition of “+“ :

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

Review E
--------
Our idea is to parameterize the classification procedure for each target data
store. Hence, if multiple partially ordered levels are appropriate, the
classification procedure can be encoded to choose the desired consistency
level based on external knowledge such as price.

### 2015-01-20 PLDI15_Response/response.txt
Edit rebuttal

@@
 level is non-trivial. Quelea incorporates transactions with a single additional
-primitive relation sametxn. We believe the apparent complexity of contracts can
+primitive relation (sametxn). We believe the apparent complexity of contracts can
 be further reduced using syntactic sugar, similar to our relational operators.
@@
 Our work varies significantly from the previous works Bailis2013b and
-Burckhardt2014. While Quelea provides a generic framework to describe new RDTs
+Burckhardt2014. <!-- While Quelea provides a generic framework to describe new RDTs
 and verifiably compile them to existing data stores, Bailis2013b is concerned
@@
 realized on existing data stores. Moreover, Bailis2013b did not incorporate
-transactions, which Quelea does.
+transactions, which Quelea does. -->
+<!-- GK: 1. I think the statement "a generic framework to describe new RDTs and
+verifiably compile them to existing data stores" does not sufficiently
+distinguish Quelea from Bailis2013b. -->
+
+Bailis2013b describes implement strategies to realize stronger consistency
+levels on top of eventually consistent stores. In contrast, Quelea provides a
+language framework to describe application-level consistency requirements and
+automatically map them to *some* implementation of appropriate store consistency
+level. As such, Quelea's contract classification is independent of how various
+consistency levels are actually implemented in the data store.  Furthermore,
+Bailis2013b did not incorporate transactions, which Quelea does. Nevertheless,
+our own implementation of strong consistency levels on top of Cassandra does
+bear some similarities with Bailis2013b.
 
@@
 Our contract classification through the solver took on average took 11.5 ms
-across the benchmarks. We expect this SMT solver overhead to remain low as the
-application scales. This is because we do not analyze the RDT/transaction
-implementation to obtain the verification conditions but rather simply analyze
-the user provided contracts. The contracts are encoded as SMT formulas. In
-particular, each of our four primitive relations is encoded as an uninterpreted
-function. Hence, the state space for the validity check remains low. Moreover,
-our language is defined carefully such that the validity check is decidable for
-which efficient decision procedures exist.
+across the benchmarks (this statistic is also available on p.11 of the paper).
+We expect this SMT solver overhead to remain low as the application scales. This
+is because we do not analyze the RDT/transaction implementation to obtain the
+verification conditions but rather simply analyze the user provided contracts.
+Further, our contract language is defined carefully such that the validity check
+is decidable for which efficient decision procedures exist. Moreover, our
+contract language artifacts (e.g: primitive relations, effects), are encoded as
+uninterpreted symbols in SMT without the need for expensive theories, such as
+linear arithmetic.  Hence, the state space for the validity check remains low.
 

### 2015-01-20 PLDI15_Response/response.txt
edits

--- 2015-01-20 PLDI15_Response/response.txt
@@
-We would like to thank the reviewers for their insightful comments. We first
-address the concerns raised my multiple reviewers first followed by individual
-concerns.
+We thank the reviewers for their detailed reviews and comments.
 
-Complexity of contracts
+Complexity of contracts  (Reviewers ...)
 -----------------------
-We tend to disagree with the idea that the difficulty of expressing the
-contracts is the same as choosing the correct store consistency level. While
-the programmer does need to translate the integrity constraints in the
-application to consistency constraints, the advantage is that the programmers
-write contracts for the abstract system model. This results in generic RDTs
-that can be realized on different data stores (as our implementation
-illustrates) with varied consistency lattices.
 
-The Quelea contract language itself is built upon three primitive,
-well-understood relations -- vis, so and sameobj, along with the usual logical
-and relational operators. Moreover, in the presence of coordination-free
-transactions with weaker isolation levels, the choice of correct consistency
-level is non-trivial. Quelea incorporates transactions with a single additional
-primitive relation sametxn. We believe the apparent complexity of contracts can
-be further reduced using syntactic sugar, similar to our relational operators.
+We believe the difficulty of writing contracts is substantially simpler than
+choosing the correct store consistency level since developers do not need to
+understand the details of the underlying data store.
 
-Comparison to previous work
+Moreover, while it is necessary to translate integrity constraints in the
+application to consistency constraints as contracts, the tremendous
+advantage that accrues is that these contracts are expressed over an
+abstract system model, resulting in generic RDT specifications that can be
+realized on different data stores (as our implementation illustrates) with
+varied consistency lattices (as pointed out by ReviewerE).
+
+The Quelea contract language itself is built upon *only* three primitive,
+well-understood relations -- vis, so and sameobj, along with standard
+propositional logical and relational operators.  In the presence of
+coordination-free transactions with weaker isolation levels, the choice of
+correct consistency level is non-trivial. Quelea incorporates transactions
+with a single additional primitive relation (sametxn).
+
+< provide an example using syntactic sugar >
+
+Comparison to previous work (Reviewers ...)
 ---------------------------
-Our work varies significantly from the previous works Bailis2013b and
-Burckhardt2014. While Quelea provides a generic framework to describe new RDTs
-and verifiably compile them to existing data stores, Bailis2013b is concerned
-with implementing stronger consistency levels on top of existing stores.
-Bailis2013b is not concerned with language abstractions, but rather an
-implementation strategy. Our implementation, while sharing similarities with
-Bailis2013b, illustrates that the Quelea programming model can be efficiently
-realized on existing data stores. Moreover, Bailis2013b did not incorporate
-transactions, which Quelea does.
+
+Our work differs significantly from Bailis2013b and Burckhardt2014.
 
@@
 comparing different forms of eventual consistency, and is not meant for
-practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
-Quelea focuses on automatic verification of contracts using SMT solver, which
-is essential for practicality. Quelea additionally incorporates support for
-programming with coordination-free transactions.
+practical programming over ECDS like Quelea. Moreover, unlike
+Burckhardt2014, Quelea focuses on *automatic* verification of RDT
+specifications, which is essential for practicality, and additionally
+incorporates support for programming with coordination-free transactions.
 
-Compile Time
-------------
+Bailis2013b describes strategies to realize stronger consistency levels on
+top of an eventually consistent store, but does not consider
+parameterization of application consistency constraints over arbitrary data
+store implementations.  In contrast, Quelea provides a language framework to
+describe application-level consistency requirements and automatically map
+them to *some* data store implementation. As such, Quelea's contract
+classification is independent of how various consistency levels are actually
+implemented in the data store.  Moreover, Bailis2013b does not incorporate
+transactions, which Quelea does.
+
+Compilation Time  (Reviewers ...)
+----------------
+
 Our contract classification through the solver took on average took 11.5 ms
-across the benchmarks. We expect this SMT solver overhead to remain low as the
-application scales. This is because we do not analyze the RDT/transaction
-implementation to obtain the verification conditions but rather simply analyze
-the user provided contracts. The contracts are encoded as SMT formulas. In
-particular, each of our four primitive relations is encoded as an uninterpreted
-function. Hence, the state space for the validity check remains low. Moreover,
-our language is defined carefully such that the validity check is decidable for
-which efficient decision procedures exist.
+across the benchmarks (this statistic is also available on p.11 of the
+paper).  We expect this SMT solver overhead to remain low as the application
+scales. This is because we do not analyze the RDT/transaction implementation
+to obtain the verification conditions but only analyze user-provided
+contracts.  Moreover, our contract language artifacts (e.g: primitive
+relations, effects), are encoded as uninterpreted symbols in SMT without the
+need for expensive theories, such as linear arithmetic.  Hence, the state
+space for the validity check remains low.
 
@@
 levels in the partially ordered lattice. For example, Doug Terry et al.'s
-Session Guarantees for Weakly Consistent Replicated Data introduced 4
-incomparable session guarantees, and the consistency level for an operation is
-any combination of the 4 session guarantees. The classification here, at worst,
-requires 4! = 24 comparisons (SMT validity tests). Since each comparison is
-very cheap, the solution remains scalable.
+Session Guarantees for Weakly Consistent Replicated Data (cite ...)
+introduced 4 incomparable session guarantees, and the consistency level for
+an operation is any combination of the 4 session guarantees. The
+classification here, at worst, requires 4! = 24 comparisons (SMT validity
+tests). Since each comparison is very cheap, the solution remains scalable.
 
@@
 --------
+
 The challenge with a dynamic strategy for enforcing contracts is the
-translation of contracts, which axiomatically describe correct execution to a
-validity check over the dynamic execution state. Such a translation is
+translation of contracts, which axiomatically describe correct execution to
+a validity check over the dynamic execution state. Such a translation is
 non-obvious.
 
-We used the same number of shim layer nodes for the results in Figure 9 a,b,c
--- 5. Shim Layer nodes are soft-state entities and can be elastically scaled on
-demand. All the benchmarks except Twissandra and RUBiS were written from
-scratch.
+We used the same number of shim layer nodes for the results in Figure 9
+a,b,c -- 5. Shim Layer nodes are soft-state entities and can be elastically
+scaled on demand. All the benchmarks except Twissandra and RUBiS were
+written from scratch.
 
@@
 --------
-Bounded staleness could be accommodated into the contract language, but the
+
+Bounded staleness could be accommodated into the contract language, but
 contract classification becomes undecidable since the theory of integer
@@
 
-RDT specification language does to some extent guards against
-malicious/incorrect code. In particular, RDT operations are side-effect free.
-While it is possible that the reduction can end up not terminating, it cannot
-do any more harm.
+The RDT specification language does to some extent guard against
+malicious/incorrect code. In particular, RDT operations are side-effect
+free.  While it is possible that the reduction can end up not terminating,
+it cannot corrupt system state.
 
@@
 
-R+ could be a total relation, but the R+ is still useful because it is not
-required to be a total relation. For example, consider causal
-consistency (CC) contract which is defined as:
+R+ could be a total relation, but R+ is still useful because it is not
+required to be a total relation. For example, consider a causal
+consistency (CC) contract defined as:
 
@@
 where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
-any superset of  transitive closure of (soo U vis). Nevertheless, all we can
-assume about hbo is that it contains the  transitive closure of (soo U vis);
-there is no proof of membership for other elements in hbo.  Consequently, by
-asserting hbo \subseteq vis, the strongest guarantee that causal consistency 
-contract is offering is that the transitive closure of (soo U vis) is contained
-in vis. An application  contract (\psi) which requires more than the transitive
-closure of (soo U vis) to be contained in vis  cannot be satisfied under CC
-because the following implication is simply not valid even under our lenient
-definition of “+“ :
+any superset of the transitive closure of (soo U vis). Nevertheless, all we
+can assume about hbo is that it contains the  transitive closure of (soo U
+vis); there is no proof of membership for other elements in hbo. 
+Consequently, by asserting hbo \subseteq vis, the strongest guarantee that
+causal consistency  contract is offering is that the transitive closure of
+(soo U vis) is contained in vis. An application  contract (\psi) which
+requires more than the transitive closure of (soo U vis) to be contained in
+vis  cannot be satisfied under CC because the following implication is
+simply not valid even under our lenient definition of “+“ :
 
@@
 --------
-Our idea is to parameterize the classification procedure for each target data
-store. Hence, if multiple partially ordered levels are appropriate, the
+
+Our idea is to parameterize the classification procedure for each target
+data store. Hence, if multiple partially ordered levels are appropriate, the
 classification procedure can be encoded to choose the desired consistency

### 2015-01-20 PLDI15_Response/response.txt
edits

@@
 We believe the difficulty of writing contracts is substantially simpler than
-choosing the correct store consistency level since developers do not need to
-understand the details of the underlying data store.
+choosing a correct store consistency level, since developers do not need to
+understand the details of the underlying data store to define consistency
+properties relevant for the application.
 
-Moreover, while it is necessary to translate integrity constraints in the
-application to consistency constraints as contracts, the tremendous
-advantage that accrues is that these contracts are expressed over an
-abstract system model, resulting in generic RDT specifications that can be
-realized on different data stores (as our implementation illustrates) with
-varied consistency lattices (as pointed out by ReviewerE).
+Moreover, while it is necessary to translate application-specific integrity
+constraints to consistency constraints in the form of contracts, the
+tremendous advantage that accrues is that these contracts are expressed over
+an abstract system model, resulting in generic RDT specifications that can
+be realized on different data stores (as our implementation illustrates)
+with varied consistency lattices (as pointed out by ReviewerE).
 
@@
 coordination-free transactions with weaker isolation levels, the choice of
-correct consistency level is non-trivial. Quelea incorporates transactions
-with a single additional primitive relation (sametxn).
+choosing the correct consistency level is non-trivial. Note also that Quelea
+incorporates transactions with a single additional primitive relation
+(sametxn).
 

### 2015-01-21 PLDI15_Response/response.txt
edits

@@
 
-Complexity of contracts  (Reviewers ...)
+Complexity of contracts  (Reviews B,C,D,E)
 -----------------------
@@
 Moreover, while it is necessary to translate application-specific integrity
-constraints to consistency constraints in the form of contracts, the
-tremendous advantage that accrues is that these contracts are expressed over
-an abstract system model, resulting in generic RDT specifications that can
-be realized on different data stores (as our implementation illustrates)
-with varied consistency lattices (as pointed out by ReviewerE).
+constraints to consistency constraints in the form of contracts, the tremendous
+advantage that accrues is that these contracts are expressed over an abstract
+system model, resulting in generic RDT specifications that can be realized on
+different data stores (as our implementation illustrates) with varied
+consistency lattices (as pointed out by ReviewerE).
 
@@
 
-< provide an example using syntactic sugar >
-
-Comparison to previous work (Reviewers ...)
+Comparison to previous work (Reviews C,D)
 ---------------------------
@@
 comparing different forms of eventual consistency, and is not meant for
-practical programming over ECDS like Quelea. Moreover, unlike
-Burckhardt2014, Quelea focuses on *automatic* verification of RDT
-specifications, which is essential for practicality, and additionally
-incorporates support for programming with coordination-free transactions.
+practical programming over ECDS like Quelea. Moreover, unlike Burckhardt2014,
+Quelea focuses on *automatic* verification of RDT specifications, which is
+essential for practicality, and additionally incorporates support for
+programming with coordination-free transactions.
 
-Bailis2013b describes strategies to realize stronger consistency levels on
-top of an eventually consistent store, but does not consider
-parameterization of application consistency constraints over arbitrary data
-store implementations.  In contrast, Quelea provides a language framework to
-describe application-level consistency requirements and automatically map
-them to *some* data store implementation. As such, Quelea's contract
-classification is independent of how various consistency levels are actually
-implemented in the data store.  Moreover, Bailis2013b does not incorporate
-transactions, which Quelea does.
+Bailis2013b describes strategies to realize stronger consistency levels on top
+of an eventually consistent store, but does not consider parameterization of
+application consistency constraints over arbitrary data store implementations.
+In contrast, Quelea provides a language framework to describe application-level
+consistency requirements and automatically map them to *some* data store
+implementation. As such, Quelea's contract classification is independent of how
+various consistency levels are actually implemented in the data store.
+Moreover, Bailis2013b does not incorporate transactions, which Quelea does.
 
-Compilation Time  (Reviewers ...)
+Compilation Time  (Reviews A,D)
 ----------------
@@
 Our contract classification through the solver took on average took 11.5 ms
-across the benchmarks (this statistic is also available on p.11 of the
-paper).  We expect this SMT solver overhead to remain low as the application
-scales. This is because we do not analyze the RDT/transaction implementation
-to obtain the verification conditions but only analyze user-provided
-contracts.  Moreover, our contract language artifacts (e.g: primitive
-relations, effects), are encoded as uninterpreted symbols in SMT without the
-need for expensive theories, such as linear arithmetic.  Hence, the state
-space for the validity check remains low.
+across the benchmarks (this statistic is also available on p.11 of the paper).
+We expect this SMT solver overhead to remain low as the application scales.
+This is because we do not analyze the RDT/transaction implementation to obtain
+the verification conditions but only analyze user-provided contracts.
+Moreover, our contract language artifacts (e.g: primitive relations, effects),
+are encoded as uninterpreted symbols in SMT without the need for expensive
+theories, such as linear arithmetic.  Hence, the state space for the validity
+check remains low.
 
 The time to classify a contract is proportional to the number of consistency
-levels in the partially ordered lattice. For example, Doug Terry et al.'s
-Session Guarantees for Weakly Consistent Replicated Data (cite ...)
-introduced 4 incomparable session guarantees, and the consistency level for
-an operation is any combination of the 4 session guarantees. The
-classification here, at worst, requires 4! = 24 comparisons (SMT validity
-tests). Since each comparison is very cheap, the solution remains scalable.
+levels in the partially ordered lattice. For example, [[Doug Terry et al.'s
+Session Guarantees for Weakly Consistent Replicated Data, PDIS '94]] introduced
+4 incomparable session guarantees, and the consistency level for an operation
+is any combination of the 4 session guarantees. The classification here, at
+worst, requires 4! = 24 comparisons (SMT validity tests). Since each comparison
+is very cheap, the solution remains scalable.
 
@@
 
-Bounded staleness could be accommodated into the contract language, but
-contract classification becomes undecidable since the theory of integer
-arithmetic is undecidable.
+Bounded staleness, which depends on wall clock time, is not incorporated in the
+contract language; as integer arithmetic, in general, is undecidable, its
+inclusion would make the classification procedure undecidable.
 
-The RDT specification language does to some extent guard against
-malicious/incorrect code. In particular, RDT operations are side-effect
-free.  While it is possible that the reduction can end up not terminating,
-it cannot corrupt system state.
-
-A crashed operation does not affect the durable log. An operation is added to
-the log only if it is successful. An operation is considered successful if it
-produces a result.
+The standard benchmarks such as TPC-C are designed for traditional relational
+database systems whose specifications dictate relational tables, joins and
+transactions with strong isolation requirements, which are not readily
+supported by NoSQL stores. As such these traditional benchmarks such as TPC are
+not used to analyze and compare ECDS performance. On the other hand, since
+Quelea was realized on top of Cassandra, we were able to obtain a point of
+comparison with respect to vanilla Cassnadra (P.10, Column.2, para.3).
 
 R+ could be a total relation, but R+ is still useful because it is not
-required to be a total relation. For example, consider a causal
-consistency (CC) contract defined as:
-
- hbo \subseteq vis
-
-where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
-any superset of the transitive closure of (soo U vis). Nevertheless, all we
-can assume about hbo is that it contains the  transitive closure of (soo U
-vis); there is no proof of membership for other elements in hbo. 
-Consequently, by asserting hbo \subseteq vis, the strongest guarantee that
-causal consistency  contract is offering is that the transitive closure of
-(soo U vis) is contained in vis. An application  contract (\psi) which
-requires more than the transitive closure of (soo U vis) to be contained in
-vis  cannot be satisfied under CC because the following implication is
-simply not valid even under our lenient definition of “+“ :
-
- \Delta |- (hbo \subseteq vis) => \psi
-
-\Delta, as usual, denotes the conjunction of basic axioms about primitive
-relations.
-
-Review E
---------
-
-Our idea is to parameterize the classification procedure for each target
-data store. Hence, if multiple partially ordered levels are appropriate, the
-classification procedure can be encoded to choose the desired consistency
-level based on external knowledge such as price.
+*required* to be a total relation. For example, consider a causal
+consistency (CC) contract defined as hbo \subseteq vis where hbo = (soo U
+vis)+. Though our definition hbo can be a superset of transitive closure (TC)
+of (soo U vis), we do not have a *proof* of inclusion of elements other than
+the ones in the TC. As such an application contract (\psi) which requires more
+than CC (say total order) would not be satisfied since the implication \Delta
+|- (hbo \subseteq vis) => \psi is not valid.

### 2015-02-18 PLDI15_Response/response.txt

@@
 Bounded staleness, which depends on wall clock time, is not incorporated in the
-contract language; as integer arithmetic, in general, is undecidable, its
+contract language as integer arithmetic, in general, is undecidable; its
 inclusion would make the classification procedure undecidable.
 
-The standard benchmarks such as TPC-C are designed for traditional relational
-database systems whose specifications dictate relational tables, joins and
-transactions with strong isolation requirements, which are not readily
-supported by NoSQL stores. As such these traditional benchmarks such as TPC are
-not used to analyze and compare ECDS performance. On the other hand, since
+The standard benchmarks such as TPC-C are designed for traditional
+relational database systems whose specifications dictate relational tables,
+joins and transactions with strong isolation requirements, which are not
+readily supported by NoSQL stores. As such these kinds of benchmarks are not
+used to analyze and compare ECDS performance. On the other hand, since
 Quelea was realized on top of Cassandra, we were able to obtain a point of

### 2015-03-17 motivation.tex
Constructing getBal contract in stages.

@@
 
-\noindent For \cf{getBalance}, we construct the following contract ($\cv_{gb}$):
+\noindent To prevent \cf{getBalance} from ever showing negative balance, it is
+necessary to prevent the scenario depicted in
+fig.~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
+emitted by the \cf{getBalance} operation. If the effect ($b$) of a withdraw
+operation is visible to $\cureff$, and the effect ($a$) of a deposit operation
+is visible to the effect ($b$) of the withdraw operation, then it must be the
+case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
+\cf{getBalance} operation precisely captures this application-level consistency
+requirement:
+
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\forall (a:\rcf{deposit}), (b:\rcf{withdraw}). \\
+\qquad\qquad (\vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff})
+\end{array}
+\end{smathpar}
+
+\noindent To prevent the missing update anomaly described in
+fig.~\ref{fig:missingUpdateAnomaly}, it is necessary for a \cf{getBalance}
+operation on a bank account to witness the effects of all previous \cf{deposit}
+and \cf{withdraw} operations performed on the same bank account in the same
+session. We can conceive an alternative contract ($\cv_{gb}^2$) for
+\cf{getBalance} to express this consistency requirement:
+
+\vspace{-1em}
+\begin{smathpar}
+\begin{array}{l}
+\forall (c: \rcf{deposit} \vee \rcf{withdraw}). \\
+\qquad \qquad((\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff})
+\end{array}
+\end{smathpar}
+
+\noindent Our contract language provides operators to compose relations. The
+syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
+above contract ($\cv_{gb}^2$) says that if a \cf{deposit} or a \cf{withdraw}
+operation precedes a \cf{getBalance} operation in session order, and is applied
+on the same object as the \cf{getBalance} operation, then it must be the case
+that the \cf{getBalance} operation witnesses the effects of the preceding
+operations.
+
+\noindent To final contract ($\cv_{gb}$) of the \cf{getBalance} operation is
+merely a conjunction of the previous two versions ($\cv_{gb}^1$ and
+$\cv_{gb}^2$):
 
@@
 
-\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
-is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
-operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
-all \cf{deposit} operations $a$ visible to $b$ should also be visible to
-$\cureff$. Intuitively, this prevents negative balance anomalies. Our contract
-language provides operators to compose relations. The syntax $(R_1 \cap
-R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
-above contract says that if a \cf{deposit} or a \cf{withdraw} operation
-precedes a \cf{getBalance} operation in session order, and is applied on the
-same object as the \cf{getBalance} operation, then it must be the case that the
-\cf{getBalance} operation witnesses the effects of the preceding operations.
+\noindent Intuitively, this prevents both the \cf{getBalance} anomalies
+described in figs.~\ref{fig:negativeBalanceAnomaly} and
+\ref{fig:missingUpdateAnomaly} from ever occurring.
 

### 2015-04-06 motivation.tex

@@
 
-\noindent To prevent \cf{getBalance} from ever showing negative balance, it is
+To prevent \cf{getBalance} from ever showing a negative balance, it is
 necessary to prevent the scenario depicted in
@@
 case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
-\cf{getBalance} operation precisely captures this application-level consistency
-requirement:
+\cf{getBalance} operation that precisely captures this application-level consistency
+requirement can be written thus:
 
@@
 and \cf{withdraw} operations performed on the same bank account in the same
-session. We can conceive an alternative contract ($\cv_{gb}^2$) for
-\cf{getBalance} to express this consistency requirement:
+session. We can express an additional contract ($\cv_{gb}^2$) for
+\cf{getBalance} that captures this consistency requirement:
 
@@
 
-\noindent To final contract ($\cv_{gb}$) of the \cf{getBalance} operation is
+The final contract ($\cv_{gb}$) of the \cf{getBalance} operation is
 merely a conjunction of the previous two versions ($\cv_{gb}^1$ and

### 2015-04-06 intro.tex
edits
Incorporated new edits
Use correct class file for the camera ready version. Remove vertical space adjustments.

@@
 To address these issues, we describe \name, a declarative programming model and
-implementation for eventually consistent geo-distributed data stores. The key
+implementation for \ecds. The key
 novelty of \name is an expressive \emph{contract language} to declare and
@@
 correctness of the mapping. The paper makes the following contributions:
-\vspace*{-.4em}
-\begin{itemize}
-\setlength{\itemsep}{2pt}
+
+\begin{itemize}[noitemsep]
 \item We introduce \name, a shallow extension of Haskell that supports the
-	description and validation of replicated data types found on eventually
-	consistent stores. Contracts are used to specify fine-grained
-	application-level consistency properties, and are statically analyzed to
-	assign the most efficient and sound store consistency level to the
-	corresponding operation.
+	description and validation of replicated data types found on \ecds. Contracts
+	are used to specify fine-grained application-level consistency properties,
+	and are statically analyzed to assign the most efficient and sound store
+	consistency level to the corresponding operation.
 \item \name supports coordination-free transactions over arbitrary datatypes.
@@
 
-\noindent The rest of the paper is organized as follows. The next section describes the
-system model.  We describe the challenges in programming with eventually
-consistent data stores, and introduce \name\ contracts as a proposed solution
-to overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
+\noindent The rest of the paper is organized as follows. The next section
+describes the system model.  We describe the challenges in programming under
+eventual consistency, and introduce \name\ contracts as a proposed solution to
+overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
 more details on the contract language, and its mapping to the store consistency
@@
 \S~\ref{sec:txns} introduces transaction contracts and their classification.
-\S~\ref{sec:impl} describes the implementation and provides details about the
-optimizations needed to make the system practical. \S~\ref{sec:results}
-discusses experimental evaluation. \S~\ref{sec:related} and~\ref{sec:concl}
-present related work and conclusions.
+\S~\ref{sec:impl} describes the implementation of \name on top of Cassandra.
+\S~\ref{sec:results} discusses experimental evaluation. \S~\ref{sec:related}
+and~\ref{sec:concl} present related work and conclusions.

### 2015-04-06 contract-language.tex
Use correct class file for the camera ready version. Remove vertical space adjustments.
Formatting changes for camera-ready

--- 2014-11-13 contract-language.tex
@@
 
-  \vspace{-1.5em}
+  \vspace{-1em}
   \begin{smathpar}
@@
 
-\item \textbf{Strong Consistency}: Strongly consistent operations may block
+\item \textbf{Strong consistency}: Strongly consistent operations may block
   indefinitely under network partitions. An example is the total-order
@@
 
-\vspace{-1.7em}
+\vspace{-1em}
 \begin{smathpar}

### 2015-04-06 transactions.tex
Use correct class file for the camera ready version. Remove vertical space adjustments.
Formatting changes for camera-ready

@@
 
-\vspace{-1.5em}
+\vspace{-1em}
 \begin{smathpar}
@@
 
-\vspace{-1.2em}
+\vspace{-1em}
 \begin{smathpar}
@@
 \centering
-\subfigure[Atomicity]{\label{fig:txn_atomicity}\includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}}
+\subfigure[Atomicity]{
+  \label{fig:txn_atomicity}
+  \includegraphics[width=0.25\columnwidth]{Figures/TxnAtomic}
+}
 \hfill
-\subfigure[Monotonic Atomic View]{\label{fig:txn_mav}\includegraphics[width=0.25\columnwidth]{Figures/TxnMAV}}
+\subfigure[Monotonic Atomic View]{
+  \label{fig:txn_mav}
+  \includegraphics[width=0.25\columnwidth] {Figures/TxnMAV}
+}
 \hfill
-\subfigure[Repeatable Read]{\label{fig:txn_rr}\includegraphics[width=0.25\columnwidth]{Figures/TxnRR}}
+\subfigure[Repeatable Read]{
+  \label{fig:txn_rr}
+  \includegraphics[width=0.25\columnwidth]{Figures/TxnRR}
+}
 \caption{Semantics of transaction contracts. $x$ and $y$ are distinct objects.
@@
 
-\vspace{-1em}
+\vspace{-0.6em}
 \begin{smathpar}
@@
 
-\vspace{-1.2em}
+\vspace{-0.6em}
 \begin{smathpar}

### 2015-04-07 contract-language.tex
Formatting edits
Bug fix in proof.

--- 2014-11-13 contract-language.tex
@@
 potential model for any first-order formula ($\cv$) expressible in our contract
-language. If $\E$ is indeed a valid model for $\cv$ (written as $\E \models
-\cv$), we say that the execution $\E$ satisfied the contract $\cv$:
+language. If $\E$ is indeed a valid model for $\cv$ (expressed using the
+model-theoretic consequence relation as $\E \models \cv$), we say that the
+execution $\E$ satisfied the contract $\cv$:
 \begin{definition}
-An axiomatic execution $\E$  satisfies a contract $\cv$ if and
+An axiomatic execution $\E$ is said to satisfy a contract $\cv$ if and
 only if $E \models \cv$.
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
+\setlength{\itemsep}{2pt}
 
@@
 	bank account example, \cf{deposit} is an eventually consistent operation.
-	While an \ecds typically offer \emph{basic} eventual consistency with all
-	possible anomalies, we assume that our store provides stronger semantics that
-	remain highly-available~\cite{BailisHAT,COPS}; the store always exposes a
-	\emph{causal cut} of the updates. This semantics can be formally captured in
-	terms of the following contract definition:
+	While eventually consistent data stores typically offer \emph{basic} eventual
+	consistency with all possible anomalies, we assume that our store provides
+	stronger semantics that remain highly-available~\cite{BailisHAT,COPS}; the
+	store always exposes a \emph{causal cut} of the updates. This semantics can
+	be formally captured in terms of the following contract definition:
 
@@
 \begin{figure}
-\vspace{-10pt}
 \begin{smathpar}
-\begin{array}{cc}
+\begin{array}{c}
+\vspace{3mm}
 \RuleTwo
 {\DDe{\cv} \le \DDe{\scc}}
-{{\sf WellFormed}(\cv)}  &
+{{\sf WellFormed}(\cv)}  \qquad
 
@@
 {\DDe{\cv} \le \DDe{\ecc}}
-{{\sf EventuallyConsistent}(\cv)} \\ \\
+{{\sf EventuallyConsistent}(\cv)} \\
 
+\vspace{3mm}
 \RuleTwo
@@
 \quad \DDe{\cv} \le \DDe{\ccc}}
-{{\sf CausallyConsistent}(\cv)} &
+{{\sf CausallyConsistent}(\cv)} \qquad
 
@@
 \end{smathpar}
-\vspace{-3mm}
+\vspace{-5mm}
 \caption{Contract classification.}
@@
 
-The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
-and $\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of
-assumptions about the nature of primitive relations. A \emph{well-formed}
-axiomatic execution ($\E$) is expected to satisfy these assumptions (i.e., $\E
-\models \Delta$).
+The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$ and
+$\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of assumptions
+about the nature of primitive relations. A \emph{well-formed} axiomatic
+execution ($\E$) is expected to satisfy these assumptions (i.e., $\E \models
+\Delta$).
 
 \begin{definition}
-An axiomatic execution $\small \E = \Exec$ is well-formed if the
+An axiomatic executions $\small \E = \Exec$ is said to be well-formed if the
 following axioms ($\Delta$) hold:
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
+\setlength{\itemsep}{2pt}
 \item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
@@
   \end{smathpar}
-  Proof follows from $H1$, $H5$, and $H4$.
-  \hfill \qed
+  Proof follows from $H1$, $H4$, and $H3$.
+  \qed
 \end{proof}

### 2015-04-07 implementation.tex
Added more implementation details.

@@
 
-Each new effect in \name is realized as a row insertion in Cassandra, and
-the state of an object is the set of all corresponding rows. The shim layer
-maintains a causally consistent in-memory snapshot of a subset of objects in
-the backing store, by explicitly tracking dependencies introduced between
-effects due to visibility, session and same transaction
-relations. Dependence tracking is similar to the techniques presented
-in~\cite{BoltOn} and~\cite{Eiger}. Because Cassandra provides durability,
-convergence, and fault tolerance, each shim layer node simply acts as a
-soft-state cache, with no inter-node communication, and can safely be
-terminated at any point. Similarly, new shim layer nodes can be spawned on
-demand.
+\subsection{Operation Consistency}
+
+The shim layer maintains a causally consistent in-memory snapshot of a subset
+of objects in the backing store, by explicitly tracking dependencies introduced
+between effects due to visibility, session and same transaction relations.
+Dependence tracking is similar to the techniques presented in~\cite{BoltOn}
+and~\cite{Eiger}. Because Cassandra provides durability, convergence, and fault
+tolerance, each shim layer node simply acts as a soft-state cache, with no
+inter-node communication, and can safely be terminated at any point. Similarly,
+new shim layer nodes can be spawned on demand.
+
+Each effect generated as a result of an effectful operation on an object
+inserts a new row $(o,\allowbreak e,\allowbreak txn,\allowbreak val,
+\allowbreak deps)$ into the backing store, where $o$ and $e$ are object and
+\emph{unique} effect identifiers, $txn$ is an optional transaction identifier,
+and $val$ is the value associated with the effect (eg: \cf{Withdraw 50}).
+$deps$ is the set of identifiers of \emph{dependencies} of this operation and
+is defined as $deps(e) = \{e_1 \ALT \vis{e_1}{e} \wedge \neg(\exists
+e_2.\vis{e_1}{e_2} \wedge \vis{e_2}{e})\}$. At any shim layer node, an effect
+is included only if all of its dependencies are also included in that node.
+This ensures that the state maintained by the shim layer node is causally
+consistent. Our dependence tracking strategy ensures that \name does not track
+every effect as the number of writes in the system grows.
 
@@
 strongly consistent operations. Strongly consistent operations are performed
-after obtaining exclusive leases on objects. The lease mechanism is
-implemented with the help of Cassandra's support for conditional updates and
-expiring columns. Cassandra does not provide general-purpose
-transactions. Since the transaction guarantees provided by \name are
-coordination-free~\cite{BailisHAT}, we realize efficient implementations by
-explicitly tracking dependencies between operations and transactions.
-Importantly, the weaker isolation semantics of transactions in \name permit
-transactions to be discharged if at least one shim layer node is reachable.
+after obtaining exclusive leases on objects. The lease mechanism is implemented
+with the help of Cassandra's support for conditional updates and expiring
+columns.
+
+\subsection{Transactions}
+
+Cassandra does not provide general-purpose transactions. Since the transaction
+guarantees provided by \name are coordination-free~\cite{BailisHAT}, we realize
+efficient implementations by explicitly tracking dependencies between
+operations and transactions. Importantly, the weaker isolation semantics of
+transactions in \name permit transactions to be discharged if at least one shim
+layer node is reachable.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.7\columnwidth]{Figures/AtomicityImpl}
+\end{center}
+\caption{Implementing atomicity semantics. Dotted circle represents effects not
+yet inserted into the backing store.}
+\label{fig:atomicity_impl}
+\end{figure}
+
+\name implements atomic visibility by exploiting shim layer causality guarantee
+-- an effect is included only if all the effects if depends on are also
+included. Consider the example given in Figure~\ref{fig:atomicity_impl}. For
+every transaction in \name, we instantiate a special transaction marker
+effect $m$. But importantly, do not insert into the backing store. $m$ is
+included as a dependence to every effect generated in the transaction. In the
+figure, the graph on the left shows the state of the store in the middle of a
+transaction. Each circle represents an effect. The dotted circle indicates that
+the effect has been instantiated, but has not yet been inserted into the store.
+Since the causally preceding effect $m$ has not yet been written to the store,
+no operation will witness $e1$ and $e2$ while the transaction in progress.
+After the transaction has finished execution, we insert $m$ into the backing
+store, marking all the effects from the transactions as a dependence for $m$.
+Now any replica which includes one of the effects from the transaction must
+include $m$, and transitively must include every effect from the transaction.
+This ensures atomicity and satisfies the RC requirement.
+
+The above scheme prevents a transaction from witnessing its own effects. This
+might conflict with the causality requirement on the operations. Hence,
+transactions piggy-back the previous effects from the same transaction for each
+request. MAV semantics is implemented by keeping track of the set of
+transaction markers $M$ witnessed by the transaction, and before performing an
+operation at some replica, ensuring that $M$ is a subset of the transaction
+markers included at that replica. If not, the missing effects are synchronously
+fetched. RR semantics is realized by capturing a optimized snapshot of the
+state of some replica; each operation from an RR transaction is applied to this
+snapshot state. Any generated effects are added to this snapshot.
+
+\subsection{Summarization}
 
@@
 the object state both in the shim layer node and the backing store, typically
-when the number of effects on an object crosses a tunable threshold. Shim layer
-summarization is straight-forward; a summarization thread takes the local lock
-on the cached object, and replaces its state with the summarized state. The
-shim layer node only remains unavailable for that particular object during
-summarization (usually a few milliseconds). Performing summarization in the
-backing store is more complicated since the whole process needs to be atomic
-from a client's perspective, but Cassandra does not provide multi-row
-transactions. We have engineered and implemented a scalable summarization
-mechanism for the backing store that permits concurrent client operations, but
-nonetheless prohibits these operations from witnessing intermediate states of
-the summarization process.
+when the number of effects on an object crosses a tunable threshold.
+Shim layer summarization is straight-forward; a summarization thread takes the
+local lock on the cached object, and replaces its state with the summarized
+state. The shim layer node only remains unavailable for that particular object
+during summarization (usually a few milliseconds).
+
+Performing summarization in the backing store is more complicated since the
+whole process needs to be atomic from a client's perspective, but Cassandra
+does not provide multi-row transactions. Summarization in the backing store
+involves deleting previously inserted rows and inserting new rows, where each
+row corresponds to an effect. It is essential that concurrent client operations
+are permitted, but are not allowed to witness the intermediate state of the
+summarization process.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{Figures/SumDisk}
+\end{center}
+\caption{Summarization in the backing store. Dotted circle represents effects
+not yet inserted into the backing store.}
+\label{fig:impl_sum_disk}
+\end{figure}
+
+To this end, we adopt a novel summarization strategy that builds on the
+causality property of the store. Figure~\ref{fig:impl_sum_disk} illustrates the
+summarization strategy. Suppose the original set of effects on an object are
+$o1$, $o2$ and $o3$. When summarized, the new effects yielded are $n1$ and
+$n2$. We first instantiate a summarization marker $s$, and similar to
+transaction marker, we do not insert it into the store immediately. We insert
+the new effects $n1$ and $n2$, with strong consistency, including $s$ as a
+dependence. Since $s$ is not yet in the store, the new effects are not made
+visible to the clients. Then we insert $s$ with strong consistency, including
+the original effects $o1$, $o2$ and $o3$ as dependence. Strongly consistent
+insertions ensure that a shim layer node witnessing $s$ on some object must
+also witness $n1$ and $n2$ on the same object. A shim layer node which
+witnesses all the effects removes the original effects from its cache since
+they are superseded by the new effects. Finally, the old effects are deleted
+from the backing store. This process ensures that clients either witness the
+old or the new effects, but not both; the summarization process appears to be
+atomic from the clients perspective.

### 2015-04-07 contract-language.tex
Added terry classification.

--- 2015-04-06 contract-language.tex
@@
 \begin{figure}
-\vspace{-7pt}
 \begin{smathpar}
+\renewcommand{\arraystretch}{1.2}
 \begin{array}{rclcl}
@@
 \begin{figure}
-\vspace{-7pt}
 \begin{smathpar}
+\renewcommand{\arraystretch}{1.2}
 \begin{array}{lclcl}
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 
@@
 
-  \vspace{-1em}
-  \begin{smathpar}
+  \vspace{-1mm}
+  \begin{cmathpar}
   \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
-  \end{smathpar}
+  \end{cmathpar}
 
@@
 
-  \vspace{-1em}
-  \begin{smathpar}
+  \vspace{-1mm}
+  \begin{cmathpar}
   \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
-  \end{smathpar}
+  \end{cmathpar}
 
@@
 
-  \vspace{-1em}
-  \begin{smathpar}
+  \vspace{-1mm}
+  \begin{cmathpar}
   \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
-  \end{smathpar}
+  \end{cmathpar}
 \end{itemize}
-\vspace{-1em}
 
@@
 \begin{figure}
-\vspace{-10pt}
 \begin{smathpar}
+\renewcommand{\arraystretch}{1.2}
 \begin{array}{cc}
@@
 \end{smathpar}
-\vspace{-3mm}
 \caption{Contract classification.}
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
@@
 
+\subsection{Generality of Contracts}
+
+It is important to note that our contract language provides a generic way to
+capture application-level consistency properties and is not tied to a
+particular store semantics. In particular, the same application-level contracts
+can easily be mapped to a different store with a varied consistency lattice. To
+illustrate this, let us consider the consistency lattice proposed by Terry et
+al.~\cite{Session} based on session guarantees. Terry et al. propose the
+following four incomparable session guarantees, whose semantics is captured in
+the contracts below:
+
+\vspace{-2mm}
+\begin{smathpar}
+\renewcommand{\arraystretch}{1.2}
+\begin{array}{rcl}
+\R{Read Your Writes (RYW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \Rightarrow \vis{a}{b} \\
+\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \soo{b}{c} \\
+ & & \quad \Rightarrow \vis{a}{c} \\
+\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b,c. ~\soo{a}{b} \wedge \vis{b}{c} \\
+ & & \quad \Rightarrow \vis{a}{c} \\
+\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c,d. ~\vis{a}{b} \wedge \vis{c}{d} \\
+ & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{d}
+\end{array}
+\end{smathpar}
+
+Here, the consistency level for an operation in this scheme is any combination
+of the above guarantees. It is easy to see that there are 24 (4!) combinations,
+which form a partially ordered consistency lattice with a height of 4.
+Classifying a contract under this scheme amounts to determining the weakest
+combination under which the contract will hold, and entails 24 comparisons at
+worst. Under this scheme, \cf{deposit} operations does not need any guarantees,
+\cf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
+\Rightarrow \cv_{gb}$), and \cf{withdraw} cannot be satisfied ($\Delta \vdash
+\R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).
+
 \subsection{Soundness of Contract Classification}
@@
 
-\vspace{-1em}
 \begin{smathpar}
+\renewcommand{\arraystretch}{1.2}
 \begin{array}{lclcl}
@@
 
-\vspace{-1em}
 \begin{smathpar}

### 2015-04-07 related.tex
Added terry classification.

@@
 
-Several conditions have been proposed to judge whether an operation on
-a replicated data object needs coordination or not. ~\cite{Calm}
-defines \emph{logical monotonicity} as a sufficient condition for
-coordination freedom, and proposes a consistency analysis that marks
-code regions performing non-monotonic reasoning (eg: aggregations,
-such as \cf{COUNT}) as potential coordination points.
-~\cite{IConfluence} and ~\cite{Sieve} define \emph{invariant
-confluence} and \emph{invariant safety}, respectively, as conditions
-for safely executing an operation without coordination. ~\cite{Sieve}
-also proposes a program analysis that conservatively marks certain
-operations as \emph{blue} (coordination not required), while marking
-the remaining as \emph{red} (coordination required). Unlike \name,
-these works focus on a coarse-grained classification of consistency as
-eventual or strong, and do not focus on finer-grained transaction
-isolation levels. However, the analyses they propose relieve
-programmers of the burden to tag operations with consistency levels.
-Indeed, we do consider automatic inference of consistency contracts
-from application-specific integrity constraints as the next step for
+Several conditions have been proposed to judge whether an operation on a
+replicated data object needs coordination or not. ~\cite{Calm} defines
+\emph{logical monotonicity} as a sufficient condition for coordination freedom,
+and proposes a consistency analysis that marks code regions performing
+non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as potential
+coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
+\emph{invariant confluence} and \emph{invariant safety}, respectively, as
+conditions for safely executing an operation without coordination.
+~\cite{Sieve} also proposes a program analysis that conservatively marks
+certain operations as \emph{blue} (coordination not required), while marking
+the remaining as \emph{red} (coordination required). Unlike \name, these works
+focus on a coarse-grained classification of consistency as eventual or strong,
+and do not focus on finer-grained transaction isolation levels. However, the
+analyses they propose relieve programmers of the burden to tag operations with
+consistency levels. Indeed, we do consider automatic inference of consistency
+contracts from application-specific integrity constraints as the next step for
 \name.

### 2015-04-07 contract-language.tex
cf to rcf. Better formatting for operation names.

@@
 classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
-simply the name of the operation (eg: \cf{withdraw}) that induced the effect.
+simply the name of the operation (eg: \rcf{withdraw}) that induced the effect.
 We admit disjunction in types to let an effect variable range over multiple
@@
 	be satisfied as long as the client can reach at least one replica. In the
-	bank account example, \cf{deposit} is an eventually consistent operation.
+	bank account example, \rcf{deposit} is an eventually consistent operation.
 	While an \ecds typically offer \emph{basic} eventual consistency with all
@@
 	different replicas $r_1$ and $r_2$, the second operation cannot be discharged
-	until the effect of $o_1$ is included in $r_2$. The \cf{getBalance} operation
+	until the effect of $o_1$ is included in $r_2$. The \rcf{getBalance} operation
 	requires causal consistency, as it requires the operations from the same
@@
   indefinitely under network partitions. An example is the total-order
-  contract on \cf{withdraw} operation. The corresponding store semantics is
+  contract on \rcf{withdraw} operation. The corresponding store semantics is
 	captured by the $\scc$ contract definition:
@@
 defines rules to judge the consistency class of a contact. For example, the
-scheme classifies the \cf{getBalance} contract ($\cv_{gb}$) from
+scheme classifies the \rcf{getBalance} contract ($\cv_{gb}$) from
 \S~\ref{sec:motivation} as a {\sf\small CausallyConsistent} contract, because
@@
 combination under which the contract will hold, and entails 24 comparisons at
-worst. Under this scheme, \cf{deposit} operations does not need any guarantees,
-\cf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
-\Rightarrow \cv_{gb}$), and \cf{withdraw} cannot be satisfied ($\Delta \vdash
+worst. Under this scheme, \rcf{deposit} operations does not need any guarantees,
+\rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
+\Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta \vdash
 \R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).

### 2015-04-07 evaluation.tex
Use correct class file for the camera ready version. Remove vertical space adjustments.
Formatting changes for camera-ready
Formatting edits
cf to rcf. Better formatting for operation names.

--- 2014-11-13 evaluation.tex
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item \textbf{LWW register}: A last-write-wins register that provides read
@@
 \setlength{\tabcolsep}{4pt}
-{\sffamily \small
+\renewcommand{\arraystretch}{1.2}
+{\sffamily
 \begin{center}
-\begin{tabular} {|l|r|r|r|r|r|r|r|r|}
+\caption{The distribution of classified contracts. \#T refers to the number of
+tables in the application. The columns 4-6 (7-9) represent operations
+(transactions) assigned to this consistency (isolation) level.}
+\label{tab:ctrts}
+\begin{tabular} {rrrrrrrrr}
 \hline
@@
 \end{center} }
-\caption{The distribution of classified contracts. \#T refers to the number of
-tables in the application. The columns 4-6 (7-9) represent operations
-(transactions) assigned to this consistency (isolation) level.}
-\label{tab:ctrts}
 \end{table}
@@
 \begin{figure*}
-  \vspace{-5pt}
   \centering
@@
 The lines marked EC and CC correspond to all operations (including
-\cf{withdraw}) being assigned EC and CC consistency levels. These levels
-compromise correctness as \cf{withdraw} has to be an SC operation. The SC line
+\rcf{withdraw}) being assigned EC and CC consistency levels. These levels
+compromise correctness as \rcf{withdraw} has to be an SC operation. The SC line
 corresponds to a configuration where all operations are strongly consistent;
@@
 due to the cost of geo-distributed coordination, whereas \name operations were
-only 2.2$\times$ slower, mainly due to the increased cost of \cf{withdraw}
-operations. Importantly, the latency of \cf{getBalance} and \cf{deposit}
+only 2.2$\times$ slower, mainly due to the increased cost of \rcf{withdraw}
+operations. Importantly, the latency of \rcf{getBalance} and \rcf{deposit}
 remained almost the same, illustrating the benefit of fine-grained contract

### 2015-04-07 motivation.tex
Use correct class file for the camera ready version. Remove vertical space adjustments.
Formatting changes for camera-ready
Formatting edits
cf to rcf. Better formatting for operation names.

--- 2014-11-13 motivation.tex
@@
 
-\subsection{RDT specification}
+\subsection{RDT Specification}
 
@@
 Let us assume that the bank account object provides three operations:
-\cf{deposit}, \cf{withdraw} and \cf{getBalance}, with the assumption that the
+\rcf{deposit}, \rcf{withdraw} and \rcf{getBalance}, with the assumption that the
 withdraw fails if the account has insufficient balance. Every operation in
@@
 \begin{figure}
-\vspace{-2.3mm}
 \begin{codehaskell}
@@
 \end{codehaskell}
+%\captionsetup{singlelinecheck=off}
 \caption{Definition of a bank account expressed in Quelea.}
@@
 every operation on the RDT is atomic, and thus amenable to sequential reasoning.
-Here, \cf{getBalance} is a read-only operation,
-\cf{deposit} always emits an effect, and \cf{withdraw} only emits an effect if
+Here, \rcf{getBalance} is a read-only operation,
+\rcf{deposit} always emits an effect, and \rcf{withdraw} only emits an effect if
 there is sufficient balance in the account. We have implemented a large corpus
@@
 
-Observe that the definition of \cf{getBalance} reduces over the \emph{entire
+Observe that the definition of \rcf{getBalance} reduces over the \emph{entire
 history} of updates to an account. If we are to realize an efficient
@@
 
-{\small
-\begin{verbatim}
+\begin{codehaskell}
  summarize hist =
-   [Deposit $ sel1 $ getBalance hist ()]
-\end{verbatim}}
+   [Deposit dollar sel1 dollar getBalance hist ()]
+\end{codehaskell}
 
@@
 account operations such that (1) the balance remains non-negative and (2) the
-\cf{getBalance} operation never incorrectly returns a negative balance.
+\rcf{getBalance} operation never incorrectly returns a negative balance.
 
 \begin{figure}[ht]
-\vspace{-2mm}
 \centering
@@
 that all operations in the figure are on the same bank account object with the
-initial balance being zero. Session 1 performs a \cf{deposit} of 100, followed
-by a \cf{withdraw} of 80 in the same session. The \cf{withdraw} operation
+initial balance being zero. Session 1 performs a \rcf{deposit} of 100, followed
+by a \rcf{withdraw} of 80 in the same session. The \rcf{withdraw} operation
 witnesses the deposit and succeeds\footnote{Although visibility and session
@@
 to relate operations, with the idea that the relations relate the effect
-emitted by those operations}. Subsequently, session 2 perform a \cf{withdraw}
+emitted by those operations}. Subsequently, session 2 perform a \rcf{withdraw}
 operation, but importantly, due to eventual consistency, only witnesses the
-\cf{deposit} from session 1, but not the subsequent withdraw. Hence, this
-\cf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
-constraint. A subsequent \cf{getBalance} operation, that happens to witness all
+\rcf{deposit} from session 1, but not the subsequent withdraw. Hence, this
+\rcf{withdraw} also \emph{incorrectly} succeeds, violating the integrity
+constraint. A subsequent \rcf{getBalance} operation, that happens to witness all
 the previous operations, would report a negative balance.
 
-It is easy to see that preventing concurrent \cf{withdraw} operations
-eliminates this anomaly. This can be done by insisting that \cf{withdraw} be
+It is easy to see that preventing concurrent \rcf{withdraw} operations
+eliminates this anomaly. This can be done by insisting that \rcf{withdraw} be
 executed as a strongly consistent operation. Despite this strengthening,
-the \cf{getBalance} operation may still incorrectly report a negative balance to
+the \rcf{getBalance} operation may still incorrectly report a negative balance to
 the user. Consider the execution shown in
 fig.~\ref{fig:negativeBalanceAnomaly}, which consists of three concurrent
-sessions performing a \cf{deposit}, a \cf{withdraw}, and a \cf{getBalance}
+sessions performing a \rcf{deposit}, a \rcf{withdraw}, and a \rcf{getBalance}
 operation, respectively, on the same bank account object. As the {\sf vis} edge
-indicates, operation \cf{withdraw(50)} in session 2 witnesses the effects of
-\cf{deposit(100)} from session 1, concludes that there is sufficient balance,
-and completes successfully. However, the \cf{getBalance} operation may only
-witness this successful withdraw, but not the causally preceding \cf{deposit},
+indicates, operation \rcf{withdraw(50)} in session 2 witnesses the effects of
+\rcf{deposit(100)} from session 1, concludes that there is sufficient balance,
+and completes successfully. However, the \rcf{getBalance} operation may only
+witness this successful withdraw, but not the causally preceding \rcf{deposit},
 and reports the balance of negative 50 to the user.
@@
 inconsistencies. Figure~\ref{fig:missingUpdateAnomaly} shows an execution where
-the \cf{getBalance} operation in a session does not witness the effects of an
-earlier \cf{withdraw} operation performed in the same session, possibly because
-it was served by a replica that has not yet merged the \cf{withdraw} effect.
-This anomaly leads the user to incorrectly conclude that the \cf{withdraw}
+the \rcf{getBalance} operation in a session does not witness the effects of an
+earlier \rcf{withdraw} operation performed in the same session, possibly because
+it was served by a replica that has not yet merged the \rcf{withdraw} effect.
+This anomaly leads the user to incorrectly conclude that the \rcf{withdraw}
 operation failed to go through.
@@
 aforementioned anomalies, finding the appropriate fixes is not readily
-apparent. Making \cf{getBalance} a strongly consistent operation is definitely
+apparent. Making \rcf{getBalance} a strongly consistent operation is definitely
 sufficient to avert anomalies, but is it really necessary? Given the cost of
@@
 In our running example, it is clear that in order to preserve the integrity
-constraint, the \cf{withdraw} operation must be strongly consistent.  That is,
-given two \cf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
+constraint, the \rcf{withdraw} operation must be strongly consistent.  That is,
+given two \rcf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
 vice-versa. We express this application-level consistency requirement as a
-contract ($\cv_w$) over \cf{withdraw}:
+contract ($\cv_w$) over \rcf{withdraw}:
 
-\vspace{-1em}
-\begin{smathpar}
+\begin{cmathpar}
 \begin{array}{l}
-\forall (a : \rcf{withdraw}).~\sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
+\forall (a : \rcf{withdraw}).\\
+\qquad \sameobj{a}{\cureff} \Rightarrow a = \cureff \vee \vis{a}{\cureff} \vee \vis{\cureff}{a}
 \end{array}
-\end{smathpar}
+\end{cmathpar}
 
-\noindent Here, $\cureff$ stands for the effect emitted by the \cf{withdraw} operation.
+\noindent Here, $\cureff$ stands for the effect emitted by the \rcf{withdraw} operation.
 The syntax $a:\rcf{withdraw}$ states that $a$ is an effect  emitted
-by a \cf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
+by a \rcf{withdraw} operation i.e., $\oper{a}{\rcf{withdraw}}$ holds.  The
 contract specifies that if the current operation emits an effect $\cureff$,
-then for any operation $a$ which was emitted by a \cf{withdraw} operation, it
+then for any operation $a$ which was emitted by a \rcf{withdraw} operation, it
 is the case that $a = \cureff$ or $a$ is visible to $\cureff$, or vice versa.
 Any execution on a bank account object that preserves the above contract for a
-\cf{withdraw} operation is said to be derived from a correct implementation of
-\cf{withdraw}.
+\rcf{withdraw} operation is said to be derived from a correct implementation of
+\rcf{withdraw}.
 
-\noindent For \cf{getBalance}, we construct the following contract ($\cv_{gb}$):
+To prevent \rcf{getBalance} from ever showing a negative balance, it is
+necessary to prevent the scenario depicted in
+fig.~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
+emitted by the \rcf{getBalance} operation. If the effect ($b$) of a withdraw
+operation is visible to $\cureff$, and the effect ($a$) of a deposit operation
+is visible to the effect ($b$) of the withdraw operation, then it must be the
+case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
+\rcf{getBalance} operation that precisely captures this application-level consistency
+requirement can be written thus:
 
-\vspace{-1em}
-\begin{smathpar}
+\begin{cmathpar}
+\begin{array}{l}
+\forall (a:\rcf{deposit}), (b:\rcf{withdraw}). \\
+\qquad (\vis{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff})
+\end{array}
+\end{cmathpar}
+
+\noindent To prevent the missing update anomaly described in
+fig.~\ref{fig:missingUpdateAnomaly}, it is necessary for a \rcf{getBalance}
+operation on a bank account to witness the effects of all previous \rcf{deposit}
+and \rcf{withdraw} operations performed on the same bank account in the same
+session. We can express an additional contract ($\cv_{gb}^2$) for
+\rcf{getBalance} that captures this consistency requirement:
+
+\begin{cmathpar}
+\begin{array}{l}
+\forall (c: \rcf{deposit} \vee \rcf{withdraw}). \\
+\qquad((\soZ \cap \sameobjZ) (c,\cureff) \Rightarrow \vis{c}{\cureff})
+\end{array}
+\end{cmathpar}
+
+\noindent Our contract language provides operators to compose relations. The
+syntax $(R_1 \cap R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The
+above contract ($\cv_{gb}^2$) says that if a \rcf{deposit} or a \rcf{withdraw}
+operation precedes a \rcf{getBalance} operation in session order, and is applied
+on the same object as the \rcf{getBalance} operation, then it must be the case
+that the \rcf{getBalance} operation witnesses the effects of the preceding
+operations.
+
+The final contract ($\cv_{gb}$) of the \rcf{getBalance} operation is
+merely a conjunction of the previous two versions ($\cv_{gb}^1$ and
+$\cv_{gb}^2$):
+
+\begin{cmathpar}
 \begin{array}{l}
@@
 \end{array}
-\end{smathpar}
+\end{cmathpar}
 
-\noindent The expression $c:\rcf{deposit} \vee \rcf{withdraw}$ states that $c$
-is an effect that was emitted either by a \cf{deposit} or a \cf{withdraw}
-operation. If a \cf{withdraw} $b$ is visible to \cf{getBalance} $\cureff$, then
-all \cf{deposit} operations $a$ visible to $b$ should also be visible to
-$\cureff$. Intuitively, this prevents negative balance anomalies. Our contract
-language provides operators to compose relations. The syntax $(R_1 \cap
-R_2)(a,b)$ is equivalent to $R_1(a,b) \wedge R_2(a,b)$. The last line of the
-above contract says that if a \cf{deposit} or a \cf{withdraw} operation
-precedes a \cf{getBalance} operation in session order, and is applied on the
-same object as the \cf{getBalance} operation, then it must be the case that the
-\cf{getBalance} operation witnesses the effects of the preceding operations.
+\noindent Intuitively, this prevents both the \rcf{getBalance} anomalies
+described in figs.~\ref{fig:negativeBalanceAnomaly} and
+\ref{fig:missingUpdateAnomaly} from ever occurring.
 
-Finally, since there are no restrictions on when or how a \cf{deposit}
+Finally, since there are no restrictions on when or how a \rcf{deposit}
 operation can execute, its contract is simply $\small \true$.
@@
 
-Notice that the contracts for \cf{withdraw} and \cf{getBalance} only express
+Notice that the contracts for \rcf{withdraw} and \rcf{getBalance} only express
 application-level consistency requirements, and make no reference to the

### 2015-04-07 transactions.tex
Formatting edits
cf to rcf. Better formatting for operation names.

@@
 
-\vspace{-1em}
-\begin{smathpar}
+\begin{cmathpar}
 \forall a,b,c.~\txn{a}{b,c} ~\wedge~ \sameobj{b}{c} ~\wedge~ \vis{b}{a}
 \Rightarrow \vis{c}{a}
-\end{smathpar}
+\end{cmathpar}
 
@@
 current ($c$) and savings ($s$). Each account provides operations
-\cf{withdraw}, \cf{deposit} and \cf{getBalance}, with the same contracts as
+\rcf{withdraw}, \rcf{deposit} and \rcf{getBalance}, with the same contracts as
 defined previously. We consider two transactions -- \cf{save(amt)}, which
@@
 of the object states. The \name code for the transactions is given below:
-\vspace{-1em}
 
@@
 the atomicity offered by the transaction, anomalies are still possible. For
-example, the two \cf{getBalance} operations in \cf{totalBalance} transactions
+example, the two \rcf{getBalance} operations in \cf{totalBalance} transactions
 might be served by different replicas with distinct set of committed \cf{save}
-transactions. If the first(second) \cf{getBalance} operation witness a
+transactions. If the first(second) \rcf{getBalance} operation witness a
 \cf{save} transaction that is not witnessed by the second(first)
-\cf{getBalance} operation, then the balance returned will be less(greater) than
+\rcf{getBalance} operation, then the balance returned will be less(greater) than
 the actual balance. It is not immediately apparent which weakest isolation
@@
 Instead, \name requires the programmer to simply state the consistency
-requirement as a contract. Since we would like both the \cf{getBalance}
+requirement as a contract. Since we would like both the \rcf{getBalance}
 operations to witness the same set of \cf{save} transactions, we define the
@@
 
-\vspace{-1em}
-\begin{smathpar}
+\begin{cmathpar}
 \begin{array}{l}
@@
 \end{array}
-\end{smathpar}
+\end{cmathpar}
 
@@
 \begin{figure}
-\vspace{-2.3mm}
 \centering
@@
 
-\vspace{-0.6em}
-\begin{smathpar}
+\begin{cmathpar}
 \begin{array}{l}

### 2015-04-08 contract-language.tex
Added lattice for session guarantees.

@@
 R				& \in & \mathtt{Relation}	& \coloneqq & \visZ \ALT \soZ
-        \ALT \sameobjZ \ALT R^+ \\
-				&			&	 &  \ALT & R \cup R \ALT R \cap R \\
+        \ALT \sameobjZ \ALT = \\
+				&			&	 &  \ALT & R \cup R \ALT R \cap R \ALT R^+ \\
 \end{array}
@@
 
-Here, the consistency level for an operation in this scheme is any combination
-of the above guarantees. It is easy to see that there are 24 (4!) combinations,
-which form a partially ordered consistency lattice with a height of 4.
-Classifying a contract under this scheme amounts to determining the weakest
-combination under which the contract will hold, and entails 24 comparisons at
-worst. Under this scheme, \rcf{deposit} operations does not need any guarantees,
-\rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
-\Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta \vdash
-\R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{Figures/lattice}
+\end{center}
+\caption{Lattice of consistency levels under session guarantees.}
+\label{fig:lattice}
+\end{figure}
+
+In this scheme, the consistency level of an operation is any combination of the
+above guarantees, which form a partially ordered consistency lattice show in
+Figure~\ref{fig:lattice}. Each element in this lattice corresponds to a
+store-consistency level, and is represented by its contract. An edge from a
+lower level element to an upper level element corresponds to logical
+implication between their corresponding contracts. Classifying a contract under
+this scheme is a directed search in the lattice, starting from the bottom, and
+determining the weakest consistency level under which the contract can be
+satisfied. Under this scheme, \rcf{deposit} operations does not need any
+guarantees, \rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge
+\R{WFR} \Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta
+\vdash \R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow
+\cv_{w}$).
 

### 2015-04-08 concl.tex

@@
 This paper presents \name, a shallow Haskell extension for declarative
-programming over eventually consistent data stores. The key idea underlying
-\name's design is the automatic classification of fine-grained consistency
-contracts on operations and distributed transactions with respect to the
-consistency and isolation levels offered by the store. Our contract language
-is carefully crafted from a decidable subset of first-order logic, enabling
-the use of automated verification tools to discharge the proof obligations
-associated with contract classification. We realize an instantiation of
-\name on top of off-the-shelf distributed store, Cassandra, and illustrate
-the benefit of fine-grained contract classification by implementing and
-evaluating several scalable applications.
+programming over \ecds. The key idea underlying \name's design is the automatic
+classification of fine-grained consistency contracts on operations and
+distributed transactions with respect to the consistency and isolation levels
+offered by the store. Our contract language is carefully crafted from a
+decidable subset of first-order logic, enabling the use of automated
+verification tools to discharge the proof obligations associated with contract
+classification. We realize an instantiation of \name on top of off-the-shelf
+distributed store, Cassandra, and illustrate the benefit of fine-grained
+contract classification by implementing and evaluating several scalable
+applications.

### 2015-04-08 contract-language.tex
Minor edit

--- 2015-04-07 contract-language.tex
@@
 \begin{figure}
+\vspace{-7pt}
 \begin{smathpar}
-\renewcommand{\arraystretch}{1.2}
 \begin{array}{rclcl}
@@
 \begin{figure}
+\vspace{-7pt}
 \begin{smathpar}
-\renewcommand{\arraystretch}{1.2}
 \begin{array}{lclcl}
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 
@@
 
-  \vspace{-1mm}
-  \begin{cmathpar}
+  \vspace{-1.5em}
+  \begin{smathpar}
   \ecc = \forall a,b. ~\hbo{a}{b} \wedge \vis{b}{\cureff} \Rightarrow \vis{a}{\cureff}
-  \end{cmathpar}
+  \end{smathpar}
 
@@
 
-  \vspace{-1mm}
-  \begin{cmathpar}
+  \vspace{-1.5em}
+  \begin{smathpar}
   \ccc = \forall a.~\hbo{a}{\cureff} \Rightarrow \vis{a}{\cureff}
-  \end{cmathpar}
+  \end{smathpar}
 
-\item \textbf{Strong consistency}: Strongly consistent operations may block
+\item \textbf{Strong Consistency}: Strongly consistent operations may block
   indefinitely under network partitions. An example is the total-order
@@
 
-  \vspace{-1mm}
-  \begin{cmathpar}
+  \vspace{-1.5em}
+  \begin{smathpar}
   \scc = \forall a.~\sameobj{a}{\cureff} \Rightarrow \vis{a}{\cureff} ~\vee~ \vis{\cureff}{a} ~\vee~ a = \cureff
-  \end{cmathpar}
+  \end{smathpar}
 \end{itemize}
+\vspace{-1em}
 
@@
 \begin{figure}
+\vspace{-10pt}
 \begin{smathpar}
-\renewcommand{\arraystretch}{1.2}
 \begin{array}{cc}
@@
 \end{smathpar}
+\vspace{-3mm}
 \caption{Contract classification.}
@@
 
-\begin{itemize}
+\begin{itemize}[noitemsep]
 \item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
@@
 
-\subsection{Generality of Contracts}
-
-It is important to note that our contract language provides a generic way to
-capture application-level consistency properties and is not tied to a
-particular store semantics. In particular, the same application-level contracts
-can easily be mapped to a different store with a varied consistency lattice. To
-illustrate this, let us consider the consistency lattice proposed by Terry et
-al.~\cite{Session} based on session guarantees. Terry et al. propose the
-following four incomparable session guarantees, whose semantics is captured in
-the contracts below:
-
-\vspace{-2mm}
-\begin{smathpar}
-\renewcommand{\arraystretch}{1.2}
-\begin{array}{rcl}
-\R{Read Your Writes (RYW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \Rightarrow \vis{a}{b} \\
-\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \soo{b}{c} \\
- & & \quad \Rightarrow \vis{a}{c} \\
-\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b,c. ~\soo{a}{b} \wedge \vis{b}{c} \\
- & & \quad \Rightarrow \vis{a}{c} \\
-\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c,d. ~\vis{a}{b} \wedge \vis{c}{d} \\
- & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{d}
-\end{array}
-\end{smathpar}
-
-Here, the consistency level for an operation in this scheme is any combination
-of the above guarantees. It is easy to see that there are 24 (4!) combinations,
-which form a partially ordered consistency lattice with a height of 4.
-Classifying a contract under this scheme amounts to determining the weakest
-combination under which the contract will hold, and entails 24 comparisons at
-worst. Under this scheme, \cf{deposit} operations does not need any guarantees,
-\cf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge \R{WFR}
-\Rightarrow \cv_{gb}$), and \cf{withdraw} cannot be satisfied ($\Delta \vdash
-\R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow \cv_{w}$).
-
 \subsection{Soundness of Contract Classification}
@@
 
+\vspace{-1em}
 \begin{smathpar}
-\renewcommand{\arraystretch}{1.2}
 \begin{array}{lclcl}
@@
 
+\vspace{-1.7em}
 \begin{smathpar}
@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
-$\ecc$ contracts in all well-formed executions, then the same store, extended
-with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
-all well-formed \name contracts. The proof of the theorem is given in the
-supplementary material\footnote{The supplementary material also provides the
-concrete reduction rules for enforcing the consistency classes.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H3$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H4$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$, $H4$, and $H3$.
+  \hfill \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2015-04-08 related.tex

--- 2014-11-13 related.tex
@@
 operations(transactions). \name relieves the developer of this burden, and
-instead expects contracts expressing declarative visibility requirements.
+instead expects programmers to declare application-specific consistency
+requirements as first-order formulas. In this regard, \name is similar to Indigo
+\cite{indigo}, which requires programmers to declare application-specific
+invariants that are expected to be met, along with pre- and post-conditions on
+operations as formulas in first-order logic. Indigo then performs a static
+analysis on formulas to identify $I$-offender sets - sets of operations,
+which, when performed concurrently, result in violation of one or more of the
+stated invariants. For each offending set of operation, if programmer chooses 
+invariant-violation avoidance over violation repair, the system employs various
+techniques, such as escrow reservation, to ensure that the offending set of
+operations are never run concurrently.
+
+While serializing operations is sufficient to ensure that the application state
+never violates invariants (e.g: \cf{balance $\ge$ 0}), there exist a range of
+desirable application-level properties, which cannot be captured in Indigo's
+specification language, or for which serialization is an overkill. For instance,
+it is neither efficient nor straightforward in Indigo to declare and enforce a
+condition to prevent the missing update anomaly
+(Figure~\ref{fig:missingUpdateAnomaly}) in the bank account application. The
+only way is to encode specifications of all bank account operations such that
+every pair of operations are identified as $I$-offending, for some invariant
+($I$), resulting in them being serialized. In contrast, \name's contract
+language allows for precise encoding of the desired property, and the contract
+classification scheme ensures that the it can be enforced without the need to
+serialize operations. Furthermore, Indigo also does not distinguish between
+various transaction isolation guarantees, which \name does.
 

### 2015-04-08 related.tex
A correction to Indigo comparision

@@
 (Figure~\ref{fig:missingUpdateAnomaly}) in the bank account application. The
-only way is to encode specifications of all bank account operations such that
-every pair of operations are identified as $I$-offending, for some invariant
-($I$), resulting in them being serialized. In contrast, \name's contract
-language allows for precise encoding of the desired property, and the contract
-classification scheme ensures that the it can be enforced without the need to
-serialize operations. Furthermore, Indigo also does not distinguish between
-various transaction isolation guarantees, which \name does.
+only way is to encode specifications of bank account operations such that pairs
+of \cf{getBalance} and \cf{withdraw}, and \cf{getBalance} and \cf{deposit}
+operations are identified as $I$-offending, resulting in them being serialized.
+In contrast, \name's contract language allows for precise encoding of the
+desired property, and the contract classification scheme ensures that the it can
+be enforced without the need to serialize operations.  Furthermore, Indigo also
+does not distinguish between various transaction isolation guarantees, which
+\name does.
 

### 2015-04-08 related.tex
Simplified comparision with Indigo

--- 2015-04-08 related.tex
@@
 operations(transactions). \name relieves the developer of this burden, and
-instead expects programmers to declare application-specific consistency
-requirements as first-order formulas. In this regard, \name is similar to Indigo
-\cite{indigo}, which requires programmers to declare application-specific
-invariants that are expected to be met, along with pre- and post-conditions on
-operations as formulas in first-order logic. Indigo then performs a static
-analysis on formulas to identify $I$-offender sets - sets of operations,
-which, when performed concurrently, result in violation of one or more of the
-stated invariants. For each offending set of operation, if programmer chooses
-invariant-violation avoidance over violation repair, the system employs various
-techniques, such as escrow reservation, to ensure that the offending set of
-operations are never run concurrently.
-
-While serializing operations is sufficient to ensure that the application state
-never violates invariants (e.g: \cf{balance $\ge$ 0}), there exist a range of
-desirable application-level properties, which cannot be captured in Indigo's
-specification language, or for which serialization is an overkill. For instance,
-it is neither efficient nor straightforward in Indigo to declare and enforce a
-condition to prevent the missing update anomaly
-(Figure~\ref{fig:missingUpdateAnomaly}) in the bank account application. The
-only way is to encode specifications of all bank account operations such that
-every pair of operations are identified as $I$-offending, for some invariant
-($I$), resulting in them being serialized. In contrast, \name's contract
-language allows for precise encoding of the desired property, and the contract
-classification scheme ensures that the it can be enforced without the need to
-serialize operations. Furthermore, Indigo also does not distinguish between
-various transaction isolation guarantees, which \name does.
+instead expects  contracts expressing declarative visibility requirements.
 
@@
 conditions for safely executing an operation without coordination.
-~\cite{Sieve} also proposes a program analysis that conservatively marks
-certain operations as \emph{blue} (coordination not required), while marking
-the remaining as \emph{red} (coordination required). Unlike \name, these works
-focus on a coarse-grained classification of consistency as eventual or strong,
-and do not focus on finer-grained transaction isolation levels. However, the
-analyses they propose relieve programmers of the burden to tag operations with
-consistency levels. Indeed, we do consider automatic inference of consistency
-contracts from application-specific integrity constraints as the next step for
-\name.
+~\cite{Sieve} also proposes a program analysis that conservatively marks certain
+operations as \emph{blue} (coordination not required), while marking the
+remaining as \emph{red} (coordination required). ~\cite{indigo} requires
+programmers to declare application semantics, and the desired
+application-specific invariants as formulas in first-order logic. It performs
+static analysis on these formulas to determine $I$-offender sets - sets of
+operations, which, when performed concurrently, result in violation of one or
+more of the stated invariants. For each offending set of operation, if
+programmer chooses invariant-violation avoidance over violation repair, the
+system employs various techniques, such as escrow reservation, to ensure that
+the offending set of operations are effectively serialized. Unlike \name, these
+works focus on a coarse-grained classification of consistency as eventual or
+strong, and do not focus on finer-grained transaction isolation levels. However,
+the analyses that some of them propose relieve programmers of the burden to tag
+operations with consistency levels. Indeed, we do consider automatic inference
+of consistency contracts from application-specific integrity constraints as the
+next step for \name.

### 2015-04-08 related.tex
Added comparision to Burckhardt's ECOOP15

@@
 
-Our contract language is inspired by the axiomatic description of RDT semantics
-proposed by~\cite{Burckhardt2014}. While they use axioms for formal
-verification of correctness of an RDT implementation, we utilize them as a
-means for the user to express the desired consistency guarantees in the
-application. Similar to their work, our contract language does not incorporate
-real (i.e., wall-clock) time. Hence, it cannot describe store semantics such as
-recency or bounded-staleness guarantees offered by certain
-stores~\cite{Pileus}.
+Our contract language and system model is inspired by the axiomatic description
+of RDT semantics proposed by~\cite{Burckhardt2014}. While they use axioms for
+formal verification of correctness of an RDT implementation, we utilize them as
+a means for the user to express the desired consistency guarantees in the
+application. Operational semantics of \name (described in the accompanying tech
+report~\cite{techrep}) describe an operational manifestation of our system
+model.  ~\cite{Burckhardt2015} also presents an operational model of a
+replicated data store that is based on the abstract system model presented in
+~\cite{Burckhardt2014}. Their claims about the expressivity and practicality of
+the system model vindicate ours.  However, both differ in the way they aim to
+empower the application programmer. In~\cite{Burckhardt2015}, they expose their
+operational model to the application programmer, along with primitives, such as
+\cf{push}, \cf{pull} and \cf{fetch}, that make it easy to \emph{implement} the
+required consistency and isolation guarantees. In contrast, \name completely
+abstracts its operational model, and instead relies on contract classification
+to \emph{choose} appropriate consistency and isolation levels.  Similar
+to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our system model does not
+incorporate real (i.e., wall-clock) time. Hence, our contract language cannot
+describe store semantics such as recency or bounded-staleness guarantees offered
+by certain stores~\cite{Pileus}.
 

### 2015-04-08 related.tex

@@
 report~\cite{techrep}) describe an operational manifestation of our system
-model.  ~\cite{Burckhardt2015} also presents an operational model of a
-replicated data store that is based on the abstract system model presented in
+model. \cite{Burckhardt2015} also presents an operational model of a replicated
+data store that is based on the abstract system model presented in
 ~\cite{Burckhardt2014}. Their claims about the expressivity and practicality of
 the system model vindicate ours. However, both differ in the way they aim to
-empower the application programmer. In~\cite{Burckhardt2015}, they expose their
+empower the application programmer. \cite{Burckhardt2015} exposes the
 operational model to the application programmer, along with primitives, such as
@@
 incorporate real (i.e., wall-clock) time. Hence, our contract language cannot
-describe store semantics such as recency or bounded-staleness guarantees offered
-by certain stores~\cite{Pileus}.
+describe store semantics such as recency or bounded-staleness guarantees
+offered by certain stores~\cite{Pileus}.
 

### 2015-04-08 PLDI15_Response/response_notes.txt

==============================================================================
Common Concerns
==============================================================================

Most important : Complexity of Contracts
—————————————————————————————————————————

(Reviewer 5; main reason for rejection) - It isn't clear that the proposed
solution is any less complicated than reasoning directly about consistency; the
evaluation does not actually evaluate this hypothesis, which is the main
motivation for the paper.

(Reviewer 5) The contract language is complex and provides an even richer set
of choices to the user than just choosing an appropriate consistency level---it
isn't clear to me that this is an improvement for the user (e.g., the example
for withdraw and getBalance on page 4 seems to require a deep understanding of
the requirements). If a realistic scenario is just three totally-ordered
consistency levels (as per this paper's evaluation), then it seems simpler to
choose the level directly rather than try to specify a contract. This problem
might be mitigated if it is expected than an application might be run on
multiple different data stores with very different sets of consistency levels;
is this the case here?

(Reviewer 2) "As the author's acknowledge, the contracts language available to
the programmer makes for verbose and low-level annotation requirements.
Moreover, there is scant evidence that writing these annotations is any easier
to get right than picking a consistency level in the first place."

(Reviewer 3) “One of my concern of the paper is the expression of contracts is
still complex. A example is the getBalance contract Page 4. While the current
contract language is expressive, it is also a little complex if the purpose is
just to specify the consistency level of a certain data structure."

(Reviewer 4) “One of the problems with new languages or new language extensions
is that they usually do not get adopted by programmers. A programmer would take
the time to learn a new language only if he feels the time spent on learning
will pay off eventually. I am not sure programmers will be willing to adopt
this extension."

** Start Response **

— Programmers write contracts for the given abstract system model, which admits
all the inconsistencies associated with eventual consistency. However, the
programmer only needs to reason about inconsistencies in this abstract model
not the intricacies of the actual store model, which are often left
unspecified. The advantage of Quelea contract is that programmer-written
contracts are automatically mapped onto the store consistency level.

** End Response **

Delta
—————

(Reviewer 3) Another issue of this paper is the delta between this paper and
the POPL paper[ Burckhardt et al. 2014] on RDT. In the POPL paper, the contract
was used as languages for verification tool while in this paper it is used to
generate distributed protocol code. There are certainly differences between
these two usage model, but the prior work does make this paper a little
incremental.

(Reviewer 4) The increment of this work over [Bailis 2013b] is not very large.
The previous work (and I don’t know if it originates from the same authors)
already proposed separation of concerns and implemented an initial scheme. The
current paper seems to extend it with a language and the use of the theorem
prover. As a non-expert, this seems to me like a nice step but not a very large
one.

Execution Time
——————————————

(Reviewer 4) Can you say how much time the execution of the theorem prover
takes? If the consistency constraints become more complex will the theorem
prover still be able to determine the appropriate levels of support
efficiently?

(Reviewer 1) I believe the compile time can still be a show stopper. Given
these applications can change a lot, isn't the long compilation time due to the
SMT solver an issue?

===============================================================================
Reviewer specific concerns
===============================================================================

Reviewer 1
——————————

I think also comparing to a dynamic strategy for enforcing contacts (mentioned
on page 5 as an alternative), would make the case stronger. Such strategy can
take advantage from transaction buffer in each replica. Can you comment?

-I believe the compile time can still be a show stopper. Given these
applications can change a lot, isn't the long compilation time due to the SMT
solver an issue?

-In your experiments did you consider the shim layer overhead? Did you use the
same number of nodes in all tests? I mean was it the case that "shim layer
nodes" + "other nodes" in Q runs was always equal to "all the nodes" in non-Q
runs? Also, how does the shim layer grow (number of nodes) as the whole system
grows?

-Are all the benchmarks written from scratch or did you convert some existing
one to QUELEA? My impression is the former is correct.

Reviewer 2
——————————

The paper doesn't provide much discussion of the limitations of the
approach---either positive or negative. For example, are there notions of
consistency that fundamentally would not fit into this model?  The end of the
paper mentions that it doesn't handle notions that depend on wall-clock time as
distinct from session order. It would be interesting to know whether notions
such as bounded staleness could be readily accomodated or not.

As the author's acknowledge, the contracts language available to the programmer
makes for verbose and low-level annotation requirements. Moreover, there is
scant evidence that writing these annotations is any easier to get right than
picking a consistency level in the first place.

The system model requires the clients to perform all the actual computation on
the data-store, including merging and summarizing. In a production setting this
would likely prove infeasible, as it relies on strong assumptions of the
clients' relationships, including identical running code and complete trust.
The possibility that a crashed operation would be durably added to the log,
only to break when a different client attempts to reduce over the log much
later, is not friendly.

** Start Response **

The RDT specification language that we have described does to some extent gaurd
against malicious/incorrect code. In particular, the operations on RDT are
described as reductions over the history, and in particular is side-efffect
free (which is ensured by the type of the operation -- section 3.1). While it
is possible that the reduction can end up not terminating, it cannot do any
more harm.

Crashed operation durably added to log?? By definition, an operation is added
to log only if it is successful; if the operation procudes a result, then it
has succeeded. Hence a crashed operation will not be added to log.

Summarization operation checks whether the summarized history is smaller than
the input history. Otherwise, the summarization operation is considered
incorrect and rejected.

** End Response **

Minor comment: Footnote 3 discusses how you "encode" transitive closure in
first-order logic. I didn't understand how this could ever be useful. Given the
constraints

	R(a,b) => R'(a,b)

	R'(a,b) /\ R'(b,c) => R'(a,c)

it seems that R' could be the total relation. How would this ever be useful?

** Start Response **

R+ could be a total relation, but the R+ is still useful because it is not
required to be a total relation. For example, consider causal
consistency (CC) contract which is defined as:

 hbo \subseteq vis

where hbo = (soo U vis)+. Because of our lenient encoding of “+", hbo can be
any superset of  transitive closure of (soo U vis). Nevertheless, all we can
assume about hbo is that it contains the  transitive closure of (soo U vis);
there is no proof of membership for other elements in hbo.  Consequently, by
asserting hbo \subseteq vis, the strongest guarantee that causal consistency 
contract is offering is that the transitive closure of (soo U vis) is contained
in vis. An application  contract (\psi) which requires more than the transitive
closure of (soo U vis) to be contained in vis  cannot be satisfied under CC
because the following implication is simply not valid even under our lenient
definition of “+“ :

 \Delta |- (hbo \subseteq vis) => \psi

\Delta, as usual, denotes the conjunction of basic axioms about primitive
relations.

** End Response **

Reviewer 3
——————————

I am wondering if there are more intuitive and succinct ways to provide a
higher level abstraction to define these contracts? One example is the "atomic"
directives in Java. Can we have something similar to that to allow easier
declaration of consistency levels?

Reviewer 5
——————————

The evaluation shows that using the weakest levels possible while still
preserving correctness does give a performance improvement, but this isn't a
surprising result. I would be more interested in seeing whether using QUELEA
leads to fewer mis-classifications than choosing levels directly, which speaks
to the main motivation of this work.

This paper finds the correct consistency level by succesively testing the
levels in order of strength in order to find the weakest one that satisfies the
contract. Is it expected that most stores would have only a few levels in total
order? If there are many levels in a partial order, is finding the correct
level still tractable? If the levels are partially ordered, how is the "best"
level defined?

** Start Response **

Most stores that are available only tend to expose a few levels, which can
indeed be partially ordered. An example is classification from [[Doug Terry’s
Session guarantees for weakly cons]stent replicated data]] which introduced 4
(incomparable) session guarantees, and the consistency level for an operation
is any combination of the 4 session guarantees. The classification here, at
worst, requires 4! = 24 comparisons (SMT validity tests). However, each
comparison is very cheap; on average, in our evaluation, each comparison took
11.5 ms. Hence, the solution remains scalable.

“Best Level?” Since the classification is parameterised, when different
incomparable consistency levels fit the bill, the RDT implementor can either
choose one of the levels, or perhaps choose a stronger level that encapsulates
both. This decision is left to the RDT implementor.

** End Response **

### 2015-04-08 contract-language.tex
Edits to bib

--- 2015-04-08 contract-language.tex
@@
 Figure~\ref{fig:lattice}. Each element in this lattice corresponds to a
-store-consistency level, and is represented by its contract. An edge from a
-lower level element to an upper level element corresponds to logical
-implication between their corresponding contracts. Classifying a contract under
+store-consistency level, and is represented by its contract. An edge from an
+upper level element to a lower level element corresponds to weaker than
+relation between their corresponding contracts. Classifying a contract under
 this scheme is a directed search in the lattice, starting from the bottom, and

### 2015-04-08 store-opsem.tex
Fixed typos.
Applied changes from paper to techrep

@@
 underlying \name.
+
+% Semantics
+% ---------
+\input{store-opsem-rules}
 
@@
 
-The state of an operational execution ($\E$) is a tuple $\Exec$ where
+The state of an operational execution ($\E$) is a tuple \Exec, where
 $\EffSoup$ is a set of effects, and $\visZ$, $\soZ$, $\sameobjZ$
@@
 visiblity.
-
-% Semantics
-% ---------
-\input{store-opsem-rules}
 
@@
   \end{smathpar}
-  Where, $\hboZ = (\Rvis \cup \Rso)^{+}$
+  Where, $\hboZ = (\Rvis \cup \sooZ)^{+}$
 \end{definition}
@@
   \ref{thm:local-state-is-cc}, and \ref{thm:weak-soundness}.
-  \qed
+  \hfill \qed
 \end{proof}

### 2015-04-08 store-opsem.tex
Fixed typos

@@
 An operation $op$ performed at replica $r$ can only witness the set of
-effects ($\Theta(r)$) visible at $r$. To avoid too many paranthesis in
+effects ($\Theta(r)$) visible at $r$. To avoid too many parentheses in
 our formal presentation, we represent operation $op$, whose contract
@@
 
-Basic gurantee provided by the store is causal visibility, which is
+Basic guarantee provided by the store is causal visibility, which is
 captured by the rule $\rulelabel{EffVis}$ as a condition for an effect
@@
 lose any generality by assuming that the store provides causal
-visiblity.
+visibility.
 

### 2015-04-10 contract-language.tex
edits

--- 2015-04-08 contract-language.tex
@@
 \emph{current operation} - the operation for which a contract is being written.
-Notice that $\cureff$ occurs free in the contract. We will fix its scope when
-classifying the contracts (\S~\ref{sec:classify}). The type of an effect is
+Notice that $\cureff$ occurs free in the contract. We fix its scope when
+classifying contracts (\S~\ref{sec:classify}). The type of an effect is
 simply the name of the operation (eg: \rcf{withdraw}) that induced the effect.
@@
 $R^{+}$ is any relation $R'$ such that forall $x$, $y$, and $z$, a) $R(x,y)
-\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$} of
+\Rightarrow R'(x,y)$, and b) $R'(x,y) \conj R'(y,z) \Rightarrow R'(x,z)$.} of
 primitive relations.  Commonly used derived relations are the \emph{same object
@@
 	until the effect of $o_1$ is included in $r_2$. The \rcf{getBalance} operation
-	requires causal consistency, as it requires the operations from the same
+	requires causal consistency, as it requires operations from the same
 	session to be visible, which cannot be guaranteed under eventual consistency.
@@
 The quantifier in the sequent binds $\cureff$ that occurs free in $\cv_{st}$
-and $\cv_{op}$. Context ($\Delta$) of the sequent is a conjunction of
+and $\cv_{op}$. The context ($\Delta$) of the sequent is a conjunction of
 assumptions about the nature of primitive relations. A \emph{well-formed}
@@
 \begin{itemize}
-\item The happens-before relation is acyclic: $\small \forall a.~\neg\hb{a}{a}$.
+\item The happens-before relation is acyclic: $\small \forall a.~\neg\hbo{a}{a}$.
 \item Visibility only relates actions on the same object:
@@
 Figure~\ref{fig:lattice}. Each element in this lattice corresponds to a
-store-consistency level, and is represented by its contract. An edge from a
-lower level element to an upper level element corresponds to logical
-implication between their corresponding contracts. Classifying a contract under
+store-consistency level, and is represented by its contract. An edge from an
+upper level element to a lower level element corresponds to a weaker-than
+relation between the corresponding contracts. Classifying a contract under
 this scheme is a directed search in the lattice, starting from the bottom, and
 determining the weakest consistency level under which the contract can be
-satisfied. Under this scheme, \rcf{deposit} operations does not need any
+satisfied. Under this scheme, \rcf{deposit} operations do not need any
 guarantees, \rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge
@@
 \S~\ref{sec:sysmod}:
-
-\begin{smathpar}
+\begin{mathpar}
 \renewcommand{\arraystretch}{1.2}
@@
 \end{array}
-\end{smathpar}
+\end{mathpar}
 
@@
 the contract classification scheme). We assume a reduction relation of form:
-
-\begin{smathpar}
+\begin{mathpar}
   \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
     {\E',\sigma \pll \Sigma}
-\end{smathpar}
+\end{mathpar}
 

### 2015-04-10 evaluation.tex
edits

--- 2015-04-07 evaluation.tex
@@
 modify and access related fields across tables. In \name, the contract
-classification process is completely performed at compile time and has no
-overheads at runtime. The proof obligations associated with contract
+classification process is completely performed at compile time and incurs no
+runtime overhead. The proof obligations associated with contract
 classification is discharged through the Z3 SMT Solver. Across our benchmarks,
@@
 within the same data center as the store, and run transactions. We deploy each
-cluster and client node on an \cf{c3.4xlarge} Amazon EC2 instance. We call this
+cluster and client node on a \cf{c3.4xlarge} Amazon EC2 instance. We call this
 a \cf{1DC} configuration. For our geo-distributed experiments (\cf{2DC}), we

### 2015-04-10 implementation.tex
Fixing figures.
edits

@@
 \begin{center}
-\includegraphics[width=0.7\columnwidth]{Figures/AtomicityImpl}
+\includegraphics[width=\columnwidth]{Figures/AtomicityImpl}
 \end{center}
@@
 
-\name implements atomic visibility by exploiting shim layer causality guarantee
+\name implements atomic visibility by exploiting shim layer causality guarantees
 -- an effect is included only if all the effects if depends on are also
@@
 every transaction in \name, we instantiate a special transaction marker
-effect $m$. But importantly, do not insert into the backing store. $m$ is
+effect $m$ that is importantly not inserted into the backing store. Marker $m$ is
 included as a dependence to every effect generated in the transaction. In the
-figure, the graph on the left shows the state of the store in the middle of a
+figure, the top graph on the left shows the state of the store in the middle of a
 transaction. Each circle represents an effect. The dotted circle indicates that
@@
 store, marking all the effects from the transactions as a dependence for $m$.
-Now any replica which includes one of the effects from the transaction must
+Now, any replica which includes one of the effects from the transaction must
 include $m$, and transitively must include every effect from the transaction.
@@
 The above scheme prevents a transaction from witnessing its own effects. This
-might conflict with the causality requirement on the operations. Hence,
+might conflict with causality requirements on the operations. Hence,
 transactions piggy-back the previous effects from the same transaction for each

### 2015-04-10 intro.tex
edits

@@
 underlying large-scale distributed systems with respect to non-uniform latency,
-availability and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
+availability, and network partitions~\cite{Brewer2000,Gilbert2002}. Indeed,
 modern web services, which aim to provide an "always-on" experience,
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item We introduce \name, a shallow extension of Haskell that supports the
-	description and validation of replicated data types found on \ecds. Contracts
-	are used to specify fine-grained application-level consistency properties,
-	and are statically analyzed to assign the most efficient and sound store
-	consistency level to the corresponding operation.
+  description and validation of replicated data types found in an
+  \ecds. Contracts are used to specify fine-grained application-level
+  consistency properties, and are statically analyzed to assign the most
+  efficient and sound store consistency level to the corresponding
+  operation.
 \item \name supports coordination-free transactions over arbitrary datatypes.
@@
 	required conditions on consistency are met.
-\item An implementation of \name as a transparent shim layer over
+\item We describe an implementation of \name as a transparent shim layer over
 	Cassandra~\cite{Cassandra}, a well-known general-purpose data store.
@@
 overcome these issues in \S~\ref{sec:motivation}. \S~\ref{sec:lang} provides
-more details on the contract language, and its mapping to the store consistency
+more details on the contract language, and its mapping to store consistency
 levels, along with meta-theory for certifying the correctness of the mapping.

### 2015-04-10 motivation.tex
edits

@@
 
-A key novelty in \name is that it allows the addition of new RDTs to the store,
-which obviates the need for coercing the application logic to utilize the store
-provided data types. In addition, \name treats the convergence semantics (i.e.,
-\emph{how} conflicting updates are resolved) of the data type separately from
-its consistency properties (i.e., \emph{when} updates become visible). This
-separation of concerns permits \emph{operational} reasoning for conflict
-resolution, and \emph{declarative} reasoning for consistency. The combination
-of these techniques enhances the programmability of the store.
+A key novelty in \name is that it allows the addition of new RDTs to the
+store, which obviates the need for coercing application logic to utilize
+store-provided data types. In addition, \name treats the convergence
+semantics (i.e., \emph{how} conflicting updates are resolved) of the data
+type separately from its consistency properties (i.e., \emph{when} updates
+become visible). This separation of concerns permits \emph{operational}
+reasoning for conflict resolution, and \emph{declarative} reasoning for
+consistency. The combination of these techniques enhances the
+programmability of the store.
 
@@
 getBalance hist _ =
-  let res = sum [x | Deposit x <- hist]
-						- sum [x | Withdraw x <- hist]
+  let res = sum [x | Deposit x <- hist] -
+            sum [x | Withdraw x <- hist]
 	in (res, Nothing)
@@
 specific to each RDT, programmers can provide a summarization
-function - (\cf{summarize}) of type \cf{[e] -> [e]} -  as a part of
+function - of type \cf{[e] -> [e]} -  as a part of
 the RDT specification. The summarization function for the bank account is:
@@
 to relate operations, with the idea that the relations relate the effect
-emitted by those operations}. Subsequently, session 2 perform a \rcf{withdraw}
+emitted by those operations.}. Subsequently, session 2 perform a \rcf{withdraw}
 operation, but importantly, due to eventual consistency, only witnesses the
@@
 aforementioned anomalies, finding the appropriate fixes is not readily
-apparent. Making \rcf{getBalance} a strongly consistent operation is definitely
-sufficient to avert anomalies, but is it really necessary? Given the cost of
-enforcing strong consistency~\cite{DynamoDB, Pileus}, it is preferable to avoid
-it unless there are no viable alternatives. Exploring the space of these
-alternatives requires understanding the subtle differences in semantics of
-various kinds of weak consistency alternatives.
+apparent. Making \rcf{getBalance} a strongly consistent operation is
+definitely sufficient to avert anomalies, but is it really necessary? Given
+the cost of enforcing strong consistency~\cite{DynamoDB, Pileus}, it is
+preferable to avoid imposing such stringent conditions unless there are no
+viable alternatives. Exploring the space of these alternatives requires
+understanding the subtle differences in semantics of various kinds of weak
+consistency alternatives.
 
@@
 
-In our running example, it is clear that in order to preserve the integrity
-constraint, the \rcf{withdraw} operation must be strongly consistent.  That is,
-given two \rcf{withdraw} operations $a$ and $b$, either $a$ is visible to $b$ or
-vice-versa. We express this application-level consistency requirement as a
-contract ($\cv_w$) over \rcf{withdraw}:
+In our running example, it is clear that in order to preserve the critical
+integrity constraint, the \rcf{withdraw} operation must be strongly
+consistent.  That is, given two \rcf{withdraw} operations $a$ and $b$,
+either $a$ is visible to $b$ or vice-versa. We express this
+application-level consistency requirement as a contract ($\cv_w$) over
+\rcf{withdraw}:
 
@@
 necessary to prevent the scenario depicted in
-fig.~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
+Figure~\ref{fig:negativeBalanceAnomaly}. Let $\cureff$ stand for the effect
 emitted by the \rcf{getBalance} operation. If the effect ($b$) of a withdraw
@@
 case that $a$ is also visible to $\cureff$. A contract ($\cv_{gb}^1$) for
-\rcf{getBalance} operation that precisely captures this application-level consistency
+\rcf{getBalance}  that precisely captures this application-level consistency
 requirement can be written thus:
@@
 \noindent To prevent the missing update anomaly described in
-fig.~\ref{fig:missingUpdateAnomaly}, it is necessary for a \rcf{getBalance}
+Figure~\ref{fig:missingUpdateAnomaly}, it is necessary for a \rcf{getBalance}
 operation on a bank account to witness the effects of all previous \rcf{deposit}
@@
 
-\noindent Intuitively, this prevents both the \rcf{getBalance} anomalies
-described in figs.~\ref{fig:negativeBalanceAnomaly} and
-\ref{fig:missingUpdateAnomaly} from ever occurring.
+\noindent Intuitively, this specification prohibits both the
+\rcf{getBalance} anomalies described in
+Figures.~\ref{fig:negativeBalanceAnomaly} and \ref{fig:missingUpdateAnomaly}
+from  occurring.
 

### 2015-04-10 related.tex
edits

@@
 implement custom solutions for durability and fault-tolerance. \name realizes
-RDTs stronger consistency models on top of off-the-shelf eventually consistent
+stronger consistency models for RDTs on top of off-the-shelf eventually consistent
 distributed stores. In this respect, \name is similar to~\cite{BoltOn} where
-causal consistency is achieved through a shim layer on Cassandra, which
+causal consistency is achieved through a shim layer layered on top of Cassandra, which
 explicitly tracks and enforces dependencies between updates.
@@
 operations(transactions). \name relieves the developer of this burden, and
-instead expects  contracts expressing declarative visibility requirements.
+instead only requires contracts to express declarative visibility requirements.
 
-Our contract language and system model is inspired by the axiomatic description
-of RDT semantics proposed by~\cite{Burckhardt2014}. While they use axioms for
-formal verification of correctness of an RDT implementation, we utilize them as
-a means for the user to express the desired consistency guarantees in the
-application. Operational semantics of \name (described in the accompanying tech
-report~\cite{techrep}) describe an operational manifestation of our system
-model. \cite{Burckhardt2015} also presents an operational model of a replicated
-data store that is based on the abstract system model presented in
-~\cite{Burckhardt2014}. Their claims about the expressivity and practicality of
-the system model vindicate ours. However, both differ in the way they aim to
-empower the application programmer. \cite{Burckhardt2015} exposes the
-operational model to the application programmer, along with primitives, such as
-\cf{push}, \cf{pull} and \cf{fetch}, that make it easy to \emph{implement} the
-required consistency and isolation guarantees. In contrast, \name completely
-abstracts its operational model, and instead relies on contract classification
-to \emph{choose} appropriate consistency and isolation levels. Similar
-to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our system model does not
-incorporate real (i.e., wall-clock) time. Hence, our contract language cannot
-describe store semantics such as recency or bounded-staleness guarantees
-offered by certain stores~\cite{Pileus}.
+Our contract language and system model is inspired by the axiomatic
+description of RDT semantics proposed in~\cite{Burckhardt2014}. While
+~\cite{Burckhardt2014} uses axioms for formal verification of
+an RDT implementation, we utilize them as a means for the user to express
+desired consistency guarantees in the application. The operational semantics
+of \name (described in the accompanying tech report~\cite{techrep}) describes
+an operational manifestation of our system model. \cite{Burckhardt2015} also
+presents an operational model of a replicated data store that is based on
+the abstract system model presented in ~\cite{Burckhardt2014}. Their claims
+about the expressivity and practicality of the system model vindicate
+ours. However, both approaches differ in the way they aim to empower the application
+programmer. \cite{Burckhardt2015} exposes the operational model to the
+programmer, along with primitives, such as \cf{push}, \cf{pull}
+and \cf{fetch}, that make it easy to \emph{implement} required
+consistency and isolation guarantees. In contrast, \name completely
+abstracts its operational model, and instead relies on contract
+classification to automatically \emph{choose} appropriate consistency and isolation
+levels. Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our
+system model does not incorporate real (i.e., wall-clock) time. Hence, our
+contract language cannot describe store semantics such as recency or
+bounded-staleness guarantees offered by certain stores~\cite{Pileus}.
 
@@
 replicated data object needs coordination or not. ~\cite{Calm} defines
-\emph{logical monotonicity} as a sufficient condition for coordination freedom,
-and proposes a consistency analysis that marks code regions performing
-non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as potential
-coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
+\emph{logical monotonicity} as a sufficient condition for coordination
+freedom, and proposes a consistency analysis that marks code regions
+performing non-monotonic reasoning (eg: aggregations, such as \cf{COUNT}) as
+potential coordination points. ~\cite{IConfluence} and ~\cite{Sieve} define
 \emph{invariant confluence} and \emph{invariant safety}, respectively, as
 conditions for safely executing an operation without coordination.
-~\cite{Sieve} also proposes a program analysis that conservatively marks certain
-operations as \emph{blue} (coordination not required), while marking the
-remaining as \emph{red} (coordination required). ~\cite{indigo} requires
+~\cite{Sieve} also proposes a program analysis that conservatively marks
+certain operations as \emph{blue} (coordination not required), while marking
+the remaining as \emph{red} (coordination required). ~\cite{indigo} requires
 programmers to declare application semantics, and the desired
-application-specific invariants as formulas in first-order logic. It performs
-static analysis on these formulas to determine $I$-offender sets - sets of
-operations, which, when performed concurrently, result in violation of one or
-more of the stated invariants. For each offending set of operation, if
-programmer chooses invariant-violation avoidance over violation repair, the
-system employs various techniques, such as escrow reservation, to ensure that
-the offending set of operations are effectively serialized. Unlike \name, these
-works focus on a coarse-grained classification of consistency as eventual or
-strong, and do not focus on finer-grained transaction isolation levels. However,
-the analyses that some of them propose relieve programmers of the burden to tag
-operations with consistency levels. Indeed, we do consider automatic inference
-of consistency contracts from application-specific integrity constraints as the
-next step for \name.
+application-specific invariants as formulas in first-order logic. It
+performs static analysis on these formulas to determine $I$-offender sets -
+sets of operations, which, when performed concurrently, result in violation
+of one or more of the stated invariants. For each offending set of
+operations, if the programmer chooses invariant-violation avoidance over
+violation repair, the system employs various techniques, such as escrow
+reservation, to ensure that the offending set is effectively
+serialized. Unlike \name, these approaches focus on a coarse-grained
+classification of consistency as eventual or strong, and do not consider
+finer-grained transaction isolation levels. However, these analyses do
+relieve programmers of the burden of tagging operations with consistency
+levels. We consider automatic inference of consistency contracts from
+application-specific integrity constraints as an obvious  next step for \name.

### 2015-04-10 implementation.tex
Minor edit.

@@
 This ensures that the state maintained by the shim layer node is causally
-consistent. Our dependence tracking strategy ensures that \name does not track
-every effect as the number of writes in the system grows.
+consistent. Hence, our dependence tracking strategy ensures that \name does not
+track every effect as the number of writes in the system grows.
 
@@
 included. Consider the example given in Figure~\ref{fig:atomicity_impl}. For
-every transaction in \name, we instantiate a special transaction marker
-effect $m$ that is importantly not inserted into the backing store. Marker $m$ is
+every transaction in \name, we instantiate a special transaction marker effect
+$m$ that is importantly not inserted into the backing store. Marker $m$ is
 included as a dependence to every effect generated in the transaction. In the
-figure, the top graph on the left shows the state of the store in the middle of a
-transaction. Each circle represents an effect. The dotted circle indicates that
-the effect has been instantiated, but has not yet been inserted into the store.
-Since the causally preceding effect $m$ has not yet been written to the store,
-no operation will witness $e1$ and $e2$ while the transaction in progress.
-After the transaction has finished execution, we insert $m$ into the backing
-store, marking all the effects from the transactions as a dependence for $m$.
-Now, any replica which includes one of the effects from the transaction must
-include $m$, and transitively must include every effect from the transaction.
-This ensures atomicity and satisfies the RC requirement.
+figure, the top graph on the left shows the state of the store in the middle of
+a transaction. Each circle represents an effect. The dotted circle indicates
+that the effect has been instantiated, but has not yet been inserted into the
+store. Since the causally preceding effect $m$ has not yet been written to the
+store, no operation will witness $e1$ and $e2$ while the transaction in
+progress. After the transaction has finished execution, we insert $m$ into the
+backing store, marking all the effects from the transactions as a dependence
+for $m$, as show in the bottom graph. Now, any replica which includes one of
+the effects from the transaction must include $m$, and transitively must
+include every effect from the transaction. This ensures atomicity and satisfies
+the RC requirement.
 

### 2015-04-13 contract-language.tex
Added references to the technical report

@@
 all well-formed \name contracts. The proof of the theorem is given in the
-supplementary material\footnote{The supplementary material also provides the
-concrete reduction rules for enforcing the consistency classes.}.
+accompanying technical report~\cite{techrep} \footnote{The technical report also
+provides the concrete reduction rules for enforcing the consistency classes.}.

### 2015-04-13 related.tex
"tech report" --> "technical report"

@@
 
-Our contract language and system model is inspired by the axiomatic
-description of RDT semantics proposed in~\cite{Burckhardt2014}. While
-~\cite{Burckhardt2014} uses axioms for formal verification of
-an RDT implementation, we utilize them as a means for the user to express
-desired consistency guarantees in the application. The operational semantics
-of \name (described in the accompanying tech report~\cite{techrep}) describes
-an operational manifestation of our system model. \cite{Burckhardt2015} also
-presents an operational model of a replicated data store that is based on
-the abstract system model presented in ~\cite{Burckhardt2014}. Their claims
-about the expressivity and practicality of the system model vindicate
-ours. However, both approaches differ in the way they aim to empower the application
-programmer. \cite{Burckhardt2015} exposes the operational model to the
-programmer, along with primitives, such as \cf{push}, \cf{pull}
-and \cf{fetch}, that make it easy to \emph{implement} required
-consistency and isolation guarantees. In contrast, \name completely
-abstracts its operational model, and instead relies on contract
-classification to automatically \emph{choose} appropriate consistency and isolation
-levels. Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our
-system model does not incorporate real (i.e., wall-clock) time. Hence, our
-contract language cannot describe store semantics such as recency or
-bounded-staleness guarantees offered by certain stores~\cite{Pileus}.
+Our contract language and system model is inspired by the axiomatic description
+of RDT semantics proposed in~\cite{Burckhardt2014}. While ~\cite{Burckhardt2014}
+uses axioms for formal verification of an RDT implementation, we utilize them as
+a means for the user to express desired consistency guarantees in the
+application. The operational semantics of \name (described in the accompanying
+technical report~\cite{techrep}) describes an operational manifestation of our
+system model. \cite{Burckhardt2015} also presents an operational model of a
+replicated data store that is based on the abstract system model presented in
+~\cite{Burckhardt2014}. Their claims about the expressivity and practicality of
+the system model vindicate ours. However, both approaches differ in the way they
+aim to empower the application programmer. \cite{Burckhardt2015} exposes the
+operational model to the programmer, along with primitives, such as \cf{push},
+\cf{pull} and \cf{fetch}, that make it easy to \emph{implement} required
+consistency and isolation guarantees. In contrast, \name completely abstracts
+its operational model, and instead relies on contract classification to
+automatically \emph{choose} appropriate consistency and isolation levels.
+Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our system model
+does not incorporate real (i.e., wall-clock) time. Hence, our contract language
+cannot describe store semantics such as recency or bounded-staleness guarantees
+offered by certain stores~\cite{Pileus}.
 

### 2015-04-15 contract-language.tex
Edits to figure atomicity implementation.

@@
 
+%  The above contract mandates that an effect $a$ must be visible to the current
+%effect $\cureff$ if some effect $b$ that causally succeeds $a$ is also visible
+%to $\cureff$. Thus, if every store replica always maintains and exposes a
+%causal cut of updates to the client, then such a store will satisfy this
+%contract. In such a system, an eventually consistent operation, such as
+%\rcf{deposit}, can be satisfied as long as some replica is reachable.
+
 \item \textbf{Causal consistency}: Causally consistent operations are required
@@
 \S~\ref{sec:motivation} as a {\sf\small CausallyConsistent} contract, because
-the sequent $\Delta \vdash \ccc \Rightarrow \cv_{gb}$ is valid in first-order
-logic (therefore, $\cv_{gb} \le \ccc$), whereas the sequent $\Delta \vdash \ecc
-\Rightarrow \cv_{gb}$ is invalid (therefore, $\cv_{gb} \not\le \ecc$). Since we
-confine our contract language to a decidable subset of the logic, validity
-of such sequents can be decided mechanically allowing us to automate the
-classification scheme in \name.
+the sequent $\Delta \vdash \forall \cureff. \ccc \Rightarrow \cv_{gb}$ is valid
+in first-order logic (therefore, $\cv_{gb} \le \ccc$), whereas the sequent
+$\Delta \vdash \forall \cureff. \ecc \Rightarrow \cv_{gb}$ is invalid
+(therefore, $\cv_{gb} \not\le \ecc$). Since we confine our contract language to
+a decidable subset of the logic, validity of such sequents can be decided
+mechanically allowing us to automate the classification scheme in \name.
 
@@
 \begin{array}{rcl}
-\R{Read Your Writes (RYW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \Rightarrow \vis{a}{b} \\
-\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \soo{b}{c} \\
- & & \quad \Rightarrow \vis{a}{c} \\
-\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b,c. ~\soo{a}{b} \wedge \vis{b}{c} \\
- & & \quad \Rightarrow \vis{a}{c} \\
-\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c,d. ~\vis{a}{b} \wedge \vis{c}{d} \\
- & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{d}
+\R{Read Your Writes (RYW)} & \coloneqq & \forall a. ~\soo{a}{\cureff} \Rightarrow \vis{a}{\cureff} \\
+\R{Monotonic Reads (MR)} & \coloneqq & \forall a,b. ~\vis{a}{b} \wedge \soo{b}{\cureff} \\
+ & & \quad \Rightarrow \vis{a}{\cureff} \\
+\R{Monotonic Writes (MW)} & \coloneqq & \forall a,b. ~\soo{a}{b} \wedge \vis{b}{\cureff} \\
+ & & \quad \Rightarrow \vis{a}{\cureff} \\
+\R{Writes Follow Reads (WFR)} & \coloneqq & \forall a,b,c. ~\vis{a}{b} \wedge \vis{c}{\cureff} \\
+ & & \wedge (\sooZ ~\cup =)(b,c) \Rightarrow \vis{a}{\cureff}
 \end{array}
@@
 upper level element to a lower level element corresponds to a weaker-than
-relation between the corresponding contracts. Classifying a contract under
-this scheme is a directed search in the lattice, starting from the bottom, and
+relation between the corresponding contracts. Classifying a contract under this
+scheme is a directed search in the lattice, starting from the bottom, and
 determining the weakest consistency level under which the contract can be
 satisfied. Under this scheme, \rcf{deposit} operations do not need any
-guarantees, \rcf{getBalance} needs RYW and WFR ($\Delta \vdash \R{RYW} \wedge
-\R{WFR} \Rightarrow \cv_{gb}$), and \rcf{withdraw} cannot be satisfied ($\Delta
-\vdash \R{RYW} \wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} \not\Rightarrow
-\cv_{w}$).
+guarantees, \rcf{getBalance} needs RYW and WFR ($\cv_{gb} \le \R{RYW} \wedge
+\R{WFR}$), and \rcf{withdraw} cannot be satisfied ($\cv_{w} \not\le \R{RYW}
+\wedge \R{MW} \wedge \R{MR} \wedge \R{WFR} $).
 

### 2015-04-15 contract-language.tex
Added explanation for EC contract.

@@
 
-%  The above contract mandates that an effect $a$ must be visible to the current
-%effect $\cureff$ if some effect $b$ that causally succeeds $a$ is also visible
-%to $\cureff$. Thus, if every store replica always maintains and exposes a
-%causal cut of updates to the client, then such a store will satisfy this
-%contract. In such a system, an eventually consistent operation, such as
-%\rcf{deposit}, can be satisfied as long as some replica is reachable.
+  The above contract mandates that an effect $a$ must be visible to the current
+  effect $\cureff$ if some effect $b$ that causally succeeds $a$ is also visible
+  to $\cureff$. Thus, if every store replica always maintains and exposes a
+  causal cut of updates to the client, then such a store will satisfy this
+  contract. In such a system, an eventually consistent operation, such as
+  \rcf{deposit}, which requires weaker guarantees than offered by the store, can
+  be satisfied as long as some replica is reachable.
 
@@
   \end{cmathpar}
+
 \end{itemize}

### 2015-04-15 contract-language.tex
Fixed Grammar.

@@
 	bank account example, \rcf{deposit} is an eventually consistent operation.
-	While an \ecds typically offer \emph{basic} eventual consistency with all
+	While an \ecds typically offers \emph{basic} eventual consistency with all
 	possible anomalies, we assume that our store provides stronger semantics that

### 2015-04-15 concl.tex

@@
 classification of fine-grained consistency contracts on operations and
-distributed transactions with respect to the consistency and isolation levels
+distributed transactions with respect various consistency and isolation levels
 offered by the store. Our contract language is carefully crafted from a
@@
 verification tools to discharge the proof obligations associated with contract
-classification. We realize an instantiation of \name on top of off-the-shelf
+classification. We realize an instantiation of \name on top of an off-the-shelf
 distributed store, Cassandra, and illustrate the benefit of fine-grained

### 2015-04-15 contract-language.tex

@@
 
-The theorem states that if a data store correctly enforces $\scc$, $\ccc$, and
-$\ecc$ contracts in all well-formed executions, then the same store, extended
-with the classification scheme shown in Figure~\ref{sem:classify}, can enforce
-all well-formed \name contracts. The proof of the theorem is given in the
-accompanying technical report~\cite{techrep} \footnote{The technical report also
-provides the concrete reduction rules for enforcing the consistency classes.}.
+\noindent The theorem states that if a data store correctly enforces $\scc$,
+$\ccc$, and $\ecc$ contracts in all well-formed executions, then the same
+store, extended with the classification scheme shown in
+Figure~\ref{sem:classify}, can enforce all well-formed \name contracts. The
+proof of the theorem is given below:
+
+\begin{proof}
+  Hypothesis:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \auxred{} {\E,\langle op,\tau \rangle;\sigma \pll \Sigma} {\eff}
+    {\E', \sigma \pll \Sigma} & H\npp\\
+    \E' \models \cv_\tau[\eff/\cureff] & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Since $\tau$ is the contract class of $\cv$, by inversion, we have
+  $\cv \le \cv_\tau$. By the definition of $\le$ relation:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \forall \cureff. \cv_\tau \Rightarrow \cv & H\npp\\
+  \end{array}
+  \end{smathpar}
+   Since $\eff$ denotes new effect, it is a fresh variable that does
+   not occur free in $\Delta$. From $H3$, after instantiating bound
+   $\cureff$ with $\eff$, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \Delta \vdash \cv_\tau[\eff/\cureff] \Rightarrow \cv[\eff/\cureff]
+      & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Due to the soundness of natural deduction for first-order logic,
+  $H4$ implies that for all models $\mathcal{M}$ such that
+  $\mathcal{M} \models \Delta$, if $\mathcal{M} \models
+  \cv_\tau[\eff/\cureff]$ then $\mathcal{M} \models
+  \cv[\eff/\cureff]$. Since $\E'$ is well-formed, we have:
+  \begin{smathpar}
+  \begin{array}{cr}
+    \E' \models \Delta & H\npp\\
+  \end{array}
+  \end{smathpar}
+  Proof follows from $H1$, $H4$, and $H3$.
+  \hfill \qed
+\end{proof}
+
+It is important to note that Theorem~\ref{thm:classification-sound}
+does not ascribe any semantics to the reduction relation
+($\xrightarrow{}$). As such, it makes no assumptions about how the
+store actually implements ${\sf ec}$, ${\sf cc}$ and ${\sf sc}$
+guarantees. The specific implementation strategy is determined by the
+operational semantics of the store, which \emph{defines} the reduction
+relation for that particular store. The following section describes
+operational semantics of the store used by the \name implementation.

### 2015-04-15 evaluation.tex
minor edit

--- 2015-04-08 evaluation.tex
@@
 that does not use transactions. Compared to this, RC is only 12\% shower in
-terms of latency with 512 clients, whereas RR is 2.3X slower. The difference
-between RC and NoTxn is due to the meta-data overhead of recording transaction
-information in the object state. For RR transactions, the cost of capturing and
-maintaining a snapshot is the biggest source of overhead.
+terms of latency with 512 clients, whereas RR is 2.3$\times$ slower. The
+difference between RC and NoTxn is due to the meta-data overhead of recording
+transaction information in the object state. For RR transactions, the cost of
+capturing and maintaining a snapshot is the biggest source of overhead.
 

### 2015-04-15 implementation.tex
Edits to Lattice figure. Edits to caption for atomicity implementation.

@@
 \begin{center}
-\includegraphics[width=\columnwidth]{Figures/AtomicityImpl}
+\includegraphics[width=0.8\columnwidth]{Figures/AtomicityImpl}
 \end{center}
-\caption{Implementing atomicity semantics. Dotted circle represents effects not
-yet inserted into the backing store.}
+\caption{Implementing atomicity semantics.}
 \label{fig:atomicity_impl}
@@
 
-\name implements atomic visibility by exploiting shim layer causality guarantees
--- an effect is included only if all the effects if depends on are also
-included. Consider the example given in Figure~\ref{fig:atomicity_impl}. For
-every transaction in \name, we instantiate a special transaction marker effect
-$m$ that is importantly not inserted into the backing store. Marker $m$ is
-included as a dependence to every effect generated in the transaction. In the
-figure, the top graph on the left shows the state of the store in the middle of
-a transaction. Each circle represents an effect. The dotted circle indicates
-that the effect has been instantiated, but has not yet been inserted into the
-store. Since the causally preceding effect $m$ has not yet been written to the
-store, no operation will witness $e1$ and $e2$ while the transaction in
-progress. After the transaction has finished execution, we insert $m$ into the
-backing store, marking all the effects from the transactions as a dependence
-for $m$, as show in the bottom graph. Now, any replica which includes one of
-the effects from the transaction must include $m$, and transitively must
-include every effect from the transaction. This ensures atomicity and satisfies
-the RC requirement.
+\name implements atomic visibility by exploiting shim layer causality
+guarantees -- an effect is included only if all the effects if depends on are
+also included. Consider the example given in Figure~\ref{fig:atomicity_impl}.
+In the figure, the graphs represent the state of the store, where the circles
+represent effects and the edges represent the dependence between effects. The
+dotted circle represents effects that are not yet inserted into the store.
+The graph on the left shows that state of the store after executing \cf{oper2}.
+For every transaction in \name, we instantiate a special transaction marker
+effect $m$ that is importantly not inserted into the backing store. Marker $m$
+is included as a dependence to every effect generated in the transaction. Since
+the causally preceding effect $m$ has not yet been written to the store, no
+operation will witness $e1$ and $e2$ while the transaction in progress. After
+the transaction has finished execution, we insert $m$ into the backing store,
+marking all the effects from the transactions as a dependence for $m$, as shown
+in the graph on the right. Now, any replica which includes one of the effects
+from the transaction must include $m$, and transitively must include every
+effect from the transaction. This ensures atomicity and satisfies the RC
+requirement.
 
@@
 \end{center}
-\caption{Summarization in the backing store. Dotted circle represents effects
-not yet inserted into the backing store.}
+\caption{Summarization in the backing store.}
 \label{fig:impl_sum_disk}

### 2015-04-15 related.tex

--- 2015-04-10 related.tex
@@
 
-Our contract language and system model is inspired by the axiomatic
-description of RDT semantics proposed in~\cite{Burckhardt2014}. While
-~\cite{Burckhardt2014} uses axioms for formal verification of
-an RDT implementation, we utilize them as a means for the user to express
-desired consistency guarantees in the application. The operational semantics
-of \name (described in the accompanying tech report~\cite{techrep}) describes
-an operational manifestation of our system model. \cite{Burckhardt2015} also
-presents an operational model of a replicated data store that is based on
-the abstract system model presented in ~\cite{Burckhardt2014}. Their claims
-about the expressivity and practicality of the system model vindicate
-ours. However, both approaches differ in the way they aim to empower the application
-programmer. \cite{Burckhardt2015} exposes the operational model to the
-programmer, along with primitives, such as \cf{push}, \cf{pull}
-and \cf{fetch}, that make it easy to \emph{implement} required
-consistency and isolation guarantees. In contrast, \name completely
-abstracts its operational model, and instead relies on contract
-classification to automatically \emph{choose} appropriate consistency and isolation
-levels. Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015}, our
-system model does not incorporate real (i.e., wall-clock) time. Hence, our
+Our contract language and system model is inspired by the axiomatic description
+of RDT semantics proposed in~\cite{Burckhardt2014}. While ~\cite{Burckhardt2014}
+uses axioms for formal verification of an RDT implementation, we utilize them as
+a means for the user to express desired consistency guarantees in the
+application. The operational semantics of \name describes an operational
+manifestation of our system model.~\cite{Burckhardt2015} also presents an
+operational model of a replicated data store that is based on the abstract
+system model presented in ~\cite{Burckhardt2014}. Their claims about the
+expressivity and practicality of the system model vindicate ours. However, both
+approaches differ in the way they aim to empower the application programmer.
+\cite{Burckhardt2015} exposes the operational model to the programmer, along
+with primitives, such as \cf{push}, \cf{pull} and \cf{fetch}, that make it easy
+to \emph{implement} required consistency and isolation guarantees. In contrast,
+\name completely abstracts its operational model, and instead relies on contract
+classification to automatically \emph{choose} appropriate consistency and
+isolation levels.  Similar to~\cite{Burckhardt2014} and~\cite{Burckhardt2015},
+our system model does not incorporate real (i.e., wall-clock) time. Hence, our
 contract language cannot describe store semantics such as recency or

### 2015-04-15 transactions.tex

@@
 to pick the correct transaction kind. This choice is further complicated by
-consistency semantics of individual operations.
+the consistency semantics of individual operations.
 
@@
 transaction isolation level. Similar to contracts on individual operations, the
-programmer associates contracts with transactions, declaratively expressing its
-consistency specification. We extend the contract language with a new term
+programmer associates contracts with transactions, declaratively expressing 
+consistency specifications. We extend the contract language with a new term
 under quantifier-free propositions - ${\small \txnZ}~S_1~S_2$, where $S_1$ and
@@
 that \cf{totalBalance} returns the result obtained from a consistent snapshot
-of the object states. The \name code for the transactions is given below:
+of the object states. The \name code for these transactions is given below:
 
@@
 While making both transactions serializable would ensure correctness,
-distributed stores rarely offer serializable transactions since it is
-unavailable and hinders scalability~\cite{BailisHAT}. As we will see, these
-transactions can be satisfied with much weaker isolation guarantees. Despite
-the atomicity offered by the transaction, anomalies are still possible. For
-example, the two \rcf{getBalance} operations in \cf{totalBalance} transactions
-might be served by different replicas with distinct set of committed \cf{save}
-transactions. If the first(second) \rcf{getBalance} operation witness a
+distributed stores rarely offer serializable transactions given their
+availability requirements and implications for
+scalability~\cite{BailisHAT}. Fortunately, these transactions can be
+satisfied with much weaker isolation guarantees. However, despite the atomicity
+offered by the transaction, anomalies are still possible. For example, the
+two \rcf{getBalance} operations in a \cf{totalBalance} transaction might be
+served by different replicas with a distinct set of committed \cf{save}
+transactions. If the first(second) \rcf{getBalance} operation witnesses a
 \cf{save} transaction that is not witnessed by the second(first)
-\rcf{getBalance} operation, then the balance returned will be less(greater) than
-the actual balance. It is not immediately apparent which weakest isolation
-guarantee will be sufficient to prevent the anomaly.
+\rcf{getBalance} operation, then the balance returned will be less(greater)
+than the actual balance. It is not immediately apparent how to choose the weakest
+isolation guarantee that would be sufficient to prevent the anomaly.
 
@@
 operations to witness the same set of \cf{save} transactions, we define the
-constraint on \cf{totalBalance} transaction $\psi_{tb}$ as:
+constraint on the \cf{totalBalance} transaction $\psi_{tb}$ as:
 
@@
 transaction, \cf{save} does not require any additional constraints, and
-$\psi_1$ is simply $\small \true$.
+$\psi_{sv}$ is simply $\small \true$.
 
@@
 and Repeatable Read (RR)~\cite{Berenson95}, and illustrate the classification
-strategy. Our technique can indeed be applied to a different isolation level
+strategy. Our technique can indeed be applied to a different isolation-level
 lattice.
@@
 updates from a transaction are available, the buffered updates are made visible
-to subsequent client requests. This ensure atomicity of transactions.
+to subsequent client requests. This ensures atomicity of transactions.
 Importantly, RC does not entail any other guarantees. As a result, a store
@@
 guarantee of transaction described in \S~\ref{sec:syn_sem_ext}. The \cf{save}
-is an example for RC transaction.
+operation is an example of an RC transaction.
 
@@
 store state. Importantly, this snapshot can be obtained from any replica, and
-hence RR is coordination-free. An example for such a transaction is the
+hence RR is also coordination-free. An example for such a transaction is the
 \cf{totalBalance} transaction. The semantics of RR is captured by the following
